### 2019-06-04 code/select_usa_aeronet.Rmd
from Nov. 2018

---
title: "Select Aeronet in USA"
author: "Johnathan Rush"
date: "11/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = 'hold')
```

```{r message=FALSE}
library(data.table)
library(rgdal)
library(sp)
library(rgeos)
library(raster)
library(rnaturalearth)
```

# Open Aeronet

This is using the updated AERONET sites.  
*Change the `aeronet_dir` to where you have the AERONET data downloaded to*

```{r aeronet2017}
aeronet_dir = "~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/aeronet/"
aeronetsites <- fread(paste0(aeronet_dir, "aeronet_locations_v3.txt"))
# rename columns to shorter names
setnames(aeronetsites, c("Site_Name", "lon", "lat", "elevm"))
rmarkdown::paged_table(aeronetsites)
```

Make the AERONET data a Spatial Points object
```{r aeronet_spatial}
aeronetData = aeronetsites[, .(Site_Name, elevm)]
#aeronetPoints = SpatialPoints(cbind(aeronetsites$lon, aeronetsites$lat))
aeronetPoints = SpatialPoints(aeronetsites[, .(lon, lat)])
aeronetSP = SpatialPointsDataFrame(aeronetPoints, aeronetData)
aeronetSP
```

Notice that we haven't set a Coordinate Reference System for the AERONET sites, yet.  
They're probably using WGS84 (the standard for GPS) datum lat/longs, so we'll set the CRS accordingly:

```{r aeronet_crs}
proj4string(aeronetSP) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

# Get a shape for the USA and for CONUS

First, download the Natural Earth data. This only has to be run once.  
*Change the `naturalEarthDir` to where you'd like to download this data.*

```{r set_ne_dir}
naturalEarthDir = "/data-belle/naturalearth/"
```
```{r dl_NE, eval=FALSE}
ne_download(scale = 10, type = "countries", category = "cultural", destdir = naturalEarthDir, load = FALSE)
```

Then load the countries file in as a Spatial Polygons object.  
```{r load_countries}
countriesSP = ne_load(scale = 10, type = "countries", category = "cultural", destdir = naturalEarthDir)
```

Plot all the aeronet stations worldwide
```{r plot_worldwide}
plot(countriesSP)
plot(aeronetSP, col = "red", add = TRUE)
```

Subset the countries to just the USA
```{r subset_usa}
usaSP = countriesSP[countriesSP$NAME == 'United States of America', "NAME"]
rm(countriesSP)
plot(usaSP)
```

Clip the USA down to just the CONUS - removing Alaska, Hawaii, PR, etc.  
```{r subset_conus}
clipmatrix = as.matrix(extent(usaSP))      # quick way to get a matrix with the right row/col names and dimensions
clipmatrix[] <- c(-126, 24.4, -66.7, 49.4) # change to actual coordinates of the CONUS bbox
clip_poly <- as(extent(as.vector(t(clipmatrix))), "SpatialPolygons") # CRS isn't set, but that's OK in this case
conusSP = gIntersection(usaSP, clip_poly, byid = TRUE)               # ignore the warning about different CRS
plot(conusSP)
```

# Select the AERONET stations in the USA

```{r aeronet_usa}
aeronet_USA = aeronetSP[usaSP, ]
aeronet_USA
plot(usaSP)
plot(aeronet_USA, col = "red", add = TRUE)
```

# Select the AERONET stations in the CONUS
```{r aeronet_conus}
aeronet_conus = aeronetSP[conusSP, ]
aeronet_conus
plot(conusSP)
plot(aeronet_conus, col = "red", add = TRUE)
```

# Export the AERONET selections

*All of these are set to export to the current working directory*. Change the `output_path` as needed.

After this are three different export verions. You may not need all of them.

```{r output_path}
output_path = getwd()
```

```{r hidden_path, echo=FALSE}
# this is just for the purpose of rendering the file to send to Meytar
output_path = "~/test"
```

First by ESRI Shapefile:  
```{r export_shapefile}
writeOGR(aeronet_USA, layer = "aeronet_usa", dsn = output_path, driver="ESRI Shapefile", overwrite_layer = TRUE)
writeOGR(aeronet_conus, layer = "aeronet_conus", dsn = output_path, driver="ESRI Shapefile", overwrite_layer = TRUE)
```

Then by Rds files for the SP objects:  
```{r export_rds}
saveRDS(aeronet_USA, file = paste0(output_path, "/aeronet_usa.Rds"))
saveRDS(aeronet_conus, file = paste0(output_path, "/aeronet_conus.Rds"))
```

Finally, by CSV files with the coordinates in the table:  
```{r export_csv}
aeronet_usa_DT = as.data.table(aeronet_USA)
aeronet_conus_DT = as.data.table(aeronet_conus)
fwrite(aeronet_usa_DT, file = paste0(output_path, "/aeronet_usa.csv"))
fwrite(aeronet_conus_DT, file = paste0(output_path, "/aeronet_conus.csv"))
```

### 2019-06-04 code/01_import_aeronet.R

@@ -1,0 +1,169 @@
+# import AERONET data
+library(drake)
+library(here)
+library(data.table)
+library(purrr)
+library(sp)
+library(maptools)
+library(dplyr)
+library(ggplot2)
+library(ggmap)
+library(raster)
+
+# https://aeronet.gsfc.nasa.gov/aeronet_locations_v3.txt
+
+# path to downloaded Aeronet data (e.g. version 3 level 2.0)
+# https://aeronet.gsfc.nasa.gov/cgi-bin/combined_data_access_v3
+
+plan <- drake_plan(
+  aeronet_stations = fread(file = file_in(here("data/aeronet_locations_v3.txt"))),
+  setnames(aeronet_stations, c("Site_Name", "lon", "lat", "elevm")),
+  
+  aeronet = fread(file_in('unzip -cq data/All_Sites_Times_All_Points_AOD20.zip')),
+  aeronet[, day := as.Date(strptime(`Date(dd:mm:yyyy)`, format = "%d:%m:%Y"))],
+  # pryr::object_size(aeronet)
+  # nrow(aeronet)
+  # subset by date
+  daterange = data.table(start = as.Date("2015-01-01"), end = as.Date("2018-11-01")),
+  aeronet_sub = aeronet[day >= daterange$start & day <= daterange$end, ],
+  # pryr::object_size(aeronet)
+  nrow(aeronet_sub)
+  
+)
+plan
+make(plan)
+
+config <- drake_config(plan)
+vis_drake_graph(config)
+
+
+
+aeronet[, day := as.Date(strptime(`Date(dd:mm:yyyy)`, format = "%d:%m:%Y"))]
+# pryr::object_size(aeronet)
+# nrow(aeronet)
+# subset by date
+daterange = data.table(start = as.Date("2015-01-01"), end = as.Date("2018-11-01"))
+aeronet = aeronet[day >= daterange$start & day <= daterange$end, ]
+# pryr::object_size(aeronet)
+# nrow(aeronet)
+
+
+setnames(aeronet, "Precipitable_Water(cm)", "aeronet_cwv") 
+
+
+##### old code ####
+
+aeronetfiles
+
+test <- fread("data/aeronet/AOT/LEV20/ALL_POINTS/920801_170325_GSFC.lev20")
+test
+dput(names(test))
+gsfc <- test[, c("Date(dd-mm-yy)", "Time(hh:mm:ss)", "Julian_Day", "AOT_440", "Water(cm)", "Solar_Zenith_Angle"), with = F]
[…]

### 2019-07-15 Code/03_import_AOD_V3.Rmd
mv files

---
title: "Import and process AODV3 Level2 AOD20 from Aeronet"
author: "Yang Liu"
date: "6/4/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
# error=TRUE allows error 
```

Borrowed code from `03_import_aeronet.R` on Jdrive  
Import and process the AOD20 V3 level 2 global Aeronet stations  
Interpolating to wavelength AOD 470nm

```{r, eval = F, include = F}
## About this script ##

```

```{r}
suppressPackageStartupMessages({
library(fst); library(data.table); library(ggplot2)
library(sf)
library(ggmap)
})

source("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/code/liuy29/funcs_for_plots.R")
source("~/liuyanguu/ECHO_PM/code/cwv_funcs.R")

```

# AERONET

Get list of stations, and NEMIA stations

```{r, eval = T}
# https://aeronet.gsfc.nasa.gov/data_push/V3/AOD/AOD_Level20_All_Points_V3.tar.gz
# load aod data
# aod20_daily_dir <- "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/DAILY/"
aod20_daily_dir <- "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
aod20_file_dir  <-  list.files(aod20_daily_dir, pattern = "*.lev20", full.names = TRUE)

# load site location file, only for limiting sites' locations to the US or to NEMIA
site_locations_2019 <- fread("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt")
aod_file_name <- gsub("^[0-9]{5,}_[0-9]{5,}_|\\.lev20", "", list.files(aod20_daily_dir, pattern = "*.lev20"))
aod_file_name[!aod_file_name%in%site_locations_2019$Site_Name]
# Among the 4 sites, Kitt-Peak_MP (Kitt Peak National Observatory) and UMBC_temp are in US.
setnames(site_locations_2019, c("Site_Name", "lon", "lat", "elev"))

aod_site_sf <- st_as_sf(site_locations_2019, coords = c("lon", "lat"), crs = 4326)
nemia_buff_aea <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/nemia_buff_aea.rds")
nemia_buff = st_transform(nemia_buff_aea, crs = 4326) 
aod_site_NE <- aod_site_sf[nemia_buff,]

# NEMIA site list
# site_list_NE <- c(aod_site_NE$Site_Name, "MVCO", "LISCO")
site_list_NE <- c(aod_site_NE$Site_Name)
# US site list
# to be done

```

create the 6.6GB global file

```{r, eval = F}
# open one file and take a look at the data
t0 <- fread(aod20_file_dir[3])
t0[1,]

grep("Angstrom_Exponent", names(t0), value = T)
# read one file
vars_aod <- intersect(grep("AOD_", names(t0), value = T), grep("nm", names(t0), value = T))
# sort the wave lengths from low to high, and update the vars_aod
x_nm <- sort(readr::parse_number(vars_aod))
vars_aod <- paste0("AOD_", x_nm,"nm")
vars_wv <- paste0("Exact_Wavelengths_of_AOD(um)_", x_nm,"nm")
vars0 <- c("Date(dd:mm:yyyy)", "Time(hh:mm:ss)", "Day_of_Year","AERONET_Site_Name",
             "AERONET_Instrument_Number", 
             "Site_Latitude(Degrees)", 
             "Site_Longitude(Degrees)", 
             "Site_Elevation(m)", "Ozone(Dobson)", "NO2(Dobson)")
vars0_new <- c("date", "time", "dayofYear", "site",
                "instrumentNumber", "lat", "lon", 
                 "elev", "ozone", "NO2")
read.aod <- function(site_dir, vars_read, vars_newname){
  t1 <- fread(site_dir, select = vars_read, col.names = vars_newname)
  t1[, date := as.Date(date, format = "%d:%m:%Y")]
  t1[, hour := format(strptime(time, "%H:%M:%S"), "%H")]
  for (i in names(t1)) t1[get(i) == -999, (i):= NA] # set NA 
  t1
}

system.time(
file_list <- lapply(aod20_file_dir, read.aod, vars_read = c(vars0,vars_aod,vars_wv),  vars_newname = c(vars0_new,vars_aod,vars_wv)) # 440s
)
aod20_data_global <- rbindlist(file_list)

print(object.size(aod20_data_global), unit = "GB") # 6.6 GB
write.fst(aod20_data_global,
          "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_global.fst",
          compress = 100)
```

## aod20_data_pred470: Interpolate AOD to 470nm wv

```{r, eval = F}
# can run from here
aod20_data_global <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_global.fst", as.data.table = T)
# overview: what percentage of wach wavelength is non-NA?
sapply(aod20_data_global[,..vars_aod], function(x)mean(!is.na(x)))

vars_aod <- grep("AOD_", names(aod20_data_global), value = T)
vars_wv <- grep("Exact_Wavelengths_of_AOD", names(aod20_data_global), value = T)
# The 5 main wavelengths
aod_main <- c("AOD_340nm", "AOD_380nm", "AOD_440nm", "AOD_500nm", "AOD_675nm")
vars_aod_sub <- vars_aod[1:22]
aod_num <- readr::parse_number(vars_aod_sub)
# Drop the last two wavelengths because they are frequently off the main trend lines
aod20_data_global <- aod20_data_global[, -c("AOD_1020nm", "AOD_1640nm"), with = F]

# % missing
# sapply(aod20_data_global[,..vars_aod_sub], function(x)mean(!is.na(x)))
# Number of AOD < 0 --> replace <0 AOD with NA
# sapply(aod20_data_global[,..vars_aod_sub], function(x)sum(x<0, na.rm=T))
aod20_data_global[, (vars_aod_sub):= lapply(.SD, function(x) replace(x, which(x<=0), NA)), .SDcol = vars_aod_sub]

# set id for obs and count non-missing AOD by row 
aod20_data_global[,obs:=.I]
setkey(aod20_data_global, obs)
aod20_data_global$N_NAAOD <- rowSums(!is.na(aod20_data_global[,..vars_aod_sub]))
# The number of available AOD to each line and their shares among all the observations:
round(table(aod20_data_global$N_NAAOD)/aod20_data_global[,.N]*100,3)
# notice that 81% have 6 non-NA points
#     1      2      3      4      5      6      7      8      9     10 
# 0.013  0.235 10.577  1.224  4.412 81.139  2.179  0.000  0.001  0.221

## Fit a poly(2) regression based on Kodi's code
# DT.m2 = melt(DT, measure = list(colA, colB), value.name = c("dob", "gender"))
# subset to aods and obs column and melt
# melt two groups of columns (AOD and Exact wavelength) in the same time: 
d <- melt(aod20_data_global[, c(vars_aod_sub,vars_wv[1:22], "obs"), with = F], measure = list(vars_aod_sub, vars_wv[1:22]),value.name = c("aod", "exact_wv"))
# choose non-NA and >0 AOD
d <- d[!is.na(aod)]
# d[, variable := readr::parse_number(levels(variable))[as.integer(variable)]]
# should have points on both sides, otherwise pick out the id
obs_os <- d[, by = obs, .(oneside = (max(exact_wv)-470)*(min(exact_wv)-470)>0)][oneside ==T, obs]
saveRDS(obs_os, "~/Intermediate/AOD_data/190621_obs_same_side")
length(obs_os)/aod20_data_global[,.N] # about 0.4% have points on one side of the 470 nm

### fit the model 
time0 <- Sys.time()
system.time(d2 <- d[, by = obs, .(AOD_470nm = exp(predict(lm(log(aod) ~  log(exact_wv) + I((log(exact_wv))^2)), newdata = data.frame(exact_wv = 470))))])
Sys.time() - time0 # 6.3 hours 6/19

setkey(d2, obs)
aod20_data_pred470 <- d2[aod20_data_global]

# If there are at least 4 non-NA, even on one side of the 470, I consider it fine,
# If there are 2 or 3 points, one side is not fine
# 1 points are set to NA, all of them are in obs_os as well, since there is only one point
length(obs_os) # 91911 
aod20_data_global[N_NAAOD <=4 & obs%in%obs_os,.N] # 91553, so most <=3 are on the same side of 470 nm anyway 
aod20_data_global[N_NAAOD ==3 & obs%in%obs_os,.N]

# because for those on the one side and requires long range extrapolating: need 5 or 6 points to be valid. For those with at least 5 points, most have close points within the allowed range: "AOD_440nm" "AOD_443nm" "AOD_490nm" "AOD_500nm"
allowed_range <- paste0("AOD_", aod_num[440<=aod_num & aod_num<=500],"nm")
aod20_data_global[obs%in%obs_os &is.na(AOD_440nm)&is.na(AOD_443nm)&is.na(AOD_490nm) &is.na(AOD_500nm), table(N_NAAOD)]
 # exterpolating with 5 or 6 might be fine 
 #    1     2     3     4     5     6 
 # 3332 37551   183    54     4     8 
aod20_data_pred470[N_NAAOD<=4 & obs%in%obs_os, AOD_470nm:=NA]
aod20_data_pred470 <- aod20_data_pred470[,.(date, time, hour, dayofYear, site, 
                                            lat, lon, N_NAAOD, AOD_470nm, AOD_440nm, AOD_500nm)]
date0 <- format(Sys.Date(), "%y%m%d")
# save results 
write.fst(aod20_data_pred470,          paste0("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAODat470nm_",date0, ".fst"), compress = 100)

```

AOD is projected to 470nm, as expected it correlates with neighbouring AOD 440nm

```{r}
# view correlation 
aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)

aod20small <- aod20_data_pred470[sample(1:nrow(aod20_data_pred470), nrow(aod20_data_pred470)/100),]
plot.scatter.diagonal(aod20small, "AOD_440nm", "AOD_470nm")
# plot.scatter.diagonal(aod20small, "AOD_500nm", "AOD_470nm")

```

# create aod20_NEMIA_idLST, link Aeronet AOT to MAIAC grid (MCD19A2)

* created aod20_NEMIA_idLST that has AERONET Site_Name, idLSTpair0

```{r, eval = F}
# NEMIA with prediction in 2018
aod20_NEMIA <- aod20_data_pred470[year(date)==2018][site%in%site_list_NE]

# load the new grid of idLST and WGS
reference_lst_fst_old = "/data-belle/LST/MODIS.LST.C6/derived/conus_lst_land_glakes_2km_tiles.fst"
reference_lst_fst ="/data-belle/LST/MODIS.LST.C6/derived/conus_grid_201906.fst"
refDT = read_fst(reference_lst_fst, as.data.table = TRUE)[inNEMIA ==TRUE]

# match aod site to grid (LST id)
# aod_site_NE was produced in chuck 1
refDT_sf <- st_as_sf(refDT, coords = c("x_wgs84", "y_wgs84"), crs = 4326)
library(nngeo)
aod_site_NE_idLST <- st_join(aod_site_NE, refDT_sf, st_nn, k = 1)

setkey(aod20_NEMIA, site)
setkey(setDT(aod_site_NE_idLST), Site_Name)

# get data 2018 NEMIA aod with lst id 
aod20_NEMIA_idLST <- aod_site_NE_idLST[,.(Site_Name, idLSTpair0, x_sinu, y_sinu)][aod20_NEMIA]
aod20_NEMIA_idLST[,N_NAAOD:=NULL]
write.fst(aod20_NEMIA_idLST, "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst")
aod20_NEMIA_idLST[,unique(Site_Name)]
```

# Join AOT to MAIAC (MCD19A2)

* rolling join with nearest time (join_time) and by (idLSTpair0)

```{r, eval = F}
aod20_NEMIA_idLST <- read.fst("~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst",as.data.table = T)
# AERONET: All times are provided as Greenwich Mean Time (GMT) for the solar day
# https://aeronet.gsfc.nasa.gov/new_web/data_description_AOD_V2.html
aod20_NEMIA_idLST[, stn_time:=as.POSIXct(paste(date, time), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")]
aod20_NEMIA_idLST[, c("date", "time"):=NULL]

lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = "mcd19_T", full.names = T)
terra0 <- read.fst(lst_files[1], as.data.table = T)
# load lst data for 2018

read.lst <- function(sat = "terra"){
  if (sat != "terra") choose = "mcd19_T" else choose = "mcd19_A" # load terra by default
  lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = choose, full.names = T)
  dt = rbindlist(lapply(lst_files, function(x)read.fst(x,as.data.table = T)))
  dt[,c("x","y", "inNEMIA"):=NULL]

  dt[, join_time:=overpass_time]
  setkey(dt, idLSTpair0, join_time)
  return(dt)
}
lst_terra <- read.lst()

# make sure both are UTC time 
# attr(aod20_NEMIA_idLST$stn_time, "tzone")
# attr(lst_terra$overpass_time, "tzone")

aod20_NEMIA_idLST[, join_time:=stn_time]
setkey(aod20_NEMIA_idLST, idLSTpair0, join_time)

# rolling join (inner join)
aod_lst_terra <- aod20_NEMIA_idLST[lst_terra, roll = 'nearest', nomatch = 0]

# time difference
aod_lst_terra[, diff_time_min:= as.numeric(overpass_time - stn_time)/60]
aod_lst_terra[, within30min_t:=0]
aod_lst_terra[abs(diff_time_min)<=30, within30min_t:=1]
table(aod_lst_terra$within30min_t)
# hist(aod_lst_terra[abs(diff_time_min)<=60, diff_time_min], 20)
hist(aod_lst_terra[within30min_t==1, diff_time_min])
mean(aod_lst_terra$within30min_t)

vars0 <- c("overpass_index", "Optical_Depth_047", "AOD_Uncertainty", "Column_WV", "AOD_QA",  "RelAZ")
vars_new <- paste0(vars0,"_t")
setnames(aod_lst_terra, vars0, vars_new)
# set NA 
aod_lst_terra[within30min_t==0, c(vars_new):= NA]

# Aqua  ~~~~ #### 
lst_aqua <- read.lst("aqua")
# rolling join
aod_lst_aqua <- aod20_NEMIA_idLST[lst_aqua, roll = 'nearest', nomatch = 0]
aod_lst_aqua[,.N]

# time difference
aod_lst_aqua[, diff_time_min:= as.numeric(overpass_time - stn_time)/60]
vars_new <- paste0(vars0,"_a")
setnames(aod_lst_aqua, vars0, vars_new)
aod_lst_aqua[abs(diff_time_min)>30, c(vars_new):= NA]

# save
saveRDS(list(aod_lst_terra = aod_lst_terra, aod_lst_aqua=aod_lst_aqua), "~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")

```

## Examine the collocated dataset

* Terra  
```{r}
# report
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra

cat("The collocated Aeronet-MAIAC Terra dataset has",aod_lst_terra[,.N],"obs\n",
    "Then limit to within 30 minutes, non-missing MAIAC AOD data have:", aod_lst_terra[!is.na(Optical_Depth_047_t),.N],"obs \n",
    "The RMSE is ", aod_lst_terra[!is.na(Optical_Depth_047_t), round(sqrt(mean((Optical_Depth_047_t-AOD_470nm)^2)),4)],"\n")

cat("Diff_AOD: MAIAC - AERONET: \n")
aod_lst_terra[, AOD_diff := Optical_Depth_047_t - AOD_470nm]
summary(aod_lst_terra[!is.na(Optical_Depth_047_t),AOD_diff])

rmarkdown::paged_table(
aod_lst_terra[!is.na(Optical_Depth_047_t),
                   .(Site_Name, stn_time, overpass_time, diff_time_min, AOD_470nm, Optical_Depth_047_t)])

# AOD diff does not associate with time difference 
# plot.scatter(aod_lst_terra[!is.na(Optical_Depth_047_t)], "diff_time_min", "AOD_diff")

```

Plot MAIAC vs. AERONET AOD470nm

```{r}
plot.scatter.diagonal(data = aod_lst_terra[abs(diff_time_min)<=30], x = "AOD_470nm", 
                      y = "Optical_Depth_047_t")
```

Plot AERONET AOD by time    
Plot the difference by time   
```{r}
plotly::ggplotly(
ggplot(data = aod_lst_terra, aes(x = stn_time, y = AOD_470nm, color = Site_Name)) + 
  geom_point(size = 0.5, alpha = 0.8) +
  theme_bw()
)

plotly::ggplotly(
ggplot(data = aod_lst_terra, aes(x = stn_time, y = AOD_diff, color = Site_Name)) + 
  geom_point(size = 0.5, alpha = 0.8) +
  theme_bw()
)

```

* Aqua

```{r}
aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
cat("The collocated Aeronet-MAIAC Terra dataset has",aod_lst_terra[,.N],"obs\n",
    "Then limit to within 30 minutes, non-missing MAIAC AOD data have:", aod_lst_aqua[!is.na(Optical_Depth_047_a),.N],"obs \n",
        "The RMSE is ", aod_lst_aqua[!is.na(Optical_Depth_047_a), round(sqrt(mean((Optical_Depth_047_a-AOD_470nm)^2)),4)],"\n")

cat("Diff_AOD: MAIAC - AERONET: \n")
aod_lst_aqua[, AOD_diff := Optical_Depth_047_a - AOD_470nm]
summary(aod_lst_aqua[!is.na(Optical_Depth_047_a),AOD_diff])

plot.scatter.diagonal(data = aod_lst_aqua[abs(diff_time_min)<=30], x = "AOD_470nm", 
                      y = "Optical_Depth_047_a")

```

# View the AOD data in US

* limit to after 2000-01-01

```{r}
aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)
# * limit to after 2000-01-01 (98% are after 2000-01-01)
aod20_data_pred470 <- aod20_data_pred470[date>as.Date("2000-01-01"),]
aod20_data_pred470[,uniqueN(site)] # 1151 globally

aod20_data_pred470[, `:=`(AODmean= mean(AOD_470nm, na.rm = T), dailyN = .N), by = .(site, date)]

# mean AOD by site and date 4% NA
aod20 <- na.omit(unique(aod20_data_pred470[,.(site, lon, lat, date, AODmean, dailyN)]))
# # of days and range of days
aod20[,`:=`(date1 = min(date), date2 = max(date), Ndays = .N), by = site] # 807690

aeronetdt <- unique(aod20[,.(site, lon, lat, date1, date2, Ndays)]) # 1151 sites
sat = "AERONET"
plot.us.map.size()

aeronetdt_sf <- sf::st_as_sf(aeronetdt, coords = c("lon", "lat"), crs = 4326)
mapview::mapview(aeronetdt_sf, zcol = "Ndays", layer.name = "N_days")

cat("limit to >1000 days: \n")
mapview::mapview(aeronetdt_sf[aeronetdt_sf$Ndays>1000,], zcol = "Ndays", layer.name = "N_days [>1K]")

cat("range of date for Mexico City: \n")
range(aod20[site == "Mexico_City", date])
# mexi <- aod20_data_pred470[site == "Mexico_City"]

```

### 2019-07-25 Code/03_import_AOD_V3.Rmd
calc. AOD buffer variables

@@
                       error = F, fig.width = 7, fig.height = 4.5)
-# error=TRUE allows error 
 ```
 
-Borrowed code from `03_import_aeronet.R` on Jdrive  
+Borrowed some code from `03_import_aeronet.R` on Jdrive  
 Import and process the AOD20 V3 level 2 global Aeronet stations  
@@
 ```{r}
-suppressPackageStartupMessages({
-library(fst); library(data.table); library(ggplot2)
-library(sf)
-library(ggmap)
-})
-
-source("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/code/liuy29/funcs_for_plots.R")
-source("~/liuyanguu/ECHO_PM/code/cwv_funcs.R")
+source("~/liuyanguu/CWV/Code/cwv_funcs.R")
+source(here::here("Code", "funcs_for_plots.R"))
 
@@
 nemia_buff = st_transform(nemia_buff_aea, crs = 4326) 
-aod_site_NE <- aod_site_sf[nemia_buff,]
-
+aod_site_NE <- aod_site_sf[nemia_buff,]  # a sf object for NEMIA
+saveRDS(aod_site_NE, here("Intermediate", "aod_site_NE"))
 # NEMIA site list
@@
 
-### fit the model 
+```
+
+## project to 470 nm
+
+```{r, eval = F}
 time0 <- Sys.time()
@@
 
+## this part obtain the binded NEMIA MCD19 data 
 lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = "mcd19_T", full.names = T)
-terra0 <- read.fst(lst_files[1], as.data.table = T)
+terra0 <- read.fst(lst_files[1], as.data.table = T, columns = NULL)
+
 # load lst data for 2018
-
-read.lst <- function(sat = "terra"){
+read.lst <- function(sat = "terra", columns0 = NULL){
   if (sat != "terra") choose = "mcd19_T" else choose = "mcd19_A" # load terra by default
   lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = choose, full.names = T)
-  dt = rbindlist(lapply(lst_files, function(x)read.fst(x,as.data.table = T)))
-  dt[,c("x","y", "inNEMIA"):=NULL]
+  dt = rbindlist(lapply(lst_files, function(x)read.fst(x, as.data.table = T, columns  = columns0)))
+  if(is.null(columns0)) dt[,c("x", "y", "inNEMIA"):=NULL]
 
@@
 lst_terra <- read.lst()
-
+print(object.size(lst_terra), unit = "GB") # 18 GB all the lst terra in 2018 (NEMIA, one year)
 # make sure both are UTC time 
@@
 
-vars0 <- c("overpass_index", "Optical_Depth_047", "AOD_Uncertainty", "Column_WV", "AOD_QA",  "RelAZ")
-vars_new <- paste0(vars0,"_t")
-setnames(aod_lst_terra, vars0, vars_new)
+# vars0 <- c("overpass_index", "Optical_Depth_047", "AOD_Uncertainty", "Column_WV", "AOD_QA",  "RelAZ")
+# vars_new <- paste0(vars0,"_t")
+# setnames(aod_lst_terra, vars0, vars_new)
 # set NA 
-aod_lst_terra[within30min_t==0, c(vars_new):= NA]
+aod_lst_terra <- aod_lst_terra[within30min_t!=0,]
 
@@
 aod_lst_aqua[, diff_time_min:= as.numeric(overpass_time - stn_time)/60]
-vars_new <- paste0(vars0,"_a")
-setnames(aod_lst_aqua, vars0, vars_new)
-aod_lst_aqua[abs(diff_time_min)>30, c(vars_new):= NA]
+# vars_new <- paste0(vars0,"_a")
+# setnames(aod_lst_aqua, vars0, vars_new)
+aod_lst_aqua <- aod_lst_aqua[abs(diff_time_min)<=30,]
 
 # save
-saveRDS(list(aod_lst_terra = aod_lst_terra, aod_lst_aqua=aod_lst_aqua), "~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")
+saveRDS(list(aod_lst_terra = aod_lst_terra, aod_lst_aqua = aod_lst_aqua), "~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")
 
@@
 * Terra
+
 ```{r}

### 2019-07-25 Code/04_AOD_join_MAIAC.Rmd
calc. AOD buffer variables

---
title: "AOD collocated with MAIAC"
author: "Yang Liu"
date: "07/16/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
```

# Derive spatial variables within buffers
  st_buffer() to create circles around the aeronet stations (radius of 300km)
  Keep track of the AERONET ID, long/lat of station
  Intersect with base grid; but note that areas off the grid (e.g. ocean) can lead to very different numbers of cells within a buffer
  Long data.table with all pairLSTid within buffer per aeronet station
  Calculate distance from each grid centroid back to station
  Calculate variables at four buffer distances of 10km, 30km, 90km, and 270km
  - % non-missing
  -   Mean AOD
  -   AOD - Mean AOD

```{r}
source(here::here("Code", "funcs_for_plots.R"))
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
```

```{r}
# Terra sites in 2018
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
site_sub <- unique(aod_lst_terra$Site_Name)
aod_site_NE <- readRDS(here("Intermediate", "aod_site_NE"))
# reproject to US atlas projection, so now the unit is meter 
aod_site_km <- st_transform(aod_site_NE[aod_site_NE$Site_Name%in%site_sub,], crs = 2163) 
st_crs(aod_site_km)

# Define radius (meter): 
radius0 = 300000
reference_lst_fst ="/data-belle/LST/MODIS.LST.C6/derived/conus_grid_201906.fst"
refDT = read_fst(reference_lst_fst, as.data.table = TRUE)[inNEMIA ==TRUE]
refDT_sf <- st_as_sf(refDT, coords = c("x_wgs84", "y_wgs84"), crs = 4326)
refDT_sf <-  st_transform(refDT_sf, crs = 2163)

# take each row (sf, df, geometry point) as the argument, create multi-point buffer 
# `aod_buffer` is a multi-buffer:
aod_buffers <- st_buffer(aod_site_km, radius0)
# mapview(aod_buffers) # if you plot this, you will see lots of circles (polygons)
# but of course, after intersect with the base grid, they won't be circles anymore. 
```

## Method

Note 7/22, for sure not the best way here, but it can work. 
1. Get geometry for buffers as a list: aod_buffers_list  
2. `st_intersects` buffers to reference grid, get a binary geometry list: join_list
3. get list of dataset of grid within each buffer: ref_list
4. Get geometry for points (site location) as a list: site_list2
5. Within each ref_list (a pie of points), Use `mapply` or a for loop to calculate distance from each grid point to the center (station)

@output: refDT_sub, AERONET sites matched to all the grid ids around it within the 300KM round buffer

```{r}
# 1. 
aod_buffers_list <- st_geometry(aod_buffers)
# 2. 
# join_list is a sparse geometry binary predicate lists, made of (empty) or 0L https://r-spatial.github.io/sf/reference/sgbp.html
join_list <- lapply(aod_buffers_list, function(x)st_intersects(refDT_sf, x))
# 3. 
# list of joined I (ID) for each Aeronet station in the grid `refDT_sf`
join_list_I <- lapply(join_list, function(x) sapply(x, function(z) if (length(z)==0) NA_integer_ else z[1]))
# ref_list is the grid points within each circle (NA removed)
ref_list <- lapply(join_list_I, function(x) refDT_sf[!is.na(x),])
# dist_list <- st_distance(ref_list, site_list[1:2])
# 4.   
site_list <- st_geometry(aod_site_km)  # for calculating distance 
# get level 1 list [i], level2 [[i]] is just a point pair
site_list2 <- lapply(1 : length(site_list), function(i) (site_list[i])) 
name_list <- as.list(aod_site_km$Site_Name) # for recording site names 
# 5. 
# select ref.grid subset by Index for each station
select.refDT.subset <- function(ref_list0, point0, name0){
  # ref_list0 = ref_list[[1]]
  # point0 = site_list2[[1]]; name0 = name_list[[1]]
  # cat(paste0("crs of ref grid is ", st_crs(ref_list0)$epsg,"; ",
  #            "crs of point is ", st_crs(point0)$epsg,"\n"))
  ref_list0$dist <- st_distance(ref_list0, point0)
  ref_list0$Site_Name <- name0
  return(as.data.table(ref_list0))
}

# # not sure why mapply doesn't work 
# refDT_sub_list <- mapply(select.refDT.subset, ref_list0 = ref_list[1], point0 = site_list2[1], name0 = name_list[1])

refDT_sub_list <- list()
for (i in 1:length(ref_list)){
  ref_list0 = ref_list[[i]];  point0 = site_list2[[i]]; name0 = name_list[[i]]
  ref_list0$dist <- st_distance(ref_list0, point0)
  ref_list0$Site_Name <- name0
  refDT_sub_list[[i]] <- ref_list0
}
refDT_sub <- rbindlist(refDT_sub_list)
refDT_sub[, dist_km:=as.numeric(dist)/1000] # change distance to km 
refDT_sub <- setDT(refDT_sub)[,.(Site_Name, idLSTpair0, dist_km)]
setnames(refDT_sub, "idLSTpair0", "idLSTpair0_in_buffer")
setkey(refDT_sub, Site_Name)
# save the dataset that links station to all the idLST in the round buffer
# refDT_sub: 3545991 x 3
saveRDS(refDT_sub, "~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")

# view a circle
refDT_Wallop <- refDT_sub_list[[2]]
mapview::mapview(refDT_Wallop[sample(1:nrow(refDT_Wallop), nrow(refDT_Wallop)/100),])

```

## Calculation

@output new aod_lst_terra and aod_lst_aqua 
The collocated Aeronet with MODIS sites, plus the new variables

```{r}
refDT_sub <- readRDS("~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")
# load the aeronet join MCD19
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
# the original MCD19 
lst_terra <- read.lst(sat = "terra", columns0 = c("idLSTpair0", "overpass_time", "Optical_Depth_047"))

# remove some variables 
aod.MODIS.newVars <- function(aod_join_MODIS, MODIS_all){
  aod_join_MODIS <- aod_join_MODIS[!is.na(Optical_Depth_047),]
  aod_join_MODIS[, day:=format(stn_time, "%Y-%m-%d")]
  var_time <- grep("time", names(aod_join_MODIS), value = T)
  aod_join_MODIS <- aod_join_MODIS[, -..var_time]

  setkey(aod_join_MODIS, Site_Name)
  aod_join_buffer <- aod_join_MODIS[refDT_sub, allow.cartesian = T]
  cat("The cartesian joined dataset between collocated Aeronet with MODIS sites data and base grid in each circles, dim is: ", paste(dim(aod_join_buffer),collapse = " x "))

  # join back to the MODIS file 
  MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"),
           skip_absent = T)
  setkey(MODIS_all, idLSTpair0_in_buffer, day)
  setkey(aod_join_buffer, idLSTpair0_in_buffer, day)
  # aod_join_buffer2 get AOD for all the buffer grid cells
  aod_join_buffer2 <- MODIS_all[,.(idLSTpair0_in_buffer, day, AOD_470_buffer)][aod_join_buffer]

  dist0 <- c(10, 30, 90, 270)
    calculate.vars.in.buffer <- function(distx){
      aod_join_buffer_new <- aod_join_buffer2[dist_km < distx, .(nonmissing = mean(!is.na(AOD_470_buffer)), AOD_buffer_mean = mean(AOD_470_buffer, na.rm = T)),
                              by =.(idLSTpair0, day)]
      setnames(aod_join_buffer_new, c("nonmissing", "AOD_buffer_mean"), 
               paste0(c("pNonNAAOD", "Mean_AOD"),distx, "km"))
      setkey(aod_join_buffer_new, idLSTpair0, day)
      return(aod_join_buffer_new)
    }
  new_varlist <- invisible(lapply(dist0, calculate.vars.in.buffer))
  new_vars <- Reduce(merge, new_varlist)
  setkey(new_vars, idLSTpair0, day)
  setkey(aod_join_MODIS, idLSTpair0, day)
  aod_join_MODIS <- new_vars[aod_join_MODIS]
  return(aod_join_MODIS)
}
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
# AOD - AOD mean, do later 
saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")

```

```{r}
# some summary
library(ggplot2)
plot.scatter.diagonal(aod_lst_terra, "Mean_AOD10km", "AOD_470nm")
cor(as.matrix(aod_lst_terra[,c(grep("Mean_AOD", names(aod_lst_terra), value = T), "AOD_470nm"), with = F]))
pmissing <- grep("pNon", names(aod_lst_terra), value = T)
summary(aod_lst_terra[,..pmissing])
```

```{r, eval = F}
# backup
# view the circle (buffer)
library(mapview)
mapview(st_buffer(aod_site_km[2,], radius0))

# distance is calculated differently under different project
st_distance(refDT_sf3[10,], aod_site_km[1,])
st_distance(st_transform(refDT_sf3[10,], crs= 4326), st_transform(aod_site_km[1,], crs= 4326))
geosphere::distm(as_Spatial(st_transform(refDT_sf3[10,], crs= 4326)), as_Spatial(aod_site_NE[1,]))

# An alternative way to join:
# refDT_sp <- as_Spatial(refDT_sf2)
# buf_poly <- as_Spatial(aod_buffer)
# sp_grid <- over(refDT_sp, buf_poly)

```

### 2019-07-25 Code/funcs_for_plots.R

@@ -1,0 +1,149 @@
+# Functions for AOD
+# Yang Liu
+
+
+# for data ----------------------------------------------------------------
+
+
+read.lst <- function(sat = "terra", columns0 = NULL){
+  if (sat != "terra") choose = "mcd19_T" else choose = "mcd19_A" # load terra by default
+  lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = choose, full.names = T)
+  dt = rbindlist(lapply(lst_files, function(x)read.fst(x, as.data.table = T, columns  = columns0)))
+  if(is.null(columns0)) dt[,c("x", "y", "inNEMIA"):=NULL]
+  
+  dt[, join_time:=overpass_time]
+  setkey(dt, idLSTpair0, join_time)
+  return(dt)
+}
+
+
+
+#  general functions for making plots -------------------------------------
+
+
+# Scatterplot -------------------------------------------------------------
+
+plot.scatter <-  function(data, x, y, size0 = 0.1, alpha0 = 0.3, dilute = F, dilute_by = 10){
+  set.seed(1234)
+  if (dilute) {
+    data <- data[sample(nrow(data), nrow(data)/dilute_by)] # dilute
+  }
+  plot0 <- ggplot(data = data, aes(x = data[[x]], y = data[[y]]))+
+    # geom_point(size = size0, alpha = alpha0) + 
+    geom_density_2d(aes(fill = ..level..), geom = "polygon") + 
+    labs(x = x, y = y) + 
+    theme_bw()
+  # add histogram
+  ggExtra::ggMarginal(plot0, type = "histogram", bins = 50, size = 10, color="white")
+}
+
+ 
+# Maps --------------------------------------------------------------------
+# us map using ggmap
+plot.us.ggmap <- function(data0){
+  # data0 contains the points to plot, plot on whole us mainland
+  map_wholeUS <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/map_us_all.RData")
+  ggmap(map_wholeUS) + 
+    geom_point(data = data0, aes(x = lon, y = lat),size = 1, color= "red") + # MOD
+    # geom_label(data = SiteName_aodid, aes(label = Site_Name), size = 2, alpha = 0.7) + 
[…]

### 2019-07-29 Code/03.2_AOD Variation.Rmd
AERONET daily error

---
title: "AODV3 Level2 AOD2.0 from Aeronet: Day Variation"
author: "Yang Liu"
date: "7/24/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 8, fig.height = 4)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
source(here::here("Code", "funcs_for_plots.R"))
# compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit#

```

# Comparing within day variation of AERONET sites AOD measurements

@output dataset has dailyN: observations with in a day (at least 20)
@d_mean and sd_rmse are daily mean and sd

```{r}
# US site list
aeronet_conus_dt <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/aeronet_conus_dt.rds")
aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)

# year > 2000
aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
# aod_conus <- aod_conus[year(date)>2000]
aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
aod_conus[,dailyN:=.N, by = .(site, date)]
aod_conus <- aod_conus[dailyN>20,]

aod_conus[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod_conus[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod_conus[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod_conus[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod_conus[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod_conus[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

# summary1: aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[,.(lat, lon, dailyN, season, mean_aod = mean(AOD_470nm), 
                                           sd_rmse = sd(AOD_470nm)), by = .(site, date)])

```

## daily SD AOD generally follows daily mean AOD, more outliers in the summer

Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date.  
Plot by time gives this figure:

```{r}
# show daily trend: AOD vs date, all data
ggplot(data = aod1, aes(x = date, y = sd_rmse, color = site)) +
  geom_point(size = 1) + 
  theme_classic() + theme(legend.position = "none") + 
  labs(y = "Daily SD(AOD) by station", x = "Date") + 
  ggtitle("Daily variation by station across time", subtitle = "(# Daily obs >20)")

```

Trend within day

```{r, fig.width = 12, fig.height = 10}
# limit to long stations
site1 <- aod_conus[,.N, by = site][N>3330, site]

d1 <- aod_conus[site%in%site1,.(lat, lon, hour, dailyN, season, AOD_470nm, 
                                MeanAOD = mean(AOD_470nm),
                                sd_rmse = sd(AOD_470nm)), by = .(site, date)]
d1[, stdAOD:=AOD_470nm - MeanAOD]
d1[, hour := as.numeric(hour)]
d1[hour%in%c(0,1,2), hour :=hour+24]
d1[, totalN := sum(dailyN), by = site]
setorder(d1, -totalN)
# data ready 
site2 <- aod_conus[,.N, by = site][N>8E4, site]
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = stdAOD, color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam") + 
  ggtitle("Daily variation (AOD - daily Mean AOD) by hour")

# sd(AOD) by hour, mean follows same trend
d1_hour <- d1[site%in%site2,.(sd_hour = sd(AOD_470nm), mean_hour = mean(AOD_470nm)), by = hour]
setkey(d1_hour, hour)
plot(x = d1_hour$hour, y = d1_hour$sd_hour, type = "o", ylab = "sd(AOD)")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")  + 
   facet_wrap(~season, ncol = 2)

# all sites

ggplot(dilute.data(d1,10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour") + 
  theme(legend.position = "none") + 
  facet_wrap(~season, ncol = 2)
```

## SD(AOD) by station

Then we calculated the median daily SD (very similar to mean) across all the days for each station.  
*N* adds up the number of all the hour-station observations _N = sum(dailyN)_  
mediansd = median(sd_rmse), meansd = mean(sd_rmse)

```{r}
aod2 <- unique(aod1[,.(mediansd = median(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site)])

setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2[N>5000 & mediansd > 0.02,])

N_label = 5E4
ggplot(data = aod2[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", subtitle = ("Half the stations *")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD(AOD)")

```

* I have removed half of the stations with low total observations, that removed only `r  round(aod2[N<3330,sum(N)]/aod2[,sum(N)])` % of the AERONET AOD measurements (hour-station observation).

Generally speaking, smaller sites have higher variation, but still there are also some long-standing sites that have higher variation than others.

```{r, fig.width = 8, fig.height = 5}

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod2[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 12) + 
        scale_colour_viridis_c() + 
        labs(x = "Longitude", y = "Latitude", color= "RMSE", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

```

## Scatterplot and map By season

```{r, fig.width = 12, fig.height = 10}
aod1[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod1[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod1[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod1[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod1[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod1[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

aod3 <- unique(aod1[,.(mediansd = median(sd_rmse), meansd = mean(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site, season)])

ggplot(data = aod3[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    facet_wrap(~ season, ncol = 2) + 
    ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", 
            subtitle = ("Showing long-lasting stations, half stations, 97% of the observations")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD by station")

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod3[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 8) + 
        scale_colour_viridis_c() + 
        facet_wrap(~ season, ncol = 2) + 
        labs(x = "Longitude", y = "Latitude", color= "SD(AOD)", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

```

## adjust for spline trend

```{r, eval=F}
library("splines")
site_selected <- aod_conus[,.N, by = site][N>3330, site]
aod3 <- aod_conus[site == site_selected[1]]
aod3[, dayint:=as.integer(date)]
aod3[, gam(AOD_470nm ~ s(dayint))]
fit0 <- smooth.spline(x = aod3$dayint, y = aod3$AOD_470nm)

plot(aod3[,.(dayint, AOD_470nm)], cex = 0.1)
lines(fit0, col = 'red')

```

### 2019-07-29 Code/03_import_AOD_V3.Rmd
AERONET daily error

@@
 nemia_buff = st_transform(nemia_buff_aea, crs = 4326)
+
 aod_site_NE <- aod_site_sf[nemia_buff,]  # a sf object for NEMIA
@@
 site_list_NE <- c(aod_site_NE$Site_Name)
+
 # US site list
-# to be done
+aeronet_conus_dt <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/aeronet_conus_dt.rds")
 
@@
 
-AOD is projected to 470nm, as expected it correlates with neighbouring AOD 440nm
+AOD is projected to 470nm globally, as expected it correlates with neighbouring AOD 440nm
 

### 2019-07-30 Code/03.2_AOD Variation.Rmd
check within day variation

@@
 ---
-title: "AODV3 Level2 AOD2.0 from Aeronet: Day Variation"
+title: "AODV3 Level2 AOD2.0 from Aeronet Day Variation"
 author: "Yang Liu"
@@
 source(here::here("Code", "funcs_for_plots.R"))
-# compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit#
+#
 
 ```
+
+compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit
 
@@
 @output dataset has dailyN: observations with in a day (at least 20)
-@d_mean and sd_rmse are daily mean and sd
+@sd_mean and sd_rmse are daily mean and sd
 
@@
 aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
+aod_conus[, year := year(date)]
+# move the hours 0, 1, 2, attach the earlier date
+aod_conus[, hour := as.numeric(hour)]
+aod_conus[hour%in%c(0,1,2), `:=`(hour =hour+24, date = date - 1)]
+
 # aod_conus <- aod_conus[year(date)>2000]
+aod_conus[,datetime:=date]
 aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
-aod_conus[,dailyN:=.N, by = .(site, date)]
-aod_conus <- aod_conus[dailyN>20,]
+aod_conus[,dailyN:=.N, by = .(site, date)] # obs in a day by site,day
+aod_conus <- aod_conus[dailyN>=20,]
+summary(aod_conus[,.N,by = .(site, date)][,N])
+aod_conus[,daysN:= uniqueN(date), by = site] # lasting days by site
 
@@
 
-# summary1: aod1 calculates the mean and sd (SD) by day 
-aod1 <- unique(aod_conus[,.(lat, lon, dailyN, season, mean_aod = mean(AOD_470nm), 
-                                           sd_rmse = sd(AOD_470nm)), by = .(site, date)])
+```
+
+* July has the most observations. Ndays are similar, what hours are shorter in winter
+
+```{r}
+# do winter have fewer observations per hour? 
+aod_conus_hour <- aod_conus[,.(Nobs = .N, Ndays = uniqueN(date)), by = .(season, hour)]  
+g1 = ggplot(aod_conus_hour, aes(x = hour, y = Ndays, color = season)) + geom_line() + theme_classic()
+g2 = ggplot(aod_conus_hour, aes(x = hour, y = Nobs, color = season)) + geom_line() + theme_classic()
+gridExtra::grid.arrange(grobs = list(g1, g2), ncol = 2, top = "By hour: similar days, fewer obs if daytime is short")
 
@@
 
-## daily SD AOD generally follows daily mean AOD, more outliers in the summer
+## Median Daily SD(AOD)
 
-Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date.  
-Plot by time gives this figure:
+### calculating using all data
+
+Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date. Then take the median.  @mediansd = median(sd_rmse)
+
+```{r,  fig.width=12, fig.height=5}
+# summary1: aod1 calculates the mean and sd (SD) by day 
+aod1 <- unique(aod_conus[,.(lat, lon, dailyN, daysN, month, season, 
+                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
+                            year = year(date)), by = .(site, date)])
+# 290 sites: aod2 [290 * 6]
+aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
+                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]
+
+# by season
+
+aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
+                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]
+
+```
+
+Among 290 stations, 87 have data from 2018 and 2019 (Figure Left).   
+@daysN: days of the station with data
+
+* The table shows daysN > 1500 stations.  
+* `r aod2[daysN>=500,.N]` with `daysN` > 500 days are plotted  
+* `r aod2[daysN>=1000,.N]` with `daysN` > 1000 days and SD(AOD)>0.013(the median)   are labelled in the map. Some other large stations are also labelled in the scatterplot.  
+* So in the map, we label the top half of the long-lasting stations with at least 1000 days, and still on at least in 2018.
+
+```{r, fig.width=12, fig.height=5}
+# print results with daysN > 500 
+setorder(aod2, -mediansd)
+rmarkdown::paged_table(aod2[daysN>1500])
+
+plotAODscatter <- function(aod_data, bySeason = F){
+  g1  <-  ggplot(data = aod_data[daysN>=500], aes(x = daysN, y = mediansd))+
+      geom_point(size = 1, alpha = 1) + 
+      theme_classic() + 
+      ggrepel::geom_label_repel(data = aod_data[(daysN>1000&mediansd>0.013)|daysN>1500], aes(x = daysN, y = mediansd, label = site)) + 
+      ggtitle("Median daily SD(AOD) vs. Total days by Station", subtitle = ("Stations with latest year >=2018")) + 
+      labs(x = "Number of Observations", y = "Daily SD(AOD)")
+  if(bySeason) {g1 <- g1 + facet_wrap(~season)}
+  g1
+}
+
+plotAODmap <- function(aod_data, bySeason = F){
+  g2 <- ggplot() +
+          geom_polygon(data = map_data("state"), 
+                       aes(x = long, y = lat, group = group), 
+                       fill = 'NA', color = "grey") + 
+          geom_point(data = aod_data[daysN>=500],
+                     aes(x=lon, y=lat, color = mediansd, size = daysN),
+                     alpha = 0.8) + 
+          # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
+          scale_size_area(max_size = 12) + 
+          scale_colour_viridis_c() + 
+          labs(x = "Longitude", y = "Latitude", color= "Daily\nSD(AOD)", 
+               size = "Number of\nDays") +
+          ggrepel::geom_label_repel(data = aod_data[daysN>1000&mediansd>0.013], 
+                                    aes(x = lon, y = lat, label = site)) + 
+          theme_minimal() +
+          ggtitle("SD by AERONET stations") +
+          theme(legend.position ='right') 
+  if(bySeason) g2 <- g2 + facet_wrap(~season)
+  return(g2)
+}
+p1 <- plotAODscatter(aod2)
+p2 <- plotAODmap(aod2)
+gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)
+
+```
+
+By season
+
+```{r, fig.width = 12, fig.height = 10}
+plotAODscatter(aod3, bySeason = T)
+plotAODmap(aod3, bySeason = T)
+```
+
+### Calculating using past 3 years: 2016 - 2019
 
 ```{r}
+# summary2: limiting the years! aod1 calculates the mean and sd (SD) by day 
+aod1 <- unique(aod_conus[year>=2016,.(lat, lon, dailyN, daysN, month, season, 
+                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
+                            year = year(date)), by = .(site, date)])
+# 290 sites: aod2 [290 * 6]
+aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
+                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]
+
+# by season
+
+aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
+                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]
+
+```
+
+```{r, fig.width=12, fig.height=5}
+# print results with daysN > 500 
+setorder(aod2, -mediansd)
+rmarkdown::paged_table(aod2)
+p1 <- plotAODscatter(aod2)
+p2 <- plotAODmap(aod2)
+gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)
+```
+
+By season
+
+```{r, fig.width = 12, fig.height = 10}
+plotAODscatter(aod3, bySeason = T)
+plotAODmap(aod3, bySeason = T)
+```
+
+## Trend within day
+
+### GSFC, Wallops (high sd) vs. "Sevilleta", "White_Sands_HELSTF" (low sd)
+
+```{r, fig.width = 12, fig.height = 10}
+# substract the daily mean -> 
+aod_conus[, dailymean := mean(AOD_470nm), by = .(site, date)]
+aod_conus[, aod_new := AOD_470nm - dailymean]
+aod_conus2 <- aod_conus[site%in%c("Wallops", "GSFC",
+                                  "Sevilleta", "White_Sands_HELSTF"),]
+ # "MD_Science_Center", "SERC",
+
+# by hour 
+ggplot(dilute.data(aod_conus2,10), aes(x = hour , y = aod_new)) + 
+  geom_point(size =1, alpha = 0.8) + 
+  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
+  facet_wrap(~site, ncol = 2) +
+  theme_classic()
+
+# by date
+ggplot(dilute.data(aod_conus2,10), aes(x = date , y = aod_new)) + 
+  geom_point(size = 1, alpha = 0.8) + 
+  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
+  facet_wrap(~site, ncol = 2) +
+  theme_classic()
+
+```
+
+```{r }
 # show daily trend: AOD vs date, all data
-ggplot(data = aod1, aes(x = date, y = sd_rmse, color = site)) +
+ggplot(data = aod1, aes(x = date, y = sd_aod, color = site)) +
   geom_point(size = 1) + 
@@
 ```
-
-Trend within day
-
-```{r, fig.width = 12, fig.height = 10}
-# limit to long stations
-site1 <- aod_conus[,.N, by = site][N>3330, site]
-
-d1 <- aod_conus[site%in%site1,.(lat, lon, hour, dailyN, season, AOD_470nm, 
-                                MeanAOD = mean(AOD_470nm),
-                                sd_rmse = sd(AOD_470nm)), by = .(site, date)]
-d1[, stdAOD:=AOD_470nm - MeanAOD]
-d1[, hour := as.numeric(hour)]
-d1[hour%in%c(0,1,2), hour :=hour+24]
-d1[, totalN := sum(dailyN), by = site]
-setorder(d1, -totalN)
-# data ready 
-site2 <- aod_conus[,.N, by = site][N>8E4, site]
-ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = stdAOD, color = site)) + 
-         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
-         geom_smooth(method = "gam") + 
-  ggtitle("Daily variation (AOD - daily Mean AOD) by hour")
-
-# sd(AOD) by hour, mean follows same trend
-d1_hour <- d1[site%in%site2,.(sd_hour = sd(AOD_470nm), mean_hour = mean(AOD_470nm)), by = hour]
-setkey(d1_hour, hour)
-plot(x = d1_hour$hour, y = d1_hour$sd_hour, type = "o", ylab = "sd(AOD)")
-
-#
-ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
-         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
-         geom_smooth(method = "gam",se = F) + 
-  ggtitle("Daily variation log(AOD) by hour")
-
-#
-ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
-         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
-         geom_smooth(method = "gam",se = F) + 
-  ggtitle("Daily variation log(AOD) by hour")  + 
-   facet_wrap(~season, ncol = 2)
-
-# all sites
-
-ggplot(dilute.data(d1,10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
-         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
-         geom_smooth(method = "gam",se = F) + 
-  ggtitle("Daily variation log(AOD) by hour") + 
-  theme(legend.position = "none") + 
-  facet_wrap(~season, ncol = 2)
-```
-
-## SD(AOD) by station
-
-Then we calculated the median daily SD (very similar to mean) across all the days for each station.  
-*N* adds up the number of all the hour-station observations _N = sum(dailyN)_  
-mediansd = median(sd_rmse), meansd = mean(sd_rmse)
-
-```{r}
-aod2 <- unique(aod1[,.(mediansd = median(sd_rmse), N = sum(dailyN), lon, lat),
-             by = .(site)])
-
-setorder(aod2, -mediansd)
-rmarkdown::paged_table(aod2[N>5000 & mediansd > 0.02,])
-
-N_label = 5E4
-ggplot(data = aod2[N>3330,], aes(x = N, y = mediansd))+
-    geom_point(size = 1, alpha = 1) + 
-    theme_classic() + 
-    ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = N, y = mediansd, label = site)) + 
-    ggtitle("SD vs. Total Obs by Station", subtitle = ("Half the stations *")) + 
-    labs(x = "Number of Observations (2000 - 2019)", y = "SD(AOD)")
-
-```
-
-* I have removed half of the stations with low total observations, that removed only `r  round(aod2[N<3330,sum(N)]/aod2[,sum(N)])` % of the AERONET AOD measurements (hour-station observation).
-
-Generally speaking, smaller sites have higher variation, but still there are also some long-standing sites that have higher variation than others.
-
-```{r, fig.width = 8, fig.height = 5}
-
-ggplot() +
-        geom_polygon(data = map_data("state"), 
-                     aes(x = long, y = lat, group = group), 
-                     fill = 'NA', color = "grey") + 
-        geom_point(data = aod2[N>3330,],
-                   aes(x=lon, y=lat, color = mediansd, size = N),
-                   alpha = 0.8) + 
-        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
-        scale_size_area(max_size = 12) + 
-        scale_colour_viridis_c() + 
-        labs(x = "Longitude", y = "Latitude", color= "RMSE", 
-             size = "Number of\nObservations") +
-        ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = lon, y = lat, label = site)) + 
-        theme_minimal() +
-        ggtitle("SD by AERONET stations") +
-        theme(legend.position ='right')
-
-```
-
-## Scatterplot and map By season
-
-```{r, fig.width = 12, fig.height = 10}
-aod1[, month := factor(months(date, abbr = T),
-                            levels = month.abb)]
-  aod1[month%in%c("Dec","Jan","Feb"), season:="Winter"]
-  aod1[month%in%c("Mar","Apr","May"), season:="Spring"]
-  aod1[month%in%c("Jun","Jul","Aug"), season:="Summer"]
-  aod1[month%in%c("Sep","Oct","Nov"), season:="Fall"]
-  aod1[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]
-
-aod3 <- unique(aod1[,.(mediansd = median(sd_rmse), meansd = mean(sd_rmse), N = sum(dailyN), lon, lat),
-             by = .(site, season)])
-
-ggplot(data = aod3[N>3330,], aes(x = N, y = mediansd))+
-    geom_point(size = 1, alpha = 1) + 
-    theme_classic() + 
-    facet_wrap(~ season, ncol = 2) + 
-    ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = N, y = mediansd, label = site)) + 
-    ggtitle("SD vs. Total Obs by Station", 
-            subtitle = ("Showing long-lasting stations, half stations, 97% of the observations")) + 
-    labs(x = "Number of Observations (2000 - 2019)", y = "SD by station")
-
-ggplot() +
-        geom_polygon(data = map_data("state"), 
-                     aes(x = long, y = lat, group = group), 
-                     fill = 'NA', color = "grey") + 
-        geom_point(data = aod3[N>3330,],
-                   aes(x=lon, y=lat, color = mediansd, size = N),
-                   alpha = 0.8) + 
-        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
-        scale_size_area(max_size = 8) + 
-        scale_colour_viridis_c() + 
-        facet_wrap(~ season, ncol = 2) + 
-        labs(x = "Longitude", y = "Latitude", color= "SD(AOD)", 
-             size = "Number of\nObservations") +
-        ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = lon, y = lat, label = site)) + 
-        theme_minimal() +
-        ggtitle("SD by AERONET stations") +
-        theme(legend.position ='right')
-
-```
-
-## adjust for spline trend
-
-```{r, eval=F}
-library("splines")
-site_selected <- aod_conus[,.N, by = site][N>3330, site]
-aod3 <- aod_conus[site == site_selected[1]]
-aod3[, dayint:=as.integer(date)]
-aod3[, gam(AOD_470nm ~ s(dayint))]
-fit0 <- smooth.spline(x = aod3$dayint, y = aod3$AOD_470nm)
-
-plot(aod3[,.(dayint, AOD_470nm)], cex = 0.1)
-lines(fit0, col = 'red')
-
-```

### 2019-07-30 Code/04_AOD_join_MAIAC.Rmd
check within day variation

@@
 ---
-title: "AOD collocated with MAIAC"
+title: "AOD Collocated with MAIAC: New Variables"
 author: "Yang Liu"
@@
 
+## Load the collocated dataset
+
 ```{r}
@@
 
-Note 7/22, for sure not the best way here, but it can work. 
+Note 7/22, for sure could be better ways, but it can work. 
 1. Get geometry for buffers as a list: aod_buffers_list  
@@
 }
+
+## Terra
 aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
@@
 
+## Aqua
 aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua

### 2019-07-30 Code/03_import_AOD_V3.Rmd
add notes

@@
 Plot the difference by time
+
 ```{r}

### 2019-08-05 Code/03_import_AOD_V3.Rmd
within day variation plot

@@
 Import and process the AOD20 V3 level 2 global Aeronet stations  
-Interpolating to wavelength AOD 470nm
+* Interpolating to wavelength AOD 470nm, the result is: 
+`aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)`
 
-```{r, eval = F, include = F}
-## About this script ##
-
-```
+* Collocate with MCD19A2
 
@@
 source("~/liuyanguu/CWV/Code/cwv_funcs.R")
-source(here::here("Code", "funcs_for_plots.R"))
+source("~/liuyanguu/AOD_CONUS/Code/funcs_for_plots.R")
 

### 2019-08-12 R/plot_predicted_aod470.Rmd
rename

---
title: "Plot predicted AOD at 470nm"
author: "Yang Liu"
date: "8/7/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")

aod20_data_global <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_global.fst", as.data.table = T)

#' prepare aer_date for interpolation of AOD470nm 
#' 
prepare_for_interpolation <- function(aod20_data_global){
  aod20_data_global <- aod20_data_global[, -c("AOD_1020nm", "AOD_1640nm"), with = F]
  aod20_data_global[,obs:=.I]
  setkey(aod20_data_global, obs)
  vars_aod_sub <- grep("AOD_", names(aod20_data_global), value = T)
  aod20_data_global$N_NAAOD <- rowSums(!is.na(aod20_data_global[,..vars_aod_sub]))
  return(aod20_data_global)
}

aod20_data_global <- prepare_for_interpolation(aod20_data_global)

```

get one side observations

```{r}
vars_wv <- grep("Exact_Wavelengths_of_AOD", names(aod20_data_global), value = T)
vars_aod_sub <- grep("AOD_", names(aod20_data_global), value = T)

d <- melt(aod20_data_global[, c(vars_aod_sub,vars_wv[1:22], "obs"), with = F], measure = list(vars_aod_sub, vars_wv[1:22]),value.name = c("aod", "exact_wv"))
# choose non-NA and >0 AOD
d <- d[!is.na(aod)]
# d[, variable := readr::parse_number(levels(variable))[as.integer(variable)]]
# should have points on both sides, otherwise pick out the id
obs_os <- d[, by = obs, .(oneside = (max(exact_wv)-0.47)*(min(exact_wv)-0.47)>0)][oneside ==T, obs]

```

# 1. Have points on both sides of 470 nm

```{r}
# to make some plots to see the fitting results 
test <- aod20_data_global[, c("obs","N_NAAOD", vars_aod_sub), with = F][!obs%in%obs_os]
table(test[,N_NAAOD])

make.test.plot.points <- function(available_points){
  print(paste0(available_points, " available points for prediction:\n"))
  test0 <- test[N_NAAOD==available_points]
  make.test.fit.plot <- function(i){
    t1 <-  test0[i,c(vars_aod_sub), with = F]
    df1 <- na.omit(melt(t1, measure.vars = c(vars_aod_sub), variable.name = "aod"))
    df1[,x:=readr::parse_number(as.character(aod))/1000]
    # refit to the dataset to recover the fitted line 
    fit0 <- lm(data = df1, log(value) ~  log(x) + I((log(x))^2))
    x <- seq(from = range(c(df1$x,0.47))[1], to = range(c(df1$x,0.47))[2], length.out = 100)
    yhat <- exp(predict(fit0, newdata = as.data.frame(x)))
    p1 = ggplot(data = data.frame(x, yhat), aes(x = x, y = yhat)) + 
      geom_line() + geom_point(data = df1, aes(x=x, y = value)) + 
      geom_point(data = data.frame(x = 0.47, y = exp(predict(fit0, newdata = data.frame(x = 0.47)))), aes(x=x, y=y), color = "red", size = 2) + theme_classic()
    p1
  }
  set.seed(1234)
  # 
  test_fig_list <- lapply(sample(nrow(test0), min(12,nrow(test0))), make.test.fit.plot)
  gridExtra::grid.arrange(grobs = test_fig_list, ncol = 4)
  # ggsave(gridExtra::grid.arrange(grobs = test_fig_list, ncol = 4), file = paste0( "figure/190806_fitted_pred470_os_n",available_points,".png"), width = 12, height = ceiling(nrow(test0)/4)*3)
}

invisible(lapply(2:7, make.test.plot.points))

```

# 2. Points on only one side of 470 nm

```{r}
test <- aod20_data_global[, c("obs","N_NAAOD", vars_aod_sub), with = F][obs%in%obs_os]
table(test[,N_NAAOD])

invisible(lapply(2:7, make.test.plot.points))

```

# 3. Points on only one side of 470 nm and no nearby points

```{r, echo = TRUE}
test <- aod20_data_global[obs%in%obs_os &is.na(AOD_440nm)&is.na(AOD_443nm)&is.na(AOD_490nm) &is.na(AOD_500nm),]

table(test[,N_NAAOD])

invisible(lapply(2:6, make.test.plot.points))

```

### 2019-08-12 R/drake_funcs.R

@@ -182,3 +182,9 @@ interpolate_aod <- function(aer_data){
   aer_data_wPred <- aer_data_wPred[,-c(vars_aod_sub, vars_wv), with = F]
   aer_data_wPred
 }
+
+
+setnames(aer_data_wPred, "AERONET_Site_Name", "Site_Name")
+setkey(aer_data_wPred, Site_Name)
+setkey(aer_nearest_NEMIA, Site_Name)
+aer_data_wPred <- aer_nearest_NEMIA[aer_data_wPred]

### 2019-08-12 R/drake_plan.R

@@ -38,10 +38,10 @@ aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # 
 aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"
   
 aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
-#date_start = "2000-01-01" # do not include in sel_data_bytime if don't want to limit
 date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
 date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
 geo_region = "NEMIA"
+year_MCD19 = "2018"
 
 mcd19path = "/data-coco/mcd19/fst/nemia" 
 # point to cache: 

### 2019-08-13 R/non_drake_code_and_md_files/03.2_AOD Variation.Rmd
organize folder

--- 2019-07-30 Code/03.2_AOD Variation.Rmd
@@
 ---
-title: "AODV3 Level2 AOD2.0 from Aeronet Day Variation"
+title: "AODV3 Level2 AOD2.0 Aeronet: investigate within-day Variation"
 author: "Yang Liu"
@@
 source(here::here("Code", "funcs_for_plots.R"))
-#
 
@@
 summary(aod_conus[,.N,by = .(site, date)][,N])
-aod_conus[,daysN:= uniqueN(date), by = site] # lasting days by site
 
@@
 # summary1: aod1 calculates the mean and sd (SD) by day 
-aod1 <- unique(aod_conus[,.(lat, lon, dailyN, daysN, month, season, 
+aod1 <- unique(aod_conus[,.(lat, lon, dailyN, month, season, 
                             mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
                             year = year(date)), by = .(site, date)])
+aod1[,daysN:= uniqueN(date), by = site] # lasting days by site 
 # 290 sites: aod2 [290 * 6]
@@
 rmarkdown::paged_table(aod2[daysN>1500])
-
-plotAODscatter <- function(aod_data, bySeason = F){
-  g1  <-  ggplot(data = aod_data[daysN>=500], aes(x = daysN, y = mediansd))+
+library(ggrepel)
+plotAODscatter <- function(aod_data, bySeason = F, daysN0 = 500){
+  g1  <-  ggplot(data = aod_data[daysN>=daysN0], aes(x = daysN, y = mediansd))+
       geom_point(size = 1, alpha = 1) + 
       theme_classic() + 
-      ggrepel::geom_label_repel(data = aod_data[(daysN>1000&mediansd>0.013)|daysN>1500], aes(x = daysN, y = mediansd, label = site)) + 
+      # ggrepel::geom_label_repel(data = aod_data[(daysN>daysN0*1.5 & mediansd> median(aod_data$mediansd))|daysN>daysN0*1.5], aes(x = daysN, y = mediansd, label = site)) + 
       ggtitle("Median daily SD(AOD) vs. Total days by Station", subtitle = ("Stations with latest year >=2018")) + 
@@
 
-plotAODmap <- function(aod_data, bySeason = F){
+plotAODmap <- function(aod_data, bySeason = F, daysN0 = 500){
   g2 <- ggplot() +
@@
                        fill = 'NA', color = "grey") + 
-          geom_point(data = aod_data[daysN>=500],
+          geom_point(data = aod_data[daysN >= daysN0],
                      aes(x=lon, y=lat, color = mediansd, size = daysN),
@@
           # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
-          scale_size_area(max_size = 12) + 
+          scale_size_area(max_size = 10) + 
           scale_colour_viridis_c() +
@@
                size = "Number of\nDays") +
-          ggrepel::geom_label_repel(data = aod_data[daysN>1000&mediansd>0.013], 
-                                    aes(x = lon, y = lat, label = site)) + 
+          # ggrepel::geom_text_repel(data = aod_data[daysN>daysN0*1.5 & mediansd>median(aod_data$mediansd)],
+          #                           aes(x = lon, y = lat, label = site),
+          #                           size =2.5, inherit.aes = F) +
           theme_minimal() +
@@
 }
+plotAODmap(aod2,daysN0 = 300)
 p1 <- plotAODscatter(aod2)
@@
 
-By season
-
-```{r, fig.width = 12, fig.height = 10}
-plotAODscatter(aod3, bySeason = T)
-plotAODmap(aod3, bySeason = T)
-```
-
-### Calculating using past 3 years: 2016 - 2019
+### Calculating using past 3 years: 2015 - 2018
 
@@
 # summary2: limiting the years! aod1 calculates the mean and sd (SD) by day 
-aod1 <- unique(aod_conus[year>=2016,.(lat, lon, dailyN, daysN, month, season, 
+aod1 <- unique(aod_conus[year>=2015,.(lat, lon, dailyN, month, season, 
                             mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
@@
 # 290 sites: aod2 [290 * 6]
+aod1[, daysN := uniqueN(date), by = site]
 aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
@@
 
+## make and save plot
+
 ```{r, fig.width=12, fig.height=5}
@@
 rmarkdown::paged_table(aod2)
-p1 <- plotAODscatter(aod2)
-p2 <- plotAODmap(aod2)
-gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)
+p_aer <- plotAODmap(aod2, daysN0 = 300) + 
+  ggtitle("Median Daily SD (AERONET AOD)", 
+                   subtitle = "2015 - 2018")
+
+# gridExtra::grid.arrange(grobs = list(p1, p_aer), ncol = 2)
+
+D2 <- readRDS(here("Intermediate/DSCOVR_DailySD_AOD_bySite.rds"))
+g_DS <- plotAODmap(D2, daysN0 = 5) + ggtitle("Median Daily SD (DSCOVR EPIC MAIAC AOD)", subtitle = "By AERONET Station, 2015 - 2018")
+
+ggsave(p_aer, file = here("Figure","190731_DailySD_AER.png"),
+      width = 6, height = 4)
+
+ggsave(g_DS, file = here("Figure","190731_DailySD_DSCOVR.png"),
+      width = 6, height = 4)
+
+ggsave(grid.arrange(p_aer, g_DS, ncol = 2), file = here("Figure","190731_DailySD_DS_AERONET.png"),
+      width = 12, height = 4)
+
 ```

### 2019-08-13 R/non_drake_code_and_md_files/03_import_AOD_V3.Rmd
organize folder

--- 2019-08-05 Code/03_import_AOD_V3.Rmd
@@
 source("~/liuyanguu/AOD_CONUS/Code/funcs_for_plots.R")
-
 ```
@@
 # load aod data
-# aod20_daily_dir <- "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/DAILY/"
-aod20_daily_dir <- "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
-aod20_file_dir  <-  list.files(aod20_daily_dir, pattern = "*.lev20", full.names = TRUE)
+aod20_file_dir  <-  list.files("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/", pattern = "*.lev20", full.names = TRUE)
 
@@
 site_locations_2019 <- fread("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt")
+
 aod_file_name <- gsub("^[0-9]{5,}_[0-9]{5,}_|\\.lev20", "", list.files(aod20_daily_dir, pattern = "*.lev20"))
+
 aod_file_name[!aod_file_name%in%site_locations_2019$Site_Name]
@@
 
+# aod_file_name[aod_file_name%in%site_list_NE]
+
 # US site list
@@
 # open one file and take a look at the data
-t0 <- fread(aod20_file_dir[3])
+t0 <- fread(aod20_file_dir[1])
 t0[1,]
@@
                  "elev", "ozone", "NO2")
+
 read.aod <- function(site_dir, vars_read, vars_newname){
@@
 }
-
 system.time(
@@
 
+# The 5 main wavelengths
+# aod_main <- c("AOD_340nm", "AOD_380nm", "AOD_440nm", "AOD_500nm", "AOD_675nm")
+vars_wv <- grep("Exact_Wavelengths_of_AOD", names(aod20_data_global), value = T)
 vars_aod <- grep("AOD_", names(aod20_data_global), value = T)
-vars_wv <- grep("Exact_Wavelengths_of_AOD", names(aod20_data_global), value = T)
-# The 5 main wavelengths
-aod_main <- c("AOD_340nm", "AOD_380nm", "AOD_440nm", "AOD_500nm", "AOD_675nm")
 vars_aod_sub <- vars_aod[1:22]
-aod_num <- readr::parse_number(vars_aod_sub)
+# aod_num <- readr::parse_number(vars_aod_sub)
 # Drop the last two wavelengths because they are frequently off the main trend lines
@@
 # should have points on both sides, otherwise pick out the id
-obs_os <- d[, by = obs, .(oneside = (max(exact_wv)-470)*(min(exact_wv)-470)>0)][oneside ==T, obs]
+obs_os <- d[, by = obs, .(oneside = (max(exact_wv)-0.47)*(min(exact_wv)-0.47)>0)][oneside ==T, obs]
 saveRDS(obs_os, "~/Intermediate/AOD_data/190621_obs_same_side")
@@
 time0 <- Sys.time()
-system.time(d2 <- d[, by = obs, .(AOD_470nm = exp(predict(lm(log(aod) ~  log(exact_wv) + I((log(exact_wv))^2)), newdata = data.frame(exact_wv = 470))))])
-Sys.time() - time0 # 6.3 hours 6/19
+system.time(d2 <- d[, by = obs, .(AOD_470nm = exp(predict(lm(log(aod) ~  log(exact_wv) + I((log(exact_wv))^2)), newdata = data.frame(exact_wv = 0.47))))])
+Sys.time() - time0 # 6.3 hours 6/19 for the world.
 
@@
 # If there are 2 or 3 points, one side is not fine
-# 1 points are set to NA, all of them are in obs_os as well, since there is only one point
+# If there is only 1 point, set to NA, all of them are in obs_os anyway, since there is only one point
 length(obs_os) # 91911
-aod20_data_global[N_NAAOD <=4 & obs%in%obs_os,.N] # 91553, so most <=3 are on the same side of 470 nm anyway 
-aod20_data_global[N_NAAOD ==3 & obs%in%obs_os,.N]
 
@@
 For those with at least 5 points, most have close points within the allowed range: "AOD_440nm" "AOD_443nm" "AOD_490nm" "AOD_500nm"
-allowed_range <- paste0("AOD_", aod_num[440<=aod_num & aod_num<=500],"nm")
-aod20_data_global[obs%in%obs_os &is.na(AOD_440nm)&is.na(AOD_443nm)&is.na(AOD_490nm) &is.na(AOD_500nm), table(N_NAAOD)]
- # exterpolating with 5 or 6 might be fine 
+
+ # exterpolating with 4, 5 or 6 might be fine 
  #    1     2     3     4     5     6 
@@
 
-# create aod20_NEMIA_idLST, link Aeronet AOT to MAIAC grid (MCD19A2)
+# Rolling join AOT to MAIAC (MCD19A2)
 
-* created aod20_NEMIA_idLST that has AERONET Site_Name, idLSTpair0
+1. created aod20_NEMIA_idLST that has AERONET Site_Name matched with idLSTpair0
 
@@
 
-# Join AOT to MAIAC (MCD19A2)
-
-* rolling join with nearest time (join_time) and by (idLSTpair0)
+2. rolling join with nearest time (join_time) and by (idLSTpair0)
 
@@
 
-# load lst data for 2018
-read.lst <- function(sat = "terra", columns0 = NULL){
-  if (sat != "terra") choose = "mcd19_T" else choose = "mcd19_A" # load terra by default
+read_lst <- function(sat = "terra"){
+  lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = "mcd19_T", full.names = T)
+  if (sat != "terra") choose = "mcd19_A" else choose = "mcd19_T" # load terra by default
   lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = choose, full.names = T)
-  dt = rbindlist(lapply(lst_files, function(x)read.fst(x, as.data.table = T, columns  = columns0)))
-  if(is.null(columns0)) dt[,c("x", "y", "inNEMIA"):=NULL]
-
+  dt = rbindlist(lapply(lst_files, function(x)read.fst(x, as.data.table = T)))
   dt[, join_time:=overpass_time]
@@
 }
-lst_terra <- read.lst()
-print(object.size(lst_terra), unit = "GB") # 18 GB all the lst terra in 2018 (NEMIA, one year)
+lst_terra <- read_lst()
+# 18 GB all the lst terra in 2018 (NEMIA, one year)
+print(object.size(lst_terra), unit = "GB") 
 # make sure both are UTC time 
-# attr(aod20_NEMIA_idLST$stn_time, "tzone")
-# attr(lst_terra$overpass_time, "tzone")
+attr(aod20_NEMIA_idLST$stn_time, "tzone")
+attr(lst_terra$overpass_time, "tzone")
 
@@
 mean(aod_lst_terra$within30min_t)
-
-# vars0 <- c("overpass_index", "Optical_Depth_047", "AOD_Uncertainty", "Column_WV", "AOD_QA",  "RelAZ")
-# vars_new <- paste0(vars0,"_t")
-# setnames(aod_lst_terra, vars0, vars_new)
-# set NA 
 aod_lst_terra <- aod_lst_terra[within30min_t!=0,]
@@
 # Aqua  ~~~~ #### 
-lst_aqua <- read.lst("aqua")
+lst_aqua <- read_lst("aqua")
 # rolling join

### 2019-08-13 Code/funcs_for_plots.R

@@ -1,149 +1,0 @@
-# Functions for AOD
-# Yang Liu
-
-
-# for data ----------------------------------------------------------------
-
-
-read.lst <- function(sat = "terra", columns0 = NULL){
-  if (sat != "terra") choose = "mcd19_T" else choose = "mcd19_A" # load terra by default
-  lst_files <- list.files(path = "/data-coco/mcd19/fst/nemia/2018", pattern = choose, full.names = T)
-  dt = rbindlist(lapply(lst_files, function(x)read.fst(x, as.data.table = T, columns  = columns0)))
-  if(is.null(columns0)) dt[,c("x", "y", "inNEMIA"):=NULL]
-  
-  dt[, join_time:=overpass_time]
-  setkey(dt, idLSTpair0, join_time)
-  return(dt)
-}
-
-
-
-#  general functions for making plots -------------------------------------
-
-
-# Scatterplot -------------------------------------------------------------
-
-plot.scatter <-  function(data, x, y, size0 = 0.1, alpha0 = 0.3, dilute = F, dilute_by = 10){
-  set.seed(1234)
-  if (dilute) {
-    data <- data[sample(nrow(data), nrow(data)/dilute_by)] # dilute
-  }
-  plot0 <- ggplot(data = data, aes(x = data[[x]], y = data[[y]]))+
-    # geom_point(size = size0, alpha = alpha0) + 
-    geom_density_2d(aes(fill = ..level..), geom = "polygon") + 
-    labs(x = x, y = y) + 
-    theme_bw()
-  # add histogram
-  ggExtra::ggMarginal(plot0, type = "histogram", bins = 50, size = 10, color="white")
-}
-
- 
-# Maps --------------------------------------------------------------------
-# us map using ggmap
-plot.us.ggmap <- function(data0){
-  # data0 contains the points to plot, plot on whole us mainland
-  map_wholeUS <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/map_us_all.RData")
-  ggmap(map_wholeUS) + 
-    geom_point(data = data0, aes(x = lon, y = lat),size = 1, color= "red") + # MOD
-    # geom_label(data = SiteName_aodid, aes(label = Site_Name), size = 2, alpha = 0.7) + 
[…]

### 2019-08-13 R/drake_funcs.R

@@ -157,9 +157,9 @@ sel_data_bytime <- function(aer_data, date_start = NULL, date_end = NULL){
 #' prepare aer_date for interpolation of AOD470nm 
 #' interpolating to wavelength 470 nm, input as 0.47 
 #' @param aer_data the dataset contains all the AOD measurements and exact wavelengths
-#' @return dataset of aer_data with pred
+#' @return dataset of aer_data with pred, also with `nearest_refgrid` next to Site_Name
 #' 
-interpolate_aod <- function(aer_data){
+interpolate_aod <- function(aer_data, aer_nearest_NEMIA){
   aer_data <- aer_data[, -c("AOD_1020nm", "AOD_1640nm"), with = F]
   aer_data[,obs:=.I]
   setkey(aer_data, obs)
@@ -176,15 +176,49 @@ interpolate_aod <- function(aer_data){
   d2 <- d[, by = obs, .(AOD_470nm = exp(predict(lm(log(aod) ~  log(exact_wv) + I((log(exact_wv))^2)), newdata = data.frame(exact_wv = 0.47))))] # notice to put 0.47 not 470
   setkey(d2, obs)
   aer_data_wPred <- d2[aer_data]
-  # set NA: 
-  aer_data_wPred[N_NAAOD<=4 & obs%in%obs_os, AOD_470nm:=NA]
+  # set NA: At least 4 observations wanted for exterpolation 
+  aer_data_wPred[N_NAAOD<4 & obs%in%obs_os, AOD_470nm:=NA]
   # remove unncessary variables: 
   aer_data_wPred <- aer_data_wPred[,-c(vars_aod_sub, vars_wv), with = F]
+  
+  # attach nearest_refgrid (also called idLSTpair0 in many our script) to Aer Site_Name
+  setnames(aer_data_wPred, "AERONET_Site_Name", "Site_Name")
+  setkey(aer_data_wPred, Site_Name)
+  aer_sites <- as.data.table(aer_nearest_NEMIA)[,c("Site_Name", "nearest_refgrid"), with = F]
+  setkey(aer_sites, Site_Name)
+  aer_data_wPred <- aer_sites[aer_data_wPred]
+  aer_data_wPred[, join_time:=stn_time]
+  setkey(aer_data_wPred, nearest_refgrid, join_time)
   aer_data_wPred
 }
 
 
-setnames(aer_data_wPred, "AERONET_Site_Name", "Site_Name")
-setkey(aer_data_wPred, Site_Name)
[…]

### 2019-08-13 R/drake_plan.R

@@ -1,5 +1,6 @@
-# beginning of script where we will have a plan to:
+# See `drake_funcs.R` for functions used in drake plan #
 
+# beginning of script where we will have a plan to:
 # select Aeronet in CONUS
 # select Aeronet in NEMIA
 # calculate nearest idLSTpair0's. May need shortcut that uses buffer around stations to reduce possibility set. 
@@ -12,21 +13,23 @@
 # restrict to 2000+ (parameter so that different subsets could be taken later)
 
 # Libraries
-library(sf)
-library(magrittr)
-library(drake)
-library(data.table)
-library(fst)
-library(future)
-#install.packages("/home/rushj03/dev/nngeo", repos = NULL, type = "source")
-# library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
-#library(sp)
-#library(rgdal)
-#library(rnaturalearth) # REPLACE WITH CONUS FILE USED IN GRID XXX
-#library(rgeos)
-library(nngeo)
-#library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
-
+suppressPackageStartupMessages({
+  library(sf)
+  library(magrittr)
+  library(drake)
+  library(data.table)
+  library(fst)
+  library(future)
+  library(here)
+  #install.packages("/home/rushj03/dev/nngeo", repos = NULL, type = "source")
+  # library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
+  #library(sp)
+  #library(rgdal)
+  #library(rnaturalearth) # REPLACE WITH CONUS FILE USED IN GRID XXX
+  #library(rgeos)
+  library(nngeo)
+  #library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
+})
 
 # Parameters ####
 
@@ -41,16 +44,15 @@ aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
 date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
 date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
 geo_region = "NEMIA"
-year_MCD19 = "2018"
-
-mcd19path = "/data-coco/mcd19/fst/nemia" 
+sats = c("terra", "aqua")
+# MCD19 data: 
[…]

### 2019-08-14 R/drake_plan.Rmd
new drake plan

---
title: "Drake Plan"
author: "Johnathan Rush, Yang Liu"
date: "08/14/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

```{r}
# See `drake_funcs.R` for functions used in drake plan #

# beginning of script where we will have a plan to:
# select Aeronet in CONUS
# select Aeronet in NEMIA
# calculate nearest idLSTpair0's. May need shortcut that uses buffer around stations to reduce possibility set.

# select from the daily data the subsets above (one of the two)
# calc columns as in old script
# as.numeric columns as necessary
# remove rows without AOD? probably not - they may have CWV
# ? which other columns to we want?
# restrict to 2000+ (parameter so that different subsets could be taken later)

# Libraries
suppressPackageStartupMessages({
  library(sf)
  library(magrittr)
  library(drake)
  library(data.table)
  library(fst)
  library(future)
  library(here)
  #install.packages("/home/rushj03/dev/nngeo", repos = NULL, type = "source")
  # library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
  #library(sp)
  #library(rgdal)
  #library(rnaturalearth) # REPLACE WITH CONUS FILE USED IN GRID XXX
  #library(rgeos)
  library(nngeo)
  #library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
})

# Parameters ####

aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"

aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # don't have time to play with Ethernet all day
# the aer data with AOD470nm
aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"

aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
geo_region = "NEMIA"
sats = c("terra", "aqua")
# MCD19 data: 
mcd19path = "/data-coco/mcd19/fst/nemia" # now there is only 2018 data 
# point to cache: 
ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake")

# Drake Plans  ------------------------------------------------------------
# load functions for drake: 
source(here("R/drake_funcs.R"))
# new plan Yang Liu
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

  # AERONET station locations file: 
  aer_stns = fread(file = file_in(aer_stn_path), col.names = c("Site_Name", "lon", "lat", "elevm")),
  conus_buff = get_conus_buff(),
  nemia_buff = get_nemia_buff(),
  aerpts = get_aer_spatial(aer_stns),               # wgs84
  conus_aer = select_points(aerpts, conus_buff),    # AERONET sites in CONUS as points
  nemia_aer = select_points(conus_aer, nemia_buff), # AERONET sites in NEMIA as points

  # get nearest MODIS cells 
  refgrid = get_ref_grid(),
  candidate_refpts = points_in_buffer(nemia_aer, refgrid),
  aer_nearest = get_nearest_cell(nemia_aer, candidate_refpts), # Aeronet matched to grid, 174x4
  aer_nearest_NEMIA = remove_site_on_water(aer_nearest), # remove 2 sites

  # measurements. Editing values of date_start and date_end should make use of the cache
  # if range was previously calc'd
  aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
                   transform = map(date_start = !!date_start, date_end = !!date_end)),

  # selected measurements, transform by same date range if not automatic in name
  nemia = target(sel_data_bystation(aer_btw, nemia_aer),
                      transform = map(aer_btw)), # use map to map to aer_btw

  # 2. Interpolation -----------------------------------------------------------
  # so instead of running all the Aeronet data, only run what is used
  # it also join with the nearest grid id 
  wPred = target(interpolate_aod(nemia, aer_nearest_NEMIA),transform = map(nemia)),

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset 
  mcd = target(read_lst(sat), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred)),

  # 4. nemia_aer_m
  s = target(aer_in_mcd19(nemia_aer, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(nemia_aer_m = s, refgrid),
             transform = map(s)),
  # sites
  p = target(get_site_list(nemia_aer_m = s), transform = map(s)),
  n = target(get_site_name(nemia_aer_m = s), transform = map(s)),

  # mapping
  # this way doesn't work:
  # sub = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub = target(purrr_map(ref = ref, p = p, n = n),
               transform = map(ref, p, n)),
  # calculate new vars 
  # for terra and aqua, using join, mcd, sub 
  new = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub),
               transform = map(join, mcd, sub))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

loadd(cache = ssd_cache)

message("To see one of the aer station with grid around it: ")
knitr::kable(sub_ref_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_p_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_n_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31[1:3,])

message("The colocated dataset with new vars")
rmarkdown::paged_table(new_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_mcd_aqua_sub_ref_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_p_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_n_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31[1:3,])
```

```{r, eval = F}
# outdated(nemia_config)
future::plan("multicore")
make(nemia_plan, cache = ssd_cache) # write update into the cache 
cached(cache = ssd_cache)
loadd(cache = ssd_cache) # load all object

# to see the updated result, run config again
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
vis_drake_graph(nemia_config)

# Join MCD19 --------------------------------------------------------------
aer_data_wPred <- readd("wPred_nemia_aer_btw_2018.01.01_2018.12.31", cache = ssd_cache) # load all object 
nemia_aer <- readd("nemia_aer", cache = ssd_cache)
refgrid <- readd("refgrid", cache = ssd_cache)

plan1 <- drake_plan(
  # 3. WPred Join MCD19 -----------------------------------------------------------
  mcd = target(read_lst(sat), transform = map(sat = c("terra", "aqua"))),
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred),
                transform = map(mcd))
)
config2 <- drake_config(plan1, cache = ssd_cache) # show the dependency
vis_drake_graph(config2)
time0 <- Sys.time()
make(plan1, cache = ssd_cache) # write update into the cache 
Sys.time() - time0
cached(cache = ssd_cache)
loadd(cache = ssd_cache) # load all object 
join_mcd_aqua = readd("join_mcd_aqua", cache = ssd_cache)

```

### 2019-08-14 R/non_drake_code_and_md_files/04_AOD_join_MAIAC.Rmd
organize folder
new drake plan

--- 2019-07-30 Code/04_AOD_join_MAIAC.Rmd
@@
 # but of course, after intersect with the base grid, they won't be circles anymore.
+
+  site_list <- st_geometry(nemia_aer_m)
+  name_list <- as.list(nemia_aer_m$Site_Name) # for recording site names
+
 ```
@@
 # # not sure why mapply doesn't work 
-# refDT_sub_list <- mapply(select.refDT.subset, ref_list0 = ref_list[1], point0 = site_list2[1], name0 = name_list[1])
+refDT_sub_list <- mapply(select.refDT.subset, ref_list, site_list2, name_list)
+# # but pmap can work. 
+refDT_sub_list <- purrr::pmap(list(ref_list, site_list2, name_list), select.refDT.subset)
 
-refDT_sub_list <- list()
+# # a loop can work 
 for (i in 1:length(ref_list)){
   ref_list0 = ref_list[[i]];  point0 = site_list2[[i]]; name0 = name_list[[i]]
-  ref_list0$dist <- st_distance(ref_list0, point0)
+  ref_list0$dist <- sf::st_distance(ref_list0, point0)
   ref_list0$Site_Name <- name0
@@
 refDT_sub <- setDT(refDT_sub)[,.(Site_Name, idLSTpair0, dist_km)]
-setnames(refDT_sub, "idLSTpair0", "idLSTpair0_in_buffer")
+setnames(refDT_sub, "idLSTpair0", "nearest_refgrid")
 setkey(refDT_sub, Site_Name)
@@
 
-# remove some variables 
+# aod_join_MODIS is the rolling-joined file, MODIS_all is like lst_terra
 aod.MODIS.newVars <- function(aod_join_MODIS, MODIS_all){
@@
   aod_join_buffer <- aod_join_MODIS[refDT_sub, allow.cartesian = T]
-  cat("The cartesian joined dataset between collocated Aeronet with MODIS sites data and base grid in each circles, dim is: ", paste(dim(aod_join_buffer),collapse = " x "))
+  cat("The cartesian joined dataset between collocated Aeronet with MODIS sites data and base grid in each circles, dim is: ", paste(dim(aod_join_buffer), collapse = " x "))
 
@@
   MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
-  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"),
-           skip_absent = T)
+  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"), skip_absent = T)
   setkey(MODIS_all, idLSTpair0_in_buffer, day)

### 2019-08-14 R/drake_funcs.R

@@ -112,12 +112,11 @@ vars0 <- c("Date(dd:mm:yyyy)", "Time(hh:mm:ss)", "Day_of_Year","AERONET_Site_Nam
 
 #' read in the Aeronet AOD measurement data from `aod20_file_dir` files.   
 #' @param aod_dir where the files located on coco.   
-#' 
+#' @return 
 #' 
 get_stn_data <- function(aod_dir){
   aer_files_dir  <-  list.files(aod_dir, pattern = "*.lev20", full.names = TRUE)
-  t0 <- fread(aer_files_dir[1])
-  
+  t0 <- fread(aer_files_dir[1]) # to get variable names: 
   vars_aod <- intersect(grep("AOD_", names(t0), value = T), grep("nm", names(t0), value = T))
   # sort the wave lengths varnames from low to high, and update the vars_aod
   x_nm <- sort(readr::parse_number(vars_aod)) # 340, 380, ... , 1640
@@ -193,7 +192,7 @@ interpolate_aod <- function(aer_data, aer_nearest_NEMIA){
 }
 
 
-# 4. join MCD19 (Terra, Aqua) -----------------------------------------------------------
+# 4. rolling join MCD19 (Terra, Aqua) -----------------------------------------------------------
 #' read MCD19
 #' this function need to be replaced in the future, as for now 
 #' we only read one-year data
@@ -213,7 +212,7 @@ read_lst <- function(sat = "terra"){
 }
 
 #' rolling join aer with MCD19
-#' @return MCD19 rolling join Aernet 
+#' @return MCD19 rolling join Aeronet 
 rolling_join <- function(mcd_sat, aer_data_wPred){
   mcd19 <- aer_data_wPred[mcd_sat, roll = 'nearest', nomatch = 0]
   # time difference
@@ -222,3 +221,125 @@ rolling_join <- function(mcd_sat, aer_data_wPred){
   mcd19[, join_time:=NULL]
   return(mcd19)
 }
+
+
+
+# 5. Calculate distance ------------------------------------------------------
+#' select aer sites in mcd19
+#' return colocated/limited aer sites, on crs 2163
+aer_in_mcd19 <- function(nemia_aer, mcd19){
+  nemia_aer_m = st_transform(nemia_aer[nemia_aer$Site_Name%in%unique(mcd19$Site_Name),], crs = 2163)   
+  return(nemia_aer_m)
+}
+
+
[…]

### 2019-08-14 R/drake_old_plan.R

@@ -1,30 +1,0 @@
-nemia_plan <- drake_plan(
-  
-  
-  # AERONET station locations file: 
-  aer_stns = fread(file = file_in(aer_stn_path), col.names = c("Site_Name", "lon", "lat", "elevm")),
-  conus_buff = get_conus_buff(),
-  nemia_buff = get_nemia_buff(),
-  aerpts = get_aer_spatial(aer_stns),               # wgs84
-  conus_aer = select_points(aerpts, conus_buff),    # AERONET sites in CONUS as points
-  nemia_aer = select_points(conus_aer, nemia_buff), # AERONET sites in NEMIA as points
-  
-  # get nearest MODIS cells 
-  refgrid = get_ref_grid(),
-  candidate_refpts = points_in_buffer(nemia_aer, refgrid),
-  aer_nearest = get_nearest_cell(nemia_aer, candidate_refpts), # Aeronet matched to grid, 174x4
-  aer_nearest_NEMIA = remove_site_on_water(aer_nearest), # remove 2 sites 
-  
-  # measurements. Editing values of date_start and date_end should make use of the cache
-  # if range was previously calc'd
-  aer_data = get_stn_data(file_in(aer_data_path), aer_columns),  # 26171178x6
-  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
-                   transform = map(date_start = !!date_start, date_end = !!date_end)),
-  
-  # selected measurements, transform by same date range if not automatic in name
-  nemia_data = target(sel_data_bystation(aer_btw, nemia_aer),
-                      transform = map(aer_btw)), # use map to map to aer_btw
-  
-  # joined mcd19 MAIAC values. transform by same date range if not automatic in name
-  
-)

### 2019-08-18 AOD_CONUS_Report.Rmd
nemia drake

---
title: "AOD CONUS Modeling Results"
author: "Yang Liu"
date: "08/17/2019"
output: html_document
---

```{r}
library(drake)

```

# Terra

```{r}
readd(print1, cache = ssd_cache)
readd(plot1, cache = ssd_cache)
```

* Top features RFE

```{r}
# top features
loadd("cv_results_new_var_join_mcd_terra", cache = ssd_cache)
# rferesults = cv_results_new_var_join_mcd_terra
cv_results_new_var_join_mcd_terra$features_rank_rfe
```

```{r}

readd(plotSHAP1_2, cache = ssd_cache)

```

# Aqua

```{r}
readd(print1_2, cache = ssd_cache)
readd(plot1_2, cache = ssd_cache)

```

* Top features RFE

```{r}
# top features
loadd("cv_results_new_var_join_mcd_aqua", cache = ssd_cache)
cv_results_new_var_join_mcd_aqua$features_rank_rfe
```

```{r}
readd(plotSHAP1_2, cache = ssd_cache)

```

### 2019-08-18 R/drake_plan_conus.Rmd
nemia drake

---
title: "Drake Plan"
author: "Johnathan Rush, Yang Liu"
date: "08/14/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

# Write the plan

```{r}
# See `drake_funcs.R` for functions used in drake plan #

# beginning of script where we will have a plan to:
# select Aeronet in CONUS
# select Aeronet in NEMIA
# calculate nearest idLSTpair0's. May need shortcut that uses buffer around stations to reduce possibility set.

# select from the daily data the subsets above (one of the two)
# calc columns as in old script
# as.numeric columns as necessary
# remove rows without AOD? probably not - they may have CWV
# ? which other columns to we want?
# restrict to 2000+ (parameter so that different subsets could be taken later)

# Libraries
suppressPackageStartupMessages({
  library(sf)
  library(magrittr)
  library(drake)
  library(data.table)
  library(fst)
  library(future)
  library(here)
  library(nngeo)
  library(ggplot2)
  library(SHAPforxgboost)
})
# ** need to revise accordingly the path to your local Just_universal
source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
source("~/liuyanguu/Just_universal/code/xgboost_cv_RFE.R")

# Parameters ####

aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"

aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # don't have time to play with Ethernet all day
# the aer data with AOD470nm
aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"

aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
geo_region = "NEMIA"
sats = c("terra", "aqua")
# MCD19 data: 
mcd19path = "/data-coco/mcd19/fst/nemia/2018/" # now there is only 2018 data 
mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
# point to cache: 
ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake")

# for reporting purpose
y_var = "diff_AOD"
y_var_pred = "diff_AOD_pred"

# CONUS Drake Plans  ------------------------------------------------------------
source(here("R/drake_funcs.R"))
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

  # AERONET station locations file: 
  aer_stns = fread(file = file_in(aer_stn_path), col.names = c("Site_Name", "lon", "lat", "elevm")),
  conus_buff = get_conus_buff(),
  nemia_buff = get_nemia_buff(),
  aerpts = get_aer_spatial(aer_stns),               # wgs84
  conus_aer = select_points(aerpts, conus_buff),    # AERONET sites in CONUS as points
  nemia_aer = select_points(conus_aer, nemia_buff), # AERONET sites in NEMIA as points

  # get nearest MODIS cells 
  refgrid = get_ref_grid(),
  # near MODIS cells within 300km
  refsub = get_near_cellsid(conus_aer, refgrid),
  # aeronet sites with cloest MODIS grid
  aer_nearest = get_nearest_cell(conus_aer, refgrid), # Aeronet matched to grid, 174x4
  aer_nearest_2 = remove_site_on_water(aer_nearest), # remove 2 sites on water

  # limit aeronet data by date (aer_btw)
  aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
                   transform = map(date_start = !!date_start, 
                                   date_end = !!date_end,
                                   .id = FALSE)),
  # limit aeronet data by station, could be NEMIA or CONUS 
  sel_aer_region = target(sel_data_bystation(aer_btw, aer_nearest_2),
                      transform = map(aer_btw), .id = FALSE),

  # 2. Interpolation -----------------------------------------------------------
  # so instead of running all the Aeronet data, only run what is used
  # it also join with the nearest grid id 
  wPred = target(interpolate_aod(sel_aer_region, aer_nearest_2), transform = map(sel_aer_region, .id = FALSE)),

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset, limit to `refsub`:nearby(300km) when readin to 
  # control total lines, unable to read in all data together, too many rows
  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS), refsub), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred, .id = mcd)),

  # select aer sites in mcd19. further limit the stations
  sel_aer = target(aer_in_mcd19(aer_nearest_2, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(region_aer_m = sel_aer, refgrid),
             transform = map(sel_aer)),
  # sites
  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),

  # mapping 
  # this way doesn't work:
  # sub_grid = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
               transform = map(ref, p, n, .id = FALSE)),
  # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
  new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub_grid),
               transform = map(join, mcd, sub_grid, .id = c(join))),

  # 5. CV -----------------------------------------------------------
  cv_results = target(run_cv(dt = new_var),transform = map(new_var)),

  # 6. Report -----------------------------------------------------------
  print1 = target(report_rmse(cv_results),
                  transform = map(cv_results,.id = FALSE)),
  plot1 = target(plot_rmse(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  plotSHAP1 = target(plot_SHAP(cv_results),
                 transform = map(cv_results,.id = FALSE)),
  report = rmarkdown::render(
    knitr_in("AOD_CONUS_Report.Rmd"),
    output_file = file_out("AOD_CONUS_Report.html"))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

```

# Run the plan

```{r, eval = F}

future::plan(future::multiprocess)
options(future.globals.maxSize= 100*1024^3) 
make(nemia_plan, cache = ssd_cache,
     parallelism = "future", jobs = 20, memory_strategy = "speed")

## Some simple analysis 
loadd("new_var_join_mcd_terra", cache = ssd_cache)
loadd("new_var_join_mcd_aqua", cache = ssd_cache)
SHAPforxgboost::scatter.plot.diagonal(new_var_join_mcd_aqua, "Optical_Depth_047",
                                      "Mean_AOD270km")
cor(as.matrix(new_var_join_mcd_terra[,c(grep("Mean_AOD", names(new_var_join_mcd_terra), value = T), "AOD_470nm"), with = F]))
## 
loadd(cache = ssd_cache) # load all object

# to see the updated graph, run config again
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
vis_drake_graph(nemia_config)

```

### 2019-08-18 R/drake_plan_nemia.Rmd
nemia drake

--- 2019-08-18 R/drake_plan_conus.Rmd
@@
 
-# CONUS Drake Plans  ------------------------------------------------------------
+# Drake Plans  ------------------------------------------------------------
+
+# load functions for drake: 
 source(here("R/drake_funcs.R"))
+# new plan Yang Liu
 nemia_plan <- drake_plan(
@@
   refgrid = get_ref_grid(),
-  # near MODIS cells within 300km
-  refsub = get_near_cellsid(conus_aer, refgrid),
-  # aeronet sites with cloest MODIS grid
-  aer_nearest = get_nearest_cell(conus_aer, refgrid), # Aeronet matched to grid, 174x4
-  aer_nearest_2 = remove_site_on_water(aer_nearest), # remove 2 sites on water
+  aer_nearest = get_nearest_cell(nemia_aer, refgrid), # Aeronet matched to grid, 174x4
+  aer_nearest_NEMIA = remove_site_on_water(aer_nearest), # remove 2 sites
 
-  # limit aeronet data by date (aer_btw)
+  # limit aer sites by date (aer_btw)
   aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
@@
                                    .id = FALSE)),
-  # limit aeronet data by station, could be NEMIA or CONUS 
-  sel_aer_region = target(sel_data_bystation(aer_btw, aer_nearest_2),
+  # limit by site 
+  nemia = target(sel_data_bystation(aer_btw, aer_nearest_NEMIA),
                       transform = map(aer_btw), .id = FALSE),
@@
   # it also join with the nearest grid id 
-  wPred = target(interpolate_aod(sel_aer_region, aer_nearest_2), transform = map(sel_aer_region, .id = FALSE)),
+  wPred = target(interpolate_aod(nemia, aer_nearest_NEMIA), transform = map(nemia)),
 
   # 3. WPred Join MCD19 -----------------------------------------------------------
-  # mcd is the large MCD19 dataset, limit to `refsub`:nearby(300km) when readin to 
-  # control total lines, unable to read in all data together, too many rows
-  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS), refsub), transform = map(sat = c("terra", "aqua"))),
+  # mcd is the large MCD19 dataset 
+  # `mcd19path` could be revised to CONUS later 
+  mcd = target(read_mcd19(sat, file_in(mcd19path)), transform = map(sat = c("terra", "aqua"))),
   # join is the rolling-joined dataset with aeronet
@@
 
-  # select aer sites in mcd19. further limit the stations
-  sel_aer = target(aer_in_mcd19(aer_nearest_2, mcd19 = join),
+  sel_aer = target(aer_in_mcd19(aer_nearest_NEMIA, mcd19 = join),
              transform = map(join)),
   # buffers 
-  ref = target(ref_in_buffer(region_aer_m = sel_aer, refgrid),
+  ref = target(ref_in_buffer(nemia_aer_m = sel_aer, refgrid),
              transform = map(sel_aer)),
   # sites
-  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
-  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),
+  p = target(get_site_list(nemia_aer_m = sel_aer), transform = map(sel_aer)),
+  n = target(get_site_name(nemia_aer_m = sel_aer), transform = map(sel_aer)),
 
@@
   # this way doesn't work:
-  # sub_grid = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
+  # sub = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
   #              transform = map(ref, p, n))
@@
                transform = map(ref, p, n, .id = FALSE)),
-  # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
+  # calculate new vars: the dataset ready for modeling 
+  # for terra and aqua, using join, mcd, sub 
   new_var = target(aod_MODIS_newVars(aod_join_MODIS = join, 
@@
 
-  # 6. Report -----------------------------------------------------------
+  # 6. MD -----------------------------------------------------------
   print1 = target(report_rmse(cv_results),
@@
   report = rmarkdown::render(
-    knitr_in("AOD_CONUS_Report.Rmd"),
-    output_file = file_out("AOD_CONUS_Report.html"))
+    knitr_in("report.Rmd"),
+    output_file = file_out("report.html"))
 )
@@
 make(nemia_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20, memory_strategy = "speed")
+     parallelism = "future", jobs = 20)
 

### 2019-08-18 R/non_drake_code_and_md_files/04_AOD_join_MAIAC.Rmd
nemia drake

@@
 # the original MCD19 
-lst_terra <- read.lst(sat = "terra", columns0 = c("idLSTpair0", "overpass_time", "Optical_Depth_047"))
+lst_terra <- read_lst(sat = "terra")
 
@@
   MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
-  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"), skip_absent = T)
+  setnames(MODIS_all, c("nearest_refgrid", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"), skip_absent = T)
   setkey(MODIS_all, idLSTpair0_in_buffer, day)
@@
 head(aod_lst_terra)
+dim(aod_lst_terra)
 # AOD - AOD mean, do later 
-saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")
+# saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")
 
@@
 head(aod_lst_aqua)
-saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")
+# saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")
 
@@
 library(ggplot2)
-plot.scatter.diagonal(aod_lst_terra, "Mean_AOD10km", "AOD_470nm")
+library(SHAPforxgboost)
+aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")
+scatter.plot.diagonal(aod_lst_terra, x = "Mean_AOD10km", y = "AOD_470nm")
 cor(as.matrix(aod_lst_terra[,c(grep("Mean_AOD", names(aod_lst_terra), value = T), "AOD_470nm"), with = F]))

### 2019-08-18 R/drake_funcs.R

@@ -63,6 +63,7 @@ select_points <- function(aerpts, reg_polygon){
 
 #' read in reference grid, and turn into geometry points
 #' @param ref_file file dir of the grid file in .fst
+#' 
 #' @return sf object of grid in WGS84
 #' 
 get_ref_grid <- function(ref_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_grid_201906.fst"){
@@ -72,29 +73,42 @@ get_ref_grid <- function(ref_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_
 
 
 
-#' find points in buffer within dist
-#' @param aerpts The NEMIA Aeronet sites points 
-#' @param refpts The loaded referecen grid points by \code{\function{get_ref_grid}}  
-#' @return candidate_refpts
+#' find points in a broad buffer, and match the nearest grid point to Aeronet sites.
 #' 
-points_in_buffer <- function(aerpts, refpts, dist = 1500){
+#' @param aerpts The NEMIA Aeronet sites points 
+#' @param refpts The loaded refereence grid points by \code{\function{get_ref_grid}}  
+#' 
+#' @return aer_nearest: aeronets sites with cloest grid point (idLSTpair0)
+#' find near cells: find grid idLSTpair0 around 
+get_nearest_cell <- function(aerpts, refpts){
   aerptsNA <- st_transform(aerpts, crs = 2163) # US National Atlas
-  aer_regions <- aerptsNA %>% st_buffer(dist = dist) %>% st_union
+  # limit a 1500m buffer, I agree this part duplicates a little bit with `ref_in_buffer`
+  aer_regions <- aerptsNA %>% st_buffer(dist = 1500) %>% st_union
   refptsNA <- st_transform(refpts, crs = 2163) # US National Atlas
   refsub <- refpts[st_intersects(refptsNA, aer_regions, sparse = FALSE), ]
-  st_transform(refsub, crs = 4326) # return in wgs84
-}
-
-
-#' match the nearest grid point to Aeronet sites  
-#' @param aerpts The NEMIA Aeronet sites points 
-#' @param candidate_refpts all the points_in_buffer
-get_nearest_cell <- function(aerpts, candidate_refpts){
+  # `candidate_refpts` is just an intermediate to speed up the process
+  # to find nearby points 
[…]

### 2019-08-19 R/drake_plan_conus.Rmd
drake conus 30days

@@
   refgrid = get_ref_grid(),
-  # near MODIS cells within 300km
-  refsub = get_near_cellsid(conus_aer, refgrid),
+
   # aeronet sites with cloest MODIS grid
@@
   sel_aer_region = target(sel_data_bystation(aer_btw, aer_nearest_2),
-                      transform = map(aer_btw), .id = FALSE),
+                      transform = map(aer_btw, .id = FALSE)),
+
+  # near MODIS cells within 300km
+  refsub = target(get_near_cellsid(conus_aer, sel_aer_region, refgrid),
+                  transform = map(sel_aer_region, .id = FALSE)),
 
@@
   # control total lines, unable to read in all data together, too many rows
-  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS), refsub), transform = map(sat = c("terra", "aqua"))),
+  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS)), transform = map(sat = c("terra", "aqua"))),
   # join is the rolling-joined dataset with aeronet
@@
 ```{r, eval = F}
+make(nemia_plan, cache = ssd_cache)
 
@@
 make(nemia_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20, memory_strategy = "speed")
+     parallelism = "future", jobs = 20, cache_log_file = TRUE)
 

### 2019-08-19 R/drake_funcs.R

@@ -16,9 +16,10 @@ modify_aer_stns <- function(aer_stns0){
 #' get_conus_buff
 #' @param conus_file location of the sf CONUS file no Great Lake -- rds file
 #' @return sf object of CONUS shapfile
-get_conus_buff <- function(conus_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_sf_noGLakes.rds"){
+get_conus_buff <- function(conus_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_GLakes_buff_sf_poly_201906.rds"){
   readRDS(conus_file)
 }
+# "/data-belle/LST/MODIS.LST.C6/derived/conus_sf_noGLakes.rds"
 
 #' get_nemia_buff
 #' @param states_file US states shp file
@@ -99,10 +100,12 @@ get_nearest_cell <- function(aerpts, refpts){
 #' get nearby id names within 300 km 
 #' 
 #' @return geom points,df,dt, varnames: "idLSTpair0" "geometry"
-get_near_cellsid <- function(aerpts, refpts){
+get_near_cellsid <- function(aerpts, sel_aer_region, refpts){
+  # limit the gridcell by useful aeronet sites in the first place:
+  aerpts = aerpts[aerpts$Site_Name%in%unique(sel_aer_region$AERONET_Site_Name),]
   aerptsNA <- st_transform(aerpts, crs = 2163) # US National Atlas
-  # limit a 1500m buffer, I agree this part duplicates a little bit with `ref_in_buffer`
-  aer_regions <- aerptsNA %>% st_buffer(dist = 300000) %>% st_union
+  # limit to useful buffer, I agree this part duplicates a little bit with `ref_in_buffer`
+  aer_regions <- aerptsNA %>% st_buffer(dist = 270000) %>% st_union
   refptsNA <- st_transform(refpts, crs = 2163) # US National Atlas
   refsub <- refpts[st_intersects(refptsNA, aer_regions, sparse = FALSE), ]
   return(refsub)
@@ -215,16 +218,17 @@ interpolate_aod <- function(aer_data, aer_nearest_2){
 #' this function need to be replaced in the future, as for now 
 #' we only read one-year data
 #' @param sat input "terra" or "aqua", if sat !="terra", load aqua
-read_mcd19 <- function(sat = "terra", filepath, refsub){
+read_mcd19 <- function(sat = "terra", filepath){
[…]

### 2019-08-19 AOD_CONUS_Report.Rmd
the version that run 30days model

@@
 library(drake) 
+loadd(plotAOD, cache = ssd_cache)
+n_vars = length(plotAOD)
+fig_height = ceiling(n_vars/2)*3.5
+```
 
+# AOD Scatterplot
+
+```{r, fig.width = 10, fig.height = fig_height}
+loadd(plotAOD, cache = ssd_cache)
+gridExtra::grid.arrange(grobs = plotAOD, ncol = 2)
 ```
@@
 # Terra
+
+## RMSE
 
@@
 
+## Scatterplot
+
+```{r, fig.width = 10, fig.height = fig_height}
+loadd(plotscatter, cache = ssd_cache)
+gridExtra::grid.arrange(grobs = plotscatter, ncol = 2)
+```
+
+## SHAP
+
 ```{r}
+readd(plotSHAP1, cache = ssd_cache)
+```
 
-readd(plotSHAP1_2, cache = ssd_cache)
-
+```{r, fig.width = 10, fig.height = fig_height}
+loadd(plotSHAP2, cache = ssd_cache)
+gridExtra::grid.arrange(grobs = plotSHAP2, ncol = 2)
 ```
@@
 
+## RMSE
+
 ```{r}
@@
 readd(plot1_2, cache = ssd_cache)
-
 ```
@@
 
+## Scatterplot
+
+```{r, fig.width = 10, fig.height = fig_height}
+loadd(plotscatter_2, cache = ssd_cache)
+gridExtra::grid.arrange(grobs = plotscatter_2, ncol = 2)
+
+```
+
+## SHAP
+
 ```{r}
 readd(plotSHAP1_2, cache = ssd_cache)
+```
+
+```{r, fig.width = 10, fig.height = fig_height}
+loadd(plotSHAP2_2, cache = ssd_cache)
+gridExtra::grid.arrange(grobs = plotSHAP2_2, ncol = 2)
 

### 2019-08-19 R/drake_plan_conus.Rmd
the version that run 30days model

@@
 y_var_pred = "diff_AOD_pred"
+# how many features to print in scatterplot
+n_vars = 4
 
@@
   # buffers 
-  ref = target(ref_in_buffer(region_aer_m = sel_aer, refgrid),
+  ref = target(ref_in_buffer(sel_aer, refgrid),
              transform = map(sel_aer)),
@@
                  transform = map(cv_results,.id = FALSE)),
-  plotSHAP1 = target(plot_SHAP(cv_results),
+  # figures
+  plotAOD = target(plot_AOD(cv_results),
                  transform = map(cv_results,.id = FALSE)),
+  plotscatter = target(plot_scatter(cv_results),
+                 transform = map(cv_results,.id = FALSE)),
+
+  shap_long = target(get_SHAP_long(cv_results),
+                  transform = map(cv_results,.id = FALSE)),
+  plotSHAP1 = target(plot_SHAP(shap_long),
+                 transform = map(shap_long,.id = FALSE)),
+  plotSHAP2 = target(plot_SHAP_scatter(cv_results, shap_long),
+                 transform = map(cv_results, shap_long,.id = FALSE)),
   report = rmarkdown::render(
@@
 
+make(nemia_plan, cache = ssd_cache)
 ```

### 2019-08-19 R/drake_funcs.R

@@ -16,7 +16,8 @@ modify_aer_stns <- function(aer_stns0){
 #' get_conus_buff
 #' @param conus_file location of the sf CONUS file no Great Lake -- rds file
 #' @return sf object of CONUS shapfile
-get_conus_buff <- function(conus_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_GLakes_buff_sf_poly_201906.rds"){
+get_conus_buff <- function(conus_file = 
+  "/data-belle/LST/MODIS.LST.C6/derived/conus_GLakes_buff_sf_poly_201906.rds"){
   readRDS(conus_file)
 }
 # "/data-belle/LST/MODIS.LST.C6/derived/conus_sf_noGLakes.rds"
@@ -76,38 +77,40 @@ get_ref_grid <- function(ref_file = "/data-belle/LST/MODIS.LST.C6/derived/conus_
 
 #' find points in a broad buffer, and match the nearest grid point to Aeronet sites.
 #' 
-#' @param aerpts The NEMIA Aeronet sites points 
-#' @param refpts The loaded refereence grid points by \code{\function{get_ref_grid}}  
+#' @param conus_aer The NEMIA Aeronet sites points 
+#' @param refgrid The loaded refereence grid points by \code{\function{get_ref_grid}}  
 #' 
-#' @return aer_nearest: aeronets sites with cloest grid point (idLSTpair0)
+#' @return conus_aer: aer_nearest, aeronets sites with cloest grid point (idLSTpair0)
 #' find near cells: find grid idLSTpair0 around 
-get_nearest_cell <- function(aerpts, refpts){
-  aerptsNA <- st_transform(aerpts, crs = 2163) # US National Atlas
-  # limit a 1500m buffer, I agree this part duplicates a little bit with `ref_in_buffer`
-  aer_regions <- aerptsNA %>% st_buffer(dist = 1500) %>% st_union
-  refptsNA <- st_transform(refpts, crs = 2163) # US National Atlas
-  refsub <- refpts[st_intersects(refptsNA, aer_regions, sparse = FALSE), ]
-  # `candidate_refpts` is just an intermediate to speed up the process
+get_nearest_cell <- function(conus_aer, refgrid){
+  conus_aerNA <- st_transform(conus_aer, crs = 2163) # US National Atlas
[…]

### 2019-08-21 R/drake_plan_conus.Rmd
[read_mcd10_one]

--- 2019-08-19 R/drake_plan_conus.Rmd
@@
 mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
+out_dir = "/data-coco/mcd19/fst/conus/2018_AOD_newVar/"
 # point to cache: 
@@
 y_var_pred = "diff_AOD_pred"
-
+# how many features to print in scatterplots
+n_vars = 4
+# how many MCD days (files) to work on:
+MCD_files_i = c(1:30)
 # CONUS Drake Plans  ------------------------------------------------------------
 source(here("R/drake_funcs.R"))
-nemia_plan <- drake_plan(
+conus_plan <- drake_plan(
   # 1.  Aeroent ---------------------------------------------------------------------
@@
   # control total lines, unable to read in all data together, too many rows
-  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS)), transform = map(sat = c("terra", "aqua"))),
+  mcd = target(read_mcd19_one(i, sat = sat, file_in(mcd19path_CONUS)), 
+               transform = cross( i = !!MCD_files_i, 
+                                  sat = c("terra", "aqua")),
+               format = "fst"),
   # join is the rolling-joined dataset with aeronet
@@
   # buffers 
-  ref = target(ref_in_buffer(region_aer_m = sel_aer, refgrid),
+  ref = target(ref_in_buffer(sel_aer, refgrid),
              transform = map(sel_aer)),
   # sites
-  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
-  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),
+  p = target(get_site_list(sel_aer), transform = map(sel_aer)),
+  n = target(get_site_name(sel_aer), transform = map(sel_aer)),
 
@@
                transform = map(join, mcd, sub_grid, .id = c(join))),
+  # deliver two rowbinded dataset: for terra and aqua 
+  # binded = target(rbindlist(new_var),
+  #                 transform = combine(new_var, .by = c(i, sat))),
+  # `.id_chr` let you save the file using target names:
+  out = target(write.fst(new_var, path = file.path(file_out(out_dir), .id_chr)), transform = map(new_var))
+)
 
-  # 5. CV -----------------------------------------------------------
-  cv_results = target(run_cv(dt = new_var),transform = map(new_var)),
-
-  # 6. Report -----------------------------------------------------------
-  print1 = target(report_rmse(cv_results),
-                  transform = map(cv_results,.id = FALSE)),
-  plot1 = target(plot_rmse(cv_results),
-                 transform = map(cv_results,.id = FALSE)),
-  plotSHAP1 = target(plot_SHAP(cv_results),
-                 transform = map(cv_results,.id = FALSE)),
-  report = rmarkdown::render(
-    knitr_in("AOD_CONUS_Report.Rmd"),
-    output_file = file_out("AOD_CONUS_Report.html"))
-)
-nemia_plan
-#drake_plan_source(nemia_plan)
-nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
+# 
+# conus_plan_cv <- drake_plan(
+#   # 5. CV -----------------------------------------------------------
+#   cv_results = target(run_cv(dt = new_var),transform = map(new_var)),
+#   
+#   # 6. Report -----------------------------------------------------------
+#   print1 = target(report_rmse(cv_results),
+#                   transform = map(cv_results,.id = FALSE)),
+#   plot1 = target(plot_rmse(cv_results),
+#                  transform = map(cv_results,.id = FALSE)),
+#   # figures
+#   plotAOD = target(plot_AOD(cv_results),
+#                  transform = map(cv_results,.id = FALSE)),
+#   plotscatter = target(plot_scatter(cv_results),
+#                  transform = map(cv_results,.id = FALSE)),
+#   
+#   shap_long = target(get_SHAP_long(cv_results),
+#                   transform = map(cv_results,.id = FALSE)),
+#   plotSHAP1 = target(plot_SHAP(shap_long),
+#                  transform = map(shap_long,.id = FALSE)),
+#   plotSHAP2 = target(plot_SHAP_scatter(cv_results, shap_long),
+#                  transform = map(cv_results, shap_long,.id = FALSE)),
+#   report = rmarkdown::render(
+#     knitr_in("AOD_CONUS_Report.Rmd"),
+#     output_file = file_out("AOD_CONUS_Report.html"))
+# )
+conus_plan
+nrow(conus_plan)
+#drake_plan_source(conus_plan)
+if(nrow(conus_plan)<100){
+nemia_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
 # vis_drake_graph(nemia_config, from = names(nemia_config$layout))
 vis_drake_graph(nemia_config, ncol_legend = 0)
-
+}
 ```
@@
 ```{r, eval = F}
-make(nemia_plan, cache = ssd_cache)
+# make(conus_plan, cache = ssd_cache)
 
+time0 <- Sys.time()
 future::plan(future::multiprocess)
 options(future.globals.maxSize= 100*1024^3) 
-make(nemia_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20, cache_log_file = TRUE)
+make(conus_plan, cache = ssd_cache,
+     parallelism = "future", jobs = 5)
+Sys.time() - time0
 
@@
 # to see the updated graph, run config again
-nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
+nemia_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
 vis_drake_graph(nemia_config)

### 2019-08-21 R/drake_funcs.R

@@ -237,9 +237,22 @@ read_mcd19 <- function(sat = "terra", filepath){
   dt = rbindlist(lapply(lst_files, readfile))
   dt[,c("x", "y", "inNEMIA"):=NULL]
   dt[, join_time:=overpass_time] # join later using `overpass_time`
-  dt[, sat := choose]
   setnames(dt, "idLSTpair0", "nearest_refgrid") # rename for join later 
   setkey(dt, nearest_refgrid, join_time)
+  return(dt)
+}
+#' read one file every time 
+read_mcd19_one <- function(sat = "terra", i, filepath){
+  
+  if (sat != "terra") choose = "A" else choose = "T" # load terra by default
+  lst_files <- list.files(path = filepath, pattern = choose, full.names = T)
+    dt = read.fst(lst_files[i], as.data.table = T)
+    # return(t[idLSTpair0%in%refsub$idLSTpair0])
+    dt[,c("x", "y", "inNEMIA"):=NULL]
+    dt[, join_time:=overpass_time] # join later using `overpass_time`
+    dt[, sat := choose]
+    setnames(dt, "idLSTpair0", "nearest_refgrid") # rename for join later 
+    setkey(dt, nearest_refgrid, join_time)
   return(dt)
 }
 

### 2019-08-22 AOD_CONUS_Report.Rmd
CONUS model using one-year data done

@@
 ---
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
+```
 
@@
 library(drake) 
-loadd(plotAOD, cache = ssd_cache)
-n_vars = length(plotAOD)
+loadd(plotscatter, cache = ssd_cache)
+n_vars = length(plotscatter)
 fig_height = ceiling(n_vars/2)*3.5
@@
 
-```{r, fig.width = 10, fig.height = fig_height}
+```{r, fig.width = 10, fig.height = 7}
 loadd(plotAOD, cache = ssd_cache)
@@
 # top features
-loadd("cv_results_new_var_join_mcd_terra", cache = ssd_cache)
+loadd("cv_results_terra", cache = ssd_cache)
 # rferesults = cv_results_new_var_join_mcd_terra
-cv_results_new_var_join_mcd_terra$features_rank_rfe
+cv_results_terra$features_rank_rfe
 ```
@@
 # top features
-loadd("cv_results_new_var_join_mcd_aqua", cache = ssd_cache)
-cv_results_new_var_join_mcd_aqua$features_rank_rfe
+loadd("cv_results_aqua", cache = ssd_cache)
+cv_results_aqua$features_rank_rfe
 ```

### 2019-08-22 R/drake_plan_conus.Rmd
CONUS model using one-year data done

--- 2019-08-19 R/drake_plan_conus.Rmd
@@
 ```{r}
-# See `drake_funcs.R` for functions used in drake plan #
+# See `drake_funcs.R` for functions used in this drake plan #
 
@@
 })
-# ** need to revise accordingly the path to your local Just_universal
+# ** need to revise accordingly the path to your local clone of Just_universal **
 source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
@@
 # Parameters ####
-
+# the aer site info. file:
 aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"
-
-aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
-aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # don't have time to play with Ethernet all day
-# the aer data with AOD470nm
-aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"
-
+# aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
+# aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' 
+# the aer data files:
 aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
@@
 date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
-geo_region = "NEMIA"
 sats = c("terra", "aqua")
-# MCD19 data: 
+
+# MCD19 data load and save output: 
 mcd19path = "/data-coco/mcd19/fst/nemia/2018/" # now there is only 2018 data 
 mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
-# point to cache: 
-ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake")
+MCD19_T_files <- list.files(path = mcd19path_CONUS, pattern = "T", full.names = T)
+MCD19_A_files <- list.files(path = mcd19path_CONUS, pattern = "A", full.names = T)
+out_dir = "/data-coco/mcd19/fst/conus/2018_AOD_newVar/"
+# how many MCD days (files) to work on:
+MCD_files_i = c(1:length(MCD19_T_files))
 
@@
 y_var_pred = "diff_AOD_pred"
-# how many features to print in scatterplot
-n_vars = 4
+# how many features to print in scatterplots
+n_vars = 10
 
+# point to cache: 
+ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake") 
+```
+
+# Plan part 1
+
+```{r}
 # CONUS Drake Plans  ------------------------------------------------------------
 source(here("R/drake_funcs.R"))
-nemia_plan <- drake_plan(
+conus_plan <- drake_plan(
   # 1.  Aeroent ---------------------------------------------------------------------
@@
 
-  # near MODIS cells within 300km
-  refsub = target(get_near_cellsid(conus_aer, sel_aer_region, refgrid),
-                  transform = map(sel_aer_region, .id = FALSE)),
-
   # 2. Interpolation -----------------------------------------------------------
@@
   # control total lines, unable to read in all data together, too many rows
-  mcd = target(read_mcd19(sat, file_in(mcd19path_CONUS)), transform = map(sat = c("terra", "aqua"))),
+  mcd = target(read_mcd19_one(i, sat = sat, file_in(mcd19path_CONUS)), 
+               transform = cross( i = !!MCD_files_i, 
+                                  sat = !!sats),
+               format = "fst"),
   # join is the rolling-joined dataset with aeronet
@@
 
-  # select aer sites in mcd19. further limit the stations
-  sel_aer = target(aer_in_mcd19(aer_nearest_2, mcd19 = join),
-             transform = map(join)),
+  # For calculating new variables
+  # selected aeronet 
+  sel_aer = target(get_conus_aer_used(conus_aer, sel_aer_region),
+                  transform = map(sel_aer_region, .id = FALSE)),
+
   # buffers 
@@
   # sites
-  p = target(get_site_list(region_aer_m = sel_aer), transform = map(sel_aer)),
-  n = target(get_site_name(region_aer_m = sel_aer), transform = map(sel_aer)),
+  p = target(get_site_list(sel_aer), transform = map(sel_aer)),
+  n = target(get_site_name(sel_aer), transform = map(sel_aer)),
 
@@
   sub_grid = target(purrr_pmap(ref = ref, p = p, n = n),
-               transform = map(ref, p, n, .id = FALSE)),
+               transform = map(ref, p, n, .id = FALSE),
+               format = "fst"),
   # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
@@
                                  MODIS_all = mcd, refDT_sub = sub_grid),
-               transform = map(join, mcd, sub_grid, .id = c(join))),
+               transform = map(join, mcd, .id = c(join)),
+               format = "fst"),
+  # deliver two rowbinded dataset: for terra and aqua 
+  # I am not sure what this way doesn't work: 
+  # binded = target(rbindlist(new_var),
+  #                 transform = combine(new_var, .by = c(i, sat))),
+  # `.id_chr` let you save the file using target names:
+  out = target(write_new_var(new_var, id = .id_chr), transform = map(new_var))
+)
 
+conus_plan
+nrow(conus_plan)
+#drake_plan_source(conus_plan)
+if(nrow(conus_plan)<100){
+nemia_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
+# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
+vis_drake_graph(nemia_config, ncol_legend = 0)
+}
+```
+
+# Plan part 2
+
+```{r}
+# drake plan to run cv and output results 
+conus_plan_cv <- drake_plan(
   # 5. CV -----------------------------------------------------------
-  cv_results = target(run_cv(dt = new_var),transform = map(new_var)),
+  new_var = target(read_new_var(sat), transform = map(sat = !!sats)),
+  cv_results = target(run_cv(dt = new_var), transform = map(new_var, .id = sat)),
 
@@
 )
-nemia_plan
-#drake_plan_source(nemia_plan)
-nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
+nemia_config <- drake_config(conus_plan_cv, cache = ssd_cache) # show the dependency
 # vis_drake_graph(nemia_config, from = names(nemia_config$layout))
-vis_drake_graph(nemia_config, ncol_legend = 0)
-
-make(nemia_plan, cache = ssd_cache)
+vis_drake_graph(nemia_config)
 ```
 
-# Run the plan
+# Run the two plans
 
 ```{r, eval = F}
-make(nemia_plan, cache = ssd_cache)
+# make(conus_plan, cache = ssd_cache)
 
+time0 <- Sys.time()
 future::plan(future::multiprocess)
 options(future.globals.maxSize= 100*1024^3) 
-make(nemia_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20, cache_log_file = TRUE)
+make(conus_plan, cache = ssd_cache,
+     parallelism = "future", jobs = 20)
+Sys.time() - time0
 
-## Some simple analysis 
-loadd("new_var_join_mcd_terra", cache = ssd_cache)
-loadd("new_var_join_mcd_aqua", cache = ssd_cache)
-SHAPforxgboost::scatter.plot.diagonal(new_var_join_mcd_aqua, "Optical_Depth_047",
-                                      "Mean_AOD270km")
-cor(as.matrix(new_var_join_mcd_terra[,c(grep("Mean_AOD", names(new_var_join_mcd_terra), value = T), "AOD_470nm"), with = F]))
-## 
-loadd(cache = ssd_cache) # load all object
-
-# to see the updated graph, run config again
-nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
-vis_drake_graph(nemia_config)
+make(conus_plan_cv, cache = ssd_cache)
 

### 2019-08-22 R/drake_funcs.R

@@ -101,19 +101,6 @@ get_nearest_cell <- function(conus_aer, refgrid){
   return(conus_aer)
 }
 
-#' get nearby id names within 300 km 
-#' 
-#' @return geom points,df,dt, varnames: "idLSTpair0" "geometry"
-get_near_cellsid <- function(conus_aer, sel_aer_region, refgrid){
-  # limit the gridcell by useful aeronet sites in the first place:
-  conus_aer = conus_aer[conus_aer$Site_Name%in%unique(sel_aer_region$AERONET_Site_Name),]
-  conus_aerNA <- st_transform(conus_aer, crs = 2163) # US National Atlas
-  aer_regions <- conus_aerNA %>% st_buffer(dist = 270000) %>% st_union
-  refgridNA <- st_transform(refgrid, crs = 2163) # US National Atlas
-  refsub <- refgrid[st_intersects(refgridNA, aer_regions, sparse = FALSE), ]
-  return(refsub)
-}
-
 
 remove_site_on_water <- function(aer_nearest){
   aer_nearest[!aer_nearest$Site_Name%in%c("MVCO", "LISCO"),]
@@ -217,38 +204,18 @@ interpolate_aod <- function(aer_data, aer_nearest_2){
 
 
 # 4. rolling join MCD19 (Terra, Aqua) -----------------------------------------------------------
-#' read MCD19 for NEMIA
-#' this function need to be replaced in the future, as for now 
-#' we only read one-year data
-#' @param sat input "terra" or "aqua", if sat !="terra", load aqua
-read_mcd19 <- function(sat = "terra", filepath){
-  if (sat != "terra") choose = "A" else choose = "T" # load terra by default
-  lst_files <- list.files(path = filepath, pattern = choose, full.names = T)
-  # lst_files <- list.files(path = mcd19path_CONUS, pattern = "T", full.names = T)
-  ### for testing:
-  lst_files <- lst_files[1:90]
-  ### 
-  readfile <- function(x){
-    t = read.fst(x, as.data.table = T)
-    # return(t[idLSTpair0%in%refsub$idLSTpair0])
-    t
-  }
-  
-  dt = rbindlist(lapply(lst_files, readfile))
-  dt[,c("x", "y", "inNEMIA"):=NULL]
-  dt[, join_time:=overpass_time] # join later using `overpass_time`
-  setnames(dt, "idLSTpair0", "nearest_refgrid") # rename for join later 
-  setkey(dt, nearest_refgrid, join_time)
-  return(dt)
-}
[…]

### 2019-08-26 R/drake_plan_conus.Rmd
slight revise[drake_plan_conus] check drake version

@@
 # Libraries
+# if needed, update drake
+if(packageVersion("drake")<7.4) install.packages("drake")
 suppressPackageStartupMessages({
@@
 # point to cache: 
-ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake") 
+ssd_cache = drake_cache(path = "/scratch/cache/aeronet_drake") 
 ```

### 2019-08-28 R/drake_plan_conus.Rmd
revise shap dependence plot

@@
 # if needed, update drake
-if(packageVersion("drake")<7.4) install.packages("drake")
+if(packageVersion("drake") < "7.4") install.packages("drake")
+if(packageVersion("SHAPforxgboost") < "0.0.2") devtools::install_github("liuyanguu/SHAPforxgboost")
+
 suppressPackageStartupMessages({

### 2019-08-28 R/drake_funcs.R

@@ -471,7 +471,7 @@ plot_SHAP <- function(shap_long){
 
 plot_SHAP_scatter <- function(rferesults, shap_long){
   var_list <- rferesults$features_rank_rfe[1:n_vars]
-  fig_list <- lapply(var_list, function(x)shap.plot.dependence.color(shap_long, x = x, y = x, 
+  fig_list <- lapply(var_list, function(x)shap.plot.dependence(shap_long, x = x, y = x, 
                                                                      color_feature = "MCD19_AOD_470nm"))
   fig_list
 }

### 2019-08-29 R/drake_plan_conus.Rmd
fix the issue that cv will always rerun itself, use hard copy

@@
 ```{r}
+source(here("R/drake_funcs.R"))
 # drake plan to run cv and output results 
@@
   new_var = target(read_new_var(sat), transform = map(sat = !!sats)),
-  cv_results = target(run_cv(dt = new_var), transform = map(new_var, .id = sat)),
+  new_dt = target(prepare_dt(new_var), transform = map(new_var, .id = FALSE)),
+  cv_results = target(run_cv(dt = new_dt), transform = map(new_dt, .id = sat)),
 
@@
 # make(conus_plan, cache = ssd_cache)
-
 time0 <- Sys.time()
@@
 
-make(conus_plan_cv, cache = ssd_cache)
+# Allowing `add_hist` in `plot_scatter` (or use plotly) requires set lock_envir to FALSE:
+make(conus_plan_cv, cache = ssd_cache, lock_envir = FALSE)
 

### 2019-08-29 R/drake_funcs.R

@@ -396,18 +396,22 @@ create_qc_vars <- function(dt){
   return(dt)
 }
 
+prepare_dt <- function(dt){
+  setnames(dt, "Optical_Depth_047", "MCD19_AOD_470nm", skip_absent=TRUE)
+  # The dependent variable: diff_AOD = MCD19 - AERONET = Optical_Depth_047 - AOD_470nm
+  dt[, diff_AOD := MCD19_AOD_470nm - AOD_470nm]
+  dt <- create_qc_vars(dt)
+  dt[, dayint:=as.integer(as.Date(day))]
+  return(dt)
+}
+
 #' run the cross-validation by station with RFE.
 #' 
 #' This function is the 4th layer wrap of Kodi's random search cv function
 #' I apologize for the confusion if anyone is planning to dig into it. 
 #' Please feel free to contact me. 
 #' 
-#' 
 run_cv <- function(dt){
-  setnames(dt, "Optical_Depth_047", "MCD19_AOD_470nm", skip_absent=TRUE)
-  # The dependent variable: diff_AOD = MCD19 - AERONET = Optical_Depth_047 - AOD_470nm
-  dt[, diff_AOD := MCD19_AOD_470nm - AOD_470nm]
-  
   # The predictors
   features = c("MCD19_AOD_470nm",
                "dayint", 
@@ -416,14 +420,10 @@ run_cv <- function(dt){
                do.call(paste0,expand.grid(
                  c("pNonNAAOD", "Mean_AOD", "diff_AOD"),
                                           paste0(c(10, 30, 90, 270),"km"))))
-  
-
-  dt <- create_qc_vars(dt)
-  dt[, dayint:=as.integer(as.Date(day))]
-  
   # cv
   cv_results <- run.k.fold.cv.rfe.wrap(
-                modeldt1 = dt, 
+    # n_rounds = 5, # this is for fast testing
+                data = dt, 
                 stn_var = "Site_Name",
                 features0 = features, 
                 sat = "", y_var = "diff_AOD", 

### 2019-08-29 R/drake_plan_conus.Rmd
add toc in rmarkdown::html_document

@@
     knitr_in("AOD_CONUS_Report.Rmd"),
-    output_file = file_out("AOD_CONUS_Report.html"))
+    output_file = file_out("AOD_CONUS_Report.html"),
+    output_format = rmarkdown::html_document(toc = TRUE))
 )

### 2019-08-29 R/drake_plan_conus.Rmd
add set.seed

@@
 ssd_cache = drake_cache(path = "/scratch/cache/aeronet_drake")
+set.seed(1234)
 ```
@@
 make(conus_plan_cv, cache = ssd_cache, lock_envir = FALSE)
-
 ```

### 2019-12-22 R/drake_plan_conus.Rmd
update path to Just_universal

@@
 # ** need to revise accordingly the path to your local clone of Just_universal **
-source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
-source("~/liuyanguu/Just_universal/code/xgboost_cv_RFE.R")
+source(here("../Just_universal/code/xgboost_cvtune.R"))
+source(here("../Just_universal/code/xgboost_cv_RFE.R"))
 

### 2019-12-26 R/drake_plan_conus.Rmd
use here for directories

@@
 conus_plan <- drake_plan(
-  # 1.  Aeroent ---------------------------------------------------------------------
+  # 1.  Aeronet ---------------------------------------------------------------------
 
@@
   report = rmarkdown::render(
-    knitr_in("AOD_CONUS_Report.Rmd"),
-    output_file = file_out("AOD_CONUS_Report.html"),
+    knitr_in(here("../AOD_CONUS_Report.Rmd")),
+    output_file = file_out(here("../AOD_CONUS_Report.html")),
     output_format = rmarkdown::html_document(toc = TRUE))

### 2019-12-26 R/drake_plan_nemia.Rmd
use here for directories

@@
 # ** need to revise accordingly the path to your local Just_universal
-source("~/liuyanguu/Just_universal/code/xgboost_cvtune.R")
-source("~/liuyanguu/Just_universal/code/xgboost_cv_RFE.R")
+source(here("../Just_universal/code/xgboost_cvtune.R"))
+source(here("../Just_universal/code/xgboost_cv_RFE.R"))
 

### 2019-12-26 R/drake_funcs.R

@@ -378,7 +378,7 @@ write_new_var <- function(new_var, id){
 read_new_var <- function(sat){
   new_files <- list.files(path = out_dir, pattern = sat, full.names = T)
   new_var <- rbindlist(lapply(new_files, function(x)read.fst(x, as.data.table = T)))
-  message("dimentions of the new_var data is ", paste(dim(new_var), collapse = ","))
+  message("dimensions of the new_var data are ", paste(dim(new_var), collapse = ","))
   return(new_var)
 }
 

### 2019-12-31 R/drake_plan_conus.Rmd
use data.table with fst

@@
                                   sat = !!sats),
-               format = "fst"),
+               format = "fst_dt"),
   # join is the rolling-joined dataset with aeronet
@@
                transform = map(ref, p, n, .id = FALSE),
-               format = "fst"),
+               format = "fst_dt"),
   # calculate new vars to obtain the dataset ready for modeling for terra and aqua, using join, mcd, sub_grid 
@@
                transform = map(join, mcd, .id = c(join)),
-               format = "fst"),
+               format = "fst_dt"),
   # deliver two rowbinded dataset: for terra and aqua 
@@
 if(nrow(conus_plan)<100){
-nemia_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
-# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
-vis_drake_graph(nemia_config, ncol_legend = 0)
+conus_config <- drake_config(conus_plan, cache = ssd_cache) # show the dependency
+# vis_drake_graph(conus_config, from = names(conus_config$layout))
+vis_drake_graph(conus_config, ncol_legend = 0)
 }
@@
   report = rmarkdown::render(
-    knitr_in(here("../AOD_CONUS_Report.Rmd")),
-    output_file = file_out(here("../AOD_CONUS_Report.html")),
+    knitr_in(here("AOD_CONUS_Report.Rmd")),
+    output_file = file_out(here("AOD_CONUS_Report.html")),
     output_format = rmarkdown::html_document(toc = TRUE))
 )
-nemia_config <- drake_config(conus_plan_cv, cache = ssd_cache) # show the dependency
-# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
-vis_drake_graph(nemia_config)
+conus_cv_config <- drake_config(conus_plan_cv, cache = ssd_cache) # show the dependency
+# vis_drake_graph(conus_cv_config, from = names(conus_cv_config$layout))
+vis_drake_graph(conus_cv_config)
 ```
@@
 # make(conus_plan, cache = ssd_cache)
-time0 <- Sys.time()
+(time0 <- Sys.time())
 future::plan(future::multiprocess)
@@
 make(conus_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20)
+     parallelism = "future", jobs = 20, garbage_collection = TRUE)
 Sys.time() - time0
@@
 # Allowing `add_hist` in `plot_scatter` (or use plotly) requires set lock_envir to FALSE:
+system.time(
   make(conus_plan_cv, cache = ssd_cache, lock_envir = FALSE)
+  )
 ```

### 2019-12-31 R/drake_funcs.R

@@ -226,6 +226,7 @@ read_mcd19_one <- function(sat = "terra", i, filepath){
 #' rolling join aer with MCD19
 #' @return MCD19 rolling join Aeronet 
 rolling_join <- function(mcd_sat, aer_data_wPred){
+  #setDT(mcd_sat, key = c("nearest_refgrid", "join_time"))
   mcd19 <- aer_data_wPred[mcd_sat, roll = 'nearest', nomatch = 0]
   # time difference
   mcd19[, diff_time_min:= as.numeric(overpass_time - stn_time)/60]
@@ -321,16 +322,19 @@ purrr_pmap <- function(ref, p, n){
 #' 
 #' 
 aod_MODIS_newVars <- function(aod_join_MODIS, MODIS_all, refDT_sub){
+  setDT(aod_join_MODIS)
   aod_join_MODIS <- aod_join_MODIS[!is.na(Optical_Depth_047),]
   aod_join_MODIS[, day:=format(stn_time, "%Y-%m-%d")]
   setkey(aod_join_MODIS, Site_Name)
   # join by site_name 
+  setDT(refDT_sub)
   setnames(refDT_sub, "nearest_refgrid", "buf_refgrid", skip_absent=TRUE)
   aod_join_buffer <- aod_join_MODIS[refDT_sub, allow.cartesian = T]
   message("The cartesian joined dataset between collocated Aeronet with MODIS sites data and base grid in each circles, dim is: ", 
         paste(dim(aod_join_buffer), collapse = " x "))
   
   # join back to the MODIS file using `buf_refgrid` in the joined dataset
+  setDT(MODIS_all)
   MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
   setnames(MODIS_all, c("nearest_refgrid", "Optical_Depth_047"), 
            c("buf_refgrid", "buffer_AOD_470"), skip_absent = T)
@@ -366,11 +370,11 @@ aod_MODIS_newVars <- function(aod_join_MODIS, MODIS_all, refDT_sub){
     setnames(aod_join_MODIS, "diff_AOD", paste0("diff_AOD",distx,"km"))
   }
   invisible(lapply(dist0, calculate.diff))
-  return(aod_join_MODIS)
+  if(nrow(aod_join_MODIS) > 0) return(aod_join_MODIS)
 }
  
 write_new_var <- function(new_var, id){
-  if (nrow(new_var)>0) write.fst(new_var, path = file.path(file_out(out_dir),id))
[…]

### 2020-01-03 R/drake_plan_conus.Rmd
2017, allows varying reporting days for T/A

@@
 # Parameters ####
+this_year = 2017
 # the aer site info. file:
@@
 aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
-date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
-date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
+date_start = "2017-01-01" # do not include in sel_data_bytime if don't want to limit
+date_end   = "2017-12-31" # do not include in sel_data_bytime if don't want to limit
 sats = c("terra", "aqua")
@@
 # MCD19 data load and save output: 
-mcd19path = "/data-coco/mcd19/fst/nemia/2018/" # now there is only 2018 data 
-mcd19path_CONUS = "/data-coco/mcd19/fst/conus/2018/"
+mcd19path = paste0("/data-coco/mcd19/fst/nemia/", this_year, "/")
+mcd19path_CONUS = paste0("/data-coco/mcd19/fst/conus/", this_year, "/")
 MCD19_T_files <- list.files(path = mcd19path_CONUS, pattern = "T", full.names = T)
 MCD19_A_files <- list.files(path = mcd19path_CONUS, pattern = "A", full.names = T)
-out_dir = "/data-coco/mcd19/fst/conus/2018_AOD_newVar/"
+out_dir = paste0("/data-coco/mcd19/fst/conus/", this_year, "_AOD_newVar/")
+if(!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
 # how many MCD days (files) to work on:
-MCD_files_i = c(1:length(MCD19_T_files))
+maxdays = as.integer(format(as.Date(paste0(this_year, "-12-31")), "%j"))
+MCD_files_i = c(1:maxdays)
 

### 2020-01-11 R/drake_plan_conus.Rmd
add memory strategy

@@
 # point to cache: 
-ssd_cache = drake_cache(path = "/scratch/cache/aeronet_drake")
+ssd_cache = drake_cache(path = "/data-coco/mcd19/cache/aeronet_drake")
 set.seed(1234)
@@
 make(conus_plan, cache = ssd_cache,
-     parallelism = "future", jobs = 20, garbage_collection = TRUE)
+     parallelism = "future", jobs = 5, garbage_collection = TRUE, 
+     keep_going = TRUE, memory_strategy = "preclean")
 Sys.time() - time0

### 2020-07-10 R/drake_plan_conus.Rmd
install/update drake & SHAPforxgboost

@@
 
+Handle checking for installed and out of date packages without attaching their namespace
+```{r check_and_install_pkgs}
+package_install <- function(pkg, version, github_source = NULL){
+  if(!is.null(github_source)){
+    remotes::install_github(github_source)
+  } else {
+    install.packages(pkg)
+  }}
+check_and_install <- function(pkg, version, github_source = NULL){
+  tryCatch(
+    if(packageVersion(pkg) < version) package_install(pkg, version, github_source), 
+  error = function(e) {
+    if(substr(e$message, 1, 26) == "there is no package called"){
+      package_install(pkg, version, github_source)
+    } else {
+      message(e)
+    }
+  })}
+```
+
 # Write the plan
-
 ```{r}
@@
 # if needed, update drake
-if(packageVersion("drake") < "7.4") install.packages("drake")
-if(packageVersion("SHAPforxgboost") < "0.0.2") devtools::install_github("liuyanguu/SHAPforxgboost")
+if(!require(remotes)) install.packages("remotes")
+check_and_install("drake", "7.4")
+check_and_install("SHAPforxgboost", "0.0.2", "liuyanguu/SHAPforxgboost")
 
@@
   library(SHAPforxgboost)
+  library(Just.universal)
 })
-# ** need to revise accordingly the path to your local clone of Just_universal **
-source(here("../Just_universal/code/xgboost_cvtune.R"))
-source(here("../Just_universal/code/xgboost_cv_RFE.R"))
 

### 2021-10-22 R/functions.R
render a basic report on all initial_cv targets

@@ -409,3 +409,17 @@ initial_cv_dart <- function(
   # output
   c(list(by_var = by_var, bin_list = bin_list, mDT_wPred = mDT), cv_results)
 }
+
+cv_reporting <- function(cv){
+  dt = cv$mDT_wPred
+  mae = function(v1, v2) mean(abs(v1 - v2))
+  mad = function(v1) mean(abs(v1 - median(v1)))
+  dt[, aod_hat := MCD19_AOD_470nm - diff_AOD_pred]
+
+  list(
+    mae_uncorrected = dt[, mae(MCD19_AOD_470nm, AOD_470nm)],
+    mae_corrected   = dt[, mae(aod_hat, AOD_470nm)],
+    mad_MCD19  = mad(dt$MCD19_AOD_470nm),
+    mad_aodhat = mad(dt$aod_hat)
+  )
+}

### 2021-10-22 _targets.R

@@ -45,7 +45,7 @@ region_values = list(regions = aoiname)
 date_table = dates_year(process_years)
 sat_values = list(sat = sats)
 
-list(
+set1 = list(
   tar_target(mcd19path, '/data-coco/mcd19/fst/conus_full'),
   tar_target(aer_stn_path, '/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt',
              format = 'file'),
@@ -96,6 +96,7 @@ list(
 
       # Model ####
       tar_map( # date range (year) mapping
+        unlist = FALSE,
         values = date_table,
         names = 'year',
         tar_target(modelinput, prepare_dt(mcd19_vars, date_range = dates),
@@ -112,3 +113,9 @@ list(
     )
   )
 )
+set1u = unlist(set1)
+combined = tar_combine(combined_cv, set1u[grep('initial_cv', names(set1u))],
+                       command = list(!!!.x))
+list(set1, combined,
+     tar_render(initial_cv_report, 'R/initial_cv_report.Rmd',
+                params = list(cv_names = names(combined_cv))))

### 2021-11-02 _targets.R
cv results in table

@@ -13,6 +13,7 @@ options(xgb.threads = 8) # used by xgboost_cv_RFE
 
 tar_option_set(
   packages = c('lubridate',
+               'stringr',
                'log4r',
                'data.table',
                'fst',

### 2021-11-05 R/functions.R
round CV summary statistics

@@ -471,12 +471,12 @@ cv_reporting <- function(cv){
   dt[, aod_hat := MCD19_AOD_470nm - diff_AOD_pred]
 
   list(
-    mae_uncorrected = dt[, mae(MCD19_AOD_470nm, AOD_470nm)],
-    mae_corrected   = dt[, mae(aod_hat, AOD_470nm)],
-    mad_MCD19  = mad(dt$MCD19_AOD_470nm),
-    mad_aodhat = mad(dt$aod_hat),
+    mae_uncorrected = round(dt[, mae(MCD19_AOD_470nm, AOD_470nm)],3),
+    mae_corrected   = round(dt[, mae(aod_hat, AOD_470nm)],3),
+    mad_MCD19  = round(mad(dt$MCD19_AOD_470nm),3),
+    mad_aodhat = round(mad(dt$aod_hat),3),
     stn_count = dt[, uniqueN(stn)],
     train_N = dt[, .N],
-    mean_daily_overpass = dt[, mean(overpass_index), by = aer_date][, mean(V1)]
+    mean_daily_overpass = round(dt[, mean(overpass_index), by = aer_date][, mean(V1)],3)
   )
 }

### 2021-11-05 _targets.R
run 2010:2019

@@ -41,7 +41,7 @@ source('R/xgboost_cv_RFE.R')
 # Targets ####
 # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-process_years = 2015:2019
+process_years = 2010:2019
 region_values = list(regions = aoiname)
 date_table = dates_year(process_years)
 sat_values = list(sat = sats)

### 2021-11-10 R/functions.R
Create cv summary data.table and load in Rmd.
Also make HTML table sortable using DT package.

@@ -464,6 +464,23 @@ initial_cv_dart <- function(
   c(list(by_var = by_var, bin_list = bin_list, mDT_wPred = mDT), cv_results)
 }
 
+#' Summarize CV statistics on all initial_cv objects
+cv_summary <- function(cv_list){
+  output = vector(mode = "list", length = length(cv_list))
+  for(i in 1:length(cv_list)){
+    cv_name = names(cv_list)[[i]]
+    cv = cv_list[[i]]
+    stats = cv_reporting(cv)
+    stats$sat <- str_extract(cv_name, 'terra|aqua')
+    stats$year <- substr(cv_name, 12, 15)
+    output[[i]] <- stats
+  }
+  outDT = rbindlist(lapply(output, as.data.table))
+  setkey(outDT, sat, year)
+  setcolorder(outDT)
+}
+
+#' Calculate CV statistics on a single initial_cv object
 cv_reporting <- function(cv){
   dt = cv$mDT_wPred
   mae = function(v1, v2) mean(abs(v1 - v2))

### 2021-11-10 _targets.R

@@ -155,9 +155,13 @@ combined_target = tar_combine(combined_cv, set1u[grep('initial_cv', names(set1u)
                               command = list(!!!.x))
 
 # Render CV report ####
-cv_report_target = tar_render(initial_cv_report, 'R/initial_cv_report.Rmd',
-                       params = list(cv_names = names(combined_cv)))
-
+report_targets = list(
+  tar_target(cv_summary_table,
+             cv_summary(combined_cv)),
+  tar_render(initial_cv_report,
+             'R/initial_cv_report.Rmd',
+             params = list(cv_names = names(combined_cv)))
+)
 
 # Final targets list ####
-list(set1_targets, combined_target, cv_report_target)
+list(set1_targets, combined_target, report_targets)

### 2022-02-03 R/functions.R
add difftime distribution to CV report

@@ -607,23 +607,33 @@ initial_cv_dart <- function(
 #' @return data.table summarizing CV statistics for each year and satellite
 cv_summary <- function(cv_list){
   cv_list = unlist(cv_list, recursive = FALSE)
-  output = vector(mode = "list", length = length(cv_list))
+  stats_list = vector(mode = "list", length = length(cv_list))
+  difftimes_list = vector(mode = "list", length = length(cv_list))
   for(i in 1:length(cv_list)){
     cv = cv_list[[i]]
     stats = cv_reporting(cv)
     stats$sat <- str_extract(names(cv_list)[[i]], 'terra|aqua')
     stats$year <- cv$mDT_wPred[1, year(aer_date)]
-    output[[i]] <- stats
+    stats_list[[i]] <- stats
+    difftimes_list[[i]] <- summary(as.numeric(abs(cv$mDT_wPred$rj_difftime)))
   }
-  outDT = rbindlist(lapply(output, as.data.table))
-  outDT[, MAE_pct_change :=
+  # prediction summary stats
+  statsDT = rbindlist(lapply(stats_list, as.data.table))
+  statsDT[, MAE_pct_change :=
           paste0(round((MAE_uncorr-MAE_corr)/MAE_uncorr, 2) * 100, '%')]
-  outDT[, MAD_pct_change :=
+  statsDT[, MAD_pct_change :=
           paste0(round((MAD_mcd19-MAD_aodhat)/MAD_mcd19, 2) * 100, '%')]
-  setkey(outDT, sat, year)
-  setcolorder(outDT, c('sat', 'year',
+  setcolorder(statsDT, c('sat', 'year',
                        'MAE_uncorr', 'MAE_corr', 'MAE_pct_change',
                        'rmse', 'MAD_mcd19', 'MAD_aodhat', 'MAD_pct_change'))
+  # difftime distribution
+  difftimeDT = rbindlist(lapply(difftimes_list, function(x) as.list(x)))
+  difftimeDT[, c('sat', 'year') := statsDT[, .(sat, year)]]
+  setcolorder(difftimeDT, c('sat', 'year'))
+
+  setkey(statsDT, sat, year)
+  setkey(difftimeDT, sat, year)
+  list(stats = statsDT, difftimes = difftimeDT)
 }
 
 #' Calculate CV statistics on a single \code{initial_cv_dart()} output list object

### 2022-02-03 _targets.R

@@ -199,11 +199,10 @@ combined_target = tar_combine(combined_cv, set1u[grep('initial_cv', names(set1u)
 
 # Render CV report ####
 report_targets = list(
-  tar_target(cv_summary_table,
+  tar_target(cv_summary_tables,
              cv_summary(combined_cv)),
   tar_render(initial_cv_report,
-             'R/initial_cv_report.Rmd',
-             params = list(cv_names = names(combined_cv)))
+             'R/initial_cv_report.Rmd')
 )
 
 # Final targets list ####

### 2022-03-03 R/functions.R
Don't split training by year.

@@ -94,25 +94,6 @@ get_stn_data <- function(aod_dir, stations, date_start = NULL, date_end = NULL){
   }
 }
 
-#' Generate a tibble with a column for year and a column with a list of every
-#' date in the year
-#'
-#' @param years vector of four-digit years
-#' @return a tibble with a "year" column and a "dates" column containing a list
-#'   of every date in the year
-dates_year <- function(years){
-  dates = lapply(years, function(y) seq.Date(as.Date(paste0(y, '-01-01')),
-                                            as.Date(paste0(y, '-12-31')), 1))
-  tibble::tibble(year = years, dates = dates)
-}
-
-dates_year_list <- function(years){
-  dates = lapply(years, function(y) seq.Date(as.Date(paste0(y, '-01-01')),
-                                            as.Date(paste0(y, '-12-31')), 1))
-  names(dates) <- years
-  dates
-}
-
 #' Assign a month index from 1970-01-01 to all observation dates.
 #'
 #' Month indexes are used to batch the extraction of training data into chunks and increase targets throughput.
@@ -474,7 +455,6 @@ create_qc_vars <- function(dt){
 prepare_dt <- function(dt, date_range = NULL){
   setnames(dt, "Optical_Depth_047", "MCD19_AOD_470nm", skip_absent=TRUE)
   if(!is.null(date_range)){
-    if(class(date_range) == 'list') date_range = date_range[[1]]
     dt = dt[aer_date %in% date_range, ]
   }
   # The dependent variable: diff_AOD = MCD19 - AERONET = Optical_Depth_047 - AOD_470nm
@@ -571,7 +551,6 @@ initial_cv_dart <- function(
 #'   satellites, and both objective functions
 #' @return data.table summarizing CV statistics for each year, loss, and satellite
 cv_summary <- function(cv_list){
-  cv_list = unlist(cv_list, recursive = FALSE)
   stats_list = vector(mode = "list", length = length(cv_list))
   difftimes_list = vector(mode = "list", length = length(cv_list))
   for(i in 1:length(cv_list)){

### 2022-03-03 _targets.R

@@ -46,8 +46,11 @@ source('R/xgboost_cv.R')
 # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
 process_years = 2003:2019
+all_dates = seq(
+    lubridate::make_date(min(process_years)),
+    lubridate::make_date(max(process_years), 12, 31),
+    by = 1)
 region_values = list(regions = aoiname)
-date_table = dates_year(process_years)
 sat_values = list(sat = sats)
 buffers_km = c(10, 30, 90, 270)
 agg_level = 10
@@ -81,9 +84,6 @@ set1_targets = list(
   tar_target(vrt_path,
              prepare_vrt_directory(intermediate.path())),
 
-  tar_target(dates_byyear,
-             dates_year_list(process_years)),
-
   # Area of interest ####
   tar_map( # region mapping
     values = region_values,
@@ -95,8 +95,6 @@ set1_targets = list(
                crop_refras_mcd(refgrid_path, mcd19path, aoiname = regions)),
 
     # Load AERONET data ####
-    tar_target(all_dates,
-               lubridate::as_date(unlist(date_table$dates))),
     tar_target(aer_nospace,
                sf::st_drop_geometry(aer)),
     tar_group_by(aer_bystation,
@@ -132,9 +130,7 @@ set1_targets = list(
 
       # Model ####
       tar_target(traindata,
-                 prepare_dt(mcd19_vars, date_range = dates_byyear),
-                 pattern = map(dates_byyear),
-                 iteration = 'list'),
+                 prepare_dt(mcd19_vars, date_range = all_dates)),
       tar_map(
         values = list(loss = c("l1", "l2")),
         tar_target(initial_cv,
@@ -143,7 +139,6 @@ set1_targets = list(
                                    y_var = "diff_AOD",
                                    features = features,
                                    stn_var = "Site_Name"),
-                   pattern = map(traindata),
                    iteration = 'list')),
       tar_target(full_model, dart_full(traindata,
                                        y_var = "diff_AOD",

### 2022-03-03 R/functions.R
Don't split CV results by year.

@@ -549,7 +549,7 @@ initial_cv_dart <- function(
 #'
 #' @param cv_list list of all CV output objects; may include every year, both
 #'   satellites, and both objective functions
-#' @return data.table summarizing CV statistics for each year, loss, and satellite
+#' @return data.table summarizing CV statistics for each loss and satellite
 cv_summary <- function(cv_list){
   stats_list = vector(mode = "list", length = length(cv_list))
   difftimes_list = vector(mode = "list", length = length(cv_list))
@@ -558,7 +558,6 @@ cv_summary <- function(cv_list){
     stats = cv_reporting(cv)
     stats$sat <- str_extract(names(cv_list)[[i]], 'terra|aqua')
     stats$loss <- str_match(names(cv_list)[[i]], 'cv_(l[12])_')[,2]
-    stats$year <- cv$mDT_wPred[1, year(aer_date)]
     stats_list[[i]] <- stats
     difftimes_list[[i]] <- round(summary(as.numeric(abs(cv$mDT_wPred$rj_difftime))),0)
   }
@@ -568,16 +567,16 @@ cv_summary <- function(cv_list){
           paste0(round((MAE_uncorr-MAE_corr)/MAE_uncorr, 2) * 100, '%')]
   statsDT[, MAD_pct_change :=
           paste0(round((MAD_mcd19-MAD_aodhat)/MAD_mcd19, 2) * 100, '%')]
-  setcolorder(statsDT, c('sat', 'year', 'loss',
+  setcolorder(statsDT, c('sat', 'loss',
                        'MAE_uncorr', 'MAE_corr', 'MAE_pct_change',
                        'rmse', 'MAD_mcd19', 'MAD_aodhat', 'MAD_pct_change'))
   # difftime distribution
   difftimeDT = rbindlist(lapply(difftimes_list, function(x) as.list(x)))
-  difftimeDT[, c('sat', 'year', 'loss') := statsDT[, .(sat, year, loss)]]
-  setcolorder(difftimeDT, c('sat', 'year', 'loss'))
+  difftimeDT[, c('sat', 'loss') := statsDT[, .(sat, loss)]]
+  setcolorder(difftimeDT, c('sat', 'loss'))
 
-  setkey(statsDT, sat, year, loss)
-  setkey(difftimeDT, sat, year, loss)
+  setkey(statsDT, sat, loss)
+  setkey(difftimeDT, sat, loss)
   list(stats = statsDT, difftimes = difftimeDT)
 }
 

### 2022-03-03 R/functions.R
Round CV results to 4 digits.

@@ -588,12 +588,13 @@ cv_reporting <- function(cv){
   mad = function(v1) mean(abs(v1 - median(v1)))
   dt[, aod_hat := MCD19_AOD_470nm - diff_AOD_pred]
 
+  r = function(x) round(x, 4)
   list(
-    MAE_uncorr = round(dt[, mae(MCD19_AOD_470nm, AOD_470nm)],3),
-    MAE_corr   = round(dt[, mae(aod_hat, AOD_470nm)],3),
-    rmse = round(cv$rmse_all_folds,3),
-    MAD_mcd19  = round(mad(dt$MCD19_AOD_470nm),3),
-    MAD_aodhat = round(mad(dt$aod_hat),3),
+    MAE_uncorr = r(dt[, mae(MCD19_AOD_470nm, AOD_470nm)]),
+    MAE_corr   = r(dt[, mae(aod_hat, AOD_470nm)]),
+    rmse = r(cv$rmse_all_folds),
+    MAD_mcd19  = r(mad(dt$MCD19_AOD_470nm)),
+    MAD_aodhat = r(mad(dt$aod_hat)),
     stn_count = dt[, uniqueN(stn)],
     train_N = dt[, .N]
   )

### 2022-03-03 R/functions.R
Clean up whitespace.

@@ -980,12 +980,12 @@ get_aoi_buffer <- function(aoiname){
     stop("Unsupported aoi name:", aoiname)
   )
   buff
-} 
+}
 
 #' get_conus_buff
 #' @param conus_file location of the sf CONUS file no Great Lake -- rds file
 #' @return sf object of CONUS shapefile
-get_conus_buff <- function(conus_file = 
+get_conus_buff <- function(conus_file =
   "/data-belle/LST/MODIS.LST.C6/derived/conus_GLakes_buff_sf_poly_201906.rds"){
   readRDS(conus_file)
 }

### 2022-04-10 R/CONUS_AOD.Rmd
update metrics; draft report with conus AOD results

---
title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basics, echo=FALSE}
library(targets)
library(data.table)
library(DT)
library(patchwork)
library(ggplot2)
library(sf)
```

```{r git_check, echo=FALSE}
{
s = function(cmd) suppressWarnings(system(cmd,
    intern = TRUE, ignore.stderr = TRUE))
commit = substr(s('git rev-parse HEAD'), 1, 8)
uncommitted = ''
if (!length(commit))
   commit = 'unknown'
else if(length(s('git diff-index HEAD --'))>0)
   uncommitted = 'with uncommitted changes'
}
```

Git commit: `r commit` `r uncommitted`

Rendered: `r Sys.time()`

## Objective

Evaluate the performance of the `Earth_obs_cleaning` software tool on MAIAC MCD19A2 for Terra and Aqua satellites across the Continental USA, 2000-2021

## Main findings from cross-validation

```{r table_cv_results, echo=FALSE}
tar_load(cv_summary_tables)
DT::datatable(cv_summary_tables$stat,
              rownames = FALSE,
              fillContainer = FALSE,
              autoHideNavigation = TRUE,
              options = list(pageLength = 50,
                             autoWidth = TRUE))
```

## Descriptives for the collocated dataset (Terra)

## Stratified results for Terra CV dataset
```{r stratified, echo=FALSE}
performance_aod <- function(dt, ...){
  performance_metrics(dt,
                      ground_truth = "AOD_470nm",
                      eo_raw = "MCD19_AOD_470nm",
                      eo_pred = "aodhat")
}
tar_load(initial_cv_l2_terra_conus)
dt <- initial_cv_l2_terra_conus$mDT_wPred
if(!"aodhat" %in% colnames(dt)){
  dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
}

dtwrapper <- function(dt){
  DT::datatable(dt,
                rownames = FALSE,
                fillContainer = FALSE,
                autoHideNavigation = TRUE,
                options = list(pageLength = 50,
                               autoWidth = TRUE))
}
```

### Performance by year
```{r strata_year, echo=FALSE}
dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
```

### Performance by month of year

```{r strata_month, echo=FALSE}
dtwrapper(dt[, performance_aod(.SD), keyby = month(aer_date)])
```

### Performance by `qa_best`

```{r strata_qa, echo=FALSE}
dtwrapper(dt[, performance_aod(.SD), by = qa_best])
```

### Performance by being over water

```{r strata_water, echo=FALSE}
dtwrapper(dt[, performance_aod(.SD), by = qa_lwsi])
```

### Performance by number of overpasses in collocation that site-day

```{r strata_overpasses, echo=FALSE}
dt[, overpasses := .N, by = c("Site_Name", "aer_date")]
dtwrapper(dt[, performance_aod(.SD), keyby = overpasses])
```

## Performance by NOAA climate regions

```{r noaa_setup, echo=FALSE}
noaa.regions = list(
  # https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-regions.php
    Central = c("IL", "IN", "KY", "MO", "OH", "TN", "WV"),
    East.North.Central = c("IA", "MI", "MN", "WI"),
    Northeast = c("DC", "CT", "DE", "ME", "MD", "MA", "NH", "NJ", "NY", "PA", "RI", "VT"),
      # Washington, DC, has no assigned NOAA climate region, so treat
      # it as part of Maryland.
    Northwest = c("ID", "OR", "WA"),
    South = c("AR", "KS", "LA", "MS", "OK", "TX"),
    Southeast = c("AL", "FL", "GA", "NC", "SC", "VA"),
    Southwest = c("AZ", "CO", "NM", "UT"),
    West = c("CA", "NV"),
    West.North.Central = c("MT", "NE", "ND", "SD", "WY"))

state.abb.plusdc = c(state.abb, "DC")
conus.states.plusdc = sort(setdiff(state.abb.plusdc, c("AK", "HI")))

stopifnot(identical(
    sort(unlist(noaa.regions, use.names = F)),
    conus.states.plusdc))

state.noaa.region = factor(
   sapply(state.abb.plusdc, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
usa = st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))
```

```{r get_noaa_regions, include=FALSE}
# assign AERONET stations to states
tar_load(aer_stations)
aer_stations_sf <- st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326")
sf::sf_use_s2(FALSE) # we were getting errors about the USA geometry otherwise
aer_stations_sf <- st_join(aer_stations_sf, usa)
dt <- merge(dt, aer_stations_sf, by = "Site_Name")
# some Site_Names don't fall within land polygons for whatever reason
dt[is.na(ID), table(Site_Name)]
# where are the sites that didn't fall within a state?
site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
ggplot() +
  geom_sf(data = usa, fill = NA) +
  geom_sf(data = site_nostate, color = "red")
# what proportion of data are not in a state?
dt[, proportions(table(is.na(ID)))]
# assign these to the closest state
matched <- st_join(site_nostate,
                   usa,
                   join=nngeo::st_nn, k=1)
# plot these matchups
ggplot() +
  geom_sf(data = usa, fill = NA) +
  geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()

# where do our data come from (by State)?
dt[, .(uniqueN(Site_Name), .N), by = ID][order(N, decreasing = T)]

# trying to get from state names to state abbreviations
dt[ , state.abb := state.abb[match(stringr::str_to_title(ID), state.name)]]
dt[is.na(state.abb), .(ID, Site_Name, geometry)][, table(Site_Name)]
dt[, state.noaa.region := factor(
   sapply(state.abb, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))]
# unmatched locations
dt[is.na(state.noaa.region), .(ID, state.abb, Site_Name, geometry)][unique(Site_Name), , mult = "first"]
```

```{r strata_noaa_regions, echo=FALSE}
# performance by NOAA climate regions
dtwrapper(dt[, performance_aod(.SD), keyby = state.noaa.region])

# map by NOAA regions

```

## Agreement Plots

*We can customize however we want*

```{r agreement_plot, echo=FALSE}
p1 <- ggplot(dt) +
  aes(AOD_470nm, MCD19_AOD_470nm) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth() +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

p2  <- ggplot(dt) +
  aes(AOD_470nm, aodhat) +
  geom_abline(linetype = "dashed") +
  geom_point(alpha = 0.01) +
  geom_smooth() +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
  theme_bw() +
  theme(aspect.ratio = 1)

p1 + p2
```

*Will add plot of agreement across bins of AOD*

### 2022-04-10 R/functions.R

@@ -555,7 +555,12 @@ cv_summary <- function(cv_list){
   difftimes_list = vector(mode = "list", length = length(cv_list))
   for(i in 1:length(cv_list)){
     cv = cv_list[[i]]
-    stats = cv_reporting(cv)
+    dt = data.table::copy(cv$mDT_wPred)
+    dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
+    stats = performance_metrics(dt,
+                                ground_truth = "AOD_470nm",
+                                eo_raw = "MCD19_AOD_470nm",
+                                eo_pred = "aodhat")
     stats$sat <- str_extract(names(cv_list)[[i]], 'terra|aqua')
     stats$loss <- str_match(names(cv_list)[[i]], 'cv_(l[12])_')[,2]
     stats_list[[i]] <- stats
@@ -563,13 +568,9 @@ cv_summary <- function(cv_list){
   }
   # prediction summary stats
   statsDT = rbindlist(lapply(stats_list, as.data.table))
-  statsDT[, MAE_change :=
-          paste0(round((MAE_uncorr-MAE_corr)/MAE_uncorr, 2) * 100, '%')]
-  statsDT[, MAD_change :=
-          paste0(round((MAD_mcd19-MAD_aodhat)/MAD_mcd19, 2) * 100, '%')]
-  setcolorder(statsDT, c('sat', 'loss',
-                       'MAE_uncorr', 'MAE_corr', 'MAE_change',
-                       'rmse', 'MAD_mcd19', 'MAD_aodhat', 'MAD_change'))
+  # setcolorder(statsDT, c('sat', 'loss',
+  #                      'MAE_uncorr', 'MAE_corr', 'MAE_change',
+  #                      'rmse', 'MAD_mcd19', 'MAD_aodhat', 'MAD_change'))
   # difftime distribution
   difftimeDT = rbindlist(lapply(difftimes_list, function(x) as.list(x)))
   difftimeDT[, c('sat', 'loss') := statsDT[, .(sat, loss)]]
@@ -580,24 +581,39 @@ cv_summary <- function(cv_list){
   list(stats = statsDT, difftimes = difftimeDT)
 }
 
-#' Calculate CV statistics on a single \code{initial_cv_dart()} output list object
[…]

### 2022-04-10 _targets.R

@@ -198,5 +198,11 @@ report_targets = list(
              'R/initial_cv_report.Rmd')
 )
 
+# Render CONUS AOD results ###
+paper_conus_targets = list(
+  tar_render(paper_conus,
+             'R/CONUS_AOD.Rmd',
+             packages = c('sf', 'DT', 'patchwork'))
+)
 # Final targets list ####
-list(set1_targets, combined_target, report_targets)
+list(set1_targets, combined_target, report_targets, paper_conus_targets)

### 2022-04-13 R/functions.R
drop water pixels from training; keep QA field

@@ -439,7 +439,6 @@ create_qc_vars <- function(dt){
   dt[qa_bits==8, qa_lwsi := "water"]
   dt[qa_bits==16, qa_lwsi := "snow"]
   dt[qa_bits==24, qa_lwsi := "ice"]
-  dt[, qa_bits:= NULL]
   return(dt)
 }
 
@@ -462,6 +461,8 @@ prepare_dt <- function(dt, date_range = NULL){
   dt <- create_qc_vars(dt)
   dt[, dayint := as.integer(as.Date(overpass_time))]
   dt = dt[!is.na(get(y_var))]
+  # drop overpass matchups for pixels classified as water
+  dt = dt[qa_lwsi != "water"]
   dt
 }
 

### 2022-04-13 R/CONUS_AOD.Rmd
update CONUS results report

@@
 library(sf)
+library(knitr)
 ```
@@
 ### Performance by year
-```{r strata_year, echo=FALSE}
+```{r strata_year, echo=FALSE, eval=TRUE}
 dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
@@
 
-## Performance by NOAA climate regions
+### Performance by NOAA climate regions
 
@@
 site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
-ggplot() +
-  geom_sf(data = usa, fill = NA) +
-  geom_sf(data = site_nostate, color = "red")
+# ggplot() +
+#   geom_sf(data = usa, fill = NA) +
+#   geom_sf(data = site_nostate, color = "red")
 # what proportion of data are not in a state?
@@
 # plot these matchups
-ggplot() +
-  geom_sf(data = usa, fill = NA) +
-  geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()
+# ggplot() +
+#   geom_sf(data = usa, fill = NA) +
+#   geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()
 
@@
 
-```{r agreement_plot, echo=FALSE}
+```{r agreement_plot, warning=FALSE, echo=FALSE}
 p1 <- ggplot(dt) +
@@
   geom_point(alpha = 0.01) +
-  geom_smooth() +
+  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
   scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
@@
 *Will add plot of agreement across bins of AOD*
+
+## SHAP interpretation ####
+In decreasing order (from Terra collocated cross-validation predictions)
+```{r shap_rank, echo=FALSE}
+# what do the SHAPs tell us?
+tar_load(initial_cv_l2_terra_conus)
+shaps <- initial_cv_l2_terra_conus$"shap_score"
+# sort by importance
+kable(sort(sapply(abs(shaps), mean), decreasing = TRUE), 
+      format = "html",
+      table.attr = "style='width:40%;'",
+      digits = 4, col.names = "Mean Abs SHAP")
+
+```

### 2022-04-19 R/CONUS_AOD.Rmd
update report for paper

@@
 output: html_document
+bibliography: references.json
 ---
@@
 ## Stratified results for Terra CV dataset
+
 ```{r stratified, echo=FALSE}
-performance_aod <- function(dt, ...){
-  performance_metrics(dt,
-                      ground_truth = "AOD_470nm",
-                      eo_raw = "MCD19_AOD_470nm",
-                      eo_pred = "aodhat")
-}
 tar_load(initial_cv_l2_terra_conus)
@@
 }
-
-dtwrapper <- function(dt){
-  DT::datatable(dt,
-                rownames = FALSE,
-                fillContainer = FALSE,
-                autoHideNavigation = TRUE,
-                options = list(pageLength = 50,
-                               autoWidth = TRUE))
-}
 ```
@@
 ### Performance by year
+
 ```{r strata_year, echo=FALSE, eval=TRUE}
@@
 
-*We can customize however we want*
-
 ```{r agreement_plot, warning=FALSE, echo=FALSE}
-p1 <- ggplot(dt) +
+p1 <- ggplot(dt[AOD_470nm < 1 & MCD19_AOD_470nm < 1]) +
   aes(AOD_470nm, MCD19_AOD_470nm) +
@@
   theme_bw() +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) + 
+  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | MCD19_AOD_470nm > 1, .N], " AOD values outside [0, 1]"))
 
-p2  <- ggplot(dt) +
+p2  <- ggplot(dt[AOD_470nm < 1 & aodhat > 0 & aodhat < 1]) +
   aes(AOD_470nm, aodhat) +
@@
   theme_bw() +
-  theme(aspect.ratio = 1)
+  theme(aspect.ratio = 1) + 
+  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | aodhat > 1 | aodhat < 0, .N], " AOD values outside [0, 1]"))
 
@@
 
-## SHAP interpretation ####
+## SHAP interpretation
+
 In decreasing order (from Terra collocated cross-validation predictions)
+
 ```{r shap_rank, echo=FALSE}
@@
       format = "html",
-      table.attr = "style='width:40%;'",
-      digits = 4, col.names = "Mean Abs SHAP")
+      table.attr = "style='width:30%;'",
+      digits = 3, col.names = "Mean Abs SHAP")
+```
 
-```
+*End of report*

### 2022-04-19 R/paper_functions.R

@@ -1,0 +1,21 @@
+# contains functions used in generating the CONUS_AOD paper
+
+#' Print a nice html table using the `DT` package
+#'
+dtwrapper <- function(dt){
+  DT::datatable(dt,
+                rownames = FALSE,
+                fillContainer = FALSE,
+                autoHideNavigation = TRUE,
+                options = list(pageLength = 50,
+                               autoWidth = TRUE))
+}
+
+#' Performance metrics for MCD19A2 AOD
+#'
+performance_aod <- function(dt, ...){
+  performance_metrics(dt,
+                      ground_truth = "AOD_470nm",
+                      eo_raw = "MCD19_AOD_470nm",
+                      eo_pred = "aodhat")
+}

### 2022-04-19 _targets.R

@@ -45,7 +45,7 @@ source('R/globals.R')
 source('R/data.R')
 source('R/functions.R')
 source('R/xgboost_cv.R')
-
+source('R/paper_functions.R')
 # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 # Targets ####
 # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

### 2022-04-19 R/CONUS_AOD.Rmd
assign non-land stations to closest state

@@
 #   geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()
+dt[as.data.table(matched, key = "Site_Name"), ID := i.ID]
 
@@
 # trying to get from state names to state abbreviations
-dt[ , state.abb := state.abb[match(stringr::str_to_title(ID), state.name)]]
-dt[is.na(state.abb), .(ID, Site_Name, geometry)][, table(Site_Name)]
+# ugly capitalization
+dt[ , state.abb := state.abb.plusdc[match(stringr::str_to_title(ID), c(state.name, "District Of Columbia"))]]
 dt[, state.noaa.region := factor(
@@
    levels = names(noaa.regions))]
-# unmatched locations
-dt[is.na(state.noaa.region), .(ID, state.abb, Site_Name, geometry)][unique(Site_Name), , mult = "first"]
 ```

### 2022-04-21 R/CONUS_AOD.Rmd
switch CONUS report to pdf

@@
 title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
-output: html_document
+header-includes:
+  - \usepackage{float}
+  - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
+output: pdf_document
+always_allow_html: true
 bibliography: references.json
@@
 library(knitr)
+library(kableExtra)
+# library(flextable)
+# set_flextable_defaults(fonts_ignore=TRUE)
 ```
@@
 tar_load(cv_summary_tables)
-DT::datatable(cv_summary_tables$stat,
-              rownames = FALSE,
-              fillContainer = FALSE,
-              autoHideNavigation = TRUE,
-              options = list(pageLength = 50,
-                             autoWidth = TRUE))
+result <- (cv_summary_tables$stat)
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 ```
@@
 ## Descriptives for the collocated dataset (Terra)
+
+*still to be added*
 
@@
 ```{r strata_year, echo=FALSE, eval=TRUE}
-dtwrapper(dt[, performance_aod(.SD), keyby = year(aer_date)])
+result <- dt[, performance_aod(.SD), keyby = year(aer_date)]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 ```
@@
 ```{r strata_month, echo=FALSE}
-dtwrapper(dt[, performance_aod(.SD), keyby = month(aer_date)])
+result <- dt[, performance_aod(.SD), keyby = month(aer_date)]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 ```
@@
 ```{r strata_qa, echo=FALSE}
-dtwrapper(dt[, performance_aod(.SD), by = qa_best])
+result <- dt[, performance_aod(.SD), by = qa_best]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 ```
@@
 
-```{r strata_water, echo=FALSE}
-dtwrapper(dt[, performance_aod(.SD), by = qa_lwsi])
+```{r strata_water, echo=FALSE, fig.pos="H"}
+result <- dt[, performance_aod(.SD), by = qa_lwsi]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 ```
@@
 
-```{r strata_overpasses, echo=FALSE}
+```{r strata_overpasses, echo=FALSE, fig.pos="H"}
 dt[, overpasses := .N, by = c("Site_Name", "aer_date")]
-dtwrapper(dt[, performance_aod(.SD), keyby = overpasses])
+result <- dt[, performance_aod(.SD), keyby = overpasses]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+
 ```
@@
 
-```{r strata_noaa_regions, echo=FALSE}
+```{r strata_noaa_regions, echo=FALSE, fig.pos="H"}
 # performance by NOAA climate regions
-dtwrapper(dt[, performance_aod(.SD), keyby = state.noaa.region])
+result <- dt[, performance_aod(.SD), keyby = state.noaa.region]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
+  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
 
@@
 
-```{r agreement_plot, warning=FALSE, echo=FALSE}
+```{r agreement_plot, echo=FALSE, dev='png', dpi=300}
 p1 <- ggplot(dt[AOD_470nm < 1 & MCD19_AOD_470nm < 1]) +
@@
 
+------------------------------------------------------------------------
+
 ## SHAP interpretation
@@
 # sort by importance
-kable(sort(sapply(abs(shaps), mean), decreasing = TRUE), 
-      format = "html",
-      table.attr = "style='width:30%;'",
-      digits = 3, col.names = "Mean Abs SHAP")
+df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
+kbl(df_shap, col.names = "Mean |SHAP|")
+# df_shap$variable <- row.names(df_shap)
+# ft_shap <- flextable(df_shap[, c(2, 1)])
+# ft_shap <- set_header_labels(ft_shap, list(variable = "Variable", meanabsshap = "Mean Absolute SHAP"))
+# ft_shap
 ```

### 2022-04-21 R/functions.R

@@ -537,7 +537,7 @@ cv_summary <- function(cv_list){
 #'
 #' Comparative cross-validation performance metrics with raw and corrected values.
 #' Takes a \code{data.table} and can be run across strata with a \code{by} statement.
-performance_metrics <- function(dt, ground_truth, eo_raw, eo_pred){
+performance_metrics <- function(dt, ground_truth, eo_raw, eo_pred, digits = 3){
   truth = quote(get(ground_truth))
   raw = quote(get(eo_raw))
   pred = quote(get(eo_pred))
@@ -546,7 +546,7 @@ performance_metrics <- function(dt, ground_truth, eo_raw, eo_pred){
   mse = function(v1, v2) mean((v1 - v2)^2)
   # rmse = function(v1, v2) sqrt(mean((v1 - v2)^2))
   mean_bias = function(v1, v2) mean(v1 - v2)
-  r = function(x) round(x, 4)
+  r = function(x) round(x, digits)
   list(
     # MAE_raw = r(mae(dt[, eval(raw)], dt[, eval(truth)])),
     # MAE_pred   = r(mae(dt[, eval(pred)], dt[, eval(truth)])),
@@ -558,7 +558,7 @@ performance_metrics <- function(dt, ground_truth, eo_raw, eo_pred){
     SD_truth   = r(sd(dt[, eval(truth)])),
     SD_pred  = r(sd(dt[, eval(pred)])),
     SD_raw   = r(sd(dt[, eval(raw)])),
-    pct_of_raw_mse = round(100 * (mse(dt[, eval(pred)], dt[, eval(truth)]) / mse(dt[, eval(raw)], dt[, eval(truth)])), 1),
+    # pct_of_raw_mse = round(100 * (mse(dt[, eval(pred)], dt[, eval(truth)]) / mse(dt[, eval(raw)], dt[, eval(truth)])), 1),
     bias_raw = r(mean_bias(dt[, eval(raw)], dt[, eval(truth)])),
     bias_pred = r(mean_bias(dt[, eval(pred)], dt[, eval(truth)])),
     r_raw = r(cor(dt[, eval(raw)], dt[, eval(truth)])),

### 2022-04-22 R/CONUS_AOD.Rmd
annotate CONUS AOD report

@@
 tar_load(cv_summary_tables)
-result <- (cv_summary_tables$stat)
+result <- (cv_summary_tables$stat)[, !"loss"]
 kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
   kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+  row_spec(0,bold=TRUE) %>% column_spec(2:12, width = "1cm")
 ```
 
+Here **raw** is MCD19A2.V006 MAIAC AOD; **truth** is AERONET AOD interpolated to 470nm; and **pred** is MAIAC after applying an XGBoost modeled correction factor.
+
 ## Descriptives for the collocated dataset (Terra)
-
-*still to be added*
-
-## Stratified results for Terra CV dataset
 
@@
 ```
+While our ultimate goal is to improve AOD, our dependent variable for predictive modeling is the difference of the AOD retrieval and the AERONET AOD which is a nicely symmetric and reasonably well-behaved distribution (an estimate of the retrieval error).
+
+```{r describe_dv, echo=FALSE, eval=TRUE}
+kable(summary(dt[, list(diff_AOD)]))
+
+ggplot(dt, aes(x = diff_AOD)) + 
+  geom_histogram(binwidth = 0.005) + 
+  scale_x_continuous("MAIAC AOD minus AERONET AOD") + 
+  coord_cartesian(xlim = c(-1, 1)) + 
+  labs(title = paste0("Histogram of difference in colocated AOD (n=", 
+                      dt[, .N], ")"),
+      caption = paste0("Not shown: ", dt[diff_AOD < -1 | diff_AOD > 1, .N], " AOD differences outside [-1, 1]")) + 
+  theme_bw()
+```
+
+*More still to be added*
+
+## Stratified results for Terra CV dataset
+Here we show our performance metrics in different subsets of the cross-validation results. These are all from the same model training, just slicing up how the performance looks in different contexts.
 
@@
 ```
+
+We expect more 2021 data to show up over time since Level 2.0 data are often quite late to be added to the AERONET datasets.
+The 2020 results (higher SD in AERONET, fewer sites, RMSE for MAIAC is higher) are a bit of an anomaly, but also, COVID could have impacted field campaigns and regular recalibration, etc.
+If we want to look at 2020 specifically, we may want to consider restricting to long-running sites.
 
@@
 
-### Performance by being over water
-
-```{r strata_water, echo=FALSE, fig.pos="H"}
-result <- dt[, performance_aod(.SD), by = qa_lwsi]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
-```
+The takeaway here is that even though the RMSE of MAIAC is substantially better for `qa_best` versus non-best pixels, the improvements in the predicted AOD make even the non-best pixels have lower RMSE than the raw MAIAC `qa_best` and in that sense, we are expanding the set of high quality data by correcting for biases related to QA flagging criteria.
 
@@
 
-In decreasing order (from Terra collocated cross-validation predictions)
+In decreasing order of variable importance (from Terra collocated cross-validation predictions)
 

### 2022-04-26 R/CONUS_AOD.Rmd
cleanup mapping to noaa regions; closes #1

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = here::here())
 ```
@@
 ```{r basics, echo=FALSE}
-library(targets)
+suppressPackageStartupMessages(
+  {library(targets)
+    library(Just.universal)
     library(data.table)
@@
     library(sf)
+    library(Just.universal)
     library(knitr)
@@
     # library(flextable)
+  })
 # set_flextable_defaults(fonts_ignore=TRUE)
@@
 # will replace with a region defined by Kodi later
-usa = st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))
+conus = get_conus()
 ```
@@
 tar_load(aer_stations)
-aer_stations_sf <- st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326")
-sf::sf_use_s2(FALSE) # we were getting errors about the USA geometry otherwise
-aer_stations_sf <- st_join(aer_stations_sf, usa)
-dt <- merge(dt, aer_stations_sf, by = "Site_Name")
+aer_stations_sf <- st_transform(crs = st_crs(conus), st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326"))
+aer_stations_sf <- st_join(aer_stations_sf, conus)
+dt <- merge(dt, aer_stations_sf[, c("Site_Name", "STUSPS")], by = "Site_Name")
 # some Site_Names don't fall within land polygons for whatever reason
-dt[is.na(ID), table(Site_Name)]
+dt[is.na(STUSPS), table(Site_Name)]
 # where are the sites that didn't fall within a state?
-site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
+site_nostate <- st_as_sf(dt[is.na(STUSPS), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"])
 # ggplot() +
-#   geom_sf(data = usa, fill = NA) +
+#   geom_sf(data = conus, fill = NA) +
 #   geom_sf(data = site_nostate, color = "red")
 # what proportion of data are not in a state?
-dt[, proportions(table(is.na(ID)))]
+dt[, proportions(table(is.na(STUSPS)))]
 # assign these to the closest state
 matched <- st_join(site_nostate,
-                   usa,
+                   conus,
                    join=nngeo::st_nn, k=1)
@@
 # ggplot() +
-#   geom_sf(data = usa, fill = NA) +
-#   geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()
-dt[as.data.table(matched, key = "Site_Name"), ID := i.ID]
+#   geom_sf(data = conus, fill = NA) +
+#   geom_sf_text(data = matched, aes(label = STUSPS), color = "red") + theme_minimal()
+dt[as.data.table(matched, key = "Site_Name"), STUSPS := i.STUSPS]
 
 # where do our data come from (by State)?
-dt[, .(uniqueN(Site_Name), .N), by = ID][order(N, decreasing = T)]
+dt[, .(uniqueN(Site_Name), .N), by = STUSPS][order(N, decreasing = T)]
 
-# trying to get from state names to state abbreviations
-# ugly capitalization
-dt[ , state.abb := state.abb.plusdc[match(stringr::str_to_title(ID), c(state.name, "District Of Columbia"))]]
+# assign NOAA region from state USPS codes
 dt[, state.noaa.region := factor(
-   sapply(state.abb, USE.NAMES = F,
+   sapply(STUSPS, USE.NAMES = F,
        function(sa) if (sa %in% c("AK", "HI")) NA else
@@
 
-# map by NOAA regions
-
+## map by NOAA regions
+# map_regions <- ggplot(conus) + geom_sf()
+# print(map_regions)
 ```

### 2022-04-26 R/CONUS_AOD.Rmd
map noaa regions

@@
   row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+```
 
-## map by NOAA regions
-# map_regions <- ggplot(conus) + geom_sf()
-# print(map_regions)
+```{r noaa_region_plot, echo=FALSE, dev='png', dpi=300}
+## map of performance by NOAA regions
+conus$state.noaa.region <- factor(
+   sapply(conus$STUSPS, USE.NAMES = F,
+       function(sa) if (sa %in% c("AK", "HI")) NA else
+           names(noaa.regions)[which(sapply(noaa.regions, function(r)
+               sa %in% r))]),
+   levels = names(noaa.regions))
+conus <- merge(conus, dt[, performance_aod(.SD), keyby = state.noaa.region], by = "state.noaa.region")
+conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE_raw", "RMSE_pred")]),
+                   id.vars = c("geometry", "state.noaa.region"))
+levels(conus_melt$variable) <- c("MCD19A2", "corrected AOD")
+map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) + 
+  geom_sf(aes(fill = value)) + 
+  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = Site_Name], aes(geometry = geometry, size = N), alpha = 0.1) + 
+  facet_wrap(~variable, ncol = 1) + 
+  scale_size("colocated\nobservations", range = c(1, 3)) + 
+  scale_fill_continuous("Regional\nRMSE") + 
+  theme_minimal()
+map_region_rmse
 ```

### 2022-04-26 R/functions.R

@@ -3,7 +3,7 @@
 #' Return AERONET points that intersect the polygon describing the area of
 #' interest.
 #'
-#' @param stations data.table including coordiantes (\code{lon, lat}) points of
+#' @param stations data.table including coordinates (\code{lon, lat}) points of
 #'   AERONET stations
 #' @param reg_polygon SF polygon defining the area of interest
 #' @param new_crs the CRS to set the result to

### 2022-04-27 R/CONUS_AOD.Rmd
add targets for paper_conus output formats

--- 2022-04-26 R/CONUS_AOD.Rmd
@@
   - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
-output: pdf_document
+  - \usepackage{helvet} \renewcommand\familydefault{\sfdefault}
+output_format: 
+  pdf_document: default
+  html_document: default
 always_allow_html: true
 bibliography: references.json
+knit: (function(inputFile, encoding) {
+  rmarkdown::render(inputFile, encoding = encoding,
+  output_dir = "output", output_format = "all") })
 ---
@@
 # set_flextable_defaults(fonts_ignore=TRUE)
+source(here::here("R/functions.R"))
+source(here::here("R/paper_functions.R"))
 ```
@@
 result <- (cv_summary_tables$stat)[, !"loss"]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:12, width = "1cm")
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:12, width = "1.1cm")
 ```
@@
 result <- dt[, performance_aod(.SD), keyby = year(aer_date)]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
 ```
@@
 result <- dt[, performance_aod(.SD), keyby = month(aer_date)]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
 ```
@@
 result <- dt[, performance_aod(.SD), by = qa_best]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
 ```
@@
 result <- dt[, performance_aod(.SD), keyby = overpasses]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
-
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
 ```
@@
 # will replace with a region defined by Kodi later
-conus = get_conus()
+conus = st_transform(crs = crs.us.atlas, get_conus())
 ```
@@
 result <- dt[, performance_aod(.SD), keyby = state.noaa.region]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
-  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+```
 
-## map by NOAA regions
-# map_regions <- ggplot(conus) + geom_sf()
-# print(map_regions)
+```{r noaa_region_plot, echo=FALSE, dev='png', dpi=300}
+## map of performance by NOAA regions
+conus$state.noaa.region <- factor(
+   sapply(conus$STUSPS, USE.NAMES = F,
+       function(sa) if (sa %in% c("AK", "HI")) NA else
+           names(noaa.regions)[which(sapply(noaa.regions, function(r)
+               sa %in% r))]),
+   levels = names(noaa.regions))
+conus <- merge(conus, dt[, performance_aod(.SD), keyby = state.noaa.region], by = "state.noaa.region")
+conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE_raw", "RMSE_pred")]),
+                   id.vars = c("geometry", "state.noaa.region"))
+levels(conus_melt$variable) <- c("MAIAC AOD", "corrected AOD")
+map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) + 
+  geom_sf(aes(fill = value)) + 
+  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = Site_Name], aes(geometry = geometry, size = N), alpha = 0.1) + 
+  facet_wrap(~variable, ncol = 1) + 
+  coord_sf(default_crs = st_crs(conus)) + 
+  scale_size("colocated\nobservations", range = c(1, 3)) + 
+  scale_fill_continuous("Regional\nRMSE") + 
+  theme_minimal() + 
+  theme(axis.text = element_blank(),
+        panel.grid.major = element_blank())
+map_region_rmse
 ```
@@
 In decreasing order of variable importance (from Terra collocated cross-validation predictions)
-
 ```{r shap_rank, echo=FALSE}
@@
 df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
-kbl(df_shap, col.names = "Mean |SHAP|")
+kbl(df_shap, col.names = "Mean |SHAP|") %>% row_spec(0, bold=TRUE)
 # df_shap$variable <- row.names(df_shap)

### 2022-04-27 _targets.R

@@ -205,9 +205,12 @@ report_targets = list(
 
 # Render CONUS AOD results ###
 paper_conus_targets = list(
-  tar_render(paper_conus,
+  tar_render(paper_conus_html,
+             'R/CONUS_AOD.Rmd', output_format = "html_document",
+             packages = c('sf', 'patchwork')),
+    tar_render(paper_conus_pdf, output_format = "pdf_document",
              'R/CONUS_AOD.Rmd',
-             packages = c('sf', 'DT', 'patchwork'))
+             packages = c('sf', 'patchwork'))
 )
 # Final targets list ####
 list(set1_targets, combined_target, report_targets, paper_conus_targets)

### 2022-06-03 R/CONUS_AOD.Rmd
add detail in CONUS_AOD

@@
     library(ggplot2)
+    library(ggridges)
+    library(ggpointdensity)
     library(sf)
@@
     library(kableExtra)
+    library(RColorBrewer)
     # library(flextable)
@@
 ```
+
 While our ultimate goal is to improve AOD, our dependent variable for predictive modeling is the difference of the AOD retrieval and the AERONET AOD which is a nicely symmetric and reasonably well-behaved distribution (an estimate of the retrieval error).
@@
 kable(summary(dt[, list(diff_AOD)])) 
+```
 
+```{r plot_dv, echo=FALSE, eval=TRUE}
 ggplot(dt, aes(x = diff_AOD)) + 
@@
 ## Stratified results for Terra CV dataset
+
 Here we show our performance metrics in different subsets of the cross-validation results. These are all from the same model training, just slicing up how the performance looks in different contexts.
@@
 
+```{r plot_by_month, echo=FALSE, eval=TRUE, dev='png', dpi=300}
+# Color by quantiles
+dt[, month := factor(month(aer_date), levels = 1:12, 
+                     labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))]
+p1 <- ggplot(dt, aes(x = diff_AOD, y = month, fill = factor(stat(quantile)))) + 
+  stat_density_ridges(
+    geom = "density_ridges_gradient", calc_ecdf = TRUE,
+    quantiles = 4, quantile_lines = FALSE
+  ) +
+  geom_vline(xintercept = 0, linetype = "dashed", color = "grey90") + 
+  scale_fill_brewer(name = "Quartiles", type = "div") + 
+  scale_x_continuous("MAIAC AOD minus AERONET AOD", expand = c(0, 0)) + 
+  coord_cartesian(xlim = c(-0.25, 0.25)) + 
+  labs(title = paste0("Histogram of difference in colocated AOD (n=", 
+                      dt[, .N], ")"),
+      caption = paste0("Not shown: ", dt[diff_AOD < -0.25 | diff_AOD > 0.25, .N], " AOD differences outside [-0.25, 0.25]")) + 
+  theme_bw()
+suppressMessages(print(p1))
+```
+
 ### Performance by `qa_best`
@@
 The takeaway here is that even though the RMSE of MAIAC is substantially better for `qa_best` versus non-best pixels, the improvements in the predicted AOD make even the non-best pixels have lower RMSE than the raw MAIAC `qa_best` and in that sense, we are expanding the set of high quality data by correcting for biases related to QA flagging criteria.
+
+### Performance for higher versus lower AOD (AERONET > 0.6)
+
+```{r strata_high_aod, echo=FALSE, fig.pos="H"}
+dt[, aod_high := AOD_470nm > 0.6]
+result <- dt[, performance_aod(.SD), by = aod_high]
+kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+    align = "l") %>% 
+  kable_styling(position = "left", 
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+```
 
@@
 
+```{r error_envelope, echo=FALSE}
+# error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
+dt[, round(100 * proportions(table(MCD19_AOD_470nm > (AOD_470nm - 0.05 - 0.1*AOD_470nm) & MCD19_AOD_470nm < (AOD_470nm + 0.05 + 0.1*AOD_470nm))))]
+# sure looks like we would have a much smaller error envelope (in both additive and multiplicative terms)
+dt[, round(100 * proportions(table(aodhat > (AOD_470nm - 0.025 - 0.055*AOD_470nm) & aodhat < (AOD_470nm + 0.025 + 0.055*AOD_470nm))))]
+
+```
+
 ```{r agreement_plot, echo=FALSE, dev='png', dpi=300}
@@
   aes(AOD_470nm, MCD19_AOD_470nm) +
+  geom_pointdensity(alpha = 0.1, size = 0.6) +
   geom_abline(linetype = "dashed") +
-  geom_point(alpha = 0.01) +
   geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
-  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
-  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1)) +
+  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
+  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
+  guides(color = "none") + 
   theme_bw() +
@@
   aes(AOD_470nm, aodhat) +
+  geom_pointdensity(alpha = 0.1, size = 0.6) +
   geom_abline(linetype = "dashed") +
-  geom_point(alpha = 0.01) +
   geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
-  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1)) +
-  scale_y_continuous("Corrected AOD 470nm", limits = c(0, 1)) +
+  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
+  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
+  guides(color = "none") + 
   theme_bw() +
@@
 
-*Will add plot of agreement across bins of AOD*
-
-------------------------------------------------------------------------
-
 ## SHAP interpretation
@@
 In decreasing order of variable importance (from Terra collocated cross-validation predictions)
+
 ```{r shap_rank, echo=FALSE}
@@
 ```
+*Will add some scatterplots of these SHAPs*
 

### 2022-06-18 R/CONUS_AOD.Rmd
CONUS AOD reporting

@@
     library(RColorBrewer)
+    library(SHAPforxgboost)
     # library(flextable)
@@
 
-Here **raw** is MCD19A2.V006 MAIAC AOD; **truth** is AERONET AOD interpolated to 470nm; and **pred** is MAIAC after applying an XGBoost modeled correction factor.
+Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **truth** is AERONET AOD interpolated to 470nm; and **pred** is MAIAC after applying an XGBoost modeled correction factor.
 
@@
 ```
-*Will add some scatterplots of these SHAPs*
+Summary visualization of SHAP contributions
+
+```{r shap_summary, echo=FALSE, dev='png', dpi=300}
+shap_long <- shap.prep(shap_contrib = initial_cv_l2_terra_conus$shap_score, 
+                       X_train = initial_cv_l2_terra_conus$mDT_wPred[, names(initial_cv_l2_terra_conus$shap_score), with = F])
+# **SHAP summary plot**
+shap.plot.summary(shap_long)
+```
+```{r shap_features, echo=FALSE, dev='png', dpi=300}
+g1 <- shap.plot.dependence(data_long = shap_long, x = 'dayint', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
+# g2 <- shap.plot.dependence(data_long = shap_long, x = 'dayint', y = 'MCD19_AOD_470nm', color_feature = 'MCD19_AOD_470nm') +  ggtitle("(B) SHAP values of MAIAC AOD vs. Time trend")
+# 
+# gridExtra::grid.arrange(g1, g2, ncol = 2)
+g1
+```
 

### 2022-06-18 R/paper_functions.R

@@ -66,7 +66,12 @@ satellite_vs_ground = function(satellite, ground)
     # Find the weighted correlation of old and new satellite values
     # with ground values.
     d[, by = .(date, cell), weight := 1 / .N]
-    sapply(c("old", "new"), function(sv_type)
+    out = lapply(c("old", "new"), function(sv_type)
         cov.wt(
             d[, .(get(paste0("satellite_value_", sv_type)), ground_value)],
-            wt = d$weight, cor = T, method = "ML")$cor[1, 2])}
+            wt = d$weight, cor = T, method = "ML")$cor[1, 2])
+    names(out) <- c("old", "new")
+    out[["nobs"]] = d[, .N]
+    out[["celldays"]] = d[, sum(weight)]
+    out
+    }

### 2022-09-30 R/CONUS_AOD.Rmd
relabel NOAA regions

@@
 Evaluate the performance of the `Earth_obs_cleaning` software tool on MAIAC MCD19A2 for Terra and Aqua satellites across the Continental USA, 2000-2021
+
+# Introduction
+
+# Methods
+
+# Results
 
@@
   # https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-regions.php
-    Central = c("IL", "IN", "KY", "MO", "OH", "TN", "WV"),
-    East.North.Central = c("IA", "MI", "MN", "WI"),
+    Ohio.Valley = c("IL", "IN", "KY", "MO", "OH", "TN", "WV"), #previously 'Central'
+    Upper.Midwest = c("IA", "MI", "MN", "WI"), #previously 'East.North.Central'
     Northeast = c("DC", "CT", "DE", "ME", "MD", "MA", "NH", "NJ", "NY", "PA", "RI", "VT"),
@@
     West = c("CA", "NV"),
-    West.North.Central = c("MT", "NE", "ND", "SD", "WY"))
+    Northern.Rockies.and.Plains = c("MT", "NE", "ND", "SD", "WY")) #previously 'West.North.Central'
 

### 2022-10-27 R/functions.R
Got the CV report working again.

@@ -634,77 +634,58 @@ initial_cv_dart <- function(
   c(list(by_var = by_var, bin_list = bin_list, mDT_wPred = mDT), cv_results)
 }
 
-#' Summarize CV statistics on all initial_cv objects
+#' Summarize CV statistics.
 #'
-#' @param cv_list list of all CV output objects; may include every year, both
-#'   satellites, and both objective functions
-#' @return data.table summarizing CV statistics for each loss and satellite
-cv_summary <- function(cv_list){
-  stats_list = vector(mode = "list", length = length(cv_list))
-  difftimes_list = vector(mode = "list", length = length(cv_list))
-  for(i in 1:length(cv_list)){
-    cv = cv_list[[i]]
-    dt = data.table::copy(cv$mDT_wPred)
-    dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
-    stats = performance_metrics(dt,
-                                ground_truth = "AOD_470nm",
-                                eo_raw = "MCD19_AOD_470nm",
-                                eo_pred = "aodhat")
-    stats$sat <- str_extract(names(cv_list)[[i]], 'terra|aqua')
-    stats$loss <- str_match(names(cv_list)[[i]], 'cv_(l[12])_')[,2]
-    stats_list[[i]] <- stats
-    difftimes_list[[i]] <- round(summary(as.numeric(abs(cv$mDT_wPred$rj_difftime))),0)
-  }
-  # prediction summary stats
-  statsDT = rbindlist(lapply(stats_list, as.data.table))
-  # setcolorder(statsDT, c('sat', 'loss',
-  #                      'MAE_uncorr', 'MAE_corr', 'MAE_change',
-  #                      'rmse', 'MAD_mcd19', 'MAD_aodhat', 'MAD_change'))
-  # difftime distribution
-  difftimeDT = rbindlist(lapply(difftimes_list, function(x) as.list(x)))
-  difftimeDT[, c('sat', 'loss') := statsDT[, .(sat, loss)]]
-  setcolorder(difftimeDT, c('sat', 'loss'))
+#' @param d The data table `mDT_wPred` returned by `initial_cv_dart`.
+cv.summary = function(d)
+   {d = d[, .(
+        site,
+        time.sat = lubridate::as_datetime(time.sat),
+        time.ground,
[…]

### 2022-10-27 _targets.R

@@ -161,4 +161,10 @@ list(
     tar_target(full_model, dart_full(
         traindata,
         y_var = "diff_AOD",
-        features = features)))
+        features = features)),
+
+    # Summarize and report on the CV
+    tar_target(cv_summary_table, cv.summary(
+        initial_cv_l2$mDT_wPred)),
+    tar_render(initial_cv_report,
+        'R/initial_cv_report.Rmd'))

### 2022-10-28 R/functions.R
Tweaked the columns returned by `initial_cv_dart`.

@@ -629,6 +629,8 @@ initial_cv_dart <- function(
                               xgb_threads = xgb_threads)
 
   mDT <- cbind(mDT, cv_results$y_pred_dt)
+  mDT[, time.sat := lubridate::as_datetime(time.sat)]
+  mDT[, y.ground.pred := y.sat - y.diff_pred]
 
   # output
   c(list(by_var = by_var, bin_list = bin_list, mDT_wPred = mDT), cv_results)
@@ -638,16 +640,9 @@ initial_cv_dart <- function(
 #'
 #' @param d The data table `mDT_wPred` returned by `initial_cv_dart`.
 cv.summary = function(d)
-   {d = d[, .(
-        site,
-        time.sat = lubridate::as_datetime(time.sat),
-        time.ground,
-        y.sat,
-        y.ground,
-        y.ground.pred = y.sat - y.diff_pred)]
     rbind(
         d[, c(list(Year = "all"), eval(performance.j))],
-        d[, keyby = .(Year = year(time.sat)), eval(performance.j)])}
+        d[, keyby = .(Year = year(time.sat)), eval(performance.j)])
 
 performance.j = quote(
    {mse = function(x, y) mean((x - y)^2)

### 2022-10-28 R/CONUS_AOD.Rmd
Clean up whitespace.
Got the paper rendering again.

@@
 source(here::here("R/paper_functions.R"))
+the.sat = tools::toTitleCase(Wf$satellite)
 ```
@@
 
-Evaluate the performance of the `Earth_obs_cleaning` software tool on MAIAC MCD19A2 for Terra and Aqua satellites across the Continental USA, 2000-2021
+Evaluate the performance of the `Earth_obs_cleaning` software tool on MAIAC MCD19A2 for `r the.sat` across the Continental USA, 2000-2021
 
@@
 ```{r table_cv_results, echo=FALSE}
-tar_load(cv_summary_tables)
-result <- (cv_summary_tables$stat)[, !"loss"]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
-    align = "l") %>% 
+kbl(pretty.table.numbers(tar_read(cv_summary_table)[1, -"Year"]), align = "l") %>%
   kable_styling(position = "left",
@@
 
-Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **truth** is AERONET AOD interpolated to 470nm; and **pred** is MAIAC after applying an XGBoost modeled correction factor.
+Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost modeled correction factor.
 
-## Descriptives for the collocated dataset (Terra)
+## Descriptives for the collocated dataset (`r the.sat`)
 
 ```{r stratified, echo=FALSE}
-tar_load(initial_cv_l2_terra_conus)
-dt <- initial_cv_l2_terra_conus$mDT_wPred
-if(!"aodhat" %in% colnames(dt)){
-  dt[, aodhat := MCD19_AOD_470nm - diff_AOD_pred]
-}
+dt <- copy(tar_read(initial_cv_l2)$mDT_wPred)
 ```
@@
 ```{r describe_dv, echo=FALSE, eval=TRUE}
-kable(summary(dt[, list(diff_AOD)])) 
+kable(summary(dt[, list(y.diff)]))
 ```
@@
 ```{r plot_dv, echo=FALSE, eval=TRUE}
-ggplot(dt, aes(x = diff_AOD)) + 
+ggplot(dt, aes(x = y.diff)) +
   geom_histogram(binwidth = 0.005) +
@@
                       dt[, .N], ")"),
-      caption = paste0("Not shown: ", dt[diff_AOD < -1 | diff_AOD > 1, .N], " AOD differences outside [-1, 1]")) + 
+      caption = paste0("Not shown: ", dt[y.diff < -1 | y.diff > 1, .N], " AOD differences outside [-1, 1]")) +
   theme_bw()
@@
 
-## Stratified results for Terra CV dataset
+## Stratified results for `r the.sat` CV dataset
 
@@
 ```{r strata_year, echo=FALSE, eval=TRUE}
-result <- dt[, performance_aod(.SD), keyby = year(aer_date)]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+kbl(pretty.table.numbers(tar_read(cv_summary_table)[-1]),
     align = "l") %>%
@@
 ```{r strata_month, echo=FALSE}
-result <- dt[, performance_aod(.SD), keyby = month(aer_date)]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
-    align = "l") %>% 
+result <- dt[,
+    keyby = .(Month = month(lubridate::as_datetime(time.sat))),
+    eval(performance.j)]
+result[, Month := month.abb[Month]]
+kbl(pretty.table.numbers(result), align = "l") %>%
   kable_styling(position = "left",
@@
 # Color by quantiles
-dt[, month := factor(month(aer_date), levels = 1:12, 
-                     labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))]
-p1 <- ggplot(dt, aes(x = diff_AOD, y = month, fill = factor(stat(quantile)))) + 
+dt[, month := factor(month(time.sat),
+                     labels = month.abb)]
+p1 <- ggplot(dt, aes(x = y.diff, y = month, fill = factor(stat(quantile)))) +
   stat_density_ridges(
@@
                       dt[, .N], ")"),
-      caption = paste0("Not shown: ", dt[diff_AOD < -0.25 | diff_AOD > 0.25, .N], " AOD differences outside [-0.25, 0.25]")) + 
+      caption = paste0("Not shown: ", dt[y.diff < -0.25 | y.diff > 0.25, .N], " AOD differences outside [-0.25, 0.25]")) +
   theme_bw()
@@
 ```{r strata_qa, echo=FALSE}
-result <- dt[, performance_aod(.SD), by = qa_best]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+kbl(pretty.table.numbers(dt[, eval(performance.j), by = qa_best]),
     align = "l") %>%
@@
 ```{r strata_high_aod, echo=FALSE, fig.pos="H"}
-dt[, aod_high := AOD_470nm > 0.6]
-result <- dt[, performance_aod(.SD), by = aod_high]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+kbl(pretty.table.numbers(dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]),
     align = "l") %>%
@@
 ```{r strata_overpasses, echo=FALSE, fig.pos="H"}
-dt[, overpasses := .N, by = c("Site_Name", "aer_date")]
-result <- dt[, performance_aod(.SD), keyby = overpasses]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
+kbl(pretty.table.numbers(dt[, eval(performance.j), keyby = overpasses]),
     align = "l") %>%
@@
 aer_stations_sf <- st_join(aer_stations_sf, conus)
-dt <- merge(dt, aer_stations_sf[, c("Site_Name", "STUSPS")], by = "Site_Name")
-# some Site_Names don't fall within land polygons for whatever reason
-dt[is.na(STUSPS), table(Site_Name)]
+colnames(aer_stations_sf)[colnames(aer_stations_sf) == "Site_Name"] = "site"
+dt <- merge(dt, aer_stations_sf[, c("site", "STUSPS")], by = "site")
+# some sites don't fall within land polygons for whatever reason
+dt[is.na(STUSPS), table(site)]
 # where are the sites that didn't fall within a state?
-site_nostate <- st_as_sf(dt[is.na(STUSPS), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"])
+site_nostate <- st_as_sf(dt[is.na(STUSPS), ][unique(site), .(site, geometry), mult = "first"])
 # ggplot() +
@@
 #   geom_sf_text(data = matched, aes(label = STUSPS), color = "red") + theme_minimal()
-dt[as.data.table(matched, key = "Site_Name"), STUSPS := i.STUSPS]
+dt[as.data.table(matched, key = "site"), STUSPS := i.STUSPS]
 
 # where do our data come from (by State)?
-dt[, .(uniqueN(Site_Name), .N), by = STUSPS][order(N, decreasing = T)]
+dt[, .(uniqueN(site), .N), by = STUSPS][order(N, decreasing = T)]
 
@@
 # performance by NOAA climate regions
-result <- dt[, performance_aod(.SD), keyby = state.noaa.region]
-kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
+by.region = dt[, eval(performance.j), keyby = state.noaa.region]
+kbl(pretty.table.numbers(by.region),
     align = "l") %>%
@@
    levels = names(noaa.regions))
-conus <- merge(conus, dt[, performance_aod(.SD), keyby = state.noaa.region], by = "state.noaa.region")
-conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE_raw", "RMSE_pred")]),
+conus <- merge(conus, by.region, by = "state.noaa.region")
+conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE, raw", "RMSE, corrected")]),
                    id.vars = c("geometry", "state.noaa.region"))
@@
   geom_sf(aes(fill = value)) +
-  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = Site_Name], aes(geometry = geometry, size = N), alpha = 0.1) + 
+  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = site], aes(geometry = geometry, size = N), alpha = 0.1) +
   facet_wrap(~variable, ncol = 1) +
@@
 # error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
-dt[, round(100 * proportions(table(MCD19_AOD_470nm > (AOD_470nm - 0.05 - 0.1*AOD_470nm) & MCD19_AOD_470nm < (AOD_470nm + 0.05 + 0.1*AOD_470nm))))]
+dt[, round(100 * proportions(table(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground))))]
 # sure looks like we would have a much smaller error envelope (in both additive and multiplicative terms)
-dt[, round(100 * proportions(table(aodhat > (AOD_470nm - 0.025 - 0.055*AOD_470nm) & aodhat < (AOD_470nm + 0.025 + 0.055*AOD_470nm))))]
+dt[, round(100 * proportions(table(y.ground.pred > (y.ground - 0.025 - 0.055*y.ground) & y.ground.pred < (y.ground + 0.025 + 0.055*y.ground))))]
 
@@
 ```{r agreement_plot, echo=FALSE, dev='png', dpi=300}
-p1 <- ggplot(dt[AOD_470nm < 1 & MCD19_AOD_470nm < 1]) +
-  aes(AOD_470nm, MCD19_AOD_470nm) +
+p1 <- ggplot(dt[y.ground < 1 & y.sat < 1]) +
+  aes(y.ground, y.sat) +
   geom_pointdensity(alpha = 0.1, size = 0.6) +
@@
   theme(aspect.ratio = 1) +
-  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | MCD19_AOD_470nm > 1, .N], " AOD values outside [0, 1]"))
+  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.sat > 1, .N], " AOD values outside [0, 1]"))
 
-p2  <- ggplot(dt[AOD_470nm < 1 & aodhat > 0 & aodhat < 1]) +
-  aes(AOD_470nm, aodhat) +
+p2  <- ggplot(dt[y.ground < 1 & y.ground.pred > 0 & y.ground.pred < 1]) +
+  aes(y.ground, y.ground.pred) +
   geom_pointdensity(alpha = 0.1, size = 0.6) +
@@
   theme(aspect.ratio = 1) +
-  labs(caption = paste0("Not shown: ", dt[AOD_470nm > 1 | aodhat > 1 | aodhat < 0, .N], " AOD values outside [0, 1]"))
+  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.ground.pred > 1 | y.ground.pred < 0, .N], " AOD values outside [0, 1]"))
 
@@
 
-In decreasing order of variable importance (from Terra collocated cross-validation predictions)
+In decreasing order of variable importance (from `r the.sat` collocated cross-validation predictions)
 
@@
 # what do the SHAPs tell us?
-tar_load(initial_cv_l2_terra_conus)
-shaps <- initial_cv_l2_terra_conus$"shap_score"
+tar_load(initial_cv_l2)
+shaps <- initial_cv_l2$"shap_score"
 # sort by importance
@@
 ```{r shap_summary, echo=FALSE, dev='png', dpi=300}
-shap_long <- shap.prep(shap_contrib = initial_cv_l2_terra_conus$shap_score, 
-                       X_train = initial_cv_l2_terra_conus$mDT_wPred[, names(initial_cv_l2_terra_conus$shap_score), with = F])
+shap_long <- shap.prep(shap_contrib = initial_cv_l2$shap_score,
+                       X_train = initial_cv_l2$mDT_wPred[, names(initial_cv_l2$shap_score), with = F])
 # **SHAP summary plot**
@@
 ```{r shap_features, echo=FALSE, dev='png', dpi=300}
-g1 <- shap.plot.dependence(data_long = shap_long, x = 'dayint', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
-# g2 <- shap.plot.dependence(data_long = shap_long, x = 'dayint', y = 'MCD19_AOD_470nm', color_feature = 'MCD19_AOD_470nm') +  ggtitle("(B) SHAP values of MAIAC AOD vs. Time trend")
+g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
+# g2 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'y.sat', color_feature = 'y.sat') +  ggtitle("(B) SHAP values of MAIAC AOD vs. Time trend")
 #

### 2022-10-28 _targets.R

@@ -167,4 +167,13 @@ list(
     tar_target(cv_summary_table, cv.summary(
         initial_cv_l2$mDT_wPred)),
     tar_render(initial_cv_report,
-        'R/initial_cv_report.Rmd'))
+        'R/initial_cv_report.Rmd'),
+
+    # Render the CONUS AOD manuscript
+    if (Wf$satellite.product == "mcd19a2") list(
+        tar_render(paper_conus_html, output_format = "html_document",
+            packages = c('sf', 'patchwork'),
+            'R/CONUS_AOD.Rmd', quiet = F),
+        tar_render(paper_conus_pdf, output_format = "pdf_document",
+            packages = c('sf', 'patchwork', 'magick'),
+            'R/CONUS_AOD.Rmd')))

### 2022-12-19 R/CONUS_AOD.Rmd
Added some maps for the AOD CONUS paper.

@@
 
+## Maps of new predictions
+
+### Median MSE day
+
+```{r pred.map.median.mse, echo = F}
+pred.map(
+    tar_read(T.median.mse.map.data)$pred, tar_read(pred_grid),
+    bg.sf = get_conus(), color.scale.name = "AOD",
+    limits = list(lon = c(-107, Inf), lat = c(-Inf, 37)),
+    quantile.cap = .99)
+```
+
+This map shows a tile on `r tar_read(T.median.mse.map.data)$date`.
+
+### Baltimore
+
+```{r pred.map.balitmore, echo = F}
+pred.map(
+    tar_read(T.baltimore.map.data), tar_read(pred_grid),
+    bg.sf = get_conus(), color.scale.name = "AOD")
+```
+
 *End of report*

### 2022-12-19 R/new_preds.R

@@ -22,6 +22,8 @@ new.preds = function(dt.start, dt.end, cells = NULL, targets = NULL)
     dt.end = lubridate::with_tz(dt.end, "UTC")
     if (is.null(cells))
         cells = which(!is.na(drop(grid$tile[])))
+    if (is.double(cells))
+        cells = as.integer(cells)
     assert(is.integer(cells) && all(
         1L <= cells & cells <= terra::ncell(grid)))
     message("Cells: ", scales::comma(length(cells)))

### 2022-12-19 R/paper_functions.R

@@ -94,3 +94,106 @@ satellite_vs_aqs = function(satellite, aqs)
     out[["celldays"]] = d[, sum(weight)]
     out
     }
+
+median.mse.map.data = function(
+      y.sat.name, the.satellite, satellite.product, n.workers,
+      d, pred_grid, buff, satellite_hdf_files, full_model)
+   {# Get the date with the median MSE.
+    date = (d
+        [, by = .(date = lubridate::as_date(time.sat, tz = "UTC")),
+            .(mse = mean((y.ground - y.ground.pred)^2))]
+        [which.min(mse - median(mse)), date])
+
+    # Get overpasses for this data.
+    sat = satellite_hdf_files[time == date]
+    sat[, sat.files.ix := .I]
+    overpasses = merge(by = "sat.files.ix", sat,
+        expand.to.overpasses(sat, sat,
+            the.satellite, satellite.product, n.workers))
+
+    # Find the local cell indices we'd want to map.
+    message("Getting raster cells")
+    overpass.local.cells = pblapply(1 : nrow(overpasses), cl = n.workers, \(i)
+       {r = with(overpasses[i], read_satellite_raster(
+            Wf$satellite.product, tile, path, overpass))[[y.sat.name]]
+        # Get only cells with non-missing values.
+        d = as.data.table(as.data.frame(r,
+            na.rm = T, cell = T, xy = T))
+        if (!nrow(d))
+            return(integer())
+        # Get only cells in the study area.
+        d[in.sf(x, y, terra::crs(r), buff), cell]})
+
+    # Choose the tile-time with the most non-missing values.
+    oi = which.max(sapply(overpass.local.cells, length))
+    # Get the prediction-grid cell for each local cell.
+    message("Getting cell indices")
+    cells = (as.data.table(as.data.frame(pred_grid, cells = T))
+        [tile == overpasses[oi, tile]]
+        [cell.local %in% overpass.local.cells[[oi]], cell])
+    assert(length(cells) == length(overpass.local.cells[[oi]]))
+
+    # Make the predictions and return.
+    list(
+        date = date,
+        time.sat = overpasses[oi, time.sat],
+        tile = overpasses[oi, tile],
+        pred = new.preds(
+            overpasses[oi, time.sat],
+            overpasses[oi, time.sat],
[…]

### 2022-12-19 R/targets.R

@@ -182,6 +182,13 @@ list(
     tar_target(aqs_comparison, satellite_vs_aqs(
         pred_at_aqs_sites, aqs_obs)),
 
+    # Data for maps
+    tar_target(T.median.mse.map.data, median.mse.map.data(
+        Wf$y.sat, Wf$satellite, Wf$satellite.product, config$n.workers,
+        initial_cv_l2$mDT_wPred, pred_grid, buff, satellite_hdf_files, full_model)),
+    tar_target(T.baltimore.map.data, baltimore.map.data(
+        pred_grid, buff, satellite_hdf_files, full_model)),
+
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2") list(
         tar_render(paper_conus_html, output_format = "html_document",

### 2022-12-19 R/functions.R
Upgraded `terra`.

@@ -745,7 +745,7 @@ read_satellite_raster = function(
         c("cosSZA", "cosVZA", "RelAZ", "Scattering_Angle", "Glint_Angle"),
         function(vname)
            {r = terra::rast(paste0(
-                str_replace(r1@ptr$filenames[1], ":grid1km:[A-Za-z0-9_]+\\Z",
+                str_replace(r1@ptr$filenames()[1], ":grid1km:[A-Za-z0-9_]+\\Z",
                     ":grid5km:"),
                 vname))
             if (!is.na(overpass))
@@ -783,8 +783,6 @@ make_pred_grid = function(satellite.product, earthdata.rows)
             r$tile = as.integer(tile)
             r$cell.local = seq_len(terra::ncell(r))
             r[[c("tile", "cell.local")]]})))
-    # We need to call `as.integer` again: https://github.com/rspatial/terra/issues/866
-    r$cell.local = as.integer(r$cell.local[])
     r$tile = factor(as.integer(drop(r$tile[])),
         labels = levels(earthdata.rows$tile))
     r}

### 2022-12-20 writing/CONUS_AOD.Rmd
Reorganized files.

--- 2022-12-19 R/CONUS_AOD.Rmd
@@
 # set_flextable_defaults(fonts_ignore=TRUE)
-source(here::here("R/functions.R"))
-source(here::here("R/paper_functions.R"))
+source(here::here("code/functions.R"))
+source(here::here("code/paper_functions.R"))
 the.sat = tools::toTitleCase(Wf$satellite)
@@
 
-```{r error_envelope, echo=FALSE}
+```{r error_envelope, echo=TRUE}
 # error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
@@
 
+# now with our error envelope function
+envelope.sat <- empirical.error.envelope(dt, pred = "y.sat") 
+head(envelope.sat)
+
+# and for the corrected data
+envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred") 
+head(envelope.ground.pred)
 ```
@@
 
-```{r shap_summary, echo=FALSE, dev='png', dpi=300}
+```{r shap_summary, echo=FALSE, dev='png', dpi=300, eval=FALSE}
 shap_long <- shap.prep(shap_contrib = initial_cv_l2$shap_score,
@@
 ```
-```{r shap_features, echo=FALSE, dev='png', dpi=300}
+
+```{r shap_features, echo=FALSE, dev='png', dpi=300, eval=FALSE}
 g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")

### 2022-12-20 code/globals.R

@@ -56,7 +56,7 @@ workflow.dir = file.path(data.dir, "workflows",
 dir.create(workflow.dir, recursive = T, showWarnings = F)
 yaml::write_yaml(file = file.path(workflow.dir, "targets.yaml"),
     list(main = list(
-        script = "R/targets.R",
+        script = "code/targets.R",
         store = file.path(workflow.dir, "targets_store"))))
 Sys.setenv(TAR_CONFIG = file.path(workflow.dir, "targets.yaml"))
 

### 2022-12-20 code/paper_functions.R

@@ -53,8 +53,8 @@ get_aqs_obs = function(years, grid)
       # PM_{2.5} from a federally approved reference or equivalent
       # method (FRM/FEM).
 
-    assert(file.exists("openssl_workaround.conf"))
-    Sys.setenv(OPENSSL_CONF = "openssl_workaround.conf")
+    assert(file.exists("code/openssl_workaround.conf"))
+    Sys.setenv(OPENSSL_CONF = "code/openssl_workaround.conf")
     d = rbindlist(lapply(years, function(the.year)
        {fname = sprintf("daily_%d_%d.zip", parameter.code, the.year)
         d = fread(

### 2022-12-20 code/targets.R

@@ -2,18 +2,18 @@
 # * Setup
 # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-source('R/globals.R')
-source('R/data.R')
-source('R/functions.R')
-source('R/xgboost_cv.R')
-source('R/new_preds.R')
-source('R/paper_functions.R')
+source('code/globals.R')
+source('code/data.R')
+source('code/functions.R')
+source('code/xgboost_cv.R')
+source('code/new_preds.R')
+source('code/paper_functions.R')
 
 library(targets)
 library(tarchetypes)
 
 tar_option_set(
-    packages = sapply(parse("R/libraries.R")[[1]][-1][[1]][-1],
+    packages = sapply(parse("code/libraries.R")[[1]][-1][[1]][-1],
         function(x) as.character(x[[2]])),
     format = 'qs',
     workspace_on_error = TRUE,
@@ -169,7 +169,7 @@ list(
     tar_target(cv_summary_table, cv.summary(
         initial_cv_l2$mDT_wPred)),
     tar_render(initial_cv_report,
-        'R/initial_cv_report.Rmd'),
+        'writing/initial_cv_report.Rmd'),
 
     # Compare AQS to our predictions
     tar_target(aqs_obs, format = "fst_dt", get_aqs_obs(
@@ -193,7 +193,7 @@ list(
     if (Wf$satellite.product == "mcd19a2") list(
         tar_render(paper_conus_html, output_format = "html_document",
             packages = c('sf', 'patchwork'),
-            'R/CONUS_AOD.Rmd', quiet = F),
+            'writing/CONUS_AOD.Rmd', quiet = F),
         tar_render(paper_conus_pdf, output_format = "pdf_document",
             packages = c('sf', 'patchwork', 'magick'),
-            'R/CONUS_AOD.Rmd')))
+            'writing/CONUS_AOD.Rmd')))

### 2022-12-20 writing/CONUS_AOD.Rmd
Removed the absolute-error CV.

@@
 ```{r stratified, echo=FALSE}
-dt <- copy(tar_read(initial_cv_l2)$mDT_wPred)
+dt <- copy(tar_read(initial_cv)$mDT_wPred)
 ```
@@
 # what do the SHAPs tell us?
-tar_load(initial_cv_l2)
-shaps <- initial_cv_l2$"shap_score"
+tar_load(initial_cv)
+shaps <- initial_cv$"shap_score"
 # sort by importance
@@
 ```{r shap_summary, echo=FALSE, dev='png', dpi=300, eval=FALSE}
-shap_long <- shap.prep(shap_contrib = initial_cv_l2$shap_score,
-                       X_train = initial_cv_l2$mDT_wPred[, names(initial_cv_l2$shap_score), with = F])
+shap_long <- shap.prep(shap_contrib = initial_cv$shap_score,
+                       X_train = initial_cv$mDT_wPred[, names(initial_cv_l2$shap_score), with = F])
 # **SHAP summary plot**

### 2022-12-20 code/functions.R

@@ -309,7 +309,6 @@ get.predictors = function(
 #' @param features character vector of column names to train on
 #' @param stn_var if binning by stations, provide the station column name
 #' @param day_var if binning by days, provide the day column name
-#' @param absolute whether to use absolute loss (rather than square loss)
 # Depends on functions in xgboost_cv_RFE.R
 initial_cv_dart <- function(
   data,
@@ -319,7 +318,6 @@ initial_cv_dart <- function(
   n_rounds = 100,
   stn_var = NULL,
   day_var = NULL,
-  absolute = FALSE,
   progress = TRUE
 ){
   xgb_threads <- get.threads()
@@ -369,8 +367,7 @@ initial_cv_dart <- function(
                               n_rounds = n_rounds,
                               y_var = y_var,
                               progress = progress,
-                              absolute = absolute,
-                              seed = (if (absolute) 400 else 1234),
+                              seed = 1234,
                               index_train = index_train,
                               index_test = index_test,
                               xgb_threads = xgb_threads)

### 2022-12-20 code/targets.R

@@ -148,13 +148,11 @@ list(
           # Using a lot more workers on Coco seems to be slower.
 
     # Modeling
-    tar_map(values = list(loss = c("l1", "l2")),
-        tar_target(initial_cv, initial_cv_dart(
-            traindata,
-            absolute = loss == "l1",
-            y_var = "y.diff",
-            features = Wf$features,
-            stn_var = "site"))),
+    tar_target(initial_cv, initial_cv_dart(
+        traindata,
+        y_var = "y.diff",
+        features = Wf$features,
+        stn_var = "site")),
     tar_target(full_model, dart_full(
         traindata,
         y_var = "y.diff",
@@ -162,7 +160,7 @@ list(
 
     # Summarize and report on the CV
     tar_target(cv_summary_table, cv.summary(
-        initial_cv_l2$mDT_wPred)),
+        initial_cv$mDT_wPred)),
     tar_render(initial_cv_report,
         'writing/initial_cv_report.Rmd'),
 
@@ -180,7 +178,7 @@ list(
     # Data for maps
     tar_target(T.median.mse.map.data, median.mse.map.data(
         Wf$y.sat, Wf$satellite, Wf$satellite.product, config$n.workers,
-        initial_cv_l2$mDT_wPred, pred_grid, buff, satellite_hdf_files, full_model)),
+        initial_cv$mDT_wPred, pred_grid, buff, satellite_hdf_files, full_model)),
     tar_target(T.baltimore.map.data, baltimore.map.data(
         pred_grid, buff, satellite_hdf_files, full_model)),
 

### 2022-12-20 code/xgboost_cv.R

@@ -8,12 +8,11 @@
 #' @param xgb_threads xgb_threads passed from parent function
 #' @param by_var string of "stn" or "day" passed from parent function
 #' @param progress whether to show progress bars, default TRUE
-#' @param absolute whether to use absolute loss (rather than square loss)
 #' @param ... other arguments
 #'
 run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
                           index_train, index_test, xgb_threads, by_var, n_rounds,
-                          progress = TRUE, seed = 1234, absolute = F, ...){
+                          progress = TRUE, seed = 1234, ...){
   y_var_pred <- paste0(y_var, "_pred") # name of the predicted y
   Y <-  dataXY_df[, ..y_var]
   data_X <- dataXY_df[, -..y_var]
@@ -38,8 +37,8 @@ run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
       # by default, gives 100 rounds, and it is enough by experience
       n.rounds = n_rounds,
       d = dataXY_df[index_train[[i]],], dv = y_var, ivs = colnames(data_X),
-      objective = (if (absolute) "logcosh" else "reg:squarederror"),
-      eval_metric = (if (absolute) "mae" else "rmse"),
+      objective = "reg:squarederror",
+      eval_metric = "rmse",
       progress = progress, nthread = xgb_threads)
     # manually select and store some params
     xgb_param_dart <- c(rsxgb0$model$params[c(1,2,4, 6:11)], nrounds = rsxgb0$model$niter)

### 2022-12-20 R/CONUS_AOD.Rmd
call error envelope in CONUS_AOD

@@
 
+# now with our error envelope function
+system.time(envelope <- empirical.error.envelope(dt)); head(envelope)
 ```

### 2022-12-20 R/CONUS_AOD.Rmd
refine error envelope

@@
 # now with our error envelope function
-system.time(envelope <- empirical.error.envelope(dt)); head(envelope)
+envelope.sat <- empirical.error.envelope(dt, pred = "y.sat") 
+head(envelope.sat)
+
+# and for the corrected data
+envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred") 
+head(envelope.ground.pred)
 ```

### 2022-12-20 R/paper_functions.R

@@ -23,18 +23,20 @@ performance_aod <- function(dt, ...){
 #' Try out empirical error envelope parameters
 #'
 empirical.error.envelope <- function(dt,
-                                     add.term = seq(0.01, 0.1, by = 0.01),
-                                     mult.term = seq(0.05, 0.3, by = 0.01),
+                                     obs = "y.ground",
+                                     pred = "y.sat",
+                                     add.term = seq(0.01, 0.1, by = 0.005),
+                                     mult.term = seq(0.05, 0.3, by = 0.005),
                                      threshold = 0.6,
                                      coverage = 2/3){
   terms <- CJ(add.term, mult.term)
   coverage.percent <- function(i, dt, ...) {
     temp <- unlist(i)
-    with(dt, mean(y.sat > y.ground - temp[["add.term"]] - temp[["mult.term"]]*y.ground &
-                            y.sat < y.ground + temp[["add.term"]] + temp[["mult.term"]]*y.ground))
+    with(dt, mean(get(pred) > get(obs) - temp[["add.term"]] - temp[["mult.term"]]*get(obs) &
+                            get(pred) < get(obs) + temp[["add.term"]] + temp[["mult.term"]]*get(obs)))
   }
-  terms[, cov.low := coverage.percent(.SD, dt[y.ground <= threshold]), by = row.names(terms)]
-  terms[, cov.high := coverage.percent(.SD, dt[y.ground > threshold]), by = row.names(terms)]
+  terms[, cov.low := coverage.percent(.SD, dt[get(obs) <= threshold]), by = row.names(terms)]
+  terms[, cov.high := coverage.percent(.SD, dt[get(obs) > threshold]), by = row.names(terms)]
   terms[, cov.overall := coverage.percent(.SD, dt), by = row.names(terms)]
   # return the parameter combo with the smallest abs diff from desired coverage
   terms <- terms[order(abs(coverage - cov.low) + abs(coverage - cov.high), decreasing = FALSE),]

### 2022-12-21 code/targets.R
Report the last-modifed date for AERONET.
Closes #30 (per conversation).

@@ -75,15 +75,15 @@ list(
     tar_target(aer_stn_path, download(
         "https://aeronet.gsfc.nasa.gov/aeronet_locations_v3.txt",
         "aeronet_stations.txt")),
+    tar_target(aer_orig_obs_path, download(
+        "https://aeronet.gsfc.nasa.gov/data_push/V3/AOD/AOD_Level20_All_Points_V3.tar.gz",
+        "aeronet_observations.tar.gz")),
     tar_target(aer_files_path,
-       {p = download(
-            "https://aeronet.gsfc.nasa.gov/data_push/V3/AOD/AOD_Level20_All_Points_V3.tar.gz",
-            "aeronet_observations.tar.gz")
-        assert(0 == unlink(intermediate.path("aeronet"),
+       {assert(0 == unlink(intermediate.path("aeronet"),
             recursive = T))
         assert(dir.create(intermediate.path("aeronet")))
         assert(0 == system2("tar", shQuote(c(
-            "--extract", "--file", p,
+            "--extract", "--file", aer_orig_obs_path,
             "--directory", intermediate.path("aeronet")))))
         intermediate.path("aeronet/AOD/AOD20/ALL_POINTS/")}),
     tar_target(aer_stations, format = 'fst_dt', fread(

### 2022-12-22 writing/CONUS_AOD.Rmd
Report the last-modifed date for AERONET.
Closes #30 (per conversation).
Simplified some inline code.
Renamed lots of variables.

@@
 
+We used a copy of the AERONET observations that was last modified `r lubridate::as_date(tz = "UTC", file.mtime(tar_read(aer_orig_obs_path)))`.
+
 # Results
@@
 ```{r table_cv_results, echo=FALSE}
-kbl(pretty.table.numbers(tar_read(cv_summary_table)[1, -"Year"]), align = "l") %>%
+kbl(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]), align = "l") %>%
   kable_styling(position = "left",
@@
 ```{r stratified, echo=FALSE}
-dt <- copy(tar_read(initial_cv)$mDT_wPred)
+dt <- copy(tar_read(cv)$mDT_wPred)
 ```
@@
 ```{r strata_year, echo=FALSE, eval=TRUE}
-kbl(pretty.table.numbers(tar_read(cv_summary_table)[-1]),
+kbl(pretty.table.numbers(tar_read(cv.summary)[-1]),
     align = "l") %>%
@@
 # what do the SHAPs tell us?
-tar_load(initial_cv)
-shaps <- initial_cv$"shap_score"
+tar_load(cv)
+shaps <- cv$"shap_score"
 # sort by importance
@@
 ```{r shap_summary, echo=FALSE, dev='png', dpi=300, eval=FALSE}
-shap_long <- shap.prep(shap_contrib = initial_cv$shap_score,
-                       X_train = initial_cv$mDT_wPred[, names(initial_cv_l2$shap_score), with = F])
+shap_long <- shap.prep(shap_contrib = cv$shap_score,
+                       X_train = cv$mDT_wPred[, names(cv$shap_score), with = F])
 # **SHAP summary plot**
@@
 pred.map(
-    tar_read(T.median.mse.map.data)$pred, tar_read(pred_grid),
+    tar_read(median.mse.map.data)$pred, tar_read(pred.grid),
     bg.sf = get_conus(), color.scale.name = "AOD",
@@
 
-This map shows a tile on `r tar_read(T.median.mse.map.data)$date`.
+This map shows a tile on `r tar_read(median.mse.map.data)$date`.
 
@@
 
-```{r pred.map.balitmore, echo = F}
+```{r pred.map.baltimore, echo = F}
 pred.map(
-    tar_read(T.baltimore.map.data), tar_read(pred_grid),
+    tar_read(baltimore.map.data), tar_read(pred.grid),
     bg.sf = get_conus(), color.scale.name = "AOD")

### 2022-12-22 code/data.R

@@ -2,7 +2,7 @@
 ## * Region geometry
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-get_aoi_buffer = function(region)
+get.region.shape = function(region)
    {buffer.size.m = 5000
 
     region = (
@@ -34,7 +34,7 @@ get_conus <- function(){
 ## * Satellite data
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 
-satellite.tiles = function(buff)
+satellite.tiles = function(region.shape)
 # Get the names of the MODIS tiles needed to cover the buffer.
    {tiles = st_zm(drop = T, read_sf(layer = "Features", download(
         "http://web.archive.org/web/2021id_/https://modis.ornl.gov/files/modis_sin.kmz",
@@ -43,13 +43,13 @@ satellite.tiles = function(buff)
         curl = "-L")))
     tiles = str_match(
         tiles[st_intersects(sparse = F, tiles,
-            st_transform(buff, crs = st_crs(tiles))),]$Name,
+            st_transform(region.shape, crs = st_crs(tiles))),]$Name,
         "h:(\\d+) v:(\\d+)")
     sort(sprintf("h%02dv%02d",
         as.integer(tiles[,2]), as.integer(tiles[,3])))}
 
 #' Compute the grid on which all predictions will be made.
-make_pred_grid = function(satellite.product, earthdata.rows)
+get.pred.grid = function(satellite.product, earthdata.rows)
    {r = do.call(terra::merge, lapply(1 : nrow(earthdata.rows),
         function(i) with(earthdata.rows[i],
            {r = read_satellite_raster(satellite.product, tile, path)[[1]]
@@ -60,7 +60,7 @@ make_pred_grid = function(satellite.product, earthdata.rows)
         labels = levels(earthdata.rows$tile))
     r}
 
-expand.to.overpasses = function(d, satellite_hdf_files, the.satellite, satellite.product, n.workers)
+expand.to.overpasses = function(d, satellite.files, the.satellite, satellite.product, n.workers)
    {if (!multipass.sat(satellite.product))
         return(cbind(d, overpass := NA_integer_))
     # Each file contains multiple overpasses and multiple
[…]

### 2022-12-22 code/globals.R

@@ -55,7 +55,7 @@ Wf$date.example = switch(Wf$satellite.product,
   # exists on all tiles.
     mcd19a2 = as.Date("2010-07-03"),
     geonexl2 = as.Date("2018-07-03"))
-if (!is.null(Wf$test_small_daterange) && Wf$test_small_daterange)
+if (!is.null(Wf$test.small.daterange) && Wf$test.small.daterange)
    {Wf$years = year(Wf$date.example)
     Wf$dates = Wf$date.example + (-1:1)}
 stopifnot(Wf$date.example %in% Wf$dates)

### 2022-12-22 code/modeling.R

@@ -26,7 +26,7 @@ get.threads <- function(){
 #' @param stn_var if binning by stations, provide the station column name
 #' @param day_var if binning by days, provide the day column name
 # Depends on functions in xgboost_cv_RFE.R
-initial_cv_dart <- function(
+cv_dart <- function(
   data,
   y_var,
   features,
@@ -228,8 +228,8 @@ helper.pack.bins = function(object.sizes, n.bins){
 
 #' Summarize CV statistics.
 #'
-#' @param d The data table `mDT_wPred` returned by `initial_cv_dart`.
-cv.summary = function(d)
+#' @param d The data table `mDT_wPred` returned by `cv_dart`.
+get.cv.summary = function(d)
     rbind(
         d[, c(list(Year = "all"), eval(performance.j))],
         d[, keyby = .(Year = year(time.sat)), eval(performance.j)])
@@ -312,12 +312,12 @@ new.preds = function(dt.start, dt.end, cells = NULL, targets = NULL)
   # - `dt.start` and `dt.end` specify a range of datetimes to make
   #   predictions for, as `POSIXct` objects.
   # - `cells`, if provided, should be a vector of cell numbers of
-  #   `pred_grid`. Otherwise, we use all cells.
+  #   `pred.grid`. Otherwise, we use all cells.
    {if (is.null(targets))
        {message("Reading targets")
-        grid = tar_read(pred_grid)
-        sat = tar_read(satellite_hdf_files)
-        model = tar_read(full_model)}
+        grid = tar_read(pred.grid)
+        sat = tar_read(satellite.files)
+        model = tar_read(model.full)}
     else
        {grid = targets[[1]]
         sat = targets[[2]]

### 2022-12-22 code/paper_functions.R

@@ -27,7 +27,7 @@ empirical.error.envelope <- function(dt,
 #' Agency's (EPA) Air Quality System. The unit is μg/m^3.
 #' File source: https://aqs.epa.gov/aqsweb/airdata/download_files.html#Daily
 #' Documentation: https://aqs.epa.gov/aqsweb/documents/about_aqs_data.html
-get_aqs_obs = function(years, grid)
+get.aqs.obs = function(years, grid)
    {aqs.url.root = "https://aqs.epa.gov/aqsweb/airdata"
     parameter.code = 88101L
       # PM_{2.5} from a federally approved reference or equivalent
@@ -58,7 +58,7 @@ get_aqs_obs = function(years, grid)
     setcolorder(d)
     d}
 
-satellite_vs_aqs = function(satellite, aqs)
+get.satellite.vs.aqs = function(satellite, aqs)
    {message("Merging with AQS")
     d = merge(
        satellite[, .(cell, y.sat.old, y.sat.new,
@@ -81,9 +81,9 @@ satellite_vs_aqs = function(satellite, aqs)
     out
     }
 
-median.mse.map.data = function(
+get.median.mse.map.data = function(
       y.sat.name, the.satellite, satellite.product, n.workers,
-      d, pred_grid, buff, satellite_hdf_files, full_model)
+      d, pred.grid, region.shape, satellite.files, model.full)
    {# Get the date with the median MSE.
     date = (d
         [, by = .(date = lubridate::as_date(time.sat, tz = "UTC")),
@@ -91,7 +91,7 @@ median.mse.map.data = function(
         [which.min(mse - median(mse)), date])
 
     # Get overpasses for this data.
-    sat = satellite_hdf_files[time == date]
+    sat = satellite.files[time == date]
     sat[, sat.files.ix := .I]
     overpasses = merge(by = "sat.files.ix", sat,
         expand.to.overpasses(sat, sat,
@@ -108,13 +108,13 @@ median.mse.map.data = function(
         if (!nrow(d))
             return(integer())
         # Get only cells in the study area.
-        d[in.sf(x, y, terra::crs(r), buff), cell]})
+        d[in.sf(x, y, terra::crs(r), region.shape), cell]})
 
     # Choose the tile-time with the most non-missing values.
     oi = which.max(sapply(overpass.local.cells, length))
     # Get the prediction-grid cell for each local cell.
     message("Getting cell indices")
[…]

### 2022-12-22 code/targets.R

@@ -70,9 +70,9 @@ list(
     tar_target(aer_stations, format = 'fst_dt', fread(
         aer_stn_path, col.names = c('Site_Name', 'lon', 'lat', 'elevm'))),
 
-    tar_target(buff, get_aoi_buffer(
+    tar_target(region.shape, get.region.shape(
         Wf$region)),
-    tar_target(satellite_hdf_files, switch(Wf$satellite.product,
+    tar_target(satellite.files, switch(Wf$satellite.product,
         mcd19a2 =
            {d = get.earthdata(
                 satellite_hdf_root,
@@ -80,7 +80,7 @@ list(
                 satellites = (if (Wf$satellite %in% c("terra", "aqua"))
                     "terra.and.aqua" else
                     stop()),
-                tiles = satellite.tiles(buff),
+                tiles = satellite.tiles(region.shape),
                 dates = Wf$dates)
             setnames(d, "date", "time")
             d},
@@ -99,9 +99,9 @@ list(
                     tile = factor(basename(dirname(paths))),
                     path = paths),
                 lubridate::as_date(time) %in% Wf$dates)})),
-    tar_target(pred_grid, format = terra.rast.fmt, make_pred_grid(
+    tar_target(pred.grid, format = terra.rast.fmt, get.pred.grid(
         Wf$satellite.product,
-        satellite_hdf_files[
+        satellite.files[
             lubridate::as_date(time) == Wf$date.example,
             by = tile,
             head(.SD, 1)])),
@@ -109,58 +109,58 @@ list(
     # AERONET processing for this region
     tar_target(aer, select_stations(
         aer_stations,
-        buff,
-        terra::crs(pred_grid))),
+        region.shape,
+        terra::crs(pred.grid))),
     tar_target(aer_filtered, format = 'fst_dt',
         get_stn_data(
             aod_dir = aer_files_path,
             stations = aer)[aer_date %in% Wf$dates]),
 
     # This step is where most of the satellite data is read.
-    tar_target(traindata, format = 'fst_dt', make_traindata(
+    tar_target(traindata, format = 'fst_dt', get.traindata(
         Wf$satellite.product, Wf$satellite,
         Wf$y.sat, Wf$features, Wf$window.radius,
         aer_filtered,
         aer,
-        satellite_hdf_files,
[…]

### 2022-12-22 writing/CONUS_AOD.Rmd
Made a factor-level assignment more predictably correct.

@@
 suppressPackageStartupMessages(
-  {library(targets)
-    library(Just.universal)
-    library(data.table)
-    library(DT)
-    library(patchwork)
-    library(ggplot2)
+   {library(patchwork)
     library(ggridges)
     library(ggpointdensity)
-    library(sf)
-    library(Just.universal)
-    library(knitr)
     library(kableExtra)
-    library(RColorBrewer)
-    library(SHAPforxgboost)
-    # library(flextable)
   })
-# set_flextable_defaults(fonts_ignore=TRUE)
-source(here::here("code/functions.R"))
-source(here::here("code/paper_functions.R"))
 the.sat = tools::toTitleCase(Wf$satellite)
@@
 # Color by quantiles
-dt[, month := factor(month(time.sat),
-                     labels = month.abb)]
+dt[, month := structure(class = "factor",
+                        month(time.sat),
+                        levels = month.abb)]
 p1 <- ggplot(dt, aes(x = y.diff, y = month, fill = factor(stat(quantile)))) +
@@
 *End of report*
+
+## Comparison with AQS
+
+We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$nobs)` observations of AQS on `r scales::comma(l$celldays)` cell-days.
+The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.

### 2022-12-22 code/targets.R
Removed some no-ops from the paper code.

@@ -165,8 +165,7 @@ list(
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2") list(
         tarchetypes::tar_render(paper_conus_html, output_format = "html_document",
-            packages = c('sf', 'patchwork'),
             'writing/CONUS_AOD.Rmd', quiet = F),
         tarchetypes::tar_render(paper_conus_pdf, output_format = "pdf_document",
-            packages = c('sf', 'patchwork', 'magick'),
+            packages = c('magick'),
             'writing/CONUS_AOD.Rmd')))

### 2022-12-26 writing/CONUS_AOD.Rmd
Reorder paper text

--- 2022-12-22 writing/CONUS_AOD.Rmd
@@
 
-## Agreement Plots
+## Error Envelope
+
+The most important difference in our comparison of MAIAC AOD with AERONET observations versus prior validation analyses is the single-pixel comparison (substantially closer spatiotemporal matchups).
+Here we only include comparisons over CONUS which restricts the range of AOD and aerosol properties.
+
+```{r error_envelope_AMT2018, echo=TRUE}
+# error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
+dt[, round(100 * proportions(table(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground))))]
+```
 
 ```{r error_envelope, echo=TRUE}
-# error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
-dt[, round(100 * proportions(table(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground))))]
-# sure looks like we would have a much smaller error envelope (in both additive and multiplicative terms)
-dt[, round(100 * proportions(table(y.ground.pred > (y.ground - 0.025 - 0.055*y.ground) & y.ground.pred < (y.ground + 0.025 + 0.055*y.ground))))]
-
 # now with our error envelope function
 envelope.sat <- empirical.error.envelope(dt, pred = "y.sat")
-head(envelope.sat)
+kable(head(envelope.sat), digits = 4)
 
@@
 envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred")
-head(envelope.ground.pred)
+kable(head(envelope.ground.pred), digits = 4)
+
 ```
+
+## Agreement Plots
 
@@
 
+This map shows a tile on `r tar_read(median.mse.map.data)$date`.
+
 ```{r pred.map.median.mse, echo = F}
@@
 
-This map shows a tile on `r tar_read(median.mse.map.data)$date`.
+### Baltimore
 
-### Baltimore
+A day with long-range transported smoke from Canadian wildfires.
 
@@
 
-*End of report*
-
 ## Comparison with AQS
@@
 The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
+
+*End of report*

### 2023-01-03 code/util.R
Ensure the return value of `pretty.table.numbers` can print

@@ -16,4 +16,4 @@ pretty.table.numbers = function(d)
             sprintf("%.03f", get(col))
         else
             get(col))]
-    d}
+    d[]}

### 2023-01-03 code/libraries.R
Added an example Quarto document.

@@ -7,4 +7,5 @@ suppressPackageStartupMessages(
     library(xgboost)
     library(data.table)
     library(pbapply)})
-loadNamespace("fst")  # Used for targets with `format = "fst_dt"`
+loadNamespace("fst")    # Used for targets with `format = "fst_dt"`
+loadNamespace("quarto") # Used by `tarchetypes::tar_quarto`

### 2023-01-03 code/targets.R

@@ -168,4 +168,6 @@ list(
             'writing/CONUS_AOD.Rmd', quiet = F),
         tarchetypes::tar_render(paper_conus_pdf, output_format = "pdf_document",
             packages = c('magick'),
-            'writing/CONUS_AOD.Rmd')))
+            'writing/CONUS_AOD.Rmd'),
+        tarchetypes::tar_quarto(quarto_example,
+            'writing/quarto_example.qmd', quiet = F)))

### 2023-01-04 writing/CONUS_AOD.Rmd
update YAML and Zotero library

@@
 ---
-title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
+title: "Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS"
 header-includes:
@@
   - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
-  - \usepackage{helvet} \renewcommand\familydefault{\sfdefault}
+  - \usepackage{helvet} 
+  - \renewcommand{\familydefault}{\sfdefault}
+bibliography: references.json
+csl: https://www.zotero.org/styles/atmospheric-measurement-techniques
 output_format:
-  pdf_document: default
   html_document: default
 always_allow_html: true
-bibliography: references.json
 knit: (function(inputFile, encoding) {

### 2023-01-04 writing/CONUS_AOD.Rmd
CONUS_AOD intro methods results draft text

@@
     library(ggpointdensity)
+    library(ggdist)
     library(kableExtra)
@@
 
-## Objective
+# Abstract
 
-Evaluate the performance of the `Earth_obs_cleaning` software tool on MAIAC MCD19A2 for `r the.sat` across the Continental USA, 2000-2021
+We present a modular and reproducible Open Source software tool that ingests remotely sensed data for supervised machine learning (ML) of the difference between Earth observation (EO) retrievals and ground truth observations in order to generate bias-corrected products.
+
+# Short summary
+
+We demonstrate a reproducible workflow to reduce retrieval error from satellite-based datasets with an application to MODIS Aerosol Optical Depth over the conterminous United States 2000-2021.
+It facilitates the often laborious task of obtaining and comparing ground-based and remotely sensed quantities for product validation and evaluating retrieval biases.
+The ML data-driven approach demonstrates impressive improvements in reducing retrieval error with real-world impacts for air quality modeling.
 
@@
 
+Many types of EO are validated using "gold standard" datasets obtained from networks of ground stations.[1--4](https://paperpile.com/c/ze3giJ/fWox+jvWM+EDSi+RF7S) When satellite observations are aligned with these ground measures in space and time, and the ground station data are considered top quality and are observations of the same phenomenon, then we can use the difference in these quantities as an operational measure of retrieval error (setting aside additional uncertainty in the measurements).
+Making these EO-to-ground comparisons is a key task in product validation and is often conducted globally and for particular space-time regions or within the conditions of particular co-occurring phenomena.
+While various metrics for these validation comparisons include an assessment of mean bias and visualizations of mean differences,[5](https://paperpile.com/c/ze3giJ/vA4s) an overall summary of these differences across a collocated dataset is the Mean Square Error (MSE).
+The process of adjusting an EO dataset to account for systematic error, conditional on covariates, is also known as bias-correction and can lead to datasets that more accurately reflect the ground truth.
+Using data from the ground-stations to define our desired output, along with covariates from the same EO dataset that explain the retrieval error, makes this process "data-driven".
+The application of a leading supervised machine-learning approach, extreme gradient boosting (XGBoost),[6](https://paperpile.com/c/ze3giJ/72SB) which can capture complex interactions and outperforms related algorithms,[@just2018] has great empirical potential so long as care is taken to avoid overfitting (e.g., capturing phenomena that would not generalize to new data).[7--9](https://paperpile.com/c/ze3giJ/NejN+4wlV+gFoW)
+
+We have previously demonstrated a closely-related data-driven analysis in the Northeast USA but lacked reproducible analytic code or a configurable and automated software tool.[7](https://paperpile.com/c/ze3giJ/NejN) Our XGBoost model empirically corrects biases related to complex interactions of acquisition/sensor and regional characteristics (e.g., aerosol composition or surface brightness) without reliance on any external data assimilation.
+Using an interpretability method, such as Shapley Additive Explanations (SHAP),[11](https://paperpile.com/c/ze3giJ/TgOX) to interpret our predictions, we see that our model accounts for complex interactions with a physical basis (e.g., challenging backscattering configurations).
+We also adapted our approach to refine MAIAC Column Water Vapor (CWV) achieving a 27% decrease in Terra CWV RMSE, with external validation at independent SuomiNet GPS receivers.[8](https://paperpile.com/c/ze3giJ/4wlV)
+
+Satellite-borne EO retrievals are physics-based and often validated with reliable ground networks.
+Building collocated datasets of EO over regions and time periods of interest is facilitated by both specialized tools for particular products (e.g., those from LP DAAC[14](https://paperpile.com/c/ze3giJ/KNHa) and ORNL DAAC[15](https://paperpile.com/c/ze3giJ/lQFH), although neither support MAIAC AOD) and structured access protocols such as OPeNDAP.[16](https://paperpile.com/c/ze3giJ/4m2u) When validating aerosol retrievals with ground-based observations, it is a common convention to spatially average retrieval products \-- even up to a 40 km \* 40 km area.[3](https://paperpile.com/c/ze3giJ/EDSi) This appears, at least in part, to be a legacy of 1) desiring to calculate local variance in products that were coarser (10 km resolution) than the more recent retrieval products like 1 km MAIAC and 2) a concern that accommodating temporal mismatches that were larger under prior AERONET sampling regimen (our analysis of AERONET 2018-2021 over CONUS finds an overall station-level median retrieval frequency of 5 minutes) in turn requires spatial averaging (since an aerosol air mass may move at 50 km/h or \~13.9 m/s).[17](https://paperpile.com/c/ze3giJ/TRsT) The consequence of these large-area averaged comparisons is that they do not directly apply to the performance of the product at the original resolution (e.g., 1 km for MAIAC).
+Leveraging data-driven tools to validate high resolution (e.g., 1 km) products enables a time-efficient and cost-effective ML-based technology for historical and near real-time validation and bias correction.
+Thus a configurable tool also offers finer spatio-temporal domains (versus commonly global validation papers) with single pixel performance.
+
+The GMAO MERRA-2[18](https://paperpile.com/c/ze3giJ/QSDA) uses a data-driven neural network retrieval (NNR) for AOD directly training on cloud-screened MODIS (as well as other sensors) observed radiances along with key modifiers of light scattering and retrieval error (e.g, view geometry, cloud fraction, and surface albedo) versus AERONET AOD.[19](https://paperpile.com/c/ze3giJ/1aAj) The NNR relies on the older MODIS AOD Collection 5 (C5) (while currently available is Collection 6.1 (C6.1)[20](https://paperpile.com/c/ze3giJ/oGky)), and does not build on the physics-based retrievals of MODIS operational products.
+The NNR lacks interpretability in how it differs from physics-based retrievals, and ultimately offers no code or reproducible tool/approach that can be adapted for other remote sensing products.
+A key distinction of our approach is that we model the difference between the retrieval and the ground observation as our target to be estimated, rather than predicting the retrieval product (e.g., AOD) directly.
+Our data-driven bias correction strategy leverages and supports the physics-based retrieval products while making the key predictors those variables that are informative on retrieval error rather than the first-principles target phenomenon.
+
+ORNL's Fixed Sites Subsets Tool[21](https://paperpile.com/c/ze3giJ/ScfT) generates collocated land product datasets; it does not include MAIAC AOD.
+NASA also has several sophisticated graphical tools and powerful APIs ranging from Worldview to EarthData search that serve as visualization and data access platforms, however these generally do not handle collocation with ground measures or any retrieval-error correction.
+Several groups, including our own, have published on retrieval error/bias correction on Low Earth orbit (for MODIS[7,22,23](https://paperpile.com/c/ze3giJ/01P2+NejN+9sah)) and Geostationary satellite-based datasets (ABI[2](https://paperpile.com/c/ze3giJ/jvWM)) demonstrating data-driven approaches in the research stage, although our publication with CWV[8](https://paperpile.com/c/ze3giJ/4wlV) is the only one with public code sharing.[24](https://paperpile.com/c/ze3giJ/1rUv) To our knowledge, we are presenting the first tool that will enable collocation and improved output data in a reproducible and performant workflow.
+Although demonstrated for CONUS, we have developed flexible user parameterization for a reproducible workflow tool that can be deployed on any Linux system.
+
 # Methods
 
-We used a copy of the AERONET observations that was last modified `r lubridate::as_date(tz = "UTC", file.mtime(tar_read(aer_orig_obs_path)))`.
+### Study Domain
+
+The study region was the conterminous United States (CONUS) and all match-ups of MODIS data were based on the 1-km grid cells in the Global sinusoidal Coordinate Reference System.
+The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`.
+The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file.
+While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra and report results for Aqua in Supplementary Data.
+
+### AERONET Data
+
+The workflow retrieved the global AERONET archive (Version 3 Direct Sun algorithm) that was last modified `r lubridate::as_date(tz = "UTC", file.mtime(tar_read(aer_orig_obs_path)))`.
+Only Level 2.0 (cloud screened and quality-assured) data were used.
+AERONET data were interpolated to 470 nm, matching the MAIAC blue-band AOD retrieval, using a prediction from a linear regression for each AERONET site-time with a quadratic term on log-transformed observations using AERONET's reported 'exact wavelengths', restricted to a subset of relevant wavelengths (≤1000 nm) where there was at least one observation less than and greater than 470 nm.
+
+MAIAC AOD retrievals from all overpasses in the overlying 1-km grid cell were matched to the closest AERONET observation in time (no more than 8 minutes).
+Because the purpose of the `Earth_obs_cleaning` workflow is to reduce retrieval error, there was no restriction based on quality codes of input MAIAC data.
+The targeted estimand in all analyses was the difference of the two AODs (*diff~AOD~*): MAIAC AOD minus AERONET observation, our empirical estimate of retrieval error.
+
+### Predictors
+
+The model to predict *diff~AOD~* included the following 12 predictors which were all extracted from the Science Dataset (SDS) of the MCD19A2 product or calculated over spatial focal windows (within tile-overpass):
+
+-   the MAIAC AOD
+
+-   the time of the satellite acquisition
+
+-   the blue-band AOD uncertainty
+
+-   cosine of the Solar Zenith Angle
+
+-   cosine of the View Zenith Angle
+
+-   Relative Azimuth
+
+-   Scattering Angle
+
+-   Glint Angle
+
+-   Column Water Vapor
+
+-   a dichotomous indicator for QA_best
+
+-   the mean AOD calculated from a square window with an 11-km side-length
+
+-   the proportion missing AOD from a square window with an-11 km side-length
+
+Thus, the `Earth_obs_cleaning` workflow uses only SDS layers and predictors derived from the original EO product and includes no external meteorology or data assimilation.
+
+### Modeling and performance metrics
+
+Modeling used XGBoost for regression with a squared error objective function and the Dropouts meet Multiple Additive Regression Trees (DART) booster.
+Briefly, 50 candidate sets of hyperparameters, which govern the flexibility of the gradient boosting model, were generated using an efficient latin hypercube sampling and evaluated within nested cross-validation (internal to the training data) and with several limits on overall model complexity for parsimony, speed, and to avoid overfitting.
+Because data splitting used a 5-fold site-level block cross validation, we thus only evaluated the corrected AOD at sites that were never used in model training.
+
+Our primary metric to evaluate the difference between MAIAC and AERONET before and after applying our correction was the root mean square error (RMSE), which is at the original scale of the AOD and can be compared in magnitude with the overall variability based on the standard deviation (SD) of the AERONET AOD.
+We also report the SD for raw and corrected MAIAC AOD to examine how our empirical adjustments impact variability.
+We calculate bias (the mean difference of satellite AOD versus AERONET observations) for the raw and corrected products.
+We also report the Pearson correlation of the raw and corrected products versus the collocated AERONET measures.
+Finally, we report on the distribution of differences between the acquisition time for each overpass-tile and the nearest AERONET observation in time.
+We plotted the agreement of raw and corrected AOD versus AERONET AOD in scatterplots for the interval [0, 1] given that almost all AERONET observations in the cross-validation set fell within that range.
+To generate empirical error envelopes with performance for both low and high AOD conditions, we checked the combination of additive and multiplicative terms closest to containing \~2/3 of observations for AERONET AOD below and above 0.6.
+
+While the model was trained using all of the collocated observations generated from within the study domain defined above, we present a number of stratified tables with the same metrics listed above to examine performance of the correction procedure in various data subsets.
+Stratification variables included time/space (by year, months of the year, and NOAA climate regions), atmospheric conditions (low versus high AOD), closeness of the temporal matchup (above versus below median time difference), and two additional factors related to MAIAC retrieval (number of overpasses per day and MAIAC quality flag).
+
+### Model interpretation with SHAP
+
+We used SHAP, generated for the cross-validated predictions, to interpret the contribution of each predictor to the predictions.
+Our summary metric of variable importance was the mean of the absolute SHAP for each predictor, and individual SHAPs were plotted as a time series and versus predictor values for exploratory visualizations.
+
+### New predictions and comparison with ground monitors
+
+After cross validation, the model was refit with the entire training data set and predictions were generated to correct AOD in new locations (i.e., grid cells without AERONET sites) on demand.
+The `Earth_obs_cleaning` workflow stores these in Parquet format for efficient data storage and retrieval.
+To demonstrate that corrected AOD still shows spatial patterns consistent with expectations we generated maps of raw and corrected MAIAC on a typical day (selected as the tile with the most non-missing AOD on the day with the median improvement in MSE for AERONET stations).
+To show the correction in known high-AOD conditions we plotted the Baltimore, MA area on June 10, 2015 which had substantial subsiding long-range transported Canadian wildfire smoke as previously documented with extensive surface and atmospheric monitoring [@dreessen2016].
+
+Beyond cross-validation performance, we also sought to evaluate the utility of applying the `Earth_obs_cleaning` workflow for the correlation of satellite AOD with surface measurements of PM~2.5~ from the broad network of air quality monitors in the Environmental Protection Agency's Air Quality System (AQS).
+We downloaded all 24-hour measures of PM~2.5~ concentration (using the Federal Reference Method or Federally Equivalent Method; Parameter Code 88101) from all AQS sites within CONUS.
+We joined each unique monitor with AOD for the overlying grid cell for each overpass occurring in the same day.
+We then reweighted by the inverse of the number of PM~2.5~\~AOD matchups in each cell-day and calculated the Pearson correlation using either MAIAC AOD or corrected AOD.
 
@@
 
-```{r table_cv_results, echo=FALSE}
-kbl(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]), align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:12, width = "1.1cm")
-```
-
-Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost modeled correction factor.
-
-## Descriptives for the collocated dataset (`r the.sat`)
-
-```{r stratified, echo=FALSE}
+```{r get_cv, echo=FALSE}
 dt <- copy(tar_read(cv)$mDT_wPred)
@@
 
-While our ultimate goal is to improve AOD, our dependent variable for predictive modeling is the difference of the AOD retrieval and the AERONET AOD which is a nicely symmetric and reasonably well-behaved distribution (an estimate of the retrieval error).
+<!--# any missingness in predictors to report?-->
+
+After collocation there were \## matched AOD retrievals from \## AERONET sites in CONUS.
+The median absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation was \## minutes (25^th^ and 75^th^ percentiles of \## and \## minutes, respectively).
+While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.
 
@@
   coord_cartesian(xlim = c(-1, 1)) +
-  labs(title = paste0("Histogram of difference in colocated AOD (n=",
-                      dt[, .N], ")"),
+  labs(title = paste0("Histogram of difference in collocated AOD (n=", scales::comma(dt[, .N]), ")"),
       caption = paste0("Not shown: ", dt[y.diff < -1 | y.diff > 1, .N], " AOD differences outside [-1, 1]")) +
@@
 
-*More still to be added*
+Table \## includes the performance metrics for the full cross-validated training data set.
+Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
 
-## Stratified results for `r the.sat` CV dataset
+```{r table_cv_results, echo=FALSE}
+kbl(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]), align = "l", caption = "Table ##. Cross validation results for the entire training set") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:12, width = "1.1cm")
+```
+
+## Stratified results for cross validation dataset
 
@@
 ```{r strata_year, echo=FALSE, eval=TRUE}
-kbl(pretty.table.numbers(tar_read(cv.summary)[-1]),
+kbl(pretty.table.numbers(tar_read(cv.summary)[-1, .SD, .SDcols = -patterns("Q[1-9]")]),
     align = "l") %>%
@@
 ```
-
-We expect more 2021 data to show up over time since Level 2.0 data are often quite late to be added to the AERONET datasets.
-The 2020 results (higher SD in AERONET, fewer sites, RMSE for MAIAC is higher) are a bit of an anomaly, but also, COVID could have impacted field campaigns and regular recalibration, etc.
-If we want to look at 2020 specifically, we may want to consider restricting to long-running sites.
 
@@
 result[, Month := month.abb[Month]]
-kbl(pretty.table.numbers(result), align = "l") %>%
+kbl(pretty.table.numbers(result[, .SD, .SDcols = -patterns("Time")]), align = "l") %>%
   kable_styling(position = "left",
@@
 
-```{r plot_by_month, echo=FALSE, eval=TRUE, dev='png', dpi=300}
+```{r plot_by_month, echo=FALSE, eval=FALSE, dev='png', dpi=300}
 # Color by quantiles
@@
   labs(title = paste0("Histogram of difference in colocated AOD (n=",
-                      dt[, .N], ")"),
+                      scales::comma(dt[, .N]), ")"),
       caption = paste0("Not shown: ", dt[y.diff < -0.25 | y.diff > 0.25, .N], " AOD differences outside [-0.25, 0.25]")) +
@@
 suppressMessages(print(p1))
-```
-
-### Performance by `qa_best`
-
-```{r strata_qa, echo=FALSE}
-kbl(pretty.table.numbers(dt[, eval(performance.j), by = qa_best]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
-```
-
-The takeaway here is that even though the RMSE of MAIAC is substantially better for `qa_best` versus non-best pixels, the improvements in the predicted AOD make even the non-best pixels have lower RMSE than the raw MAIAC `qa_best` and in that sense, we are expanding the set of high quality data by correcting for biases related to QA flagging criteria.
-
-### Performance for higher versus lower AOD (AERONET > 0.6)
-
-```{r strata_high_aod, echo=FALSE, fig.pos="H"}
-kbl(pretty.table.numbers(dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
-```
-
-### Performance by number of overpasses in collocation that site-day
-
-```{r strata_overpasses, echo=FALSE, fig.pos="H"}
-dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
-kbl(pretty.table.numbers(dt[, eval(performance.j), keyby = overpasses]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
 ```
@@
 ```{r get_noaa_regions, include=FALSE}
-# assign AERONET stations to states
+# assign AERONET sites to states
 tar_load(aer_stations)
@@
 by.region = dt[, eval(performance.j), keyby = state.noaa.region]
-kbl(pretty.table.numbers(by.region),
+kbl(pretty.table.numbers(by.region[, .SD, .SDcols = -patterns("Time")]),
     align = "l") %>%
@@
   coord_sf(default_crs = st_crs(conus)) +
-  scale_size("colocated\nobservations", range = c(1, 3)) +
+  scale_size("collocated\nobservations", range = c(1, 3)) +
   scale_fill_continuous("Regional\nRMSE") +
+  guides(fill = guide_legend(order = 1), size = guide_legend(order = 2)) + 
   theme_minimal() +
@@
 
+### Performance for higher versus lower AOD (AERONET \> 0.6)
+
+```{r strata_high_aod, echo=FALSE, fig.pos="H"}
+by.highaod = dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]
+kbl(pretty.table.numbers(by.highaod[, .SD, .SDcols = -patterns("Time")]),
+    align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+```
+
+### Performance by restriction of matchup calipers
+
+```{r strata_timecalipers, echo=FALSE, fig.pos="H"}
+
+```
+
+### Performance by number of overpasses in collocation that site-day
+
+```{r strata_overpasses, echo=FALSE, fig.pos="H"}
+dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
+by.overpasses = dt[, eval(performance.j), keyby = overpasses]
+kbl(pretty.table.numbers(by.overpasses[, .SD, .SDcols = -patterns("Time")]),
+    align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+```
+
+### Performance by `qa_best`
+
+```{r strata_qa, echo=FALSE}
+by.qabest = dt[, eval(performance.j), by = qa_best]
+kbl(pretty.table.numbers(by.qabest[, .SD, .SDcols = -patterns("Time")]),
+    align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+```
+
+The takeaway here is that even though the RMSE of MAIAC is substantially better for `qa_best` versus non-best pixels, the improvements in the predicted AOD make even the non-best pixels have lower RMSE than the raw MAIAC `qa_best` and in that sense, we are expanding the set of high quality data by correcting for biases related to QA flagging criteria.
+
 ## Error Envelope
 
-The most important difference in our comparison of MAIAC AOD with AERONET observations versus prior validation analyses is the single-pixel comparison (substantially closer spatiotemporal matchups).
-Here we only include comparisons over CONUS which restricts the range of AOD and aerosol properties.
+The most important difference in our comparison of MAIAC AOD with AERONET observations versus prior validation analyses is the single-pixel comparison whereas prior analyses use averaging over much larger space and time regions.
+In this manuscript, we only include comparisons over CONUS which limits the diversity of aerosol properties and most training data are in relatively low AOD conditions.
 
-```{r error_envelope_AMT2018, echo=TRUE}
+```{r error_envelope_AMT2018, echo=FALSE}
 # error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
-dt[, round(100 * proportions(table(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground))))]
+dt[, round(100 * mean(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground)))][1]
 ```
 
-```{r error_envelope, echo=TRUE}
+```{r error_envelope, echo=FALSE}
 # now with our error envelope function
 envelope.sat <- empirical.error.envelope(dt, pred = "y.sat")
-kable(head(envelope.sat), digits = 4)
+kable(head(envelope.sat, 1), digits = 4)
 
@@
 envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred")
-kable(head(envelope.ground.pred), digits = 4)
+kable(head(envelope.ground.pred, 1), digits = 4)
 
@@
 ```
+
 Summary visualization of SHAP contributions
@@
 
-```{r pred.map.median.mse, echo = F}
+```{r pred.map.median.mse, echo = F, eval=FALSE}
 pred.map(
@@
 
-```{r pred.map.baltimore, echo = F}
+```{r pred.map.baltimore, echo = F, eval=FALSE}
 pred.map(
@@
 
-*End of report*
+# Discussion
+
+This project arose from a need to subset AOD datasets to exclude improbable values that were contaminating statistical relationships with ground-level observations of fine particulate matter (PM~2.5~) and a desire to derive data-driven rules for cleaning EO for Health and Air Quality applications.
+
+We hypothesized that the time-series methods of MAIAC might benefit from having multiple overpasses in the same day and indeed saw lower bias and lower MAIAC RMSE for observations from site-days with multiple overpasses, that was still substantially improved by our correction.
+We did not investigate any potential geophysical basis for the regional variation that we saw in raw MAIAC AOD RMSE, but note that the substantially higher SD in the Northwest US region is likely driven by the contribution of episodic wildfire-related higher AOD conditions.
+Applying our correction decreased the differences in RMSE between regions, diminishing spatial variation in retrieval error.
+Similarly, although the `QA_best` quality control flag is strongly related to AOD RMSE, we found that applying the `Earth_obs_cleaning` workflow made the corrected low-quality AOD retrievals have a lower RMSE (better accuracy) than the RMSE for the MAIAC high-quality AOD retrievals.
+
+An interesting consequence of empirical tuning with our workflow is that addressing sensor-specific sources of retrieval error can lead to more harmonious datasets between platforms, a key goal for NASA's long-term Earth System Data Records.
+For example, we previously showed that MAIAC CWV from the aging Terra MODIS had worse agreement with AERONET CWV than that of Aqua MODIS, particularly in more recent years, yet the remaining retrieval error after our bias-correction for Terra and Aqua CWV was virtually the same.[8](https://paperpile.com/c/ze3giJ/4wlV) Thus, empirical correction of aerosol observations from, for example, the VIIRS instruments that continues the legacy of MODIS,[13](https://paperpile.com/c/ze3giJ/zr7Y) would be expected to result in greater consistency of the aerosols Earth System Data Records.
+Our workflow uses ancillary and derived quality control features for a data-driven refinement of retrievals to make EO more reliable and useful to a broader community of end users.
+
+The premise that satellite retrievals representing an area are comparable to point-based observations is common in product validation and among the most frequent uses of the often cited AERONET products.
+We use an unusually stringent space-time matchup criteria (single pixel with most observations within a few minutes) in order to better match up the air mass being observed.
+The reliance on the equivalence with ground-based observations might be a greater limitation for Earth observations with larger pixel sizes.
+
+Our workflow treats all cloud-screened and quality-assured (Level 2.0) observations from AERONET instruments as reference data and does not require a minimum number of days of observation from a given site for inclusion (as might be desired in an analysis of long-term MODIS sensor drift).
+The restriction to this highest quality subset of AERONET means that the training set may have few or no examples of tricky retrieval conditions (e.g., cloud edges).
+Similarly, nearly all AERONET stations are over land and thus may not adequately represent retrieval errors over water and mixed pixels, however the MAIAC algorithm itself is applicable over land and inland water (and not the global ocean).
+We leave for a future analysis the applicability of a trained `Earth_obs_cleaning` model to other space-time regions (with similar or different aerosol properties).
+Although here we present the `Earth_obs_cleaning` trained on almost 100,000 collocated observations, we use machine-learning methods likely to perform well with substantially smaller training datasets such as would be available for more recently launched sensors or over regions with more sparse ground network coverage.
+
+Our workflow is not a gap-filling method and thus corrected results have the same coverage as the original EO product (although we rescue retrievals flagged as low quality), however such approaches which include assimilation and spatial interpolation could be applied after our correction method (which involves minimal borrowing of spatial information no more than 5 km from each grid cell).
+
+# Data availability
+
+# Supplement
+
+# Author contributions
+
+# Competing interests
+
+The authors declare that they have no conflict of interest.
+
+# Acknowledgments
+
+We thank the AERONET federation principal investigators and their staff for establishing and maintaining the 328 sun photometer sites used in this investigation.
+
+# Financial support
+
+This research has been supported by the National Institutes of Health Office of the Director (grant no.
+UH3 OD023337), the National Institutes of Health National Institute of Environmental Health Sciences (grant nos.
+P30 ES023515 and R01 ES031295).
+
+# References

### 2023-01-05 writing/CONUS_AOD.Rmd
Removed an unused `library` call.

@@
     library(ggpointdensity)
-    library(ggdist)
     library(kableExtra)

### 2023-01-05 writing/CONUS_AOD.Rmd
Converted the CONUS_AOD paper to Quarto.

@@
 title: "Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS"
-header-includes:
-  - \usepackage{float}
-  - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
-  - \usepackage{helvet} 
-  - \renewcommand{\familydefault}{\sfdefault}
+
 bibliography: references.json
 csl: https://www.zotero.org/styles/atmospheric-measurement-techniques
-output_format:
-  html_document: default
-always_allow_html: true
-knit: (function(inputFile, encoding) {
-  rmarkdown::render(inputFile, encoding = encoding,
-  output_dir = "output", output_format = "all") })
+
+df-print: kable
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_knit$set(root.dir = here::here())
-```
-
 ```{r basics, echo=FALSE}
+source.files = "globals libraries util data modeling paper_functions"
+for (filename in strsplit(source.files, " ")[[1]])
+     source(sprintf("code/%s.R", filename))
 suppressPackageStartupMessages(

### 2023-01-05 code/targets.R

@@ -164,10 +164,7 @@ list(
 
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2") list(
-        tarchetypes::tar_render(paper_conus_html, output_format = "html_document",
+        tarchetypes::tar_quarto(paper_conus_html,
             'writing/CONUS_AOD.Rmd', quiet = F),
-        tarchetypes::tar_render(paper_conus_pdf, output_format = "pdf_document",
-            packages = c('magick'),
-            'writing/CONUS_AOD.Rmd'),
         tarchetypes::tar_quarto(quarto_example,
             'writing/quarto_example.qmd', quiet = F)))

### 2023-01-05 code/targets.R
Use the file extension `.qmd` for Quarto.

@@ -165,6 +165,6 @@ list(
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2") list(
         tarchetypes::tar_quarto(paper_conus_html,
-            'writing/CONUS_AOD.Rmd', quiet = F),
+            'writing/CONUS_AOD.qmd', quiet = F),
         tarchetypes::tar_quarto(quarto_example,
             'writing/quarto_example.qmd', quiet = F)))

### 2023-01-05 writing/CONUS_AOD.qmd
Set `echo` at the top of `CONUS_AOD.qmd`.

--- 2023-01-05 writing/CONUS_AOD.Rmd
@@
 
+echo: FALSE
 df-print: kable
@@
 
-```{r basics, echo=FALSE}
+```{r basics}
 source.files = "globals libraries util data modeling paper_functions"
@@
 
-```{r git_check, echo=FALSE}
+```{r git_check}
 {
@@
 
-```{r get_cv, echo=FALSE}
+```{r get_cv}
 dt <- copy(tar_read(cv)$mDT_wPred)
@@
 
-```{r describe_dv, echo=FALSE, eval=TRUE}
+```{r describe_dv, eval=TRUE}
 kable(summary(dt[, list(y.diff)]))
@@
 
-```{r plot_dv, echo=FALSE, eval=TRUE}
+```{r plot_dv, eval=TRUE}
 ggplot(dt, aes(x = y.diff)) +
@@
 
-```{r table_cv_results, echo=FALSE}
+```{r table_cv_results}
 kbl(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]), align = "l", caption = "Table ##. Cross validation results for the entire training set") %>%
@@
 
-```{r strata_year, echo=FALSE, eval=TRUE}
+```{r strata_year, eval=TRUE}
 kbl(pretty.table.numbers(tar_read(cv.summary)[-1, .SD, .SDcols = -patterns("Q[1-9]")]),
@@
 
-```{r strata_month, echo=FALSE}
+```{r strata_month}
 result <- dt[,
@@
 
-```{r plot_by_month, echo=FALSE, eval=FALSE, dev='png', dpi=300}
+```{r plot_by_month, eval=FALSE, dev='png', dpi=300}
 # Color by quantiles
@@
 
-```{r noaa_setup, echo=FALSE}
+```{r noaa_setup}
 noaa.regions = list(
@@
 
-```{r strata_noaa_regions, echo=FALSE, fig.pos="H"}
+```{r strata_noaa_regions, fig.pos="H"}
 # performance by NOAA climate regions
@@
 
-```{r noaa_region_plot, echo=FALSE, dev='png', dpi=300}
+```{r noaa_region_plot, dev='png', dpi=300}
 ## map of performance by NOAA regions
@@
 
-```{r strata_high_aod, echo=FALSE, fig.pos="H"}
+```{r strata_high_aod, fig.pos="H"}
 by.highaod = dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]
@@
 
-```{r strata_timecalipers, echo=FALSE, fig.pos="H"}
+```{r strata_timecalipers, fig.pos="H"}
 
@@
 
-```{r strata_overpasses, echo=FALSE, fig.pos="H"}
+```{r strata_overpasses, fig.pos="H"}
 dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
@@
 
-```{r strata_qa, echo=FALSE}
+```{r strata_qa}
 by.qabest = dt[, eval(performance.j), by = qa_best]
@@
 
-```{r error_envelope_AMT2018, echo=FALSE}
+```{r error_envelope_AMT2018}
 # error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
@@
 
-```{r error_envelope, echo=FALSE}
+```{r error_envelope}
 # now with our error envelope function
@@
 
-```{r agreement_plot, echo=FALSE, dev='png', dpi=300}
+```{r agreement_plot, dev='png', dpi=300}
 p1 <- ggplot(dt[y.ground < 1 & y.sat < 1]) +
@@
 
-```{r shap_rank, echo=FALSE}
+```{r shap_rank}
 # what do the SHAPs tell us?
@@
 
-```{r shap_summary, echo=FALSE, dev='png', dpi=300, eval=FALSE}
+```{r shap_summary, dev='png', dpi=300, eval=FALSE}
 shap_long <- shap.prep(shap_contrib = cv$shap_score,
@@
 
-```{r shap_features, echo=FALSE, dev='png', dpi=300, eval=FALSE}
+```{r shap_features, dev='png', dpi=300, eval=FALSE}
 g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
@@
 
-```{r pred.map.median.mse, echo = F, eval=FALSE}
+```{r pred.map.median.mse, eval=FALSE}
 pred.map(
@@
 
-```{r pred.map.baltimore, echo = F, eval=FALSE}
+```{r pred.map.baltimore, eval=FALSE}
 pred.map(

### 2023-01-05 writing/CONUS_AOD.qmd
Evaluate code blocks unconditionally in `CONUS_AOD.qmd`.

@@
 
-```{r describe_dv, eval=TRUE}
+```{r describe_dv}
 kable(summary(dt[, list(y.diff)]))
@@
 
-```{r plot_dv, eval=TRUE}
+```{r plot_dv}
 ggplot(dt, aes(x = y.diff)) +
@@
 
-```{r strata_year, eval=TRUE}
+```{r strata_year}
 kbl(pretty.table.numbers(tar_read(cv.summary)[-1, .SD, .SDcols = -patterns("Q[1-9]")]),
@@
 
-```{r plot_by_month, eval=FALSE, dev='png', dpi=300}
+```{r plot_by_month, dev='png', dpi=300}
 # Color by quantiles
@@
 
-```{r shap_summary, dev='png', dpi=300, eval=FALSE}
+```{r shap_summary, dev='png', dpi=300}
 shap_long <- shap.prep(shap_contrib = cv$shap_score,
@@
 
-```{r shap_features, dev='png', dpi=300, eval=FALSE}
+```{r shap_features, dev='png', dpi=300}
 g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
@@
 
-```{r pred.map.median.mse, eval=FALSE}
+```{r pred.map.median.mse}
 pred.map(
@@
 
-```{r pred.map.baltimore, eval=FALSE}
+```{r pred.map.baltimore}
 pred.map(

### 2023-01-05 writing/CONUS_AOD.qmd
Removed plotting limits from a map.

@@
     bg.sf = get_conus(), color.scale.name = "AOD",
-    limits = list(lon = c(-107, Inf), lat = c(-Inf, 37)),
     quantile.cap = .99)

### 2023-01-05 writing/CONUS_AOD.qmd
Converted a code block to inline code.

@@
 
-```{r error_envelope_AMT2018}
-# error envelope for MAIAC is equivalent to that reported in AMT MAIAC publication
-dt[, round(100 * mean(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground)))][1]
-```
+The original error envelope (reported in the AMT MAIAC publication) had `r dt[, round(100 * mean(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground)))][1]`% coverage in our data.
 

### 2023-08-15 writing/CONUS_AOD.qmd
edit yaml for self-contained html
Deleted an empty section of `CONUS_AOD.qmd`.
Deleted the monthly plot.

@@
 csl: https://www.zotero.org/styles/atmospheric-measurement-techniques
+
+format: 
+  html:
+    embed-resources: true
 
@@
 
-```{r plot_by_month, dev='png', dpi=300}
-# Color by quantiles
-dt[, month := structure(class = "factor",
-                        month(time.sat),
-                        levels = month.abb)]
-p1 <- ggplot(dt, aes(x = y.diff, y = month, fill = factor(stat(quantile)))) +
-  stat_density_ridges(
-    geom = "density_ridges_gradient", calc_ecdf = TRUE,
-    quantiles = 4, quantile_lines = FALSE
-  ) +
-  geom_vline(xintercept = 0, linetype = "dashed", color = "grey90") +
-  scale_fill_brewer(name = "Quartiles", type = "div") +
-  scale_x_continuous("MAIAC AOD minus AERONET AOD", expand = c(0, 0)) +
-  coord_cartesian(xlim = c(-0.25, 0.25)) +
-  labs(title = paste0("Histogram of difference in colocated AOD (n=",
-                      scales::comma(dt[, .N]), ")"),
-      caption = paste0("Not shown: ", dt[y.diff < -0.25 | y.diff > 0.25, .N], " AOD differences outside [-0.25, 0.25]")) +
-  theme_bw()
-suppressMessages(print(p1))
-```
-
 ### Performance by NOAA climate regions
@@
   row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
-```
-
-### Performance by restriction of matchup calipers
-
-```{r strata_timecalipers, fig.pos="H"}
-
 ```

### 2023-08-15 writing/CONUS_AOD.qmd
Moved around some sections of `CONUS_AOD.qmd`.
Put time differences in text instead of tables.

@@
 After collocation there were \## matched AOD retrievals from \## AERONET sites in CONUS.
-The median absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation was \## minutes (25^th^ and 75^th^ percentiles of \## and \## minutes, respectively).
+The median absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation was `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["50%"]` s (25^th^ and 75^th^ percentiles of `r tqs["25%"]` and `r tqs["75%"]` s, respectively).
 While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.
@@
 Here we show our performance metrics in different subsets of the cross-validation results. These are all from the same model training, just slicing up how the performance looks in different contexts.
-
-### Performance by year
-
-```{r strata_year}
-kbl(pretty.table.numbers(tar_read(cv.summary)[-1, .SD, .SDcols = -patterns("Q[1-9]")]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
-```
-
-### Performance by month of year
-
-```{r strata_month}
-result <- dt[,
-    keyby = .(Month = month(lubridate::as_datetime(time.sat))),
-    eval(performance.j)]
-result[, Month := month.abb[Month]]
-kbl(pretty.table.numbers(result[, .SD, .SDcols = -patterns("Time")]), align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
-```
 
@@
 by.region = dt[, eval(performance.j), keyby = state.noaa.region]
-kbl(pretty.table.numbers(by.region[, .SD, .SDcols = -patterns("Time")]),
+kbl(pretty.table.numbers(by.region),
     align = "l") %>%
@@
 
+### Performance by month of year
+
+```{r strata_month}
+result <- dt[,
+    keyby = .(Month = month(lubridate::as_datetime(time.sat))),
+    eval(performance.j)]
+result[, Month := month.abb[Month]]
+kbl(pretty.table.numbers(result), align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+```
+
 ### Performance for higher versus lower AOD (AERONET \> 0.6)
@@
 by.highaod = dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]
-kbl(pretty.table.numbers(by.highaod[, .SD, .SDcols = -patterns("Time")]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
-```
-
-### Performance by number of overpasses in collocation that site-day
-
-```{r strata_overpasses, fig.pos="H"}
-dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
-by.overpasses = dt[, eval(performance.j), keyby = overpasses]
-kbl(pretty.table.numbers(by.overpasses[, .SD, .SDcols = -patterns("Time")]),
+kbl(pretty.table.numbers(by.highaod),
     align = "l") %>%
@@
 by.qabest = dt[, eval(performance.j), by = qa_best]
-kbl(pretty.table.numbers(by.qabest[, .SD, .SDcols = -patterns("Time")]),
+kbl(pretty.table.numbers(by.qabest),
     align = "l") %>%
@@
 
+## Performance by year
+
+```{r strata_year}
+kbl(pretty.table.numbers(tar_read(cv.summary)[-1]),
+    align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+```
+
+## Performance by number of overpasses in collocation that site-day
+
+```{r strata_overpasses, fig.pos="H"}
+dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
+by.overpasses = dt[, eval(performance.j), keyby = overpasses]
+kbl(pretty.table.numbers(by.overpasses),
+    align = "l") %>%
+  kable_styling(position = "left",
+                latex_options=c("HOLD_position","scale_down")) %>%
+  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+```
+
 # Author contributions

### 2023-08-15 code/modeling.R

@@ -236,10 +236,6 @@ get.cv.summary = function(d)
 performance.j = quote(
    {mse = function(x, y) mean((x - y)^2)
     sdn = function(x) sqrt(mse(x, mean(x)))
-    tqs = unname(quantile(
-        abs(as.numeric(difftime(units = "mins",
-            time.sat, time.ground))),
-        c(.25, .5, .75)))
     list(
         "Cases" = .N,
         "Sites" = length(unique(site)),
@@ -253,10 +249,7 @@ performance.j = quote(
         "Bias, raw" = mean(y.sat - y.ground),
         "Bias, corrected" = mean(y.ground.pred - y.ground),
         "Correlation, raw" = cor(y.sat, y.ground),
-        "Correlation, corrected" = cor(y.ground.pred, y.ground),
-        "Time difference, Q1" = tqs[1],
-        "Time difference, median" = tqs[2],
-        "Time difference, Q3" = tqs[3])})
+        "Correlation, corrected" = cor(y.ground.pred, y.ground))})
 
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 ## * New predictions

### 2023-08-15 writing/CONUS_AOD.qmd
Filled in some numbers in `CONUS_AOD.qmd`.
Adjusted the regional results.

@@
 
-After collocation there were \## matched AOD retrievals from \## AERONET sites in CONUS.
+After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS.
 The median absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation was `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["50%"]` s (25^th^ and 75^th^ percentiles of `r tqs["25%"]` and `r tqs["75%"]` s, respectively).
@@
 # performance by NOAA climate regions
-by.region = dt[, eval(performance.j), keyby = state.noaa.region]
+by.region = (dt
+    [, eval(performance.j), by = state.noaa.region]
+    [order(`Proportion of raw MSE`)])
 kbl(pretty.table.numbers(by.region),
@@
   geom_sf(aes(fill = value)) +
-  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = site], aes(geometry = geometry, size = N), alpha = 0.1) +
   facet_wrap(~variable, ncol = 1) +
   coord_sf(default_crs = st_crs(conus)) +
-  scale_size("collocated\nobservations", range = c(1, 3)) +
   scale_fill_continuous("Regional\nRMSE") +
-  guides(fill = guide_legend(order = 1), size = guide_legend(order = 2)) + 
   theme_minimal() +

### 2023-08-16 writing/CONUS_AOD.qmd
Reworded the sentence about quartiles of time differences.

@@
 After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS.
-The median absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation was `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["50%"]` s (25^th^ and 75^th^ percentiles of `r tqs["25%"]` and `r tqs["75%"]` s, respectively).
+The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["25%"]` s, and `r tqs["75%"]` s.
 While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.

### 2023-08-16 writing/CONUS_AOD.qmd
Removed an obsolete comment.
Deduplicated the code for agreement plots.

@@
 # Contiguous US state boundaries
-# will replace with a region defined by Kodi later
 conus = st_transform(crs = crs.us.atlas, get_conus())
@@
 
-The original error envelope (reported in the AMT MAIAC publication) had `r dt[, round(100 * mean(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground)))][1]`% coverage in our data.
-
-```{r error_envelope}
-# now with our error envelope function
-envelope.sat <- empirical.error.envelope(dt, pred = "y.sat")
-kable(head(envelope.sat, 1), digits = 4)
-
-# and for the corrected data
-envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred")
-kable(head(envelope.ground.pred, 1), digits = 4)
-
-```
+We use Nelder-Mead optimization to construct new error envelopes with coverage as close to 2/3 as possible for each of two subsets of the data: observations less than or equal to 0.6, and observations greater than 0.6.
+The results are an envelope with `r f = \(pred) {ps = empirical.error.envelope(obs = dt$y.ground, pred)$par; sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; f(dt$y.sat)` for the raw satellite values, and an envelope with `r f(dt$y.ground.pred)` for our predictions.
+The original error envelope (reported in the AMT MAIAC publication), applied to the raw satellite values, has coverage `r x = envelope.tester(obs = dt$y.ground, pred = dt$y.sat)(.05, .1); sprintf("%.03f", x["lo"])` in the low subset and `r sprintf("%.03f", x["hi"])` in the high subset.
 
@@
 ```{r agreement_plot, dev='png', dpi=300}
-p1 <- ggplot(dt[y.ground < 1 & y.sat < 1]) +
-  aes(y.ground, y.sat) +
+agreement.plot = \(vname.pred)
+   {i = dt[, between(y.ground, 0, 1) & between(get(vname.pred), 0, 1)]
+    ggplot(dt[i]) +
+        aes(y.ground, get(vname.pred)) +
         geom_pointdensity(alpha = 0.1, size = 0.6) +
@@
         theme(aspect.ratio = 1) +
-  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.sat > 1, .N], " AOD values outside [0, 1]"))
+        labs(caption = paste0("Not shown: ", dt[!i, .N], " AOD values outside [0, 1]"))}
 
-p2  <- ggplot(dt[y.ground < 1 & y.ground.pred > 0 & y.ground.pred < 1]) +
-  aes(y.ground, y.ground.pred) +
-  geom_pointdensity(alpha = 0.1, size = 0.6) +
-  geom_abline(linetype = "dashed") +
-  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
-  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
-  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
-  guides(color = "none") +
-  theme_bw() +
-  theme(aspect.ratio = 1) +
-  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.ground.pred > 1 | y.ground.pred < 0, .N], " AOD values outside [0, 1]"))
-
-p1 + p2
+agreement.plot("y.sat") + agreement.plot("y.ground.pred")
 ```

### 2023-08-16 code/paper_functions.R
Overhauled the error envelopes.

@@ -1,27 +1,22 @@
 # contains functions used in generating the CONUS_AOD paper
 
-#' Try out empirical error envelope parameters
-#'
-empirical.error.envelope <- function(dt,
-                                     obs = "y.ground",
-                                     pred = "y.sat",
-                                     add.term = seq(0.01, 0.05, by = 0.0025),
-                                     mult.term = seq(0.125, 0.2, by = 0.0025),
-                                     threshold = 0.6,
-                                     coverage = 2/3){
-  envelope.terms <- CJ(add.term, mult.term)
-  coverage.percent <- function(i, dt) {
-    temp <- unlist(i)
-    with(dt, mean(get(pred) > get(obs) - temp[["add.term"]] - temp[["mult.term"]]*get(obs) &
-                            get(pred) < get(obs) + temp[["add.term"]] + temp[["mult.term"]]*get(obs)))
-  }
-  envelope.terms[, cov.low := coverage.percent(.SD, dt[get(obs) <= threshold]), by = row.names(envelope.terms)]
-  envelope.terms[, cov.high := coverage.percent(.SD, dt[get(obs) > threshold]), by = row.names(envelope.terms)]
-  envelope.terms[, cov.overall := coverage.percent(.SD, dt), by = row.names(envelope.terms)]
-  # return the parameter combo with the smallest abs diff from desired coverage
-  envelope.terms <- envelope.terms[order(abs(coverage - cov.low) + abs(coverage - cov.high), decreasing = FALSE),]
-  envelope.terms
-}
+#' Find an error envelope (defined with an additive and multiplicative
+#' term) around some predicted values `pred` that cover a proportion
+#' `target.coverage` of the observed values `obs` above and below `threshold`.
+empirical.error.envelope = \(..., target.coverage = 2/3)
+   {f = envelope.tester(...)
+    optim(c(.05, .1), \(p)
+        sum(abs(target.coverage - f(p[1], p[2]))))}
+
+envelope.tester = \(obs, pred, threshold = 0.6)
+   {get.coverage = \(slice, add, mult)
+        slice[, mean(abs(pred - obs) <= add + mult*pred)]
+    d = data.table(obs, pred)
+    d.lo = d[obs <= threshold]
+    d.hi = d[obs > threshold]
+    \(add, mult) c(
+        lo = get.coverage(d.lo, add, mult),
+        hi = get.coverage(d.hi, add, mult))}
 
[…]

### 2023-08-17 code/modeling.R
Show the hyperparameters selected for the full model.

@@ -123,7 +123,7 @@ run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
   y_pred_dt <- data.table(y_pred = rep(NA_real_, n_row))
   shap_score <- as.data.table(matrix(rep(NA_real_, n_row*n_col), ncol = n_col))
   names(shap_score) <- names(data_X)
-  xgb_param_list1 <- xgb_param_list2 <- list()
+  hyperparams = data.table()
   BIAS0 <- rep(NA_real_, k_fold)
   # loop for each fold
 
@@ -137,8 +137,7 @@ run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
       objective = "reg:squarederror",
       eval_metric = "rmse",
       progress = progress, nthread = xgb_threads)
-    # manually select and store some params
-    xgb_param_dart <- c(rsxgb0$model$params[c(1,2,4, 6:11)], nrounds = rsxgb0$model$niter)
+    hyperparams = rbind(hyperparams, rsxgb0$hyperparams)
     xgbmod <- rsxgb0$model
 
     # fit model
@@ -150,9 +149,6 @@ run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
     BIAS0[i] <- first(shap_pred$BIAS)
     shap_pred[, BIAS := NULL]
     shap_score[index_test[[i]],] <- shap_pred # record the SHAP values (for whole model)
-
-    xgb_param_list1[[paste0(by_var, i)]] <- unlist(xgb_param_dart)
-    xgb_param_list2[[paste0(by_var, i)]] <- xgb_param_dart
   })
   cat("loop finished\n")
   rmse_all_folds <- sqrt(mean((y_pred_dt$y_pred - dataXY_df[[y_var]])^2))
@@ -165,8 +161,7 @@ run.k.fold.cv <- function(k_fold, dataXY_df, y_var,
               shap_score = shap_score,
               y_pred_dt = y_pred_dt,
               BIAS = BIAS0,
-              xgb_param_list1 = xgb_param_list1,
-              xgb_param_list2 = xgb_param_list2))
+              hyperparams = hyperparams))
 }
 
 #' bin the data into the specified number of folds by day or by station.
@@ -296,6 +291,7 @@ dart_full <- function(
        y_preds = preds,
        shap_pred = shap_pred,
        shap_bias = shap_bias,
+       hyperparams = xdc_out$hyperparams,
[…]

### 2023-08-17 writing/CONUS_AOD.qmd
Show the hyperparameters selected for the full model.
Made error envelopes more like they used to be.

--- 2023-08-16 writing/CONUS_AOD.qmd
@@
 We use Nelder-Mead optimization to construct new error envelopes with coverage as close to 2/3 as possible for each of two subsets of the data: observations less than or equal to 0.6, and observations greater than 0.6.
-The results are an envelope with `r f = \(pred) {ps = empirical.error.envelope(obs = dt$y.ground, pred)$par; sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; f(dt$y.sat)` for the raw satellite values, and an envelope with `r f(dt$y.ground.pred)` for our predictions.
+The results are an envelope with `r f = \(pred) {o = empirical.error.envelope(obs = dt$y.ground, pred); stopifnot(o$convergence == 0); sprintf("additive part %.03f and multiplicative part %.03f", o$par[1], o$par[2])}; f(dt$y.sat)` for the raw satellite values, and an envelope with `r f(dt$y.ground.pred)` for our predictions.
 The original error envelope (reported in the AMT MAIAC publication), applied to the raw satellite values, has coverage `r x = envelope.tester(obs = dt$y.ground, pred = dt$y.sat)(.05, .1); sprintf("%.03f", x["lo"])` in the low subset and `r sprintf("%.03f", x["hi"])` in the high subset.
@@
 
-## Maps of new predictions
+## New predictions
+
+For the model trained on all the data, the selected hyperparameters were `r v = tar_read(model.full)$hyperparams; paste0(collapse = ", ", sprintf("%s = %s", names(v), v))`.
 

### 2023-08-17 code/paper_functions.R

@@ -1,16 +1,17 @@
 # contains functions used in generating the CONUS_AOD paper
 
-#' Find an error envelope (defined with an additive and multiplicative
-#' term) around some predicted values `pred` that cover a proportion
-#' `target.coverage` of the observed values `obs` above and below `threshold`.
+#' Find an "error envelope" with an additive and multiplicative
+#' term in the fashion of Dark Target validation
+#' https://darktarget.gsfc.nasa.gov/validation
 empirical.error.envelope = \(..., target.coverage = 2/3)
    {f = envelope.tester(...)
     optim(c(.05, .1), \(p)
         sum(abs(target.coverage - f(p[1], p[2]))))}
 
 envelope.tester = \(obs, pred, threshold = 0.6)
-   {get.coverage = \(slice, add, mult)
-        slice[, mean(abs(pred - obs) <= add + mult*pred)]
+   {get.coverage = \(slice, add, mult) slice[, mean(
+        pred >= obs - add - mult*obs &
+        pred <= obs + add + mult*obs)]
     d = data.table(obs, pred)
     d.lo = d[obs <= threshold]
     d.hi = d[obs > threshold]

### 2023-08-17 writing/CONUS_AOD.qmd
Draw error envelopes in the agreement plot.

@@
 agreement.plot = \(vname.pred)
-   {i = dt[, between(y.ground, 0, 1) & between(get(vname.pred), 0, 1)]
-    ggplot(dt[i]) +
-        aes(y.ground, get(vname.pred)) +
-        geom_pointdensity(alpha = 0.1, size = 0.6) +
+   {d = dt[, .(y.ground, y.pred = get(vname.pred))]
+    i = d[, between(y.ground, 0, 1) & between(y.pred, 0, 1)]
+    sample.obs = data.table(y.ground = seq(0, 1, len = 1000))
+    ps = empirical.error.envelope(obs = d$y.ground, d$y.pred)$par
+    sample.obs[, envelope.hi := y.ground + ps[1] + ps[2]*y.ground]
+    sample.obs[, envelope.lo := y.ground - ps[1] - ps[2]*y.ground]
+    ggplot() +
+        geom_pointdensity(data = d, aes(y.ground, y.pred),
+            alpha = 0.1, size = 0.6) +
         geom_abline(linetype = "dashed") +
-        geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
-        scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
-        scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
+        geom_line(data = sample.obs, aes(y.ground, envelope.hi)) +
+        geom_line(data = sample.obs, aes(y.ground, envelope.lo)) +
+        xlab("AERONET AOD 470nm") +
+        ylab("MAIAC AOD 470nm") +
         guides(color = "none") +
         theme_bw() +
-        theme(aspect.ratio = 1) +
-        labs(caption = paste0("Not shown: ", dt[!i, .N], " AOD values outside [0, 1]"))}
+        coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
+        labs(caption = paste0("Not shown: ", d[!i, .N], " AOD values outside [0, 1]"))}
 

### 2023-08-18 writing/CONUS_AOD.qmd
Added mean error-envelope widths.

@@
 We use Nelder-Mead optimization to construct new error envelopes with coverage as close to 2/3 as possible for each of two subsets of the data: observations less than or equal to 0.6, and observations greater than 0.6.
-The results are an envelope with `r f = \(pred) {o = empirical.error.envelope(obs = dt$y.ground, pred); stopifnot(o$convergence == 0); sprintf("additive part %.03f and multiplicative part %.03f", o$par[1], o$par[2])}; f(dt$y.sat)` for the raw satellite values, and an envelope with `r f(dt$y.ground.pred)` for our predictions.
+The results are an envelope with `r eee = \(pred) {o = empirical.error.envelope(obs = dt$y.ground, pred); stopifnot(o$convergence == 0); o$par}; show.eee = \(pred) {ps = eee(pred); sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; show.eee(dt$y.sat)` for the raw satellite values, and an envelope with `r show.eee(dt$y.ground.pred)` for our predictions.
+The mean widths of the envelopes across all observations are `r me = \(pred) {ps = eee(pred); sprintf("%.03f", 2*ps[1] + 2*mean(dt$y.ground))}; me(dt$y.sat)` for the raw values and `r me(dt$y.ground.pred)` for our predictions.
 The original error envelope (reported in the AMT MAIAC publication), applied to the raw satellite values, has coverage `r x = envelope.tester(obs = dt$y.ground, pred = dt$y.sat)(.05, .1); sprintf("%.03f", x["lo"])` in the low subset and `r sprintf("%.03f", x["hi"])` in the high subset.

### 2023-08-21 writing/CONUS_AOD.qmd
Cleaned up a plot axis.

@@
 ```{r shap_features, dev='png', dpi=300}
-g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') + ggtitle("(A) SHAP values of Column WV vs. Time trend")
+g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') +
+    aes(x = lubridate::as_datetime(x_feature)) +
+    xlab("time.sat") +
+    ggtitle("(A) SHAP values of Column WV vs. Time trend")
 # g2 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'y.sat', color_feature = 'y.sat') +  ggtitle("(B) SHAP values of MAIAC AOD vs. Time trend")

### 2023-08-21 writing/CONUS_AOD.qmd
Replaced a use of `patchwork` with facets.

@@
 ```{r agreement_plot, dev='png', dpi=300}
-agreement.plot = \(vname.pred)
-   {d = dt[, .(y.ground, y.pred = get(vname.pred))]
+(\()
+   {d = melt(dt[, .(y.ground, y.ground.pred, y.sat)],
+        id = "y.ground", variable.name = "comparison", value.name = "y.pred")
     i = d[, between(y.ground, 0, 1) & between(y.pred, 0, 1)]
-    sample.obs = data.table(y.ground = seq(0, 1, len = 1000))
-    ps = empirical.error.envelope(obs = d$y.ground, d$y.pred)$par
-    sample.obs[, envelope.hi := y.ground + ps[1] + ps[2]*y.ground]
-    sample.obs[, envelope.lo := y.ground - ps[1] - ps[2]*y.ground]
+    sample.obs = `[`(
+        CJ(
+            comparison = unique(d$comparison),
+            y.ground = seq(0, 1, len = 1000)),
+        by = comparison,
+        j =
+           {ps = with(d[comparison == .BY$comparison],
+                empirical.error.envelope(obs = y.ground, y.pred)$par)
+            punl(
+                y.ground,
+                envelope.hi = y.ground + ps[1] + ps[2]*y.ground,
+                envelope.lo = y.ground - ps[1] - ps[2]*y.ground)})
     ggplot() +
@@
         guides(color = "none") +
+        facet_wrap(vars(comparison)) +
         theme_bw() +
         coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
-        labs(caption = paste0("Not shown: ", d[!i, .N], " AOD values outside [0, 1]"))}
-
-agreement.plot("y.sat") + agreement.plot("y.ground.pred")
+        labs(caption = sprintf("Not shown: %s %s and %s %s",
+           d[!i][comparison == "y.ground.pred", .N],
+           "points outside left panel",
+           d[!i][comparison == "y.sat", .N],
+           "points outside right panel"))})()
 ```

### 2023-08-22 writing/CONUS_AOD.qmd
Changed the appearance of the agreement plot.

--- 2023-08-21 writing/CONUS_AOD.qmd
@@
 ```{r agreement_plot, dev='png', dpi=300}
-agreement.plot = \(vname.pred)
-   {d = dt[, .(y.ground, y.pred = get(vname.pred))]
-    i = d[, between(y.ground, 0, 1) & between(y.pred, 0, 1)]
-    sample.obs = data.table(y.ground = seq(0, 1, len = 1000))
-    ps = empirical.error.envelope(obs = d$y.ground, d$y.pred)$par
-    sample.obs[, envelope.hi := y.ground + ps[1] + ps[2]*y.ground]
-    sample.obs[, envelope.lo := y.ground - ps[1] - ps[2]*y.ground]
+(\()
+   {d = melt(dt[, .(y.ground, y.ground.pred, y.sat)],
+        id = "y.ground", variable.name = "comparison", value.name = "y.pred")
+    d[, comparison := factor(comparison, levels = rev(levels(comparison)))]
+    d[, in.region := between(y.ground, 0, 1) & between(y.pred, 0, 1)]
+    sample.obs = `[`(
+        CJ(
+            comparison = unique(d$comparison),
+            y.ground = seq(0, 1, len = 1000)),
+        by = comparison,
+        j =
+           {ps = with(d[comparison == .BY$comparison],
+                empirical.error.envelope(obs = y.ground, y.pred)$par)
+            punl(
+                y.ground,
+                envelope.hi = y.ground + ps[1] + ps[2]*y.ground,
+                envelope.lo = y.ground - ps[1] - ps[2]*y.ground)})
     ggplot() +
-        geom_pointdensity(data = d, aes(y.ground, y.pred),
-            alpha = 0.1, size = 0.6) +
-        geom_abline(linetype = "dashed") +
-        geom_line(data = sample.obs, aes(y.ground, envelope.hi)) +
-        geom_line(data = sample.obs, aes(y.ground, envelope.lo)) +
+        geom_pointdensity(data = d[(in.region)], aes(y.ground, y.pred),
+            size = 1/4, adjust = 1/2) +
+        scale_color_gradient(low = "#dddddd", high = "black") +
+        geom_abline(linetype = "dashed", color = "#6666aa") +
+        geom_line(data = sample.obs, aes(y.ground, envelope.hi),
+            color = "#6666aa") +
+        geom_line(data = sample.obs, aes(y.ground, envelope.lo),
+            color = "#6666aa") +
         xlab("AERONET AOD 470nm") +
@@
         guides(color = "none") +
-        theme_bw() +
+        facet_wrap(vars(comparison)) +
+        theme_classic() +
         coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
-        labs(caption = paste0("Not shown: ", d[!i, .N], " AOD values outside [0, 1]"))}
-
-agreement.plot("y.sat") + agreement.plot("y.ground.pred")
+        labs(caption = sprintf("Not shown: %s %s and %s %s",
+           d[!in.region & comparison == "y.sat", .N],
+           "points outside left panel",
+           d[!in.region & comparison == "y.ground.pred", .N],
+           "points outside right panel"))})()
 ```

### 2023-08-23 writing/CONUS_AOD.qmd
Corrected the computation of envelope widths.

@@
 The results are an envelope with `r eee = \(pred) {o = empirical.error.envelope(obs = dt$y.ground, pred); stopifnot(o$convergence == 0); o$par}; show.eee = \(pred) {ps = eee(pred); sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; show.eee(dt$y.sat)` for the raw satellite values, and an envelope with `r show.eee(dt$y.ground.pred)` for our predictions.
-The mean widths of the envelopes across all observations are `r me = \(pred) {ps = eee(pred); sprintf("%.03f", 2*ps[1] + 2*mean(dt$y.ground))}; me(dt$y.sat)` for the raw values and `r me(dt$y.ground.pred)` for our predictions.
+The mean widths of the envelopes across all observations are `r me = \(pred) {ps = eee(pred); sprintf("%.03f", 2*ps[1] + 2*ps[2]*mean(dt$y.ground))}; me(dt$y.sat)` for the raw values and `r me(dt$y.ground.pred)` for our predictions.
 The original error envelope (reported in the AMT MAIAC publication), applied to the raw satellite values, has coverage `r x = envelope.tester(obs = dt$y.ground, pred = dt$y.sat)(.05, .1); sprintf("%.03f", x["lo"])` in the low subset and `r sprintf("%.03f", x["hi"])` in the high subset.

### 2023-08-23 writing/CONUS_AOD.qmd
Targetized a plot.

@@
     library(ggridges)
-    library(ggpointdensity)
     library(kableExtra)
@@
 
-```{r agreement_plot, dev='png', dpi=300}
-(\()
-   {d = melt(dt[, .(y.ground, y.ground.pred, y.sat)],
-        id = "y.ground", variable.name = "comparison", value.name = "y.pred")
-    d[, comparison := factor(comparison, levels = rev(levels(comparison)))]
-    d[, in.region := between(y.ground, 0, 1) & between(y.pred, 0, 1)]
-    sample.obs = `[`(
-        CJ(
-            comparison = unique(d$comparison),
-            y.ground = seq(0, 1, len = 1000)),
-        by = comparison,
-        j =
-           {ps = with(d[comparison == .BY$comparison],
-                empirical.error.envelope(obs = y.ground, y.pred)$par)
-            punl(
-                y.ground,
-                envelope.hi = y.ground + ps[1] + ps[2]*y.ground,
-                envelope.lo = y.ground - ps[1] - ps[2]*y.ground)})
-    ggplot() +
-        geom_pointdensity(data = d[(in.region)], aes(y.ground, y.pred),
-            size = 1/4, adjust = 1/2) +
-        scale_color_gradient(low = "#dddddd", high = "black") +
-        geom_abline(linetype = "dashed", color = "#6666aa") +
-        geom_line(data = sample.obs, aes(y.ground, envelope.hi),
-            color = "#6666aa") +
-        geom_line(data = sample.obs, aes(y.ground, envelope.lo),
-            color = "#6666aa") +
-        xlab("AERONET AOD 470nm") +
-        ylab("MAIAC AOD 470nm") +
-        guides(color = "none") +
-        facet_wrap(vars(comparison)) +
-        theme_classic() +
-        coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
-        labs(caption = sprintf("Not shown: %s %s and %s %s",
-           d[!in.region & comparison == "y.sat", .N],
-           "points outside left panel",
-           d[!in.region & comparison == "y.ground.pred", .N],
-           "points outside right panel"))})()
+```{r agreement_plot}
+knitr::include_graphics(targets::tar_read(agreement.plot.path))
 ```

### 2023-08-23 code/paper_functions.R

@@ -1,5 +1,44 @@
 # contains functions used in generating the CONUS_AOD paper
 
+agreement.plot = \(d)
+   {d = melt(d[, .(y.ground, y.ground.pred, y.sat)],
+        id = "y.ground", variable.name = "comparison", value.name = "y.pred")
+    d[, comparison := factor(comparison, levels = rev(levels(comparison)))]
+    d[, in.region := between(y.ground, 0, 1) & between(y.pred, 0, 1)]
+    sample.obs = `[`(
+        CJ(
+            comparison = unique(d$comparison),
+            y.ground = seq(0, 1, len = 1000)),
+        by = comparison,
+        j =
+           {ps = with(d[comparison == .BY$comparison],
+                empirical.error.envelope(obs = y.ground, y.pred)$par)
+            punl(
+                y.ground,
+                envelope.hi = y.ground + ps[1] + ps[2]*y.ground,
+                envelope.lo = y.ground - ps[1] - ps[2]*y.ground)})
+    ggplot() +
+        ggpointdensity::geom_pointdensity(
+            data = d[(in.region)], aes(y.ground, y.pred),
+            size = 1/4, adjust = 1/2) +
+        scale_color_gradient(low = "#dddddd", high = "black") +
+        geom_abline(linetype = "dashed", color = "#6666aa") +
+        geom_line(data = sample.obs, aes(y.ground, envelope.hi),
+            color = "#6666aa") +
+        geom_line(data = sample.obs, aes(y.ground, envelope.lo),
+            color = "#6666aa") +
+        xlab("AERONET AOD 470nm") +
+        ylab("MAIAC AOD 470nm") +
+        guides(color = "none") +
+        facet_wrap(vars(comparison)) +
+        theme_classic() +
+        coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
+        labs(caption = sprintf("Not shown: %s %s and %s %s",
+           d[!in.region & comparison == "y.sat", .N],
+           "points outside left panel",
+           d[!in.region & comparison == "y.ground.pred", .N],
+           "points outside right panel"))}
+
 #' Find an "error envelope" with an additive and multiplicative
 #' term in the fashion of Dark Target validation
 #' https://darktarget.gsfc.nasa.gov/validation

### 2023-08-23 code/targets.R

@@ -44,6 +44,8 @@ terra.rast.fmt = tar_format(
                     as.data.frame(r[[lyr]], na.rm = F)[[1]]),
                 names(r))))})
 
+tar_file = tarchetypes::tar_file
+
 pbapply::pboptions(type = "timer")
 
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
@@ -162,6 +164,15 @@ list(
         cv$mDT_wPred, pred.grid, region.shape, satellite.files, model.full)),
     tar_target(baltimore.map.data, get.baltimore.map.data(
         pred.grid, region.shape, satellite.files, model.full)),
+
+    # Manuscript graphics
+    tar_target(ipath,
+       {idir = file.path(workflow.dir, "img")
+        dir.create(idir, showWarnings = F)
+        \(name) file.path(idir, paste0(name, ".png"))}),
+    tar_file(agreement.plot.path, ggsave(
+        ipath("agreement_plot"), dpi = 300, width = 7, height = 4,
+        agreement.plot(cv$mDT_wPred))),
 
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2")

### 2023-08-23 writing/CONUS_AOD.qmd
Targetized more plots.

@@
 the.sat = tools::toTitleCase(Wf$satellite)
+IG = knitr::include_graphics
 ```
@@
 ```{r agreement_plot}
-knitr::include_graphics(targets::tar_read(agreement.plot.path))
+IG(tar_read(agreement.plot.path))
 ```
@@
 
-```{r shap_summary, dev='png', dpi=300}
-shap_long <- shap.prep(shap_contrib = cv$shap_score,
-                       X_train = cv$mDT_wPred[, names(cv$shap_score), with = F])
-# **SHAP summary plot**
-shap.plot.summary(shap_long)
+```{r shap_summary}
+IG(tar_read(shap.summary.plot.path))
 ```
 
-```{r shap_features, dev='png', dpi=300}
-g1 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') +
-    aes(x = lubridate::as_datetime(x_feature)) +
-    xlab("time.sat") +
-    ggtitle("(A) SHAP values of Column WV vs. Time trend")
-# g2 <- shap.plot.dependence(data_long = shap_long, x = 'time.sat', y = 'y.sat', color_feature = 'y.sat') +  ggtitle("(B) SHAP values of MAIAC AOD vs. Time trend")
-#
-# gridExtra::grid.arrange(g1, g2, ncol = 2)
-g1
+```{r shap_features}
+IG(tar_read(shap.features.plot.path))
 ```
@@
 ```{r pred.map.median.mse}
-pred.map(
-    tar_read(median.mse.map.data)$pred, tar_read(pred.grid),
-    bg.sf = get_conus(), color.scale.name = "AOD",
-    quantile.cap = .99)
+IG(tar_read(pred.map.median.mse.path))
 ```
@@
 ```{r pred.map.baltimore}
-pred.map(
-    tar_read(baltimore.map.data), tar_read(pred.grid),
-    bg.sf = get_conus(), color.scale.name = "AOD")
+IG(tar_read(pred.map.baltimore.path))
 ```

### 2023-08-23 code/targets.R

@@ -173,6 +173,29 @@ list(
     tar_file(agreement.plot.path, ggsave(
         ipath("agreement_plot"), dpi = 300, width = 7, height = 4,
         agreement.plot(cv$mDT_wPred))),
+    tar_target(shap_long, shap.prep(
+        shap_contrib = cv$shap_score,
+        X_train = cv$mDT_wPred[, mget(names(cv$shap_score))])),
+    tar_file(shap.summary.plot.path, ggsave(
+        ipath("shap_summary"), dpi = 300, width = 7, height = 7,
+        shap.plot.summary(shap_long))),
+    tar_file(shap.features.plot.path, ggsave(
+        ipath("shap_features"), dpi = 300, width = 7, height = 7,
+        shap.plot.dependence(data_long = tar_read(shap_long), x = 'time.sat', y = 'Column_WV', color_feature = 'Column_WV') +
+            aes(x = lubridate::as_datetime(x_feature)) +
+            xlab("time.sat") +
+            ggtitle("(A) SHAP values of Column WV vs. Time trend"))),
+    tar_file(pred.map.median.mse.path, ggsave(
+        ipath("pred_map_median_mse"), width = 7, height = 7,
+        pred.map(
+            median.mse.map.data$pred, pred.grid,
+            bg.sf = get_conus(), color.scale.name = "AOD",
+            quantile.cap = .99))),
+    tar_file(pred.map.baltimore.path, ggsave(
+        ipath("pred_map_baltimore"), width = 7, height = 7,
+        pred.map(
+            baltimore.map.data, pred.grid,
+            bg.sf = get_conus(), color.scale.name = "AOD"))),
 
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2")

### 2023-08-24 code/modeling.R
Don't include so many columns in `performance.j`.

@@ -230,7 +230,6 @@ get.cv.summary = function(d)
 
 performance.j = quote(
    {mse = function(x, y) mean((x - y)^2)
-    sdn = function(x) sqrt(mse(x, mean(x)))
     list(
         "Cases" = .N,
         "Sites" = length(unique(site)),
@@ -238,13 +237,8 @@ performance.j = quote(
         "RMSE, corrected" = sqrt(mse(y.ground, y.ground.pred)),
         "Proportion of raw MSE" = mse(y.ground, y.ground.pred) / mse(y.ground, y.sat),
         "Median, ground" = median(y.ground),
-        "SD, ground" = sdn(y.ground),
-        "SD, raw" = sdn(y.sat),
-        "SD, corrected" = sdn(y.ground.pred),
         "Bias, raw" = mean(y.sat - y.ground),
-        "Bias, corrected" = mean(y.ground.pred - y.ground),
-        "Correlation, raw" = cor(y.sat, y.ground),
-        "Correlation, corrected" = cor(y.ground.pred, y.ground))})
+        "Bias, corrected" = mean(y.ground.pred - y.ground))})
 
 ## ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
 ## * New predictions

### 2023-08-24 writing/CONUS_AOD.qmd
Support Word output with `flextable`.

@@
     embed-resources: true
+  docx: {}
 
 echo: FALSE
-df-print: kable
 ---
@@
     library(ggridges)
-    library(kableExtra)
+    library(flextable)
   })
@@
 ```{r describe_dv}
-kable(summary(dt[, list(y.diff)]))
+summary(dt[, list(y.diff)])
 ```
@@
 
-```{r table_cv_results}
-kbl(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]), align = "l", caption = "Table ##. Cross validation results for the entire training set") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:12, width = "1.1cm")
+```{r}
+flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))
 ```
@@
 
-```{r strata_noaa_regions, fig.pos="H"}
+```{r}
 # performance by NOAA climate regions
@@
     [order(`Proportion of raw MSE`)])
-kbl(pretty.table.numbers(by.region),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+flextable(pretty.table.numbers(by.region))
 ```
@@
 
-```{r strata_month}
+```{r}
 result <- dt[,
@@
 result[, Month := month.abb[Month]]
-kbl(pretty.table.numbers(result), align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+flextable(pretty.table.numbers(result))
 ```
@@
 
-```{r strata_high_aod, fig.pos="H"}
+```{r}
 by.highaod = dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]
-kbl(pretty.table.numbers(by.highaod),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+flextable(pretty.table.numbers(by.highaod))
 ```
@@
 
-```{r strata_qa}
+```{r}
 by.qabest = dt[, eval(performance.j), by = qa_best]
-kbl(pretty.table.numbers(by.qabest),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+flextable(pretty.table.numbers(by.qabest))
 ```
@@
 
-```{r shap_rank}
+```{r}
 # what do the SHAPs tell us?
@@
 df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
-kbl(df_shap, col.names = "Mean |SHAP|") %>% row_spec(0, bold=TRUE)
+flextable(df_shap)
 # df_shap$variable <- row.names(df_shap)
@@
 
-```{r strata_year}
-kbl(pretty.table.numbers(tar_read(cv.summary)[-1]),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
+```{r}
+flextable(pretty.table.numbers(tar_read(cv.summary)[-1]))
 ```
@@
 
-```{r strata_overpasses, fig.pos="H"}
+```{r}
 dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]
 by.overpasses = dt[, eval(performance.j), keyby = overpasses]
-kbl(pretty.table.numbers(by.overpasses),
-    align = "l") %>%
-  kable_styling(position = "left",
-                latex_options=c("HOLD_position","scale_down")) %>%
-  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
+flextable(pretty.table.numbers(by.overpasses))
 ```

### 2023-08-24 code/targets.R

@@ -199,5 +199,5 @@ list(
 
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2")
-        tarchetypes::tar_quarto(paper_conus_html,
+        tarchetypes::tar_quarto(paper_conus,
             'writing/CONUS_AOD.qmd', quiet = F))

### 2023-08-25 writing/CONUS_AOD.qmd
Move the counts of hidden points into a caption.

@@
 ```{r agreement_plot}
+#| fig-cap: !expr caption
+caption = with(list(d = agreement.plot.data(dt)), sprintf(
+    "Not shown: %s %s and %s %s",
+    d[!in.region & comparison == "y.sat", .N],
+    "points outside left panel",
+    d[!in.region & comparison == "y.ground.pred", .N],
+    "points outside right panel"))
 IG(tar_read(agreement.plot.path))

### 2023-08-25 code/paper_functions.R

@@ -1,10 +1,14 @@
 # contains functions used in generating the CONUS_AOD paper
 
-agreement.plot = \(d, base_size = 11)
+agreement.plot.data = \(d)
    {d = melt(d[, .(y.ground, y.ground.pred, y.sat)],
         id = "y.ground", variable.name = "comparison", value.name = "y.pred")
     d[, comparison := factor(comparison, levels = rev(levels(comparison)))]
     d[, in.region := between(y.ground, 0, 1) & between(y.pred, 0, 1)]
+    d}
+
+agreement.plot = \(d, base_size = 11)
+   {d = agreement.plot.data(d)
     sample.obs = `[`(
         CJ(
             comparison = unique(d$comparison),
@@ -37,12 +41,7 @@ agreement.plot = \(d, base_size = 11)
         theme(panel.spacing = unit(1.5, "lines"),
           plot.margin = margin(half_line,
             base_size, half_line, half_line)) +
-        coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F) +
-        labs(caption = sprintf("Not shown: %s %s and %s %s",
-           d[!in.region & comparison == "y.sat", .N],
-           "points outside left panel",
-           d[!in.region & comparison == "y.ground.pred", .N],
-           "points outside right panel"))}
+        coord_equal(xlim = c(0, 1), ylim = c(0, 1), expand = F)}
 
 #' Find an "error envelope" with an additive and multiplicative
 #' term in the fashion of Dark Target validation

### 2023-08-25 writing/CONUS_AOD.qmd
Correct reporting of time difference.

@@
 After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS.
-The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["25%"]` s, and `r tqs["75%"]` s.
+The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.
 While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.

### 2023-08-26 writing/CONUS_AOD.qmd
Caption DV density figure, plot descriptives, and caption overall CV table.

@@
    {library(patchwork)
-    library(ggridges)
+    library(gridExtra)
     library(flextable)
@@
 
-<!--# any missingness in predictors to report?-->
-
 After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS.
 The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.
-While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian.
+While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian (see @fig-dv-density).
 
-```{r describe_dv}
-summary(dt[, list(y.diff)])
+```{r plot_dv}
+#| label: fig-dv-density
+#| fig-cap: !expr caption
+caption = with(list(d = dt), sprintf(
+    "Histogram of the difference in collocated AOD (n=%s). Not shown: %s %s",
+    scales::comma(d[, .N]),
+    d[y.diff < -1 | y.diff > 1, .N],
+    "AOD differences outside [-1, 1]"))
+p1 <- ggplot(dt, aes(x = y.diff)) +
+  geom_histogram(binwidth = 0.005) +
+  coord_cartesian(xlim = c(-1, 1), expand = FALSE) +
+  xlab("MAIAC AOD minus AERONET AOD") +
+  theme_classic()
+t1 <- summary(dt[, list(y.diff)], digits = 2)
+colnames(t1) <- ""
+t1 <- tableGrob(t1, theme=ttheme_minimal(), rows=NULL)  # transform into a tableGrob
+p1 + inset_element(t1, left = 0.02, bottom = 0.6, right = 0.25, top = .98)
 ```
 
-```{r plot_dv}
-ggplot(dt, aes(x = y.diff)) +
-  geom_histogram(binwidth = 0.005) +
-  scale_x_continuous("MAIAC AOD minus AERONET AOD") +
-  coord_cartesian(xlim = c(-1, 1)) +
-  labs(title = paste0("Histogram of difference in collocated AOD (n=", scales::comma(dt[, .N]), ")"),
-      caption = paste0("Not shown: ", dt[y.diff < -1 | y.diff > 1, .N], " AOD differences outside [-1, 1]")) +
-  theme_bw()
-```
-
-Table \## includes the performance metrics for the full cross-validated training data set.
-Here **raw** is MCD19A2.V006 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
+@tbl-cv-overall includes the performance metrics for the full cross-validated training data set.
+Here **raw** is MCD19A2.V061 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
 
 ```{r}
+#| label: tbl-cv-overall
+#| tbl-cap: Overall performance from cross-validation.
 flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))

### 2023-08-26 writing/CONUS_AOD.qmd
Label more figures and tweak RMSE map.

@@
 @tbl-cv-overall includes the performance metrics for the full cross-validated training data set.
-Here **raw** is MCD19A2.V061 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
+Here **raw** is Collection 6.1 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
 
@@
 Here we show our performance metrics in different subsets of the cross-validation results. These are all from the same model training, just slicing up how the performance looks in different contexts.
-
-### Performance by NOAA climate regions
 
@@
 ```{r}
-# performance by NOAA climate regions
+#| label: tbl-cv-noaa-regions
+#| tbl-cap: Performance by NOAA climate regions. Sorted by decreasing relative improvement in MSE after correction.
 by.region = (dt
@@
 ```{r noaa_region_plot, dev='png', dpi=300}
-## map of performance by NOAA regions
+#| label: fig-cv-noaa-regions
+#| fig-cap: Map of performance (RMSE) by NOAA regions.
 conus$state.noaa.region <- factor(
@@
                    id.vars = c("geometry", "state.noaa.region"))
-levels(conus_melt$variable) <- c("MAIAC AOD", "corrected AOD")
+levels(conus_melt$variable) <- c("MAIAC AOD", "Corrected AOD")
 map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) +
@@
   coord_sf(default_crs = st_crs(conus)) +
-  scale_fill_continuous("Regional\nRMSE") +
+  scale_fill_continuous("Regional\nRMSE", trans = 'reverse') +
   theme_minimal() +
   theme(axis.text = element_blank(),
-        panel.grid.major = element_blank())
+        panel.grid.major = element_blank(),
+        strip.text = element_text(size = 13))
 map_region_rmse
 ```
-
-### Performance by month of year
+@fig-cv-noaa-regions shows that performance improves in all regions and leads to greater consistency in the RMSE across CONUS with slightly higher RMSE remaining in the Northwest and West regions.
 
 ```{r}
+#| label: tbl-cv-month
+#| tbl-cap: Performance by month of year.
 result <- dt[,
@@
 ```{r agreement_plot}
+#| label: fig-agreement-plot
 #| fig-cap: !expr caption

### 2023-08-26 writing/CONUS_AOD.qmd
Minor text updates and add captions.

@@
 
-We demonstrate a reproducible workflow to reduce retrieval error from satellite-based datasets with an application to MODIS Aerosol Optical Depth over the conterminous United States 2000-2021.
+We demonstrate a reproducible workflow to reduce retrieval error from satellite-based datasets with an application to MODIS Aerosol Optical Depth over the conterminous United States 2000-2022.
 It facilitates the often laborious task of obtaining and comparing ground-based and remotely sensed quantities for product validation and evaluating retrieval biases.
@@
 The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`.
-The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file.
-While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra and report results for Aqua in Supplementary Data.
+The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6.1[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file.
+While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra.
 
@@
 We also report the Pearson correlation of the raw and corrected products versus the collocated AERONET measures.
-Finally, we report on the distribution of differences between the acquisition time for each overpass-tile and the nearest AERONET observation in time.
 We plotted the agreement of raw and corrected AOD versus AERONET AOD in scatterplots for the interval [0, 1] given that almost all AERONET observations in the cross-validation set fell within that range.
-To generate empirical error envelopes with performance for both low and high AOD conditions, we checked the combination of additive and multiplicative terms closest to containing \~2/3 of observations for AERONET AOD below and above 0.6.
+To generate empirical error envelopes with performance for both low and high AOD conditions, we optimized the combination of additive and multiplicative terms closest to containing \~2/3 of observations for AERONET AOD below and above 0.6, a threshold previously used when evaluating MAIAC AOD.
 
 While the model was trained using all of the collocated observations generated from within the study domain defined above, we present a number of stratified tables with the same metrics listed above to examine performance of the correction procedure in various data subsets.
-Stratification variables included time/space (by year, months of the year, and NOAA climate regions), atmospheric conditions (low versus high AOD), closeness of the temporal matchup (above versus below median time difference), and two additional factors related to MAIAC retrieval (number of overpasses per day and MAIAC quality flag).
+Stratification variables included time/space (NOAA climate regions, months of the year, and by year), atmospheric conditions (low versus high AOD), and two additional factors related to MAIAC retrieval (MAIAC quality flag and number of overpasses per day).
 
@@
 
-### Performance for higher versus lower AOD (AERONET \> 0.6)
-
 ```{r}
+#| label: tbl-cv-high-aod
+#| tbl-cap: Performance for higher versus lower AOD (AERONET \> 0.6).
 by.highaod = dt[, eval(performance.j), by = .(aod_high = y.ground > 0.6)]
@@
 
-### Performance by `qa_best`
-
 ```{r}
+#| label: tbl-cv-qa-best
+#| tbl-cap: Performance by quality flag `qa_best`.
 by.qabest = dt[, eval(performance.j), by = qa_best]
@@
 caption = with(list(d = agreement.plot.data(dt)), sprintf(
-    "Not shown: %s %s and %s %s",
+    "Agreement plots for satellite-based and AERONET AOD. Straight lines show empirical error envelopes containing 2/3 of retrievals. Not shown: %s %s and %s %s",
     d[!in.region & comparison == "y.sat", .N],
@@
 
-## Comparison with AQS
+## Comparison with AQS PM~2.5~
 
-We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$nobs)` observations of AQS on `r scales::comma(l$celldays)` cell-days.
+We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$nobs)` observations of AQS PM~2.5~ on `r scales::comma(l$celldays)` cell-days.
 The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
@@
 We leave for a future analysis the applicability of a trained `Earth_obs_cleaning` model to other space-time regions (with similar or different aerosol properties).
-Although here we present the `Earth_obs_cleaning` trained on almost 100,000 collocated observations, we use machine-learning methods likely to perform well with substantially smaller training datasets such as would be available for more recently launched sensors or over regions with more sparse ground network coverage.
+Although here we present the `Earth_obs_cleaning` trained on over 100,000 collocated observations, we use machine-learning methods likely to perform well with substantially smaller training datasets such as would be available for more recently launched sensors or over regions with more sparse ground network coverage.
 

### 2023-08-26 writing/CONUS_AOD.qmd
Update citation style.
Change AERONET acknowledgment to snippet with count.

@@
 bibliography: references.json
-csl: https://www.zotero.org/styles/atmospheric-measurement-techniques
+csl: https://raw.githubusercontent.com/citation-style-language/styles/795ad0c77258cb7e01f3413123b5b556b4cb6a98/dependent/journal-of-geophysical-research-atmospheres.csl
 
@@
 
-We thank the AERONET federation principal investigators and their staff for establishing and maintaining the 328 sun photometer sites used in this investigation.
+We thank the AERONET federation principal investigators and their staff for establishing and maintaining the `r uniqueN(dt$site)` sun photometer sites used in this investigation.
 

### 2023-08-26 writing/CONUS_AOD.qmd
Remove SHAP table, add captions, remove hyphens in distances.

@@
 
-The study region was the conterminous United States (CONUS) and all match-ups of MODIS data were based on the 1-km grid cells in the Global sinusoidal Coordinate Reference System.
+The study region was the conterminous United States (CONUS) and all match-ups of MODIS data were based on the 1 km grid cells in the Global sinusoidal Coordinate Reference System.
 The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`.
@@
 
-MAIAC AOD retrievals from all overpasses in the overlying 1-km grid cell were matched to the closest AERONET observation in time (no more than 8 minutes).
+MAIAC AOD retrievals from all overpasses in the overlying 1 km grid cell were matched to the closest AERONET observation in time (no more than 8 minutes).
 Because the purpose of the `Earth_obs_cleaning` workflow is to reduce retrieval error, there was no restriction based on quality codes of input MAIAC data.
@@
 
--   the mean AOD calculated from a square window with an 11-km side-length
+-   the mean AOD calculated from a square window with an 11 km side-length
 
--   the proportion missing AOD from a square window with an-11 km side-length
+-   the proportion missing AOD from a square window with an 11 km side-length
 
@@
 After collocation there were `r scales::comma(nrow(dt))` matched AOD retrievals from `r uniqueN(dt$site)` AERONET sites in CONUS.
-The quartiles of the absolute difference from the labeled overpass time (one per 1200-km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.
+The quartiles of the absolute difference from the labeled overpass time (one per 1200 km wide tile) and the time of the matched AERONET observation were `r tqs = round(quantile(dt[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.
 While the distribution of AOD is skewed, our estimate of retrieval error and the dependent variable for our predictive model, *diff~AOD~*, was fairly symmetric and Gaussian (see @fig-dv-density).
@@
 
-In decreasing order of variable importance (from `r the.sat` collocated cross-validation predictions)
-
-```{r}
-# what do the SHAPs tell us?
-tar_load(cv)
-shaps <- cv$"shap_score"
-# sort by importance
-df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
-flextable(df_shap)
-# df_shap$variable <- row.names(df_shap)
-# ft_shap <- flextable(df_shap[, c(2, 1)])
-# ft_shap <- set_header_labels(ft_shap, list(variable = "Variable", meanabsshap = "Mean Absolute SHAP"))
-# ft_shap
-```
-
-Summary visualization of SHAP contributions
-
 ```{r shap_summary}
+#| label: fig-shap-summary
+#| fig-cap: Summary of SHAP contributions. Predictors ordered by decending mean absolute SHAP (printed along left).
 IG(tar_read(shap.summary.plot.path))
@@
 ```{r shap_features}
+#| label: fig-shap-timeseries
+#| fig-cap: Time series of SHAPs for Column Water Vapor (CWV). SHAPs for observations with missing values for CWV are shown in grey.
 IG(tar_read(shap.features.plot.path))
@@
 
-### Median MSE day
+### Mapping a typical day
 
-This map shows a tile on `r tar_read(median.mse.map.data)$date`.
+@fig-map-median-mse shows a tile on `r tar_read(median.mse.map.data)$date`.
 
 ```{r pred.map.median.mse}
+#| label: fig-map-median-mse
+#| fig-cap: Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites.
 IG(tar_read(pred.map.median.mse.path))
@@
 
-### Baltimore
-
-A day with long-range transported smoke from Canadian wildfires.
+### Mapping a documented high AOD event
 
 ```{r pred.map.baltimore}
+#| label: fig-map-baltimore-smoke
+#| fig-cap: Maps of MAIAC and Corrected AOD over the Mid-Atlantic US during a documented exceedence event with downmixing of long-range transported smoke from Canadian wildfires.
 IG(tar_read(pred.map.baltimore.path))
@@
 
-## Performance by year
-
 ```{r}
+#| label: tbl-cv-by-year
+#| tbl-cap: Cross-validation performance by year.
 flextable(pretty.table.numbers(tar_read(cv.summary)[-1]))
@@
 
-## Performance by number of overpasses in collocation that site-day
-
 ```{r}
+#| label: tbl-cv-overpasses
+#| tbl-cap: Cross-validation performance by the number of overpasses in collocation for that site-day.
 dt[, overpasses := .N, by = .(site, lubridate::as_date(time.ground))]

### 2023-08-26 writing/CONUS_AOD.qmd
Minor typos.

@@
 The study period included all retrievals from `r min(Wf$dates)` through `r max(Wf$dates)`.
-The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6.1[@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file.
+The `Earth_obs_cleaning` workflow automatically downloaded all intersecting tiles of MCD19A2 Collection 6.1 [@lyapustinalexei2018] based on a NASA-supplied MODIS tile boundary file.
 While the MCD19A2 product includes data from both Aqua and Terra MODIS instruments, we focus our analyses on the longer running Terra.
@@
 To demonstrate that corrected AOD still shows spatial patterns consistent with expectations we generated maps of raw and corrected MAIAC on a typical day (selected as the tile with the most non-missing AOD on the day with the median improvement in MSE for AERONET stations).
-To show the correction in known high-AOD conditions we plotted the Baltimore, MA area on June 10, 2015 which had substantial subsiding long-range transported Canadian wildfire smoke as previously documented with extensive surface and atmospheric monitoring [@dreessen2016].
+To show the correction in known high-AOD conditions we plotted the Baltimore, MD area on June 10, 2015 which had substantial subsiding long-range transported Canadian wildfire smoke as previously documented with extensive surface and atmospheric monitoring [@dreessen2016].
 

### 2023-08-28 writing/CONUS_AOD.qmd
Show the name of the center state in `pred.map.median.mse`.

@@
 #| label: fig-map-median-mse
-#| fig-cap: Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites.
+#| fig-cap: !expr caption
+caption = sprintf("Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites. The map is centered on %s.",
+    tar_read(median.mse.map)$center.name)
 IG(tar_read(pred.map.median.mse.path))

### 2023-08-28 code/paper_functions.R

@@ -212,14 +212,22 @@ pred.map = function(
     if (!is.null(quantile.cap))
         d[, value := pmin(quantile(value, quantile.cap, na.rm = T), value)]
 
-    ggplot() +
-        geom_raster(aes(lon, lat, fill = value), data = d) +
-        scale_fill_distiller(name = color.scale.name,
-            palette = "Spectral", na.value = "transparent") +
-        geom_sf(data = bg.sf, fill = NA, size = .1) +
-        facet_grid(rows = "variable", labeller = labeller(variable =
-            c(y.sat.old = "MAIAC AOD", y.sat.new = "Corrected AOD"))) +
-        coord_sf(expand = F,
-            xlim = range(d$lon), ylim = range(d$lat)) +
-        theme_void() +
-        theme(strip.text = element_text(size = 13))}
+    list(
+        center.name = bg.sf[as.integer(st_intersects(
+            st_as_sf(
+                d[, .(
+                    mean(c(max(lon), min(lon))),
+                    mean(c(max(lat), min(lat))))],
+                coords = c(1, 2), crs = crs.lonlat),
+            st_transform(bg.sf, crs = crs.lonlat))),]$NAME,
+        plot = ggplot() +
+            geom_raster(aes(lon, lat, fill = value), data = d) +
+            scale_fill_distiller(name = color.scale.name,
+                palette = "Spectral", na.value = "transparent") +
+            geom_sf(data = bg.sf, fill = NA, size = .1) +
+            facet_grid(rows = "variable", labeller = labeller(variable =
+                c(y.sat.old = "MAIAC AOD", y.sat.new = "Corrected AOD"))) +
+            coord_sf(expand = F,
+                xlim = range(d$lon), ylim = range(d$lat)) +
+            theme_void() +
+            theme(strip.text = element_text(size = 13)))}

### 2023-08-28 code/targets.R

@@ -164,6 +164,13 @@ list(
         cv$mDT_wPred, pred.grid, region.shape, satellite.files, model.full)),
     tar_target(baltimore.map.data, get.baltimore.map.data(
         pred.grid, region.shape, satellite.files, model.full)),
+    tar_target(median.mse.map, pred.map(
+        median.mse.map.data$pred, pred.grid,
+        bg.sf = get_conus(), color.scale.name = "AOD",
+        quantile.cap = .99)),
+    tar_target(baltimore.map, pred.map(
+        baltimore.map.data, pred.grid,
+        bg.sf = get_conus(), color.scale.name = "AOD")),
 
     # Manuscript graphics
     tar_target(ipath,
@@ -188,15 +195,10 @@ list(
                   panel.grid.minor.x = element_blank()))),
     tar_file(pred.map.median.mse.path, ggsave(
         ipath("pred_map_median_mse"), width = 7, height = 7,
-        pred.map(
-            median.mse.map.data$pred, pred.grid,
-            bg.sf = get_conus(), color.scale.name = "AOD",
-            quantile.cap = .99))),
+        median.mse.map$plot)),
     tar_file(pred.map.baltimore.path, ggsave(
         ipath("pred_map_baltimore"), width = 7, height = 7,
-        pred.map(
-            baltimore.map.data, pred.grid,
-            bg.sf = get_conus(), color.scale.name = "AOD"))),
+        baltimore.map$plot)),
 
     # Render the CONUS AOD manuscript
     if (Wf$satellite.product == "mcd19a2")

### 2023-08-28 writing/CONUS_AOD.qmd
Added a table of IVs.

@@
 p1 + inset_element(t1, left = 0.02, bottom = 0.6, right = 0.25, top = .98)
+```
+
+@tbl-ivs shows the distribution of each predictor in the CV.
+
+```{r tbl-ivs}
+#| tbl-cap: !expr 'sprintf("`time.sat` is represented in days since %s.", as.character(min(dt$time.sat), usetz = T))'
+flextable(rbindlist(lapply(Wf$features, \(iv)
+   {nv = dt[[iv]]
+    v = na.omit(nv[!is.na(nv)])
+    r = \(x) round(x, 3)
+    if (iv == "time.sat")
+        v = as.numeric(difftime(v, min(v), units = "days"))
+    data.table(
+        Variable = iv,
+        missing = r(mean(is.na(nv))),
+        min = r(min(v)),
+        max = r(max(v)),
+        mean = r(mean(v)),
+        sd = r(sd(v)))})))
 ```

### 2023-08-29 writing/CONUS_AOD.qmd
Moved the IVs table to the paper supplement.
Added some inline code to replace deleted table columns.

@@
 
-@tbl-ivs shows the distribution of each predictor in the CV.
-
-```{r tbl-ivs}
-#| tbl-cap: !expr 'sprintf("`time.sat` is represented in days since %s.", as.character(min(dt$time.sat), usetz = T))'
-flextable(rbindlist(lapply(Wf$features, \(iv)
-   {nv = dt[[iv]]
-    v = na.omit(nv[!is.na(nv)])
-    r = \(x) round(x, 3)
-    if (iv == "time.sat")
-        v = as.numeric(difftime(v, min(v), units = "days"))
-    data.table(
-        Variable = iv,
-        missing = r(mean(is.na(nv))),
-        min = r(min(v)),
-        max = r(max(v)),
-        mean = r(mean(v)),
-        sd = r(sd(v)))})))
-```
-
 @tbl-cv-overall includes the performance metrics for the full cross-validated training data set.
-Here **raw** is Collection 6.1 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.
+`r sdn = \(x) sqrt(mean((x - mean(x))^2)); f = \(x) sprintf("(SD %.03f)", sdn(dt[[x]])); "";` Here **raw** is Collection 6.1 MAIAC AOD 470nm `r f("y.sat")`; **ground** is AERONET AOD interpolated to 470nm `r f("y.ground")`; and **corrected** is MAIAC after applying an XGBoost-modeled correction term `r f("y.ground.pred")`.
+The ground values are correlated `r dt[, sprintf("%+.03f", cor(y.ground, y.sat))]` with the raw values and `r dt[, sprintf("%+.03f", cor(y.ground, y.ground.pred))]` with the corrected values.
 
@@
 
+@tbl-ivs shows the distribution of each predictor in the CV.
+
+```{r tbl-ivs}
+#| tbl-cap: !expr 'sprintf("`time.sat` is represented in days since %s.", as.character(min(dt$time.sat), usetz = T))'
+flextable(rbindlist(lapply(Wf$features, \(iv)
+   {nv = dt[[iv]]
+    v = na.omit(nv[!is.na(nv)])
+    r = \(x) round(x, 3)
+    if (iv == "time.sat")
+        v = as.numeric(difftime(v, min(v), units = "days"))
+    data.table(
+        Variable = iv,
+        missing = r(mean(is.na(nv))),
+        min = r(min(v)),
+        max = r(max(v)),
+        mean = r(mean(v)),
+        sd = r(sd(v)))})))
+```
+
 ```{r}

### 2024-01-29 writing/paper_mcd19a2.qmd
Renamed a paper.

--- 2023-08-29 writing/CONUS_AOD.qmd
@@
 ```
-
-```{r git_check}
-{
-s = function(cmd) suppressWarnings(system(cmd,
-    intern = TRUE, ignore.stderr = TRUE))
-commit = substr(s('git rev-parse HEAD'), 1, 8)
-uncommitted = ''
-if (!length(commit))
-   commit = 'unknown'
-else if(length(s('git diff-index HEAD --'))>0)
-   uncommitted = 'with uncommitted changes'
-}
-```
-
-Git commit: `r commit` `r uncommitted`
 

### 2024-01-29 code/targets.R

@@ -212,12 +212,11 @@ list(
         ipath("pred_map_baltimore"), width = 7, height = 7,
         baltimore.map$plot)),
 
-    # Render the CONUS AOD manuscript
-    if (Wf$satellite.product == "mcd19a2") list(
-        tarchetypes::tar_quarto(paper.conus.render,
-            'writing/CONUS_AOD.qmd', quiet = F),
-        tar_target(paper.conus,
-            file.copy(paper.conus.render[1], writing.out.dir))),
+    # Render the manuscript.
+    tarchetypes::tar_quarto(paper.render,
+        sprintf('writing/paper_%s.qmd', Wf$satellite.product), quiet = F),
+    tar_target(paper,
+        file.copy(paper.render[1], writing.out.dir)),
 
     # Do a supplementary analysis for AODC.
     if (Wf$satellite.product == "aodc") list(

### 2024-01-30 writing/paper_aodc.qmd
Added a paper for AODC.

---
title: aodc

bibliography: references.json
csl: https://raw.githubusercontent.com/citation-style-language/styles/795ad0c77258cb7e01f3413123b5b556b4cb6a98/dependent/journal-of-geophysical-research-atmospheres.csl

format:
  html: {embed-resources: true}

echo: FALSE
---

```{r}
source.files = "globals libraries util data modeling paper_functions"
for (filename in strsplit(source.files, " ")[[1]])
    source(sprintf("code/%s.R", filename))
suppressPackageStartupMessages(
    library(flextable))
```

# Results

```{r}
d = copy(tar_read(cv)$mDT_wPred)
```

After collocation there were `r scales::comma(nrow(d))` matched AOD retrievals from `r uniqueN(d$site)` AERONET sites in CONUS. The quartiles of the absolute difference between the satellite time and the time of the matched AERONET observation were `r tqs = round(quantile(d[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.

@tbl-cv-overall includes the performance metrics for the full cross-validated training data set. `r sdn = \(x) sqrt(mean((x - mean(x))^2)); f = \(x) sprintf("(SD %.03f)", sdn(d[[x]])); "";` Here **raw** is … `r f("y.sat")`; **ground** is … `r f("y.ground")`; and **corrected** is … `r f("y.ground.pred")`. The ground values are correlated `r d[, sprintf("%+.03f", cor(y.ground, y.sat))]` with the raw values and `r d[, sprintf("%+.03f", cor(y.ground, y.ground.pred))]` with the corrected values.

```{r}
#| label: tbl-cv-overall
#| tbl-cap: Overall performance from cross-validation.
flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))
```

```{r}
#| label: tbl-cv-month
#| tbl-cap: Performance by month of year.
out = d[,
    keyby = .(Month = month(time.sat)),
    eval(performance.j)]
out[, Month := month.abb[Month]]
flextable(pretty.table.numbers(out))
```

```{r}
#| label: tbl-cv-hour
#| tbl-cap: Performance by hour of the day.
flextable(pretty.table.numbers(d[,
    keyby = .(Hour = hour(lubridate::with_tz(time.sat, "Etc/GMT+6"))),
    eval(performance.j)]))
```

```{r}
#| label: tbl-cv-high-aod
#| tbl-cap: Performance for higher versus lower AOD.
threshold = .6
flextable(pretty.table.numbers(d[,
    by = .("Ground value" = ifelse(y.ground > threshold,
        sprintf("> %.1f", threshold),
        sprintf("≤ %.1f", threshold))),
    eval(performance.j)]))
```

```{r}
#| label: tbl-cv-quality
#| tbl-cap: Performance by value of the quality variable `DQF`.
flextable(pretty.table.numbers(d[,
    keyby = .(DQF = ifelse(is.na(DQF), "missing", DQF)),
    eval(performance.j)]))
```

### 2024-01-30 code/data.R
Use an approximate solar time for `seconds.since.midnight`.

@@ -426,12 +426,17 @@ get.traindata = function(
     d = d[!is.na(y.ground)]
 
     if ("seconds.since.midnight" %in% features)
-        d[, seconds.since.midnight := as.integer(difftime(
+        # Calculate seconds since midnight in an approximation of
+        # local solar time.
+       {s.utc = d[, as.integer(difftime(
             time.sat,
-            lubridate::floor_date(
-                lubridate::with_tz(time.sat, "Etc/GMT+6"),
-                "day"),
+            lubridate::floor_date(time.sat, "day"),
             units = "secs"))]
+        one.day = 24L * 60L * 60L
+        lon = aer_stn[.(d$site), lon]
+        d[, seconds.since.midnight :=
+            (s.utc + as.integer(round((lon / 360) * one.day))) %%
+                one.day]}
 
     # Calculate the difference.
     d[, y.diff := y.sat - y.ground]

### 2024-02-05 writing/paper_aodc.qmd
Use an approximate solar time for `seconds.since.midnight`.
Added AODC error envelopes.

@@
     library(flextable))
+IG = knitr::include_graphics
 ```
@@
 flextable(pretty.table.numbers(d[,
-    keyby = .(Hour = hour(lubridate::with_tz(time.sat, "Etc/GMT+6"))),
+    keyby = .(Hour = seconds.since.midnight %/% (60L * 60L)),
     eval(performance.j)]))
@@
 ```
+
+## Error envelopes
+
+We use … The results are an envelope with `r eee = \(pred) {o = empirical.error.envelope(obs = d$y.ground, pred); stopifnot(o$convergence == 0); o$par}; show.eee = \(pred) {ps = eee(pred); sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; show.eee(d$y.sat)` for the raw satellite values, and an envelope with `r show.eee(d$y.ground.pred)` for our predictions.
+The mean widths of the envelopes across all observations are `r me = \(pred) {ps = eee(pred); sprintf("%.03f", 2*ps[1] + 2*ps[2]*mean(d$y.ground))}; me(d$y.sat)` for the raw values and `r me(d$y.ground.pred)` for our predictions.
+
+```{r agreement_plot}
+#| label: fig-agreement-plot
+#| fig-cap: !expr caption
+caption = with(list(d = agreement.plot.data(d)), sprintf(
+    "Agreement plots for satellite-based and AERONET AOD. Straight lines show empirical error envelopes containing 2/3 of retrievals. Not shown: %s %s and %s %s",
+    d[!in.region & comparison == "y.sat", .N],
+    "points outside left panel",
+    d[!in.region & comparison == "y.ground.pred", .N],
+    "points outside right panel"))
+IG(tar_read(agreement.plot.path))
+```

### 2024-02-07 writing/paper_mcd19a2.qmd
Map the day with median improvement, not median MSE.

@@
 
-@fig-map-median-mse shows a tile on `r tar_read(median.mse.map.data)$date`.
+@fig-map-improve-mse shows a tile on `r tar_read(improve.mse.map.data)$date`.
 
-```{r pred.map.median.mse}
-#| label: fig-map-median-mse
+```{r pred.map.median.improve}
+#| label: fig-map-median-improve
 #| fig-cap: !expr caption
 caption = sprintf("Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites. The map is centered on %s.",
-    tar_read(median.mse.map)$center.name)
-IG(tar_read(pred.map.median.mse.path))
+    tar_read(median.improve.map)$center.name)
+IG(tar_read(pred.map.median.improve.path))
 ```

### 2024-02-07 code/paper_functions.R

@@ -120,14 +120,16 @@ get.satellite.vs.aqs = function(satellite, aqs)
     out
     }
 
-get.median.mse.map.data = function(
+get.median.improve.map.data = function(
       y.sat.name, the.satellite, satellite.product, n.workers,
       d, pred.grid, region.shape, satellite.files, model.full)
-   {# Get the date with the median MSE.
+   {# Get the date with the median improvement in MSE.
     date = (d
-        [, by = .(date = lubridate::as_date(time.sat, tz = "UTC")),
-            .(mse = mean((y.ground - y.ground.pred)^2))]
-        [which.min(mse - median(mse)), date])
+        [, by = .(date = lubridate::as_date(time.sat, tz = tz)),
+            .(improve =
+                mean((y.ground - y.ground.pred)^2) /
+                mean((y.ground - y.sat)^2))]
+        [which.min(improve - median(improve)), date])
 
     # Get overpasses for this data.
     sat = satellite.files[time == date]

### 2024-02-07 code/targets.R

@@ -171,13 +171,13 @@ list(
         pred.at.aqs.sites, aqs.obs)),
 
     # Data for maps
-    tar_target(median.mse.map.data, get.median.mse.map.data(
+    tar_target(median.improve.map.data, get.median.improve.map.data(
         Wf$y.sat, Wf$satellite, Wf$satellite.product, config$n.workers,
         cv$mDT_wPred, pred.grid, region.shape, satellite.files, model.full)),
     tar_target(baltimore.map.data, get.baltimore.map.data(
         pred.grid, region.shape, satellite.files, model.full)),
-    tar_target(median.mse.map, pred.map(
-        median.mse.map.data$pred, pred.grid,
+    tar_target(median.improve.map, pred.map(
+        median.improve.map.data$pred, pred.grid,
         bg.sf = get_conus(), color.scale.name = "AOD",
         quantile.cap = .99)),
     tar_target(baltimore.map, pred.map(
@@ -205,9 +205,9 @@ list(
             ylab("SHAP value for MAIAC CWV") +
             theme(axis.title.x = element_blank(),
                   panel.grid.minor.x = element_blank()))),
-    tar_file(pred.map.median.mse.path, ggsave(
-        ipath("pred_map_median_mse"), width = 7, height = 7,
-        median.mse.map$plot)),
+    tar_file(pred.map.median.improve.path, ggsave(
+        ipath("pred_map_median_improve"), width = 7, height = 7,
+        median.improve.map$plot)),
     tar_file(pred.map.baltimore.path, ggsave(
         ipath("pred_map_baltimore"), width = 7, height = 7,
         baltimore.map$plot)),

### 2024-02-07 code/data.R
Added a median-improvement map for AODC.

@@ -114,7 +114,8 @@ expand.to.overpasses = function(d, satellite.files, the.satellite, satellite.pro
 
 get.predictors = function(
         d, satellite.files,
-        satellite.product, y.sat.name, features, window.radius, n.workers)
+        satellite.product, y.sat.name, features, window.radius,
+        crs.sat, n.workers)
    {vnames = intersect(
         str_replace(features, "qa_best", "AOD_QA"),
         feature.raster.layers)
@@ -132,7 +133,7 @@ get.predictors = function(
     message("Reading predictors from satellite data")
     d = rbindlist(xlapply(parallel.outside, chunks, \(chunk)
       tryCatch(
-        chunk[, c("y.sat", vnames, "y.sat.mean", "y.sat.present") :=
+        chunk[, c("sat.coords.x", "sat.coords.y", "y.sat", vnames, "y.sat.mean", "y.sat.present") :=
            {r = read_satellite_raster(
                 satellite.product,
                 satellite.files[chunk$sat.files.ix[1], tile],
@@ -140,6 +141,7 @@ get.predictors = function(
                 overpass[1])
             y.sat.values = drop(suppressWarnings(r[[y.sat.name]][]))
             cbind(
+                terra::xyFromCell(r, cell.local),
                 y.sat.values[cell.local],
                 r[[vnames]][cell.local],
                 rbindlist(xlapply(!parallel.outside, cell.local, \(cell)
@@ -184,6 +186,21 @@ get.predictors = function(
             levels = c("Background", "Smoke", "Dust"))]
         assert(d[, all(as.integer(qa_model) <= 3L)])
         d[, AOD_QA := NULL]}
+
+    if ("seconds.since.midnight" %in% features)
+      # Calculate seconds since midnight in an approximation of local
+      # solar time.
+       {s.utc = d[, as.integer(difftime(
+            time.sat,
+            lubridate::floor_date(time.sat, "day"),
+            units = "secs"))]
+        one.day = 24L * 60L * 60L
+        lon = convert.crs(
+            d[, .(sat.coords.x, sat.coords.y)],
+            crs.sat, crs.lonlat)$x
+        d[, seconds.since.midnight :=
+            (s.utc + as.integer(round((lon / 360) * one.day))) %%
+                one.day]}
 
     d}
 
@@ -416,7 +433,8 @@ get.traindata = function(
         # Read in the values of the appropriate satellite files.
         get.predictors(
             d, satellite.files,
[…]

### 2024-02-07 code/modeling.R

@@ -321,7 +321,7 @@ new.preds = function(dt.start, dt.end, cells = NULL, targets = NULL)
     message("Selecting satellite files")
     # Keep only enough `sat` files to cover the requested time range.
     sat = sat[
-        if (class(time) == "Date")
+        if ("Date" %in% class(time))
             lubridate::as_date(dt.start) - 1 <= time &
                 time <= lubridate::as_date(dt.end) + 1
         else
@@ -347,7 +347,8 @@ new.preds = function(dt.start, dt.end, cells = NULL, targets = NULL)
     d = get.predictors(d, sat,
         Wf$satellite.product, Wf$y.sat,
         Wf$features, Wf$window.radius,
-        config$n.workers)
+        terra::crs(grid), config$n.workers)
+
     message(sprintf("Result: %s rows", scales::comma(nrow(d))))
     if (!nrow(d))
         return(data.table())

### 2024-02-07 code/paper_functions.R

@@ -123,7 +123,9 @@ get.satellite.vs.aqs = function(satellite, aqs)
 get.median.improve.map.data = function(
       y.sat.name, the.satellite, satellite.product, n.workers,
       d, pred.grid, region.shape, satellite.files, model.full)
-   {# Get the date with the median improvement in MSE.
+   {tz = "Etc/GMT+6" # I.e., UTC-6, or Central Standard Time
+
+    # Get the date with the median improvement in MSE.
     date = (d
         [, by = .(date = lubridate::as_date(time.sat, tz = tz)),
             .(improve =
@@ -131,18 +133,20 @@ get.median.improve.map.data = function(
                 mean((y.ground - y.sat)^2))]
         [which.min(improve - median(improve)), date])
 
-    # Get overpasses for this data.
-    sat = satellite.files[time == date]
+    # Get overpasses for this date.
+    sat = satellite.files[lubridate::as_date(time, tz = tz) == date]
     sat[, sat.files.ix := .I]
-    overpasses = merge(by = "sat.files.ix", sat,
-        expand.to.overpasses(sat, sat,
-            the.satellite, satellite.product, n.workers))
+    overpasses = (if (multipass.sat(satellite.product))
+        merge(by = "sat.files.ix", sat,
+            expand.to.overpasses(sat, sat,
+                the.satellite, satellite.product, n.workers)) else
+        setnames(cbind(sat, overpass = NA), "time", "time.sat"))
 
     # Find the local cell indices we'd want to map.
     message("Getting raster cells")
     overpass.local.cells = pblapply(1 : nrow(overpasses), cl = n.workers, \(i)
        {r = with(overpasses[i], read_satellite_raster(
-            Wf$satellite.product, tile, path, overpass))[[y.sat.name]]
+            satellite.product, tile, path, overpass))[[y.sat.name]]
         # Get only cells with non-missing values.
         d = as.data.table(as.data.frame(r,
             na.rm = T, cell = T, xy = T))

### 2024-02-07 writing/paper_aodc.qmd
Added a median-improvement map for AODC.
Added another map for AODC.

@@
 ```
+
+## Mapping a typical day
+
+@fig-map-median-improve shows a tile on `r tar_read(median.improve.map.data)$date`.
+
+```{r pred.map.median.improve}
+#| label: fig-map-median-improve
+#| fig-cap: foofy
+IG(tar_read(pred.map.median.improve.path))
+```
+
+### Mapping a documented high-AOD event
+
+```{r pred.map.special.time}
+#| label: fig-map-special-time
+#| fig-cap: foofy2
+IG(tar_read(pred.map.special.time.path))
+```

### 2024-02-07 code/targets.R

@@ -176,12 +176,36 @@ list(
         cv$mDT_wPred, pred.grid, region.shape, satellite.files, model.full)),
     tar_target(baltimore.map.data, get.baltimore.map.data(
         pred.grid, region.shape, satellite.files, model.full)),
+    tar_target(special.time.map.data,
+       {time.min = lubridate::as_datetime("2023-06-07T12:00:00-0400")
+        time.max = lubridate::as_datetime("2023-06-07T18:00:00-0400")
+          # A time of intense smoke from Canadian wildfires in the US.
+        corners = convert.crs(
+            cbind(
+                c(-81, -66),  # To the SW of Pennsylvania
+                c( 39,  48)), # To the NE of CONUS
+            crs.lonlat,
+            terra::crs(pred.grid))
+        cells = as.data.frame(terra::crop(tar_read(pred.grid), with(corners,
+            terra::ext(x[1], x[2], y[1], y[2]))))$cell.local
+        sat = satellite.files[time.min <= time & time <= time.max]
+        ns = pbapply::pbsapply(cl = n.workers, sat$path, \(p)
+            sum(!is.na(`[`(
+                as.data.frame(terra::rast(p)$AOD, na.rm = F),
+                cells, "AOD"))))
+        dt = sat[which.max(ns), time]
+        new.preds(
+            dt, dt, cells = cells,
+            targets = list(pred.grid, satellite.files, model.full))}),
     tar_target(median.improve.map, pred.map(
         median.improve.map.data$pred, pred.grid,
         bg.sf = get_conus(), color.scale.name = "AOD",
         quantile.cap = .99)),
     tar_target(baltimore.map, pred.map(
         baltimore.map.data, pred.grid,
+        bg.sf = get_conus(), color.scale.name = "AOD")),
+    tar_target(special.time.map, pred.map(
+        special.time.map.data, pred.grid,
         bg.sf = get_conus(), color.scale.name = "AOD")),
 
     # Manuscript graphics
@@ -211,6 +235,9 @@ list(
     tar_file(pred.map.baltimore.path, ggsave(
         ipath("pred_map_baltimore"), width = 7, height = 7,
         baltimore.map$plot)),
+    tar_file(pred.map.special.time.path, ggsave(
+        ipath("pred_map_special_time"), width = 7, height = 7,
+        special.time.map$plot)),
 
     # Render the manuscript.
     tarchetypes::tar_quarto(paper.render,

### 2024-02-09 writing/paper_aodc.qmd
Improved the AODC map captions a bit.

--- 2024-02-05 writing/paper_aodc.qmd
@@
 ```
+
+## Mapping a typical day
+
+@fig-map-median-improve shows a tile on `r tar_read(median.improve.map.data)$date`.
+
+```{r pred.map.median.improve}
+#| label: fig-map-median-improve
+#| fig-cap: !expr caption
+caption = sprintf("From %s.",
+    tar_read(median.improve.map.data)$date)
+IG(tar_read(pred.map.median.improve.path))
+```
+
+### Mapping a documented high-AOD event
+
+```{r pred.map.special.time}
+#| label: fig-map-special-time
+#| fig-cap: !expr caption
+caption = sprintf("From %s.", format(
+    lubridate::with_tz(tar_read(special.time.map.data)[1, time.sat],
+        "America/New_York"),
+    "%Y-%m-%dT%H:%M:%S%z"))
+IG(tar_read(pred.map.special.time.path))
+```

### 2024-02-12 code/paper_functions.R
Got the MCD19A2 paper rendering again.

@@ -240,7 +240,7 @@ pred.map = function(
                 palette = "Spectral", na.value = "transparent") +
             geom_sf(data = bg.sf, fill = NA, size = .1) +
             ggspatial::annotation_scale(data = cbind(
-                data.frame(variable = "y.sat.old"),
+                data.table(variable = "y.sat.old"),
                 (if (include.quality) list(qualities = "all qualities")))) +
             facet_grid(
                 rows = vars(variable),

### 2024-02-21 writing/paper_aodc.qmd
Added a SHAP figure for AODC.
Extended the AQS comparsion to AODC.

@@
 
+## SHAP interpretation
+
+```{r shap_summary}
+#| label: fig-shap-summary
+#| fig-cap: Summary of SHAP contributions. Predictors ordered by decending mean absolute SHAP (printed along left).
+IG(tar_read(shap.summary.plot.path))
+```
+
 ## Mapping a typical day
@@
 ```
+
+## Comparison with AQS PM~2.5~
+
+We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$n.aqs.obs)` observations of AQS PM~2.5~ in `r scales::comma(l$n.cell.times)` cell-hours.
+The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.

### 2024-02-21 writing/paper_mcd19a2.qmd
Got the MCD19A2 paper rendering again.
Extended the AQS comparsion to AODC.

@@
 
-@fig-map-improve-mse shows a tile on `r tar_read(improve.mse.map.data)$date`.
+@fig-map-median-improve shows a tile on `r tar_read(median.improve.map.data)$date`.
 
@@
 
-We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$nobs)` observations of AQS PM~2.5~ on `r scales::comma(l$celldays)` cell-days.
+We examined `r l = tar_read(satellite.vs.aqs); scales::comma(l$n.aqs.obs)` observations of AQS PM~2.5~ on `r scales::comma(l$n.cell.times)` cell-days.
 The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.

### 2024-02-21 code/paper_functions.R

@@ -64,9 +64,10 @@ envelope.tester = \(obs, pred, threshold = 0.6)
 
 #' Get observations of PM_{2.5} from the Environmental Protection
 #' Agency's (EPA) Air Quality System. The unit is μg/m^3.
+#' `daily` is T for daily data or F for hourly data.
 #' File source: https://aqs.epa.gov/aqsweb/airdata/download_files.html#Daily
 #' Documentation: https://aqs.epa.gov/aqsweb/documents/about_aqs_data.html
-get.aqs.obs = function(years, grid)
+get.aqs.obs = function(daily, years, grid)
    {aqs.url.root = "https://aqs.epa.gov/aqsweb/airdata"
     parameter.code = 88101L
       # PM_{2.5} from a federally approved reference or equivalent
@@ -75,50 +76,71 @@ get.aqs.obs = function(years, grid)
     assert(file.exists("code/openssl_workaround.conf"))
     Sys.setenv(OPENSSL_CONF = "code/openssl_workaround.conf")
     d = rbindlist(lapply(years, function(the.year)
-       {fname = sprintf("daily_%d_%d.zip", parameter.code, the.year)
+       {message("AQS ", the.year)
+        fname = sprintf("%s_%d_%d.zip",
+            (if (daily) "daily" else "hourly"),
+            parameter.code,
+            the.year)
         d = fread(
             cmd = paste("unzip -p ", shQuote(download(
                 paste0(aqs.url.root, "/", fname),
                 file.path("aqs", fname)))),
-            select = c(
-                "Date Local", "Longitude", "Latitude",
-                "Event Type", "Sample Duration", "Arithmetic Mean"))
+            select = c("Longitude", "Latitude", (
+                if (daily)
+                   c("Date Local", "Event Type", "Sample Duration", "Arithmetic Mean")
+                else
+                   c("Date GMT", "Time GMT", "Sample Measurement"))))
         setnames(d, str_replace_all(names(d), " ", "."))
-        d[
+        if (daily) d[
             Sample.Duration %in% c("24 HOUR", "24-HR BLK AVG") &
                 Event.Type != "Excluded",
-            .(date = Date.Local, lon = Longitude, lat = Latitude,
-                value = Arithmetic.Mean)]}))
+            .(time = Date.Local,
+                lon = Longitude, lat = Latitude,
+                value = Arithmetic.Mean)]
+        else d[, .(
+            time = `-`(
[…]

### 2024-02-21 code/targets.R

@@ -137,14 +137,18 @@ list(
 
     # Compare AQS to our predictions
     tar_target(aqs.obs, format = "fst_dt", get.aqs.obs(
-        Wf$years, pred.grid)),
+        daily.sat(), Wf$years, pred.grid)),
     tar_target(pred.at.aqs.sites, format = "fst_dt", new.preds.compact(
-        dt.start = lubridate::as_datetime(min(aqs.obs$date) - 1),
-        dt.end = lubridate::as_datetime(max(aqs.obs$date) + 1),
+        dt.start = (if (daily.sat())
+            lubridate::as_datetime(min(aqs.obs$time) - 1) else
+            min(aqs.obs$time)),
+        dt.end = (if (daily.sat())
+            lubridate::as_datetime(max(aqs.obs$time) + 1) else
+            max(aqs.obs$time) + lubridate::hours(1)),
         cells = sort(unique(aqs.obs$cell)),
         targets = list(pred.grid, satellite.files, model.full))),
     tar_target(satellite.vs.aqs, get.satellite.vs.aqs(
-        pred.at.aqs.sites, aqs.obs)),
+        daily.sat(), pred.at.aqs.sites, aqs.obs)),
 
     # Data for maps
     tar_target(median.improve.map.data, get.median.improve.map.data(

### 2024-02-23 writing/paper_aodc.qmd
For AODC, provide an hourly breakdown of the AQS comparison.

--- 2024-02-09 writing/paper_aodc.qmd
@@
 ```
+
+## Comparison with AQS PM~2.5~
+
+```{r}
+#| label: tbl-aqs
+#| tbl-cap: Correlations of satellite values with hourly AQS PM~2.5~.
+flextable(pretty.table.numbers(tar_read(satellite.vs.aqs)))
+```

### 2024-02-23 code/paper_functions.R

@@ -132,15 +132,24 @@ get.satellite.vs.aqs = function(daily, satellite, aqs)
     # satellite value, more than one AQS value, or both. We consider
     # all combinations and weight them to sum to 1 per cell-time.
     message("Computing statistics")
-    d[, by = .(time, cell), weight := 1 / .N]
-    c(
-        list(
-            n.aqs.obs = d[, uniqueN(aqs.i)],
-            n.cell.times = uniqueN(d[, .(time, cell)])),
-        sapply(c("old", "new"), simplify = F, function(sv.type)
-            cov.wt(
-                d[, .(get(paste0("y.sat.", sv.type)), y.aqs)],
-                wt = d$weight, cor = T, method = "ML")$cor[1, 2]))}
+    f = \(d)
+       {d = copy(d)
+        d[, by = .(time, cell), weight := 1 / .N]
+        cbind(
+            data.table(
+                n.aqs.obs = d[, uniqueN(aqs.i)],
+                n.cell.times = uniqueN(d[, .(time, cell)])),
+            as.data.table(sapply(c("old", "new"), simplify = F, function(sv.type)
+                cov.wt(
+                    d[, .(get(paste0("y.sat.", sv.type)), y.aqs)],
+                    wt = d$weight, cor = T, method = "ML")$cor[1, 2])))}
+    if (daily)
+        f(d)
+    else
+        rbind(
+            cbind(hour = "all", f(d)),
+            d[, keyby = hour(lubridate::with_tz(time, "Etc/GMT+5")),
+                f(.SD), .SDcols = colnames(d)])}
 
 get.median.improve.map.data = function(
       y.sat.name, the.satellite, satellite.product, n.workers,

### 2024-02-26 code/paper_functions.R
Added an annual comparison with AQS for MCD19A2.

@@ -64,49 +64,61 @@ envelope.tester = \(obs, pred, threshold = 0.6)
 
 #' Get observations of PM_{2.5} from the Environmental Protection
 #' Agency's (EPA) Air Quality System. The unit is μg/m^3.
-#' `daily` is T for daily data or F for hourly data.
 #' File source: https://aqs.epa.gov/aqsweb/airdata/download_files.html#Daily
 #' Documentation: https://aqs.epa.gov/aqsweb/documents/about_aqs_data.html
-get.aqs.obs = function(daily, years, grid)
-   {aqs.url.root = "https://aqs.epa.gov/aqsweb/airdata"
+get.aqs.obs = function(time.unit, years, grid)
+   {assert(time.unit %in% c("hour", "day", "year"))
+    aqs.url.root = "https://aqs.epa.gov/aqsweb/airdata"
     parameter.code = 88101L
       # PM_{2.5} from a federally approved reference or equivalent
       # method (FRM/FEM).
+    min.days.for.annual = 100L
 
     assert(file.exists("code/openssl_workaround.conf"))
     Sys.setenv(OPENSSL_CONF = "code/openssl_workaround.conf")
     d = rbindlist(lapply(years, function(the.year)
        {message("AQS ", the.year)
-        fname = sprintf("%s_%d_%d.zip",
-            (if (daily) "daily" else "hourly"),
-            parameter.code,
-            the.year)
-        d = fread(
-            cmd = paste("unzip -p ", shQuote(download(
-                paste0(aqs.url.root, "/", fname),
-                file.path("aqs", fname)))),
-            select = c("Longitude", "Latitude", (
-                if (daily)
-                   c("Date Local", "Event Type", "Sample Duration", "Arithmetic Mean")
-                else
-                   c("Date GMT", "Time GMT", "Sample Measurement"))))
+        fname = switch(time.unit,
+            hour = sprintf("hourly_%d_%d.zip",
+                parameter.code,
+                the.year),
+            day = sprintf("daily_%d_%d.zip",
+                parameter.code,
+                the.year),
+            year = sprintf("annual_conc_by_monitor_%d.zip",
+                the.year))
+        d = fread(cmd = paste("unzip -p ", shQuote(download(
+            paste0(aqs.url.root, "/", fname),
+            file.path("aqs", fname)))))
         setnames(d, str_replace_all(names(d), " ", "."))
-        if (daily) d[
-            Sample.Duration %in% c("24 HOUR", "24-HR BLK AVG") &
[…]

### 2024-02-26 code/targets.R

@@ -137,7 +137,9 @@ list(
 
     # Compare AQS to our predictions
     tar_target(aqs.obs, format = "fst_dt", get.aqs.obs(
-        daily.sat(), Wf$years, pred.grid)),
+        (if (daily.sat()) "day" else "hour"),
+        Wf$years,
+        pred.grid)),
     tar_target(pred.at.aqs.sites, format = "fst_dt", new.preds.compact(
         dt.start = (if (daily.sat())
             lubridate::as_datetime(min(aqs.obs$time) - 1) else
@@ -148,7 +150,14 @@ list(
         cells = sort(unique(aqs.obs$cell)),
         targets = list(pred.grid, satellite.files, model.full))),
     tar_target(satellite.vs.aqs, get.satellite.vs.aqs(
-        daily.sat(), pred.at.aqs.sites, aqs.obs)),
+        (if (daily.sat()) "day" else "hour"),
+        pred.at.aqs.sites,
+        aqs.obs)),
+    if (Wf$satellite.product == "mcd19a2") list(
+        tar_target(aqs.obs.yearly, format = "fst_dt", get.aqs.obs(
+            "year", Wf$years, pred.grid)),
+        tar_target(satellite.vs.aqs.yearly, get.satellite.vs.aqs(
+            "year", pred.at.aqs.sites, aqs.obs.yearly))),
 
     # Data for maps
     tar_target(median.improve.map.data, get.median.improve.map.data(

### 2024-02-28 writing/paper_mcd19a2.qmd
Added an annual comparison with AQS for MCD19A2.
Put code etc. in the bind mount
so it can be updated without rebuilding the image.

@@
 IG = knitr::include_graphics
+knitr::opts_knit$set(root.dir = "/data/src/writing")
 ```
@@
 The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
+In a year-level comparison, against `r l = tar_read(satellite.vs.aqs.yearly); scales::comma(l$n.cell.times)` cell-years of AQS annual means, we found a correlation with the original satellite values of `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
 

### 2024-02-28 code/docker_Rprofile_extra.R

@@ -7,5 +7,6 @@ if (interactive())
     .Last <- function() try(savehistory("/data/R_history"))}
 
 if (file.exists("/data/R-packages-installed"))
-   {invisible(utils::capture.output({renv::load(); renv::restore(prompt = F)}))
+   {dir.create("renv", showWarnings = F)
+    invisible(utils::capture.output({renv::load(); renv::restore(prompt = F)}))
     source("code/globals.R")}

### 2024-02-28 code/globals.R

@@ -7,8 +7,6 @@ suppressPackageStartupMessages(
 
 data.dir = "/data"
 stopifnot(dir.exists(data.dir))
-writing.out.dir = file.path(data.dir, "writing")
-dir.create(writing.out.dir, showWarnings = F)
 intermediate.path = function(...)
    file.path(data.dir, 'intermediate', ...)
 dir.create(intermediate.path(), showWarnings = F)

### 2024-02-28 code/targets.R

@@ -130,10 +130,8 @@ list(
     # Summarize and report on the CV
     tar_target(cv.summary, get.cv.summary(
         cv$mDT_wPred)),
-    tarchetypes::tar_render(cv.report.render,
+    tarchetypes::tar_render(cv.report,
         'writing/cv_report.Rmd'),
-    tar_target(cv.report,
-        file.copy(cv.report.render[1], writing.out.dir, overwrite = T)),
 
     # Compare AQS to our predictions
     tar_target(aqs.obs, format = "fst_dt", get.aqs.obs(
@@ -199,7 +197,7 @@ list(
 
     # Manuscript graphics
     tar_target(ipath,
-       {idir = file.path(workflow.dir, "img")
+       {idir = file.path(getwd(), "writing/img")
         dir.create(idir, showWarnings = F)
         \(name) file.path(idir, paste0(name, ".png"))}),
     tar_file(agreement.plot.path, ggsave(
@@ -229,10 +227,8 @@ list(
         special.time.map$plot)),
 
     # Render the manuscript.
-    tarchetypes::tar_quarto(paper.render,
+    tarchetypes::tar_quarto(paper,
         sprintf('writing/paper_%s.qmd', Wf$satellite.product), quiet = F),
-    tar_target(paper,
-        file.copy(paper.render[1], writing.out.dir, overwrite = T)),
 
     # Do a supplementary analysis for AODC.
     if (Wf$satellite.product == "aodc") list(

### 2024-03-01 writing/paper_mcd19a2.qmd
Made the annual AQS comparison more sophisticated.

@@
 The correlation of the observations with the original satellite values was `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
-In a year-level comparison, against `r l = tar_read(satellite.vs.aqs.yearly); scales::comma(l$n.cell.times)` cell-years of AQS annual means, we found a correlation with the original satellite values of `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
+
+`r l = tar_read(satellite.vs.aqs.yearly); ""` In a year-level comparison, we took cell-years with at least `r min.days.for.annual.aqs` days of AQS observations and `r l$min.days.in.month` days per month of satellite observations for at least `r l$min.months` months.
+Obtaining `r scales::comma(l$n.cell.times)` cell-years, we compared official AQS annual means to the yearly means of satellite observations, computed with daily means weighted according to how many days of the year to which each day was closest.
+The result was a correlation with the original satellite values of `r sprintf("%.03f", l$old)`, compared to `r sprintf("%.03f", l$new)` with our corrected values.
 

### 2024-03-01 code/paper_functions.R

@@ -66,13 +66,13 @@ envelope.tester = \(obs, pred, threshold = 0.6)
 #' Agency's (EPA) Air Quality System. The unit is μg/m^3.
 #' File source: https://aqs.epa.gov/aqsweb/airdata/download_files.html#Daily
 #' Documentation: https://aqs.epa.gov/aqsweb/documents/about_aqs_data.html
+min.days.for.annual.aqs = 100L
 get.aqs.obs = function(time.unit, years, grid)
    {assert(time.unit %in% c("hour", "day", "year"))
     aqs.url.root = "https://aqs.epa.gov/aqsweb/airdata"
     parameter.code = 88101L
       # PM_{2.5} from a federally approved reference or equivalent
       # method (FRM/FEM).
-    min.days.for.annual = 100L
 
     assert(file.exists("code/openssl_workaround.conf"))
     Sys.setenv(OPENSSL_CONF = "code/openssl_workaround.conf")
@@ -113,7 +113,7 @@ get.aqs.obs = function(time.unit, years, grid)
             year = d[
                 Parameter.Code == parameter.code &
                     Pollutant.Standard == "PM25 Annual 2012" &
-                    Required.Day.Count >= min.days.for.annual &
+                    Required.Day.Count >= min.days.for.annual.aqs &
                     Completeness.Indicator == "Y",
                 .(
                     time = the.year,
@@ -166,23 +166,49 @@ get.satellite.vs.aqs = function(time.unit, satellite, aqs)
         f(d)}
 
 get.satellite.vs.aqs.yearly = function(sat, aqs)
-   {min.days.in.month = 10L
+   {tz = "Etc/GMT+6"
+    min.days.in.month = 10L
     min.months = 12L
 
     # Summarize the satellite predictions into year-cells.
     sat = sat[cell %in% unique(aqs$cell)]
-    sat[, date := lubridate::as_date(time.sat, tz = "Etc/GMT+6")]
-    sat = sat[, by = .(year(date), month(date), cell), if (uniqueN(date) >= min.days.in.month) .(
-        y.sat.old = mean(y.sat.old / 10^Wf$pred.round.digits),
-        y.sat.new = mean(y.sat.new / 10^Wf$pred.round.digits))]
-    sat = sat[, by = .(time = year, cell), if (uniqueN(month) >= min.months) .(
-        y.sat.old = mean(y.sat.old),
-        y.sat.new = mean(y.sat.new))]
[…]

### 2025-06-18 code/paper_functions.R
Rejiggered some plot-drawing targets.
Serialized ggplot2 objects were getting massive, like 22 G for `baltimore.map`.

@@ -319,28 +319,20 @@ pred.map = function(
              # For DQF interpretation, see PDF p. 334 of
              # http://web.archive.org/web/20230915201900if_/https://www.goes-r.gov/products/docs/PUG-L2+-vol5.pdf
 
-    list(
-        center.name = bg.sf[as.integer(st_intersects(
-            st_as_sf(
-                d[, .(
-                    mean(c(max(lon), min(lon))),
-                    mean(c(max(lat), min(lat))))],
-                coords = c(1, 2), crs = crs.lonlat),
-            st_transform(bg.sf, crs = crs.lonlat))),]$NAME,
-        plot = ggplot() +
-            geom_raster(aes(lon, lat, fill = value), data = d) +
-            scale_fill_distiller(name = color.scale.name,
-                palette = "Spectral", na.value = "transparent") +
-            geom_sf(data = bg.sf, fill = NA, size = .1) +
-            ggspatial::annotation_scale(data = cbind(
-                data.table(variable = "y.sat.old"),
-                (if (include.quality) list(qualities = "all qualities")))) +
-            facet_grid(
-                rows = vars(variable),
-                cols = (if (include.quality) vars(qualities)),
-                labeller = labeller(variable =
-                    c(y.sat.old = "Raw", y.sat.new = "Corrected"))) +
-            coord_sf(expand = F,
-                xlim = range(d$lon), ylim = range(d$lat)) +
-            theme_void() +
-            theme(strip.text = element_text(size = 13)))}
+    ggplot() +
+        geom_raster(aes(lon, lat, fill = value), data = d) +
+        scale_fill_distiller(name = color.scale.name,
+            palette = "Spectral", na.value = "transparent") +
+        geom_sf(data = bg.sf, fill = NA, size = .1) +
+        ggspatial::annotation_scale(data = cbind(
+            data.table(variable = "y.sat.old"),
+            (if (include.quality) list(qualities = "all qualities")))) +
+        facet_grid(
+            rows = vars(variable),
+            cols = (if (include.quality) vars(qualities)),
+            labeller = labeller(variable =
+                c(y.sat.old = "Raw", y.sat.new = "Corrected"))) +
+        coord_sf(expand = F,
+            xlim = range(d$lon), ylim = range(d$lat)) +
+        theme_void() +
+        theme(strip.text = element_text(size = 13))}

### 2025-06-18 code/targets.R

@@ -162,6 +162,16 @@ list(
     tar_target(median.improve.map.data, get.median.improve.map.data(
         Wf$y.sat, Wf$satellite, Wf$satellite.product, config$n.workers,
         cv$mDT_wPred, pred.grid, region.shape, satellite.files, model.full)),
+    tar_target(median.improve.map.center,
+        get_conus()[as.integer(st_intersects(
+            st_centroid(st_as_sfc(st_bbox(st_transform(crs = crs.lonlat,
+                st_as_sf(
+                     median.improve.map.data$pred[,
+                         .(sat.coords.x, sat.coords.y)],
+                     coords = c(1, 2),
+                     crs = st_crs(tar_read(pred.grid))))))),
+            st_transform(get_conus(), crs = crs.lonlat))),]),
+
     tar_target(baltimore.map.data, get.baltimore.map.data(
         pred.grid, region.shape, satellite.files, model.full)),
     tar_target(special.time.map.data,
@@ -185,16 +195,6 @@ list(
         new.preds(
             dt, dt, cells = cells,
             targets = list(pred.grid, satellite.files, model.full))}),
-    tar_target(median.improve.map, pred.map(
-        median.improve.map.data$pred, pred.grid,
-        bg.sf = get_conus(), color.scale.name = "AOD",
-        quantile.cap = .99)),
-    tar_target(baltimore.map, pred.map(
-        baltimore.map.data, pred.grid,
-        bg.sf = get_conus(), color.scale.name = "AOD")),
-    tar_target(special.time.map, pred.map(
-        special.time.map.data, pred.grid,
-        bg.sf = get_conus(), color.scale.name = "AOD")),
 
     # Manuscript graphics
     tar_target(ipath,
@@ -219,13 +219,20 @@ list(
                   panel.grid.minor.x = element_blank()))),
     tar_file(pred.map.median.improve.path, ggsave(
         ipath("pred_map_median_improve"), width = 7, height = 7,
-        median.improve.map$plot)),
+        pred.map(
+            median.improve.map.data$pred, pred.grid,
+            bg.sf = get_conus(), color.scale.name = "AOD",
+            quantile.cap = .99))),
     tar_file(pred.map.baltimore.path, ggsave(
         ipath("pred_map_baltimore"), width = 7, height = 7,
-        baltimore.map$plot)),
+        pred.map(
+            baltimore.map.data, pred.grid,
[…]

### 2025-08-05 writing/paper_mcd19a2.qmd
Rejiggered some plot-drawing targets.
Serialized ggplot2 objects were getting massive, like 22 G for `baltimore.map`.
Convert dead citations from grant proposal language to integrated Zotero citations

@@
 
-Many types of EO are validated using "gold standard" datasets obtained from networks of ground stations.[1--4](https://paperpile.com/c/ze3giJ/fWox+jvWM+EDSi+RF7S) When satellite observations are aligned with these ground measures in space and time, and the ground station data are considered top quality and are observations of the same phenomenon, then we can use the difference in these quantities as an operational measure of retrieval error (setting aside additional uncertainty in the measurements).
+Many types of EO are validated using "gold standard" datasets obtained from networks of ground stations.[@sayer2019a; @zhang2022; @jethva2019; @verhoelst2021] When satellite observations are aligned with these ground measures in space and time, and the ground station data are considered top quality and are observations of the same phenomenon, then we can use the difference in these quantities as an operational measure of retrieval error (setting aside additional uncertainty in the measurements).
 Making these EO-to-ground comparisons is a key task in product validation and is often conducted globally and for particular space-time regions or within the conditions of particular co-occurring phenomena.
-While various metrics for these validation comparisons include an assessment of mean bias and visualizations of mean differences,[5](https://paperpile.com/c/ze3giJ/vA4s) an overall summary of these differences across a collocated dataset is the Mean Square Error (MSE).
+While various metrics for these validation comparisons include an assessment of mean bias and visualizations of mean differences,[@giavarina2015] an overall summary of these differences across a collocated dataset is the Mean Square Error (MSE).
 The process of adjusting an EO dataset to account for systematic error, conditional on covariates, is also known as bias-correction and can lead to datasets that more accurately reflect the ground truth.
 Using data from the ground-stations to define our desired output, along with covariates from the same EO dataset that explain the retrieval error, makes this process "data-driven".
-The application of a leading supervised machine-learning approach, extreme gradient boosting (XGBoost),[6](https://paperpile.com/c/ze3giJ/72SB) which can capture complex interactions and outperforms related algorithms,[@just2018] has great empirical potential so long as care is taken to avoid overfitting (e.g., capturing phenomena that would not generalize to new data).[7--9](https://paperpile.com/c/ze3giJ/NejN+4wlV+gFoW)
+The application of a leading supervised machine-learning approach, extreme gradient boosting (XGBoost),[@chen2016] which can capture complex interactions and outperforms related algorithms, has great empirical potential so long as care is taken to avoid overfitting (e.g., capturing phenomena that would not generalize to new data).[@just2018c; @just2020.1c; @just2020c]
 
-We have previously demonstrated a closely-related data-driven analysis in the Northeast USA but lacked reproducible analytic code or a configurable and automated software tool.[7](https://paperpile.com/c/ze3giJ/NejN) Our XGBoost model empirically corrects biases related to complex interactions of acquisition/sensor and regional characteristics (e.g., aerosol composition or surface brightness) without reliance on any external data assimilation.
-Using an interpretability method, such as Shapley Additive Explanations (SHAP),[11](https://paperpile.com/c/ze3giJ/TgOX) to interpret our predictions, we see that our model accounts for complex interactions with a physical basis (e.g., challenging backscattering configurations).
-We also adapted our approach to refine MAIAC Column Water Vapor (CWV) achieving a 27% decrease in Terra CWV RMSE, with external validation at independent SuomiNet GPS receivers.[8](https://paperpile.com/c/ze3giJ/4wlV)
+We have previously demonstrated a closely-related data-driven analysis in the Northeast USA but lacked reproducible analytic code or a configurable and automated software tool.[@just2018b]Our XGBoost model empirically corrects biases related to complex interactions of acquisition/sensor and regional characteristics (e.g., aerosol composition or surface brightness) without reliance on any external data assimilation.
+Using an interpretability method, such as Shapley Additive Explanations (SHAP),[@lundberg2020] to interpret our predictions, we see that our model accounts for complex interactions with a physical basis (e.g., challenging backscattering configurations).
+We also adapted our approach to refine MAIAC Column Water Vapor (CWV) achieving a 27% decrease in Terra CWV RMSE, with external validation at independent SuomiNet GPS receivers.[@just2020d]
 
 Satellite-borne EO retrievals are physics-based and often validated with reliable ground networks.
-Building collocated datasets of EO over regions and time periods of interest is facilitated by both specialized tools for particular products (e.g., those from LP DAAC[14](https://paperpile.com/c/ze3giJ/KNHa) and ORNL DAAC[15](https://paperpile.com/c/ze3giJ/lQFH), although neither support MAIAC AOD) and structured access protocols such as OPeNDAP.[16](https://paperpile.com/c/ze3giJ/4m2u) When validating aerosol retrievals with ground-based observations, it is a common convention to spatially average retrieval products \-- even up to a 40 km \* 40 km area.[3](https://paperpile.com/c/ze3giJ/EDSi) This appears, at least in part, to be a legacy of 1) desiring to calculate local variance in products that were coarser (10 km resolution) than the more recent retrieval products like 1 km MAIAC and 2) a concern that accommodating temporal mismatches that were larger under prior AERONET sampling regimen (our analysis of AERONET 2018-2021 over CONUS finds an overall station-level median retrieval frequency of 5 minutes) in turn requires spatial averaging (since an aerosol air mass may move at 50 km/h or \~13.9 m/s).[17](https://paperpile.com/c/ze3giJ/TRsT) The consequence of these large-area averaged comparisons is that they do not directly apply to the performance of the product at the original resolution (e.g., 1 km for MAIAC).
+Building collocated datasets of EO over regions and time periods of interest is facilitated by both specialized tools for particular products (e.g., those from LP DAAC[@appeears]and ORNL DAAC[@daac2017], although neither support MAIAC AOD) and structured access protocols such as OPeNDAP.[@cornillon2003] When validating aerosol retrievals with ground-based observations, it is a common convention to spatially average retrieval products -- even up to a 40 km \* 40 km area.[@jethva2019a] This appears, at least in part, to be a legacy of 1) desiring to calculate local variance in products that were coarser (10 km resolution) than the more recent retrieval products like 1 km MAIAC and 2) a concern that accommodating temporal mismatches that were larger under prior AERONET sampling regimen (our analysis of AERONET 2018-2021 over CONUS finds an overall station-level median retrieval frequency of 5 minutes) in turn requires spatial averaging (since an aerosol air mass may move at 50 km/h or \~13.9 m/s).[@ichoku2002] The consequence of these large-area averaged comparisons is that they do not directly apply to the performance of the product at the original resolution (e.g., 1 km for MAIAC).
 Leveraging data-driven tools to validate high resolution (e.g., 1 km) products enables a time-efficient and cost-effective ML-based technology for historical and near real-time validation and bias correction.
-Thus a configurable tool also offers finer spatio-temporal domains (versus commonly global validation papers) with single pixel performance.
+Thus a configurable tool also offers finer spatiotemporal domains (versus commonly global validation papers) with single pixel performance.
 
-The GMAO MERRA-2[18](https://paperpile.com/c/ze3giJ/QSDA) uses a data-driven neural network retrieval (NNR) for AOD directly training on cloud-screened MODIS (as well as other sensors) observed radiances along with key modifiers of light scattering and retrieval error (e.g, view geometry, cloud fraction, and surface albedo) versus AERONET AOD.[19](https://paperpile.com/c/ze3giJ/1aAj) The NNR relies on the older MODIS AOD Collection 5 (C5) (while currently available is Collection 6.1 (C6.1)[20](https://paperpile.com/c/ze3giJ/oGky)), and does not build on the physics-based retrievals of MODIS operational products.
+The GMAO MERRA-2[@gelaro2017] uses a data-driven neural network retrieval (NNR) for AOD directly training on cloud-screened MODIS (as well as other sensors) observed radiances along with key modifiers of light scattering and retrieval error (e.g, view geometry, cloud fraction, and surface albedo) versus AERONET AOD.[@randles2017] The NNR relies on the older MODIS AOD Collection 5 (C5) (while currently available is Collection 6.1 (C6.1)[@wei2019]), and does not build on the physics-based retrievals of MODIS operational products.
 The NNR lacks interpretability in how it differs from physics-based retrievals, and ultimately offers no code or reproducible tool/approach that can be adapted for other remote sensing products.
@@
 
-ORNL's Fixed Sites Subsets Tool[21](https://paperpile.com/c/ze3giJ/ScfT) generates collocated land product datasets; it does not include MAIAC AOD.
+ORNL's Fixed Sites Subsets Tool[@terrestr] generates collocated land product datasets; it does not include MAIAC AOD.
 NASA also has several sophisticated graphical tools and powerful APIs ranging from Worldview to EarthData search that serve as visualization and data access platforms, however these generally do not handle collocation with ground measures or any retrieval-error correction.
-Several groups, including our own, have published on retrieval error/bias correction on Low Earth orbit (for MODIS[7,22,23](https://paperpile.com/c/ze3giJ/01P2+NejN+9sah)) and Geostationary satellite-based datasets (ABI[2](https://paperpile.com/c/ze3giJ/jvWM)) demonstrating data-driven approaches in the research stage, although our publication with CWV[8](https://paperpile.com/c/ze3giJ/4wlV) is the only one with public code sharing.[24](https://paperpile.com/c/ze3giJ/1rUv) To our knowledge, we are presenting the first tool that will enable collocation and improved output data in a reproducible and performant workflow.
+Several groups, including our own, have published on retrieval error/bias correction on Low Earth orbit (for MODIS[@just2018d; @lipponen2021; @albayrak2013]) and Geostationary satellite-based datasets (ABI[@zhang2020a]) demonstrating data-driven approaches in the research stage, although our publication with CWV[@just2020e] is the only one with public code sharing.
+To our knowledge, we are presenting the first tool that will enable collocation and improved output data in a reproducible and performant workflow.
 Although demonstrated for CONUS, we have developed flexible user parameterization for a reproducible workflow tool that can be deployed on any Linux system.
@@
 We also report the Pearson correlation of the raw and corrected products versus the collocated AERONET measures.
-We plotted the agreement of raw and corrected AOD versus AERONET AOD in scatterplots for the interval [0, 1] given that almost all AERONET observations in the cross-validation set fell within that range.
+We plotted the agreement of raw and corrected AOD versus AERONET AOD in scatterplots for the interval \[0, 1\] given that almost all AERONET observations in the cross-validation set fell within that range.
 To generate empirical error envelopes with performance for both low and high AOD conditions, we optimized the combination of additive and multiplicative terms closest to containing \~2/3 of observations for AERONET AOD below and above 0.6, a threshold previously used when evaluating MAIAC AOD.
@@
 ```
+
 @fig-cv-noaa-regions shows that performance improves in all regions and leads to greater consistency in the RMSE across CONUS with slightly higher RMSE remaining in the Northwest and West regions.
@@
 caption = sprintf("Maps of typical MAIAC and Corrected AOD. Day selected as having the median improvement in MSE at AERONET sites. The map is centered on %s.",
-    tar_read(median.improve.map)$center.name)
+    tar_read(median.improve.map.center)$NAME)
 IG(tar_read(pred.map.median.improve.path))
@@
 An interesting consequence of empirical tuning with our workflow is that addressing sensor-specific sources of retrieval error can lead to more harmonious datasets between platforms, a key goal for NASA's long-term Earth System Data Records.
-For example, we previously showed that MAIAC CWV from the aging Terra MODIS had worse agreement with AERONET CWV than that of Aqua MODIS, particularly in more recent years, yet the remaining retrieval error after our bias-correction for Terra and Aqua CWV was virtually the same.[8](https://paperpile.com/c/ze3giJ/4wlV) Thus, empirical correction of aerosol observations from, for example, the VIIRS instruments that continues the legacy of MODIS,[13](https://paperpile.com/c/ze3giJ/zr7Y) would be expected to result in greater consistency of the aerosols Earth System Data Records.
+For example, we previously showed that MAIAC CWV from the aging Terra MODIS had worse agreement with AERONET CWV than that of Aqua MODIS, particularly in more recent years, yet the remaining retrieval error after our bias-correction for Terra and Aqua CWV was virtually the same.[@just2020f] Thus, empirical correction of aerosol observations from, for example, the VIIRS instruments that continues the legacy of MODIS,[@sawyer2020] would be expected to result in greater consistency of the aerosols Earth System Data Records.
 Our workflow uses ancillary and derived quality control features for a data-driven refinement of retrievals to make EO more reliable and useful to a broader community of end users.
